### 2017-08-11 paper/aldebaran.tex
Paper skeleton added.

% !TEX TS-program = pdflatexmk

\documentclass[modern]{aastex61}
\usepackage[utf8]{inputenc}
\begin{document}

\title{Something About Ground Based Asteroseismology}

\author[0000-0003-1540-8562]{Will M. Farr}

\affiliation{Birmingham Institute for Gravitational Wave Astronomy,
  University of Birmingham, Birmingham, B15 2TT, United Kingdom}

\affiliation{School of Physics and Astronomy, University of
  Birmingham, Birmingham, B15 2TT, United Kingdom}

\author{Guy Davies}

\affiliation{School of Physics and Astronomy, University of
  Birmingham, Birmingham, B15 2TT, United Kingdom}

\author{Benjamin Pope}
\affiliation{Oxford or NYU or ??}

\author{Thomas North}
\affiliation{School of Physics and Astronomy, University of Birmingham, Birmingham, B15 2TT, United Kingdom}

\email{w.farr@bham.ac.uk, G.R.Davies@bham.ac.uk, Benjamin.Pope@physics.ox.ac.uk, txn016@student.bham.ac.uk}

\begin{abstract}
Stuff.
\end{abstract}

\section{Introduction}

\section{Time-Domain Models}

\section{Aldebaran}

\subsection{SONG Observations}

\subsection{Stellar Parameters}

\section{Conclusions}

\acknowledgments

We thank the academy....

\bibliography{aldebaran}

\end{document}

### 2017-09-06 paper/aldebaran.tex
Paper skeleton (author names, couple of references).

@@
 
-\author{Guy Davies}
+\author{Guy R. Davies}
 
@@
 
-\email{w.farr@bham.ac.uk, G.R.Davies@bham.ac.uk, Benjamin.Pope@physics.ox.ac.uk, txn016@student.bham.ac.uk}
+\author{James Barrett}
+\affiliation{Birmingham Institute for Gravitational Wave Astronomy,
+  University of Birmingham, Birmingham, B15 2TT, United Kingdom}
+
+\affiliation{School of Physics and Astronomy, University of
+  Birmingham, Birmingham, B15 2TT, United Kingdom}
+
+\email{w.farr@bham.ac.uk, G.R.Davies@bham.ac.uk, Benjamin.Pope@physics.ox.ac.uk, txn016@student.bham.ac.uk, jimbarrett27@gmail.com}
 
@@
 \section{Time-Domain Models}
+
+Reference \citet{Kelly2014}, compare to \citet{Foreman-Mackey2017}.
 

### 2017-09-07 paper/aldebaran.tex
placeholder figures; light curve; a bit of bibliography; rubbish first stab at getting some words on paper

@@
 \usepackage[utf8]{inputenc}
+\usepackage{hyperref}
+\usepackage{natbib}
+\usepackage{graphicx}
+\usepackage{xspace}
+
+\include{macros}
+% Journals
+
+% Astronomical abbreviations (thanks to Dan Huber)
+\newcommand{\numax}{\mbox{$\nu_{\rm max}$}\xspace}
+\newcommand{\Dnu}{\mbox{$\Delta \nu$}\xspace}
+\newcommand{\dnu}{\mbox{$\delta \nu$}\xspace}
+\newcommand{\muHz}{\mbox{$\mu$Hz}\xspace}
+\newcommand{\teff}{\mbox{$T_{\rm eff}$}\xspace}
+\newcommand{\logg}{\mbox{$\log g$}\xspace}
+\newcommand{\feh}{\mbox{$\rm{[Fe/H]}$}\xspace}
+\newcommand{\msun}{\mbox{$\mathrm{M}_{\odot}$}\xspace}
+\newcommand{\mearth}{\mbox{$\mathrm{M}_{\oplus}$}\xspace}
+\newcommand{\rsun}{\mbox{$\mathrm{R}_{\odot}$}\xspace}
+\newcommand{\kepler}{\emph{Kepler}\xspace}
+\newcommand{\hipparcos}{\emph{Hipparcos}\xspace}
+\newcommand{\gaia}{\emph{Gaia}\xspace}
+% \newcommand{\ktwo}{\textit{K2}\xspace}
+\newcommand{\ktwo}{\emph{K2}\xspace}
+\newcommand{\ksc}{{\sc k2sc}\xspace}
+\newcommand{\ksf}{{\sc k2sff}\xspace}
+\newcommand{\kms}{\,km\,s$^{-1}$} % kilometres per second
+\newcommand{\bibtex}{\textsc{Bib}\!\TeX} % bibtex. Not quite the correct typesetting, but close enough
+
 \begin{document}
@@
 
-\author{Benjamin Pope}
-\affiliation{Oxford or NYU or ??}
+\author[0000-0003-2595-9114]{Benjamin Pope}
+\affiliation{Sydney Institute for Astronomy, School of Physics, University of Sydney, Sydney NSW 2006, Australia} % for the moment - if it is a couple of months til this is published I should say NYU
 
@@
 
-\email{w.farr@bham.ac.uk, G.R.Davies@bham.ac.uk, Benjamin.Pope@physics.ox.ac.uk, txn016@student.bham.ac.uk, jimbarrett27@gmail.com}
+\email{w.farr@bham.ac.uk, G.R.Davies@bham.ac.uk, b.pope@sydney.edu.au, txn016@student.bham.ac.uk, jimbarrett27@gmail.com}
 
@@
 
+\subsection{K2 Observations}
+
+Aldebaran was observed \ktwo under Guest Observer Program 130471 in Campaign~13, from 2017-03-08 to 2017-05-27.
+The \kepler Space Telescope \citep{2010sci...327..977b} originally surveyed a region of the sky in the direction of Cygnus and Lyra, with the aim both to find transiting planets, and to study stellar interiors with asteroseismology.
+After 3.5 years of service, \kepler suffered a critical reaction wheel failure in May 2013, which made it impossible to maintain a stable pointing and therefore continue its nominal mission.
+It was revived as \ktwo, balancing the third axis uncontrolled by its reaction wheels by orienting carefully perpendicular to the Sun.
+This unusual observing strategy requires that \ktwo observes fields in the Ecliptic in $\sim~80$~d Campaigns.
+
+As Aldebaran is extremely bright, it saturates the \kepler detector and it is therefore not possible to use standard photometry pipelines to extract a K2 lightcurve.
+We therefore use halo photometry \citep[as originally implemented in]{White2017}, whereby unsaturated pixels from the outer part of the large and complicated halo of scattered light around bright stars are used to reconstruct a light curve.
+The brightness of this halo varies in the same way as that of the primary star, and we therefore construct our halo light curve $\mathbf{f}$ as a weighted linear combination of individual pixel-level time series $\mathbf{p}_j$, i.e.
+
+\begin{equation}
+	f_i = \sum_{j=1}^{M} w_j p_{ij}.
+\end{equation}
+
+\noindent We choose the weights $w_j$ such that they lie between 0~and~1, add to 1, and minimize the Total Variation (TV) of the weighted light curve.
+In the continuous case, $n$-th order TV is defined as the integral of the absolute value of the $n$-th derivative of a function; in the discrete case, replacing the derivative with finite differences, first-order TV becomes
+
+\begin{equation}
+\text{TV} = \dfrac{\sum_{i=1}^{N} |{f}_i - {f}_{i-1}|}{\sum_{j=1}^{N} {f}_j}
+\end{equation}
+
+\noindent and likewise second-order TV the equivalent expression in second-order finite differences.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.75\textwidth]{Aldebaran_lc.png}
+\caption{\ktwo lightcurve of Aldebaran.}
+\label{k2_lightcurve}
+\end{figure}
+
 \subsection{Stellar Parameters}

### 2017-09-07 paper/aldebaran.tex
more stuff

@@
 
+It is easy to observe Aldebaran and similarly-bright stars with ground-based spectroscopic instruments, typically only requiring short exposures that can be obtained even under adverse observing conditions.
+There is indeed a considerable archive of such observations already, as a legacy of radial velocity (RV) surveys conducted to find exoplanets.
+In most cases, however, these have not so far been useful for asteroseismology, because these RV data are sparsely and irregularly-sampled.
+Because we have to pause observations during the day, during poor weather conditions, or simply when targets of higher priority are being observed, we get time series of only a few points and which may have significant and uneven gaps.
+This introduces a window-function effect: the power spectrum as constructed for example by a Fourier transform, or a Lomb-Scargle Periodogram \citep{lomb,scargle}, is convolved with the Fourier transform of the window function, introducing strong sidelobes adjacent to real frequency peaks and causing crosstalk between adjacent frequency channels.
+This imposes significant limitations both on the signal-to-noise and frequency resolution of power spectra derived from linear methods such as the Lomb-Scargle periodogram, and in practice makes asteroseismology difficult or impossible from the ground for stars with oscillation frequencies ranging from $\sim 12$~h to $\sim$~a few days.
+
 Reference \citet{Kelly2014}, compare to \citet{Foreman-Mackey2017}.
@@
 
-Aldebaran was observed \ktwo under Guest Observer Program 130471 in Campaign~13, from 2017-03-08 to 2017-05-27.
-The \kepler Space Telescope \citep{2010sci...327..977b} originally surveyed a region of the sky in the direction of Cygnus and Lyra, with the aim both to find transiting planets, and to study stellar interiors with asteroseismology.
-After 3.5 years of service, \kepler suffered a critical reaction wheel failure in May 2013, which made it impossible to maintain a stable pointing and therefore continue its nominal mission.
-It was revived as \ktwo, balancing the third axis uncontrolled by its reaction wheels by orienting carefully perpendicular to the Sun.
+In order to verify the results of the novel analysis presented above, we sought to obtain an independent detection of the oscillations of Aldebaran and compare the frequencies determined with the two methods.
+
+Aldebaran was therefore observed with \ktwo under Guest Observer Program 130471 in Campaign~13, from 2017-03-08 to 2017-05-27.
+The \kepler Space Telescope \citep{2010sci...327..977b} %originally surveyed a region of the sky in the direction of Cygnus and Lyra, with the aim both to find transiting planets, and to study stellar interiors with asteroseismology.
+After 3.5 years of service, \kepler
+suffered a critical reaction wheel failure in May 2013, which made it impossible to maintain a stable pointing and therefore continue its nominal mission.
+It was revived as \ktwo \citep{howell14}, balancing the third axis uncontrolled by its reaction wheels by orienting carefully perpendicular to the Sun.
 This unusual observing strategy requires that \ktwo observes fields in the Ecliptic in $\sim~80$~d Campaigns.
@@
 We therefore use halo photometry \citep[as originally implemented in]{White2017}, whereby unsaturated pixels from the outer part of the large and complicated halo of scattered light around bright stars are used to reconstruct a light curve.
-The brightness of this halo varies in the same way as that of the primary star, and we therefore construct our halo light curve $\mathbf{f}$ as a weighted linear combination of individual pixel-level time series $\mathbf{p}_j$, i.e.
+The brightness of this halo varies in the same way as that of the primary star, and we therefore obtain data in a region of 20~pixel radius around the mean position of Aldebaran, and discard saturated pixels.
+We construct our halo light curve as a weighted linear combination of individual pixel-level time series, as described in Appendix~\ref{halo}.
+We then additionally correct the halo light curve using \textsc{k2sc} \citep{k2sc}, though we find that this has a minor effect.
+There is somewhat higher than usual residual noise at multiples of $4 c/d$ (the satellite thruster firing frequency), but this is nevertheless very small in comparison to the signal from Aldebaran, and may be ascribed to the large fraction of the pixel mask occupied by the bleed column from this extremely bright star.
 
-\begin{equation}
-	f_i = \sum_{j=1}^{M} w_j p_{ij}.
-\end{equation}
-
-\noindent We choose the weights $w_j$ such that they lie between 0~and~1, add to 1, and minimize the Total Variation (TV) of the weighted light curve.
-In the continuous case, $n$-th order TV is defined as the integral of the absolute value of the $n$-th derivative of a function; in the discrete case, replacing the derivative with finite differences, first-order TV becomes
-
-\begin{equation}
-\text{TV} = \dfrac{\sum_{i=1}^{N} |{f}_i - {f}_{i-1}|}{\sum_{j=1}^{N} {f}_j}
-\end{equation}
-
-\noindent and likewise second-order TV the equivalent expression in second-order finite differences.
+Using the \textsc{k2ps} planet-search code \citep{k2ps,Pope2016} to examine this light curve, we search for transits across a wide range of periods, and find no evidence either of planetary transits or an eclipsing stellar companion.
 
@@
 
+\appendix
+
+\section{Halo Photometry}
+\label{halo}
+
+The halo photometry method proceeds as in \citet{White2017}. \textcolor{red}{More text here to explain in words.}
+The flux $f_i$ at each cadence $i$ is constructed as a weighted sum of pixel values $p_{ij}$:
+
+\begin{equation}
+	f_i = \sum_{j=1}^{M} w_j p_{ij}.
+\end{equation}
+
+\noindent We choose the weights $w_j$ such that they lie between 0~and~1, add to 1, and minimize the Total Variation (TV) of the weighted light curve.
+In the continuous case, $n$-th order TV is defined as the integral of the absolute value of the $n$-th derivative of a function; in the discrete case, replacing the derivative with finite differences, first-order TV becomes
+
+\begin{equation}
+\text{TV} = \dfrac{\sum_{i=1}^{N} |{f}_i - {f}_{i-1}|}{\sum_{j=1}^{N} {f}_j}
+\end{equation}
+
+\noindent and likewise second-order TV the equivalent expression in second-order finite differences.
+
+In an improvement since \citet{White2017}, we use the \textsc{Theano} library \citep{theano} to calculate analytic derivatives for this objective function, which reduces the computational time for a single halo light curve on a commercial laptop from tens of minutes to tens of seconds.
+
 \bibliography{aldebaran}

### 2017-09-11 paper/aldebaran.tex
small changes before I go to bed
fix maths
acknowledgements stub

@@
 \usepackage{xspace}
+\usepackage{amsmath}
+\usepackage{amssymb}
 
@@
 
+If we apply nonlinear statistical inference methods...
+
 Reference \citet{Kelly2014}, compare to \citet{Foreman-Mackey2017}.
@@
 
+This research has made use of the SIMBAD database, operated at CDS, Strasbourg, France, and NASA's Astrophysics Data System.
+The \kepler data presented in this paper were obtained from the Mikulski Archive for Space Telescopes (MAST).
+STScI is operated by the Association of Universities for Research in Astronomy, Inc., under NASA contract NAS5-26555.
+Support for MAST for non-HST data is provided by the NASA Office of Space Science via grant NNX13AC07G and by other grants and contracts.
+This work made use of the IPython package \citep{PER-GRA:2007} This research made use of SciPy \citep{jones_scipy_2001};  \textsc{matplotlib}, a \textsc{Python} library for publication quality graphics \citep{Hunter:2007}; \textsc{Astropy}, a community-developed core \textsc{Python} package for astronomy \citep{2013A&A...558A..33A}; and \textsc{NumPy} \citep{van2011numpy}.
+Some of the acknowledgements were compiled using the Astronomy Acknowledgement Generator.
+
 \appendix
@@
 
-The halo photometry method proceeds as in \citet{White2017}. \textcolor{red}{More text here to explain in words.}
+The halo photometry method proceeds as in \citet{White2017}, with only minor changes. \textcolor{red}{More text here to explain in words.}
 The flux $f_i$ at each cadence $i$ is constructed as a weighted sum of pixel values $p_{ij}$:

### 2017-09-11 paper/aldebaran.tex
minor fix

@@
 Support for MAST for non-HST data is provided by the NASA Office of Space Science via grant NNX13AC07G and by other grants and contracts.
-This work made use of the IPython package \citep{PER-GRA:2007} This research made use of SciPy \citep{jones_scipy_2001};  \textsc{matplotlib}, a \textsc{Python} library for publication quality graphics \citep{Hunter:2007}; \textsc{Astropy}, a community-developed core \textsc{Python} package for astronomy \citep{2013A&A...558A..33A}; and \textsc{NumPy} \citep{van2011numpy}.
+This work made use of the \textsc{IPython} package \citep{PER-GRA:2007}; SciPy \citep{jones_scipy_2001};  \textsc{matplotlib}, a \textsc{Python} library for publication quality graphics \citep{Hunter:2007}; \textsc{Astropy}, a community-developed core \textsc{Python} package for astronomy \citep{2013A&A...558A..33A}; and \textsc{NumPy} \citep{van2011numpy}.
 Some of the acknowledgements were compiled using the Astronomy Acknowledgement Generator.

### 2017-09-18 paper/aldebaran.tex
full appendix on halo photometry

@@
 
-Aldebaran was therefore observed with \ktwo under Guest Observer Program 130471 in Campaign~13, from 2017-03-08 to 2017-05-27.
-The \kepler Space Telescope \citep{2010sci...327..977b} %originally surveyed a region of the sky in the direction of Cygnus and Lyra, with the aim both to find transiting planets, and to study stellar interiors with asteroseismology.
-After 3.5 years of service, \kepler
+Aldebaran was therefore observed with \ktwo under Guest Observer Program 130471 in Campaign~13, from 2017-03-08 to 2017-05-27. The \kepler Space Telescope \citep{2010sci...327..977b}  
 suffered a critical reaction wheel failure in May 2013, which made it impossible to maintain a stable pointing and therefore continue its nominal mission.
-It was revived as \ktwo \citep{howell14}, balancing the third axis uncontrolled by its reaction wheels by orienting carefully perpendicular to the Sun.
-This unusual observing strategy requires that \ktwo observes fields in the Ecliptic in $\sim~80$~d Campaigns.
+It was revived as \ktwo \citep{howell14}, balanced by orienting perpendicular to the Sun.
+This requires that \ktwo observes fields in the Ecliptic in $\sim~80$~d Campaigns.
 
@@
 
+BP is grateful for funding from the Breakthrough Foundation. % possibly update to say Sagan
+
 This research has made use of the SIMBAD database, operated at CDS, Strasbourg, France, and NASA's Astrophysics Data System.
@@
 
-The halo photometry method proceeds as in \citet{White2017}, with only minor changes. \textcolor{red}{More text here to explain in words.}
+The \kepler detector saturates for stars brighter than the $\sim~11^\text{th}$ magnitude.
+Nevertheless, the excess flux deposited in a saturated pixel spills conservatively up and down the pixel column, such that it is possible to sum this `bleed column' for bright stars and still obtain precise photometry, such as was done for the brightest star in the nominal \kepler mission, $\theta$~Cyg \citep{} ($V = 4.48$).
+There are two main reasons why this is not possible for all bright stars in general.
+First, because the downlink bandwidth from \kepler is very limited, it is often not desirable to download the large pixel stamps as are required for such bright stars; these must be even larger in \ktwo to account for the reduced image stability.
+Second, if the bleed column for a sufficiently bright star reaches the edge of the chip, flux spills over and is not conserved, imposing a hard brightness limit that depends on the distance to the detector edge.
+Collateral `smear' data, which are collected to help calibrate the photometric bias from stars sharing the same column as a target, can be used to reconstruct light curves for un-downloaded bright stars and thereby avoid bandwidth constraints \citep{k2smear}, but these data are still rendered unusable if the bleed column falls off the edge of the chip and contaminates the smear rows.
+
+Bright stars have a wide, complicated, position-dependent PSF arising from diffraction and scattering from secondary and higher-order reflections inside the instrument, with the result that they may contaminate thousands of nearby pixels with significant flux.
+We can therefore use this `halo' of unsaturated pixels for photometry.
+In this paper we proceed as in \citet{White2017}, with only minor changes.
+
 The flux $f_i$ at each cadence $i$ is constructed as a weighted sum of pixel values $p_{ij}$:
@@
 
-\noindent We choose the weights $w_j$ such that they lie between 0~and~1, add to 1, and minimize the Total Variation (TV) of the weighted light curve.
+\noindent We choose the weights $w_j$ such that they lie between 0~and~1, add to unity, and minimize the Total Variation (TV) of the weighted light curve.
 In the continuous case, $n$-th order TV is defined as the integral of the absolute value of the $n$-th derivative of a function; in the discrete case, replacing the derivative with finite differences, first-order TV becomes
@@
 
+As a final step to reduce residual uncorrected systematics, we apply the \textsc{k2sc} \citep{k2sc} Gaussian Process-based systematics correction code to the initial halo light curve, but the effect of this in the present case for Aldebaran is minimal.
+
 \bibliography{aldebaran}

### 2017-09-18 paper/aldebaran.tex
theta cyg

@@
 The \kepler detector saturates for stars brighter than the $\sim~11^\text{th}$ magnitude.
-Nevertheless, the excess flux deposited in a saturated pixel spills conservatively up and down the pixel column, such that it is possible to sum this `bleed column' for bright stars and still obtain precise photometry, such as was done for the brightest star in the nominal \kepler mission, $\theta$~Cyg \citep{} ($V = 4.48$).
+Nevertheless, the excess flux deposited in a saturated pixel spills conservatively up and down the pixel column, such that it is possible to sum this `bleed column' for bright stars and still obtain precise photometry, such as was done for the brightest star in the nominal \kepler mission, $\theta$~Cyg \citep{guzik2011,thetacygwhite,guzik2016} ($V = 4.48$).
 There are two main reasons why this is not possible for all bright stars in general.

### 2017-09-18 paper/aldebaran.tex
text
psf

@@
 
-Bright stars have a wide, complicated, position-dependent PSF arising from diffraction and scattering from secondary and higher-order reflections inside the instrument, with the result that they may contaminate thousands of nearby pixels with significant flux.
+Bright stars have a wide, complicated, position-dependent point spread function (PSF) arising from diffraction and scattering from secondary and higher-order reflections inside the instrument, with the result that they may contaminate thousands of nearby pixels with significant flux.
 We can therefore use this `halo' of unsaturated pixels for photometry.
-In this paper we proceed as in \citet{White2017}, with only minor changes.
+In this paper we proceed as in \citet{White2017}, in which the method was demonstrated on the seven bright Pleiades, with only minor changes.
 

### 2017-09-18 paper/aldebaran.tex
breakthrough
pope affiliations

@@
 \affiliation{Sydney Institute for Astronomy, School of Physics, University of Sydney, Sydney NSW 2006, Australia} % for the moment - if it is a couple of months til this is published I should say NYU
+\affiliation{Center for Cosmology and Particle Physics, Department of Physics, New York University, 4 Washington Place, New York, NY 10003, USA}
+\affiliation{NASA Sagan Fellow}
 
@@
 
-BP is grateful for funding from the Breakthrough Foundation. % possibly update to say Sagan
+BP is grateful for funding from the Breakthrough Prize Foundation.
+This work was performed in part under contract with the Jet Propulsion Laboratory (JPL) funded by NASA through the Sagan Fellowship Program executed by the NASA Exoplanet Science Institute.
+ % possibly update to say Sagan
 

### 2017-09-18 paper/aldebaran.tex
figure, affiliations, text

@@
 
-\author[0000-0003-2595-9114]{Benjamin Pope}
+\author[0000-0003-2595-9114]{Benjamin J. S. Pope}
 \affiliation{Sydney Institute for Astronomy, School of Physics, University of Sydney, Sydney NSW 2006, Australia} % for the moment - if it is a couple of months til this is published I should say NYU
@@
 Support for MAST for non-HST data is provided by the NASA Office of Space Science via grant NNX13AC07G and by other grants and contracts.
-This work made use of the \textsc{IPython} package \citep{PER-GRA:2007}; SciPy \citep{jones_scipy_2001};  \textsc{matplotlib}, a \textsc{Python} library for publication quality graphics \citep{Hunter:2007}; \textsc{Astropy}, a community-developed core \textsc{Python} package for astronomy \citep{2013A&A...558A..33A}; and \textsc{NumPy} \citep{van2011numpy}.
+This work made use of the \textsc{IPython} package \citep{PER-GRA:2007}; SciPy \citep{scipy};  \textsc{matplotlib}, a \textsc{Python} library for publication quality graphics \citep{Hunter:2007}; \textsc{Astropy}, a community-developed core \textsc{Python} package for astronomy \citep{2013A&A...558A..33A}; and \textsc{NumPy} \citep{van2011numpy}.
 Some of the acknowledgements were compiled using the Astronomy Acknowledgement Generator.

### 2017-09-18 paper/aldebaran.tex
use eps
comments

@@
 \author[0000-0003-2595-9114]{Benjamin J. S. Pope}
-\affiliation{Sydney Institute for Astronomy, School of Physics, University of Sydney, Sydney NSW 2006, Australia} % for the moment - if it is a couple of months til this is published I should say NYU
+\affiliation{Sydney Institute for Astronomy, School of Physics, University of Sydney, Sydney NSW 2006, Australia} 
 \affiliation{Center for Cosmology and Particle Physics, Department of Physics, New York University, 4 Washington Place, New York, NY 10003, USA}
@@
 \centering
-\includegraphics[width=0.75\textwidth]{Aldebaran_lc.png}
+% \includegraphics[width=0.75\textwidth]{Aldebaran_lc.png}
+\includegraphics[width=0.75\textwidth]{Aldebaran_lc.eps}
 \caption{\ktwo lightcurve of Aldebaran.}
@@
 This work was performed in part under contract with the Jet Propulsion Laboratory (JPL) funded by NASA through the Sagan Fellowship Program executed by the NASA Exoplanet Science Institute.
- % possibly update to say Sagan
 

### 2017-09-20 paper/aldebaran.tex
candidate title
some references

@@
 
-\title{Something About Ground Based Asteroseismology}
+\title{Detection of Oscillations in Aldebaran with Ground-Based Observations}
 
@@
 
+Aldebaran, or $\alpha$~Tauri, is one of the brightest and most easily recognized naked-eye stars, and has long been the subject of astronomical investigations.
+
 \section{Time-Domain Models}
@@
 \section{Aldebaran}
+Aldebaran is one of the nearest red giant stars, at a distance of only $19.96 \pm 0.38$~pc \citep{hipparcos}, and its position near the Ecliptic permits the determination of its angular diameter by lunar occultations \citep[$20.58 \pm 0.03$ mas;][]{richichi2005,1979ApJ...228L.111B,brown1979,panek1980}.
 

### 2017-09-20 paper/aldebaran.tex
history

--- 2017-09-18 paper/aldebaran.tex
@@
 
+Aldebaran, or $\alpha$~Tauri, is one of the brightest and most easily recognized naked-eye stars, and has long been the subject of astronomical investigations.
+Following the proposal of \citet{struverv} several groups began to search for planets by the radial velocity (RV) method, looking for Doppler shifts from the star's reflex motion around the common centre of mass, culminating in the discovery of 51~Peg~b \citep{51peg}, the first exoplanet to be recognized as such.
+Before this, \citet{hatzes1993} noted RV variations in Pollux \citep[$\beta$~Gem; subsequently confirmed as a planet:][]{betgemconf,betgemconf2}, Arcturus ($\alpha$~Boo, unconfirmed), and Aldebaran.
+
 \section{Time-Domain Models}
@@
 \section{Aldebaran}
+Aldebaran is one of the nearest red giant stars, at a distance of only $19.96 \pm 0.38$~pc \citep{hipparcos}, and its position near the Ecliptic permits the determination of its angular diameter by lunar occultations and by interferometry \citep[$20.58 \pm 0.03$ mas;][]{richichi2005,1979ApJ...228L.111B,brown1979,panek1980}.
 

### 2017-09-20 paper/aldebaran.tex
intro

@@
 
-Aldebaran, or $\alpha$~Tauri, is one of the brightest and most easily recognized naked-eye stars, and has long been the subject of astronomical investigations.
+Aldebaran, or $\alpha$~Tauri, is a well-known first-magnitude naked-eye red giant star, and has long been the subject of astronomical investigations.
 Following the proposal of \citet{struverv} several groups began to search for planets by the radial velocity (RV) method, looking for Doppler shifts from the star's reflex motion around the common centre of mass, culminating in the discovery of 51~Peg~b \citep{51peg}, the first exoplanet to be recognized as such.
 Before this, \citet{hatzes1993} noted RV variations in Pollux \citep[$\beta$~Gem; subsequently confirmed as a planet:][]{betgemconf,betgemconf2}, Arcturus ($\alpha$~Boo, unconfirmed), and Aldebaran.
+After further investigation by \citet{Hatzes1998}, \citet{Hatzes2015} now claim a firm RV detection of a planetary-mass companion Aldebaran~b, with a period of $628.96 \pm 0.90$~d.
+
+In this paper, we present a re-analysis of these original RV data in which we not only confirm this signal, but detect acoustic oscillations in Aldebaran for the first time.
+We validate this method and its result with new RV observations with the SONG~Telescope, and photometry from the \ktwo Mission.
+By measuring the average frequency \numax of these \emph{p}-mode oscillations we asteroseismically determine the mass of Aldebaran to be $1.17 \pm 0.05$~\msun.
+This measured stellar mass allows us to calculate that any moons of this giant planet, although they are now likely to be very hot, may have had equilibrium temperatures comparable to that of the Earth when Aldebaran was on the main sequence, raising the possibility that tehy may have once been habitable.
+
+Our new approach to asteroseismic data analysis, based on Continuous Auto-Regressive Moving Average (CARMA) models, can extract exoplanet signals together with measures of \numax from sparse and irregularly-sampled time series.
+An all-sky survey planetary companions to precisely measure the masses of all nearby red giant stars is feasible with this new approach, and the required data either already exist in large exoplanet surveys, or are easy to obtain with ground-based telescopes.
 
@@
 \section{Aldebaran}
-Aldebaran is one of the nearest red giant stars, at a distance of only $19.96 \pm 0.38$~pc \citep{hipparcos}, and its position near the Ecliptic permits the determination of its angular diameter by lunar occultations and by interferometry \citep[$20.58 \pm 0.03$ mas;][]{richichi2005,1979ApJ...228L.111B,brown1979,panek1980}.
+Aldebaran is a red giant star with spectral type K5, one of the nearest such stars at a distance of only $19.96 \pm 0.38$~pc \citep{hipparcos}.
+Its position near the Ecliptic permits the determination of its angular diameter by lunar occultations and by interferometry \citep[$20.58 \pm 0.03$ mas;][]{richichi2005,1979ApJ...228L.111B,brown1979,panek1980}.
 

### 2017-09-20 paper/aldebaran.tex
wording

@@
 
-Using the \textsc{k2ps} planet-search code \citep{k2ps,Pope2016} to examine this light curve, we search for transits across a wide range of periods, and find no evidence either of planetary transits or an eclipsing stellar companion.
+Using the \textsc{k2ps} planet-search code \citep{k2ps,Pope2016} to examine this light curve, we search for transits across a wide range of periods, and find no evidence either of short-period planetary transits or an eclipsing stellar companion.
 

### 2017-09-20 paper/aldebaran.tex
move most of k2 to the appendix

@@
 
-Aldebaran was therefore observed with \ktwo under Guest Observer Program 130471 in Campaign~13, from 2017-03-08 to 2017-05-27. The \kepler Space Telescope \citep{2010sci...327..977b}  
-suffered a critical reaction wheel failure in May 2013, which made it impossible to maintain a stable pointing and therefore continue its nominal mission.
-It was revived as \ktwo \citep{howell14}, balanced by orienting perpendicular to the Sun.
-This requires that \ktwo observes fields in the Ecliptic in $\sim~80$~d Campaigns.
+Aldebaran was therefore observed with the \ktwo Mission \citep{howell14}, a two-wheeled revival of the \kepler Mission \citep{2010sci...327..977b}, under Guest Observer Program 130471 in Campaign~13, from 2017-03-08 to 2017-05-27.
 
 As Aldebaran is extremely bright, it saturates the \kepler detector and it is therefore not possible to use standard photometry pipelines to extract a K2 lightcurve.
-We therefore use halo photometry \citep[as originally implemented in]{White2017}, whereby unsaturated pixels from the outer part of the large and complicated halo of scattered light around bright stars are used to reconstruct a light curve.
-The brightness of this halo varies in the same way as that of the primary star, and we therefore obtain data in a region of 20~pixel radius around the mean position of Aldebaran, and discard saturated pixels.
-We construct our halo light curve as a weighted linear combination of individual pixel-level time series, as described in Appendix~\ref{halo}.
-We then additionally correct the halo light curve using \textsc{k2sc} \citep{k2sc}, though we find that this has a minor effect.
-There is somewhat higher than usual residual noise at multiples of $4 c/d$ (the satellite thruster firing frequency), but this is nevertheless very small in comparison to the signal from Aldebaran, and may be ascribed to the large fraction of the pixel mask occupied by the bleed column from this extremely bright star.
+We instead use halo photometry \citep[as originally implemented in][]{White2017}, whereby unsaturated pixels from the outer part of the large and complicated halo of scattered light around bright stars are used to reconstruct a light curve as a weighted linear combination of individual pixel-level time series, as described in Appendix~\ref{halo}.
 
@@
 
+The \kepler Space Telescope \citep{2010sci...327..977b}  
+suffered a critical reaction wheel failure in May 2013, which made it impossible to maintain a stable pointing and therefore continue its nominal mission.
+It was revived as \ktwo \citep{howell14}, balanced by orienting perpendicular to the Sun.
+This requires that \ktwo observes fields in the Ecliptic in $\sim~80$~d Campaigns.
+
 The \kepler detector saturates for stars brighter than the $\sim~11^\text{th}$ magnitude.
@@
 We can therefore use this `halo' of unsaturated pixels for photometry.
+The brightness of this halo varies in the same way as that of the primary star, and we therefore obtain data in a region of 20~pixel radius around the mean position of Aldebaran, and discard saturated pixels.
 In this paper we proceed as in \citet{White2017}, in which the method was demonstrated on the seven bright Pleiades, with only minor changes.
@@
 As a final step to reduce residual uncorrected systematics, we apply the \textsc{k2sc} \citep{k2sc} Gaussian Process-based systematics correction code to the initial halo light curve, but the effect of this in the present case for Aldebaran is minimal.
+There is somewhat higher than usual residual noise at multiples of $4 c/d$ (the satellite thruster firing frequency), but this is nevertheless very small in comparison to the signal from Aldebaran, and may be ascribed to the large fraction of the pixel mask occupied by the bleed column from this extremely bright star.
 

### 2017-09-21 paper/aldebaran.tex
habitability stub
habitability

@@
 
+We note that with the determination of the stellar mass it is possible to infer the parameters Aldebaran had while it was on the main sequence.
+We conduct a Monte Carlo simulation, drawing masses randomly from a Gaussian distribution $M \sim \mathscr{N}(1.17,0.05)$ to predict the luminosity of the main-sequence progenitor, and the semi-major axis of the planet's orbit, using the scaling relations of \citep{eker2015}, finding that it would have had a luminosity $L_* = 2.0 \pm 0.37~\lsun$ on the main sequence.
+
+We note that this implies that the planet at a semimajor axis of $1.512 \pm 0.021$~AU would have been subject to an insolation of $0.87 \pm 0.14$ times that of the Earth.
+Subject as well as this to the great uncertainties of the planet's orbital evolution and albedo, this planet and any of its moons may well have had temperate environments now long-since destroyed by their host star's evolution away from the main sequence.
+
 \section{Conclusions}

### 2017-09-22 paper/aldebaran.tex
mesa models

--- 2017-09-20 paper/aldebaran.tex
@@
 \newcommand{\msun}{\mbox{$\mathrm{M}_{\odot}$}\xspace}
+\newcommand{\lsun}{\mbox{$\mathrm{L}_{\odot}$}\xspace}
 \newcommand{\mearth}{\mbox{$\mathrm{M}_{\oplus}$}\xspace}
@@
 
+We note that with the determination of the stellar mass it is possible to infer the parameters Aldebaran had while it was on the main sequence.
+We conduct a Monte Carlo simulation, drawing masses randomly from a Gaussian distribution $M \sim \mathscr{N}(1.17,0.05)$ to predict the luminosity of the main-sequence progenitor, and the semi-major axis of the planet's orbit.
+%Using the scaling relations of \citep{eker2015}, we find that Aldebaran would have had a luminosity $L_* = 2.0 \pm 0.37~\lsun$ on the main sequence.
+Using the Mesa Isochrones and Stellar Tracks \citep[MIST;][]{mist0,mist1}, models, we compute the luminosity and radius evolution of the planet along the main sequence with uncertainty both in stellar mass (this work) and metallicity \citep[$-0.15 \pm 0.2$; ][]{decin2003}.
+$L$ evolves from $2.03 \pm 0.56 \lsun$ at 0.5~Gyr to $2.74 \pm 0.93 \lsun$ at 2.5~Gyr.
+
+We note that this implies that the planet at a semimajor axis of $1.512 \pm 0.021$~AU would have been subject to an insolation comparable to that of the Earth, evolving from 0.5~Gyr to 2.5~Gyr from $0.89 \pm 0.22$ $1.19 \pm 0.37$ times that of the Earth today.
+Subject as well as this to the great uncertainties of the planet's orbital evolution and albedo, this planet and any of its moons may well have had temperate environments now long-since destroyed by their host star's evolution away from the main sequence.
+
 \section{Conclusions}

### 2017-09-22 paper/aldebaran.tex
mathscr
typos

@@
 \usepackage{amssymb}
+\usepackage{mathrsfs}
 
@@
 %Using the scaling relations of \citep{eker2015}, we find that Aldebaran would have had a luminosity $L_* = 2.0 \pm 0.37~\lsun$ on the main sequence.
-Using the Mesa Isochrones and Stellar Tracks \citep[MIST;][]{mist0,mist1}, models, we compute the luminosity and radius evolution of the planet along the main sequence with uncertainty both in stellar mass (this work) and metallicity \citep[$-0.15 \pm 0.2$; ][]{decin2003}.
-$L$ evolves from $2.03 \pm 0.56 \lsun$ at 0.5~Gyr to $2.74 \pm 0.93 \lsun$ at 2.5~Gyr.
+Using the Mesa Isochrones and Stellar Tracks \citep[MIST:][]{mist0,mist1}, models, we compute the luminosity and radius evolution of the planet along the main sequence with uncertainty both in stellar mass (this work) and metallicity \citep[$-0.15 \pm 0.2$; ][]{decin2003}.
+$L$ evolves from $2.03 \pm 0.56$~\lsun at 0.5~Gyr to $2.74 \pm 0.93$~\lsun at 2.5~Gyr.
 
-We note that this implies that the planet at a semimajor axis of $1.512 \pm 0.021$~AU would have been subject to an insolation comparable to that of the Earth, evolving from 0.5~Gyr to 2.5~Gyr from $0.89 \pm 0.22$ $1.19 \pm 0.37$ times that of the Earth today.
+We note that this implies that the planet at a semimajor axis of $1.512 \pm 0.021$~AU would have been subject to an insolation comparable to that of the Earth, evolving from 0.5~Gyr to 2.5~Gyr from $0.89 \pm 0.22$ to $1.19 \pm 0.37$ times that of the Earth today.
 Subject as well as this to the great uncertainties of the planet's orbital evolution and albedo, this planet and any of its moons may well have had temperate environments now long-since destroyed by their host star's evolution away from the main sequence.

### 2017-09-22 paper/aldebaran.tex
Added Seismic text

@@
 
+\newcommand{\guy}[1]{{\bf \color{blue} #1}}
+
 \begin{document}
@@
 
-\author{Guy R. Davies}
+\author[0000-0002-4290-7351]{Guy R. Davies}
 
@@
   Birmingham, Birmingham, B15 2TT, United Kingdom}
+\affiliation{Stellar Astrophysics Centre, Department of Physics 
+and Astronomy, Aarhus University, Ny Munkegade 120, DK-8000 Aarhus C, Denmark}
 
@@
 
+\author{Tim White}
+\affiliation{Stellar Astrophysics Centre, Department of Physics 
+and Astronomy, Aarhus University, Ny Munkegade 120, DK-8000 Aarhus C, Denmark}
+
 \author{James Barrett}
@@
 
+\affiliation{School of Physics and Astronomy, University of
+  Birmingham, Birmingham, B15 2TT, United Kingdom}
+
+\author{Andrea Miglio}
+
+\affiliation{School of Physics and Astronomy, University of
+  Birmingham, Birmingham, B15 2TT, United Kingdom}
+\affiliation{Stellar Astrophysics Centre, Department of Physics 
+and Astronomy, Aarhus University, Ny Munkegade 120, DK-8000 Aarhus C, Denmark}
+
+\author{Mikkel N. Lund}
+
+\affiliation{Stellar Astrophysics Centre, Department of Physics 
+and Astronomy, Aarhus University, Ny Munkegade 120, DK-8000 Aarhus C, Denmark}
 \affiliation{School of Physics and Astronomy, University of
@@
 We validate this method and its result with new RV observations with the SONG~Telescope, and photometry from the \ktwo Mission.
-By measuring the average frequency \numax of these \emph{p}-mode oscillations we asteroseismically determine the mass of Aldebaran to be $1.17 \pm 0.05$~\msun.
-This measured stellar mass allows us to calculate that any moons of this giant planet, although they are now likely to be very hot, may have had equilibrium temperatures comparable to that of the Earth when Aldebaran was on the main sequence, raising the possibility that tehy may have once been habitable.
+By measuring the frequency of maximum power ,\numax, of these \emph{p}-mode oscillations we asteroseismically determine the mass of Aldebaran to be $1.16 \pm 0.07$~\msun.
+This measured stellar mass allows us to calculate that any moons of this giant planet, although they are now likely to be very hot, may have had equilibrium temperatures comparable to that of the Earth when Aldebaran was on the main sequence, raising the possibility that they may have once been habitable.
 
@@
 An all-sky survey planetary companions to precisely measure the masses of all nearby red giant stars is feasible with this new approach, and the required data either already exist in large exoplanet surveys, or are easy to obtain with ground-based telescopes.
+
+\section{Asteroseismology of Red Giants}
+Asteroseismology is a powerful tool for the charaterisation of red giant stars \citep[see][for a detailed review]{2013ARA&A..51..353C}.
+Red giants are solar-type pulsators, they exhibit acoustic modes of oscillation that are both excited and damped by stellar convection.
+In the frequency-power spectrum of either radial velocity of photometric observations, these modes of oscillation form an often intricate pattern of sharp peaks where the properties of the pattern can be mapped back to intinsic stellar properties \citep[e.g.,][]{2016AN....337..774D}.
+The easiest properties of the pattern to determine are the the frequency of maximum amplitude \numax and the so-called large separation \Dnu \citep{1995A&A...293...87K}.
+The two observables are often referred to as the global average seismic parameters.
+
+Stellar properties can be determined with the average seismic measures and effective temperature \teff.
+A set of scaling relations exist that illustrate the constraint the global asteroseismic measures provide.
+These scaling relations are understood to be approximate but never the less are commonly used for radius $R$, mass $M$, and surface gravity $g$:
+\begin{eqnarray}
+\left( \frac{R}{\mathrm{R_{\odot}}} \right) \simeq & \left( \frac{\nu_{\rm max}}{\nu_{\rm max, \odot}} \right) \, 
+\left( \frac{\Delta \nu}{\Delta \nu_{\odot}} \right)^{-2} \, \left( \frac{T_{\rm eff}}{T_{\rm eff, \odot}} \right)^{0.5}, \\
+\left( \frac{M}{\mathrm{M_{\odot}}} \right) \simeq & \left( \frac{\nu_{\rm max}}{\nu_{\rm max, \odot}} \right)^{3} \, 
+\left( \frac{\Delta \nu}{\Delta \nu_{\odot}} \right)^{-4} \, \left( \frac{T_{\rm eff}}{T_{\rm eff, \odot}} \right)^{1.5}, \\
+\left( \frac{g}{\mathrm{g_{\odot}}} \right) \simeq & \left( \frac{\nu_{\rm max}}{\nu_{\rm max, \odot}} \right) \, 
+ \left( \frac{T_{\rm eff}}{T_{\rm eff, \odot}} \right)^{0.5}. \\
+\end{eqnarray}
+While the accuracy of these scaling relations is still a matter of ongoing work \citep[e.g.,][]{2017ApJ...844..102H}, stellar parameters can be estimated by comparing observables to parmaeter from models of stellar evolution.
+
+As can be seen from the scaling relations, asteroseismology can provide constraint on Mass, Radius, and surface gravity if \Dnu, \numax, and \teff can be measured.
+In practise, for the higher luminosity red giants it is more straight forward to measure \numax than \Dnu.
+Typical values for \numax might range from $0.1$ to $20 \rm \mu Hz$ for the luminous giants and $20$ to $50 \rm \mu Hz$ for red giants with luminosities similar to the red clump .
+For the same sets of stars, \Dnu would take ranges from $0.02$ to $3 \rm \mu Hz$ and $3$ to $7 \rm \mu Hz$ depending on stellar mass and \teff (or metallicity which is considered more fundamental) \citep[e.g.,][]{2011A&A...525L...9M, 2013A&A...559A.137M}.
+Hence, \numax requires less frequency resolution than \Dnu to establish a good measurement.
+
+The ability to measure \numax and \Dnu, therefore, depends on the length and sampling rate of a data set.
+For regularly spaced data, typically analysed using an estimate of the frequency-power spectrum, the Nyquist limit and frequency resolution are trivial to estimate.
+For the historic data sets proposed here the typical length is \guy{blah} and typical sampling is \guy{blah}.
+This in principle gives us access to measures of \numax for \guy{these types of stars} and \Dnu $\,$ \guy{for those types of stars}.
 
@@
 % \includegraphics[width=0.75\textwidth]{Aldebaran_lc.png}
-\includegraphics[width=0.75\textwidth]{Aldebaran_lc.eps}
+\includegraphics[width=0.75\textwidth]{aldebaran_lc.eps}
 \caption{\ktwo lightcurve of Aldebaran.}
@@
 \subsection{Stellar Parameters}
+
+We have determined the stellar properties of Aldebaran with the additional constraint of \numax determined from our analysis here.
+The details of the stellar modelling are presented in detail in \ref{sec:mod}.
+We have considered the impact of our additional \numax constraint and run our analysis for different sets of spectroscopic estimates.
+For final values we adopt the Sheffield et al.
+(2012) b spectrosocpic solution as being 'middle of the road' estimates.
+For the analysis without constraint on \numax we find estimates of the stellar propeties as $M = 1.43^{+0.24}_{-0.2} \, \mathrm{M_{\odot}}$ and age $4.86^{3.56}_{2.04} \, \rm Gyr$.
+With the inclusion of \numax we find $M = 1.16^{+0.07}_{-0.07} \, \mathrm{M_{\odot}}$ and age $6.38^{1.42}_{1.12} \, \rm Gyr$.
+It is clear that the addition of \numax provides substatially more precise estimates of Mass and Age.
 
@@
 
+\include{aldeb_North}
+
 \bibliography{aldebaran}

### 2017-09-22 paper/aldebaran.tex
fix typo, affiliation, acknowlegement

@@
 
-\author{Tim White}
+\author[0000-0002-6980-3392]{Timothy R. White}
 \affiliation{Stellar Astrophysics Centre, Department of Physics 
@@
 Our new approach to asteroseismic data analysis, based on Continuous Auto-Regressive Moving Average (CARMA) models, can extract exoplanet signals together with measures of \numax from sparse and irregularly-sampled time series.
-An all-sky survey planetary companions to precisely measure the masses of all nearby red giant stars is feasible with this new approach, and the required data either already exist in large exoplanet surveys, or are easy to obtain with ground-based telescopes.
+An all-sky survey to find planetary companions and to precisely measure the masses of all nearby red giant stars is feasible with this new approach, and the required data either already exist in large exoplanet surveys, or are easy to obtain with ground-based telescopes.
 
@@
 This work was performed in part under contract with the Jet Propulsion Laboratory (JPL) funded by NASA through the Sagan Fellowship Program executed by the NASA Exoplanet Science Institute.
-
+Funding for the Stellar Astrophysics Centre is provided by The Danish National Research Foundation. TRW acknowledges the support of the Villum Foundation (research grant 10118).
 This research has made use of the SIMBAD database, operated at CDS, Strasbourg, France, and NASA's Astrophysics Data System.

### 2017-09-22 paper/aldebaran.tex
typos and added affiliation

@@
 \affiliation{School of Physics and Astronomy, University of Birmingham, Birmingham, B15 2TT, United Kingdom}
+\affiliation{Stellar Astrophysics Centre, Department of Physics 
+and Astronomy, Aarhus University, Ny Munkegade 120, DK-8000 Aarhus C, Denmark}
 
@@
 Red giants are solar-type pulsators, they exhibit acoustic modes of oscillation that are both excited and damped by stellar convection.
-In the frequency-power spectrum of either radial velocity of photometric observations, these modes of oscillation form an often intricate pattern of sharp peaks where the properties of the pattern can be mapped back to intinsic stellar properties \citep[e.g.,][]{2016AN....337..774D}.
+In the frequency-power spectrum of either radial velocity of photometric observations, these modes of oscillation form an often intricate pattern of sharp peaks where the properties of the pattern can be mapped back to intrinsic stellar properties \citep[e.g.,][]{2016AN....337..774D}.
 The easiest properties of the pattern to determine are the the frequency of maximum amplitude \numax and the so-called large separation \Dnu \citep{1995A&A...293...87K}.
@@
 \end{eqnarray}
-While the accuracy of these scaling relations is still a matter of ongoing work \citep[e.g.,][]{2017ApJ...844..102H}, stellar parameters can be estimated by comparing observables to parmaeter from models of stellar evolution.
+While the accuracy of these scaling relations is still a matter of ongoing work \citep[e.g.,][]{2017ApJ...844..102H}, stellar parameters can be estimated by comparing observables to parameter from models of stellar evolution.
 
@@
 For final values we adopt the Sheffield et al.
-(2012) b spectrosocpic solution as being 'middle of the road' estimates.
-For the analysis without constraint on \numax we find estimates of the stellar propeties as $M = 1.43^{+0.24}_{-0.2} \, \mathrm{M_{\odot}}$ and age $4.86^{3.56}_{2.04} \, \rm Gyr$.
+(2012) b spectroscopic solution as being 'middle of the road' estimates.
+For the analysis without constraint on \numax we find estimates of the stellar properties as $M = 1.43^{+0.24}_{-0.2} \, \mathrm{M_{\odot}}$ and age $4.86^{3.56}_{2.04} \, \rm Gyr$.
 With the inclusion of \numax we find $M = 1.16^{+0.07}_{-0.07} \, \mathrm{M_{\odot}}$ and age $6.38^{1.42}_{1.12} \, \rm Gyr$.
-It is clear that the addition of \numax provides substatially more precise estimates of Mass and Age.
+It is clear that the addition of \numax provides substantially more precise estimates of Mass and Age.
 
@@
 
-We note that this implies that the planet at a semimajor axis of $1.512 \pm 0.021$~AU would have been subject to an insolation comparable to that of the Earth, evolving from 0.5~Gyr to 2.5~Gyr from $0.89 \pm 0.22$ to $1.19 \pm 0.37$ times that of the Earth today.
+We note that this implies that the planet at a semi-major axis of $1.512 \pm 0.021$~AU would have been subject to an insolation comparable to that of the Earth, evolving from 0.5~Gyr to 2.5~Gyr from $0.89 \pm 0.22$ to $1.19 \pm 0.37$ times that of the Earth today.
 Subject as well as this to the great uncertainties of the planet's orbital evolution and albedo, this planet and any of its moons may well have had temperate environments now long-since destroyed by their host star's evolution away from the main sequence.

### 2017-09-23 paper/aldeb_North.tex

\section{Stellar Modelling}\label{sec:mod}
\subsection{Stellar Models}\label{sec:stell_mod}
We used our determination of \numax and several combinations of the asteroseismic and spectroscopic parameters, along with luminosity, to estimate the fundamental stellar parameters, via fitting to stellar models. We used \textsc{MESA} models \citep{2011Paxton,2013Paxton} in conjunction with the Bayesian code \textsc{PARAM} \citep{2006dasilva, 2017Rod}.  A summary of our selected ``benchmark'' options is as follows; 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{itemize}%GUY THIS BIT COULD PROBABLY BE DROPPED AND REFERENCED OUT TO \citep{2017Rod}
\item Heavy element partitioning from \cite{1993Grevesse}.
\item OPAL equation of state \citep{2002Rogers} along with OPAL opacities \citep{1996Iglesias}, with complementary values at low temperatures from \cite{2005Ferguson}.
\item Nuclear reaction rates from NACRE \citep{1999Angulo}.
\item The atmosphere model is taken according to \cite{1966Kris}.
\item The mixing length theory was used to describe convection (a solar-calibrated parameter $\alpha_{\textrm{MLT}} =1.9657$ was adopted).
\item Convective overshooting on the main sequence is set to $\alpha_{\textrm{ov}}=0.2H_{p}$, with $H_{p}$ the pressure scale height at the border of the convective core. Overshooting was applied according to the \cite{1975Maeder} step function
scheme.
\item No rotational mixing or diffusion is included.
\end{itemize}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Note that the precision on our estimate of \numax is orders of magnitude less that the point at which we would have needed to correct for the line-of-sight doppler shift \citep{2014MNRAS.445L..94D}.
\subsection{Additional modelling inputs}\label{sec:addmod}
In addition to the asteroseismic parameters, temperature and metallicity values are needed. There exist multiple literature values for Aldebaran. We chose to use a selection of sources as a separate modelling variation, to investigate what variation this would produce in final stellar properties.

To ensure the values are self-consistent, when a literature value was chosen for temperature, we took the stellar metallicity from the same source i.e. matched pairs of temperature and metallicity. The final constraint is the stellar luminosity, which may be estimated as follows (e.g. see \citealt{pijpers2003}):
\begin{multline}
\log_{10} \frac{L}{L_{\odot}} = 4.0+
0.4 M_{{\textrm{bol}},\odot} -2.0 \log_{10} {\pi [{\textrm{mas}]}} \\-0.4(V-A_V + BC(V)).
\label{eqn:lum}
\end{multline}
The solar bolometric magnitude $M_{\textrm{bol},\odot}=4.73$ is taken from \cite{Torres2010}, from which we also take the polynomial expression for the bolometric correction $BC(V)$. Extinction $A_V$ was assumed to be zero.

The final constraint available for Aldebaran is the angular diameter of the star, measured by long baseline interferometry, combined with the parallax to produce a physical radius constraint of $R_{\textrm{int}}=44.2\pm0.9\textrm{R}_{\odot}$ \citep{2005Richichi}.

% Example table
\begin{table*}
	\centering
	\caption{Spectroscopic from each literature source, along with calculated luminosity. All temperature uncertainties assumed to be 50K, 0.2 dex in $\log{g}$, and 0.1 dex in [FeH]. For the two \protect{\cite{2012Sheffield}} results, the reason for discrepancy between the two sets of metallicity results is not discussed.}
	\label{tab:spec}
	\begin{tabular}{lllll} % 5 columns, alignment for each
		\hline
		Spectrosocpy Source & $T_{\textrm{eff}}$ (K) & $\log{g}$ (dex) & [FeH] (dex) & Luminosity (L$_{\odot}$)\\
		\hline
		\cite{2012Sheffield}$_{a}$	&	3900	&	1.3	&	0.17	&	480\\
		\cite{2012Sheffield}$_{b}$	&	3900	&	1.3	&	0.05	&	480\\
		\cite{2011Prugniel}	&	3870 & 1.66 & -0.04 & 507\\
		\cite{2008Massarotti} & 3936 & 1 & -0.34 & 456\\
		\cite{2009Frasca} & 3850 & 0.55 & -0.1 & 526\\
		\hline
	\end{tabular}
\end{table*}

As Table \ref{tab:spec} shows, the spectroscopic parameters of Aldebaran are somewhat unclear, particularly $\log{g}$ and [FeH], which may have an impact on the recovered stellar properties when fitting to models. To explore what impact each parameter is having on the final stellar properties, multiple \textsc{PARAM} runs were performed, using different constraints. Two constraints potentially in tension were $\nu_{\textrm{max}}$ and $\log{g}$. $\nu_{\textrm{max}}$ has been shown to scale with $\log{g}$ \citep{Kjeldsen95, 2011A&A...530A.142B},
\begin{equation}
\frac{\nu_{\textrm{max}}}{\nu_{\textrm{max},\odot}}=\frac{\log{g}}{\log{g},\odot}\left(\frac{T_{\textrm{eff}}}{T_{\textrm{eff},\odot}}\right)^{-1/2}.
\label{eqn:numax}
\end{equation}

Using Eq \ref{eqn:numax} with the values in Table \ref{tab:spec} predicts $\nu_{\textrm{max}}$ in the range $0.5-6\mu$Hz. Reversing the equation to produce a predicted $\log{g}$ from the observed $\nu_{\textrm{max,obs}}=2.23\pm0.1\mu$Hz results in a predicted $\log{g}\sim1.2$ dex, using an assumed temperature of 3900K. The solar calibration values used here are $g_{\odot}=274$ms$^{-2}$, $\nu_{\textrm{max},\odot}=3150\mu$Hz and $T_{\textrm{eff},\odot}=5777$K.

Table \ref{tab:res_mass} shows the results, for all modelling variations, both different inputs and different constraints. It shows that results with the addition of $\nu_{\textrm{max}}$ as a constraint exhibit in general smaller uncertainties, with or without the addition of $\log{g}$ as a constraint.

Recovering the mass without the use of asteroseismic constraints produces considerable scatter on the results ($0.96-1.5\textrm{M}_{\odot}$), whilst the use of asteroseismology brings the mass estimates into closer agreement, with the exception of the very low metallicity solution of \cite{2008Massarotti}. Any systematic offset between asteroseismic masses and spectroscopic masses is sensitive to chosen reference mass, in agreement with \cite{2017North}.

Within the results of Table \ref{tab:res_mass}, the parallax of Aldebaran is contained within the angular diameter and the luminosity of the star. To avoid this, \textsc{PARAM} was also ran without luminosity constraint, the average absolute mass offset between the two sets of results was $<0.04\textrm{M}_{\odot}$ for the runs without $\nu_{\textrm{max}}$ constraint, and $<0.02\textrm{M}_{\odot}$ for the runs with $\nu_{\textrm{max}}$.

% Example table
\begin{table*}
	\centering

	\caption{Recovered stellar properties from \textsc{PARAM} using various constraints. Uncertainties quoted are the 68\% credible interval.}
	\label{tab:res_mass}
	\begin{tabular}{llll} % 5 columns, alignment for each
		\hline
		Spectroscopy Source & Mass ($\textrm{M}_{\odot}$) & Radius ($\textrm{R}_{\odot}$) & Age (Gyr)\\
		\hline
		\multicolumn{4}{c}{$\nu_{\textrm{max}}$, $\log{g}$, $T_{\textrm{eff}}$, $R_\textrm{int}$, L and [FeH].} \\
		\hline
		\cite{2012Sheffield}$_{a}$ &$1.17^{+0.07}_{-0.07}$&$43.9^{+0.9}_{-0.9}$&$6.47^{+1.39}_{-1.1}$\\
		\cite{2012Sheffield}$_{b}$&$1.17^{+0.07}_{-0.07}$&$43.8^{+0.8}_{-0.9}$&$6.35^{+1.4}_{-1.11}$\\
		\cite{2011Prugniel}&$1.17^{+0.07}_{-0.07}$&$43.9^{+0.9}_{-0.9}$&$6.23^{+1.4}_{-1.11}$\\
		\cite{2008Massarotti}&$1.13^{+0.07}_{-0.07}$&$43.5^{+0.9}_{-0.9}$&$5.95^{+1.37}_{-1.07}$\\
		\cite{2009Frasca}&$1.02^{+0.04}_{-0.04}$&$44.0^{+0.8}_{-0.8}$&$10.26^{+1.54}_{-1.45}$\\
		\hline
		\multicolumn{4}{c}{$\log{g}$, $T_{\textrm{eff}}$, $R_\textrm{int}$, L and [FeH].} \\
		\hline
\cite{2012Sheffield}$_{a}$&$1.43^{+0.26}_{-0.24}$&$43.8^{+0.9}_{-0.9}$&$3.53^{+2.72}_{-1.44}$\\
\cite{2012Sheffield}$_{b}$&$1.27^{+0.24}_{-0.2}$&$43.8^{+0.9}_{-0.9}$&$4.86^{+3.56}_{-2.04}$\\
\cite{2011Prugniel}&$1.25^{+0.22}_{-0.19}$&$43.8^{+0.9}_{-0.9}$&$5.02^{+3.47}_{-2.02}$\\
\cite{2008Massarotti}&$0.95^{+0.11}_{-0.05}$&$43.8^{+0.9}_{-0.9}$&$10.24^{+2.36}_{-3.08}$\\
\cite{2009Frasca}&$0.96^{+0.04}_{-0.04}$&$44.4^{+0.8}_{-0.8}$&$11.64^{+1.43}_{-1.8}$\\
		\hline
		\multicolumn{4}{c}{$\nu_{\textrm{max}}$, $T_{\textrm{eff}}$, $R_\textrm{int}$, L and [FeH].} \\
		\hline
\cite{2012Sheffield}$_{a}$&$1.17^{+0.07}_{-0.07}$&$43.9^{+0.9}_{-0.9}$&$6.51^{+1.41}_{-1.11}$\\
\cite{2012Sheffield}$_{b}$&$1.16^{+0.07}_{-0.07}$&$43.8^{+0.8}_{-0.9}$&$6.38^{+1.42}_{-1.12}$\\
\cite{2011Prugniel}&$1.16^{+0.07}_{-0.07}$&$43.9^{+0.9}_{-0.9}$&$6.43^{+1.48}_{-1.15}$\\
\cite{2008Massarotti}&$1.13^{+0.07}_{-0.07}$&$43.5^{+0.9}_{-0.9}$&$5.86^{+1.35}_{-1.05}$\\
\cite{2009Frasca}&$1.15^{+0.07}_{-0.07}$&$43.9^{+0.9}_{-0.8}$&$6.47^{+1.51}_{-1.17}$\\
		\hline
	\end{tabular}

\end{table*}

### 2017-09-23 paper/aldebaran.tex

@@
 (2012) b spectroscopic solution as being 'middle of the road' estimates.
-For the analysis without constraint on \numax we find estimates of the stellar properties as $M = 1.43^{+0.24}_{-0.2} \, \mathrm{M_{\odot}}$ and age $4.86^{3.56}_{2.04} \, \rm Gyr$.
+For the analysis without constraint on \numax we find estimates of the stellar properties as $M = 1.27^{+0.24}_{-0.2} \, \mathrm{M_{\odot}}$ and age $4.86^{3.56}_{2.04} \, \rm Gyr$.
 With the inclusion of \numax we find $M = 1.16^{+0.07}_{-0.07} \, \mathrm{M_{\odot}}$ and age $6.38^{1.42}_{1.12} \, \rm Gyr$.

### 2017-09-23 paper/aldebaran.tex
misc changes

@@
 (2012) b spectroscopic solution as being 'middle of the road' estimates.
-For the analysis without constraint on \numax we find estimates of the stellar properties as $M = 1.27^{+0.24}_{-0.2} \, \mathrm{M_{\odot}}$ and age $4.86^{3.56}_{2.04} \, \rm Gyr$.
+For the analysis without constraint on \numax we find estimates of the stellar properties as $M = 1.27^{+0.24}_{-0.2} \, \mathrm{M_{\odot}}$ and age $4.86^{+3.56}_{-2.04} \, \rm Gyr$.
 With the inclusion of \numax we find $M = 1.16^{+0.07}_{-0.07} \, \mathrm{M_{\odot}}$ and age $6.38^{1.42}_{1.12} \, \rm Gyr$.
@@
 We note that with the determination of the stellar mass it is possible to infer the parameters Aldebaran had while it was on the main sequence.
-We conduct a Monte Carlo simulation, drawing masses randomly from a Gaussian distribution $M \sim \mathscr{N}(1.17,0.05)$ to predict the luminosity of the main-sequence progenitor, and the semi-major axis of the planet's orbit.
+We conduct a Monte Carlo simulation, drawing masses randomly from a Gaussian distribution $M \sim \mathscr{N}(1.16,0.07)$ and metallicitiies $[\text{Fe}/\text{H}] \sim \mathscr{N}(-0.15,0.2)$ \citep{decin2003}  to predict the luminosity of the main-sequence progenitor, and the semi-major axis of the planet's orbit.
 %Using the scaling relations of \citep{eker2015}, we find that Aldebaran would have had a luminosity $L_* = 2.0 \pm 0.37~\lsun$ on the main sequence.
-Using the Mesa Isochrones and Stellar Tracks \citep[MIST:][]{mist0,mist1}, models, we compute the luminosity and radius evolution of the planet along the main sequence with uncertainty both in stellar mass (this work) and metallicity \citep[$-0.15 \pm 0.2$; ][]{decin2003}.
-$L$ evolves from $2.03 \pm 0.56$~\lsun at 0.5~Gyr to $2.74 \pm 0.93$~\lsun at 2.5~Gyr.
-
-We note that this implies that the planet at a semi-major axis of $1.512 \pm 0.021$~AU would have been subject to an insolation comparable to that of the Earth, evolving from 0.5~Gyr to 2.5~Gyr from $0.89 \pm 0.22$ to $1.19 \pm 0.37$ times that of the Earth today.
-Subject as well as this to the great uncertainties of the planet's orbital evolution and albedo, this planet and any of its moons may well have had temperate environments now long-since destroyed by their host star's evolution away from the main sequence.
+Using the Mesa Isochrones and Stellar Tracks \citep[MIST:][]{mist0,mist1}, models, we compute the luminosity and radius evolution of the planet along the main sequence.
+$L$ evolves from $2.0 \pm 0.7$~\lsun at 0.5~Gyr to $3.5 \pm 2.3$~\lsun at 2.5~Gyr.
+We note that this implies that the planet at a semi-major axis of $1.507 \pm 0.03$~AU would have been subject to an insolation comparable to that of the Earth, evolving from 0.5~Gyr to 2.5~Gyr from $0.86 \pm 0.27$ to $1.55 \pm 1.0$ times that of the Earth today.
+Subject as well as this to the great uncertainties of the planet's orbital evolution and albedo, this planet and any of its moons may well have had temperate environments for some of their previous evolution, now long-since destroyed by their host star's evolution away from the main sequence.
 

### 2017-09-23 paper/aldebaran.tex
misc changes

@@
 We have considered the impact of our additional \numax constraint and run our analysis for different sets of spectroscopic estimates.
-For final values we adopt the Sheffield et al.
-(2012) b spectroscopic solution as being 'middle of the road' estimates.
+For final values we adopt the \citet{2012Sheffield} spectroscopic solution as being `middle of the road' estimates.
 For the analysis without constraint on \numax we find estimates of the stellar properties as $M = 1.27^{+0.24}_{-0.2} \, \mathrm{M_{\odot}}$ and age $4.86^{+3.56}_{-2.04} \, \rm Gyr$.
@@
 It is clear that the addition of \numax provides substantially more precise estimates of Mass and Age.
+% re sheffield - I have substituted the citation from the .bib but you said 'b'; afaik she only published one paper in 2012?
+- Ben
 
@@
 %Using the scaling relations of \citep{eker2015}, we find that Aldebaran would have had a luminosity $L_* = 2.0 \pm 0.37~\lsun$ on the main sequence.
-Using the Mesa Isochrones and Stellar Tracks \citep[MIST:][]{mist0,mist1}, models, we compute the luminosity and radius evolution of the planet along the main sequence.
+Using the Mesa Isochrones and Stellar Tracks \citep[MIST:][]{mist0,mist1} models, we compute the luminosity and radius evolution of the planet along the main sequence.
 $L$ evolves from $2.0 \pm 0.7$~\lsun at 0.5~Gyr to $3.5 \pm 2.3$~\lsun at 2.5~Gyr.

### 2017-09-23 paper/aldebaran.tex
longer time base

@@
 Using the Mesa Isochrones and Stellar Tracks \citep[MIST:][]{mist0,mist1} models, we compute the luminosity and radius evolution of the planet along the main sequence.
-$L$ evolves from $2.0 \pm 0.7$~\lsun at 0.5~Gyr to $3.5 \pm 2.3$~\lsun at 2.5~Gyr.
-We note that this implies that the planet at a semi-major axis of $1.507 \pm 0.03$~AU would have been subject to an insolation comparable to that of the Earth, evolving from 0.5~Gyr to 2.5~Gyr from $0.86 \pm 0.27$ to $1.55 \pm 1.0$ times that of the Earth today.
+$L$ evolves from $2.0 \pm 0.7$~\lsun at 0.5~Gyr to $3.5 \pm 2.3$~\lsun at 4.5~Gyr.
+We note that this implies that the planet at a semi-major axis of $1.507 \pm 0.03$~AU would have been subject to an insolation comparable to that of the Earth, evolving from 0.5~Gyr to 4.5~Gyr from $0.86 \pm 0.27$ to $1.55 \pm 1.0$ times that of the Earth today.
 Subject as well as this to the great uncertainties of the planet's orbital evolution and albedo, this planet and any of its moons may well have had temperate environments for some of their previous evolution, now long-since destroyed by their host star's evolution away from the main sequence.

### 2017-09-25 paper/aldebaran.tex
brewer and stello

@@
 This introduces a window-function effect: the power spectrum as constructed for example by a Fourier transform, or a Lomb-Scargle Periodogram \citep{lomb,scargle}, is convolved with the Fourier transform of the window function, introducing strong sidelobes adjacent to real frequency peaks and causing crosstalk between adjacent frequency channels.
-This imposes significant limitations both on the signal-to-noise and frequency resolution of power spectra derived from linear methods such as the Lomb-Scargle periodogram, and in practice makes asteroseismology difficult or impossible from the ground for stars with oscillation frequencies ranging from $\sim 12$~h to $\sim$~a few days.
+This imposes significant limitations both on the signal-to-noise and frequency resolution of power spectra derived from linear methods such as the Lomb-Scargle periodogram, and in practice makes asteroseismology with these conventional approaches difficult or impossible from the ground for stars with oscillation frequencies ranging from $\sim 12$~h to $\sim$~a few days.
 
-If we apply nonlinear statistical inference methods...
+If we apply nonlinear statistical inference methods this situation can be improved.
+\citet{brewer2009} show that a system of driven, damped harmonic oscillators such as we encounter in asteroseismology can be statistically modelled as a Gaussian Process, with a covariance kernel consisting of a sum of damped sinusoids.
+The hyperparameters of this process can be used to retrieve very accurate power spectral density distributions.
+Unfortunately, this is impractical for long time series, as the computational cost of evaluating this likelihood function scales as $\mathcal{O}(n^3)$.
 

### 2017-09-25 paper/aldebaran.tex
font

@@
 We note that with the determination of the stellar mass it is possible to infer the parameters Aldebaran had while it was on the main sequence.
-We conduct a Monte Carlo simulation, drawing masses randomly from a Gaussian distribution $M \sim \mathscr{N}(1.16,0.07)$ and metallicitiies $[\text{Fe}/\text{H}] \sim \mathscr{N}(-0.15,0.2)$ \citep{decin2003}  to predict the luminosity of the main-sequence progenitor, and the semi-major axis of the planet's orbit.
+We conduct a Monte Carlo simulation, drawing masses randomly from a Gaussian distribution $M \sim \mathcal{N}(1.16,0.07)$ and metallicitiies $[\text{Fe}/\text{H}] \sim \mathcal{N}(-0.15,0.2)$ \citep{decin2003}  to predict the luminosity of the main-sequence progenitor, and the semi-major axis of the planet's orbit.
 %Using the scaling relations of \citep{eker2015}, we find that Aldebaran would have had a luminosity $L_* = 2.0 \pm 0.37~\lsun$ on the main sequence.

### 2017-09-25 paper/aldebaran.tex
formatting

@@
 For the analysis without constraint on \numax we find estimates of the stellar properties as $M = 1.27^{+0.24}_{-0.2} \, \mathrm{M_{\odot}}$ and age $4.86^{+3.56}_{-2.04} \, \rm Gyr$.
-With the inclusion of \numax we find $M = 1.16^{+0.07}_{-0.07} \, \mathrm{M_{\odot}}$ and age $6.38^{1.42}_{1.12} \, \rm Gyr$.
-It is clear that the addition of \numax provides substantially more precise estimates of Mass and Age.
+With the inclusion of \numax we find $M = 1.16^{+0.07}_{-0.07} \, \mathrm{M_{\odot}}$ and age $6.38^{+1.42}_{-1.12} \, \rm Gyr$.
+It is clear that the addition of \numax provides substantially more precise estimates of mass and age.
 % re sheffield - I have substituted the citation from the .bib but you said 'b'; afaik she only published one paper in 2012?
@@
 The \kepler detector saturates for stars brighter than the $\sim~11^\text{th}$ magnitude.
-Nevertheless, the excess flux deposited in a saturated pixel spills conservatively up and down the pixel column, such that it is possible to sum this `bleed column' for bright stars and still obtain precise photometry, such as was done for the brightest star in the nominal \kepler mission, $\theta$~Cyg \citep{guzik2011,thetacygwhite,guzik2016} ($V = 4.48$).
+Nevertheless, the excess flux deposited in a saturated pixel spills conservatively up and down the pixel column, such that it is possible to sum this `bleed column' for bright stars and still obtain precise photometry, such as was done for the brightest star in the nominal \kepler mission, $\theta$~Cyg \citep[$V = 4.48$;][]{guzik2011,thetacygwhite,guzik2016}.
 There are two main reasons why this is not possible for all bright stars in general.

### 2017-09-25 paper/aldebaran.tex
edited in North section

@@
 
-\include{aldeb_North}
+\section{Stellar Modelling}\label{sec:mod}
+\subsection{Stellar Models}\label{sec:stell_mod}
+We used our determination of \numax and several combinations of the asteroseismic and spectroscopic parameters, along with luminosity, to estimate the fundamental stellar parameters, via fitting to stellar models.
+We used \textsc{MESA} models \citep{2011Paxton,2013Paxton} in conjunction with the Bayesian code \textsc{PARAM} \citep{2006dasilva, 2017Rod}.
+A summary of our selected ``benchmark'' options is as follows;
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{itemize}%GUY THIS BIT COULD PROBABLY BE DROPPED AND REFERENCED OUT TO \citep{2017Rod}
+\item Heavy element partitioning from \cite{1993Grevesse}.
+\item OPAL equation of state \citep{2002Rogers} along with OPAL opacities \citep{1996Iglesias}, with complementary values at low temperatures from \cite{2005Ferguson}.
+\item Nuclear reaction rates from NACRE \citep{1999Angulo}.
+\item The atmosphere model is taken according to \cite{1966Kris}.
+\item The mixing length theory was used to describe convection (a solar-calibrated parameter $\alpha_{\textrm{MLT}} =1.9657$ was adopted).
+\item Convective overshooting on the main sequence is set to $\alpha_{\textrm{ov}}=0.2H_{p}$, with $H_{p}$ the pressure scale height at the border of the convective core.
+Overshooting was applied according to the \cite{1975Maeder} step function
+scheme.
+\item No rotational mixing or diffusion is included.
+\end{itemize}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Note that the precision on our estimate of \numax is orders of magnitude less that the point at which we would have needed to correct for the line-of-sight doppler shift \citep{2014MNRAS.445L..94D}.
+\subsection{Additional modelling inputs}\label{sec:addmod}
+In addition to the asteroseismic parameters, temperature and metallicity values are needed.
+There exist multiple literature values for Aldebaran.
+We chose to use a selection of sources as a separate modelling variation, to investigate what variation this would produce in final stellar properties.
+
+To ensure the values are self-consistent, when a literature value was chosen for temperature, we took the stellar metallicity from the same source i.e.
+matched pairs of temperature and metallicity.
+The final constraint is the stellar luminosity, which may be estimated as follows (e.g.
+see \citealt{pijpers2003}):
+\begin{multline}
+\log_{10} \frac{L}{L_{\odot}} = 4.0+
+0.4 M_{{\textrm{bol}},\odot} -2.0 \log_{10} {\pi [{\textrm{mas}]}} \\-0.4(V-A_V + BC(V)).
+\label{eqn:lum}
+\end{multline}
+The solar bolometric magnitude $M_{\textrm{bol},\odot}=4.73$ is taken from \cite{Torres2010}, from which we also take the polynomial expression for the bolometric correction $BC(V)$.
+Extinction $A_V$ was assumed to be zero.
+
+The final constraint available for Aldebaran is the angular diameter of the star, measured by long baseline interferometry, combined with the parallax to produce a physical radius constraint of $R_{\textrm{int}}=44.2\pm0.9\textrm{R}_{\odot}$ \citep{richichi2005}.
+
+% Example table
+\begin{table*}
+	\centering
+	\caption{Spectroscopic from each literature source, along with calculated luminosity.
+All temperature uncertainties assumed to be 50K, 0.2 dex in $\log{g}$, and 0.1 dex in [FeH].
+For the two \protect{\cite{2012Sheffield}} results, the reason for discrepancy between the two sets of metallicity results is not discussed.}
+	\label{tab:spec}
+	\begin{tabular}{lllll} % 5 columns, alignment for each
+		\hline
+		Spectrosocpy Source & $T_{\textrm{eff}}$ (K) & $\log{g}$ (dex) & [FeH] (dex) & Luminosity (L$_{\odot}$)\\
+		\hline
+		\cite{2012Sheffield}$_{a}$	&	3900	&	1.3	&	0.17	&	480\\
+		\cite{2012Sheffield}$_{b}$	&	3900	&	1.3	&	0.05	&	480\\
+		\cite{2011Prugniel}	&	3870 & 1.66 & -0.04 & 507\\
+		\cite{2008Massarotti} & 3936 & 1 & -0.34 & 456\\
+		\cite{2009Frasca} & 3850 & 0.55 & -0.1 & 526\\
+		\hline
+	\end{tabular}
+\end{table*}
+
+As Table \ref{tab:spec} shows, the spectroscopic parameters of Aldebaran are somewhat unclear, particularly $\log{g}$ and [FeH], which may have an impact on the recovered stellar properties when fitting to models.
+To explore what impact each parameter is having on the final stellar properties, multiple \textsc{PARAM} runs were performed, using different constraints.
+Two constraints potentially in tension were $\nu_{\textrm{max}}$ and $\log{g}$.
+$\nu_{\textrm{max}}$ has been shown to scale with $\log{g}$ \citep{Kjeldsen95, 2011A&A...530A.142B},
+\begin{equation}
+\frac{\nu_{\textrm{max}}}{\nu_{\textrm{max},\odot}}=\frac{\log{g}}{\log{g},\odot}\left(\frac{T_{\textrm{eff}}}{T_{\textrm{eff},\odot}}\right)^{-1/2}.
+\label{eqn:numax}
+\end{equation}
+
+Using Eq \ref{eqn:numax} with the values in Table \ref{tab:spec} predicts $\nu_{\textrm{max}}$ in the range $0.5-6\mu$Hz.
+Reversing the equation to produce a predicted $\log{g}$ from the observed $\nu_{\textrm{max,obs}}=2.23\pm0.1\mu$Hz results in a predicted $\log{g}\sim1.2$ dex, using an assumed temperature of 3900K.
+The solar calibration values used here are $g_{\odot}=274$ms$^{-2}$, $\nu_{\textrm{max},\odot}=3150\mu$Hz and $T_{\textrm{eff},\odot}=5777$K.
+
+Table \ref{tab:res_mass} shows the results, for all modelling variations, both different inputs and different constraints.
+It shows that results with the addition of $\nu_{\textrm{max}}$ as a constraint exhibit in general smaller uncertainties, with or without the addition of $\log{g}$ as a constraint.
+
+Recovering the mass without the use of asteroseismic constraints produces considerable scatter on the results ($0.96-1.5\textrm{M}_{\odot}$), whilst the use of asteroseismology brings the mass estimates into closer agreement, with the exception of the very low metallicity solution of \cite{2008Massarotti}.
+Any systematic offset between asteroseismic masses and spectroscopic masses is sensitive to chosen reference mass, in agreement with \cite{2017North}.
+
+Within the results of Table \ref{tab:res_mass}, the parallax of Aldebaran is contained within the angular diameter and the luminosity of the star.
+To avoid this, \textsc{PARAM} was also ran without luminosity constraint, the average absolute mass offset between the two sets of results was $<0.04\textrm{M}_{\odot}$ for the runs without $\nu_{\textrm{max}}$ constraint, and $<0.02\textrm{M}_{\odot}$ for the runs with $\nu_{\textrm{max}}$.
+
+% Example table
+\begin{table*}
+	\centering
+
+	\caption{Recovered stellar properties from \textsc{PARAM} using various constraints. Uncertainties quoted are the 68\% credible interval.}
+	\label{tab:res_mass}
+	\begin{tabular}{llll} % 5 columns, alignment for each
+		\hline
+		Spectroscopy Source & Mass ($\textrm{M}_{\odot}$) & Radius ($\textrm{R}_{\odot}$) & Age (Gyr)\\
+		\hline
+		\multicolumn{4}{c}{$\nu_{\textrm{max}}$, $\log{g}$, $T_{\textrm{eff}}$, $R_\textrm{int}$, L and [FeH].} \\
+		\hline
+		\cite{2012Sheffield}$_{a}$ &$1.17^{+0.07}_{-0.07}$&$43.9^{+0.9}_{-0.9}$&$6.47^{+1.39}_{-1.1}$\\
+		\cite{2012Sheffield}$_{b}$&$1.17^{+0.07}_{-0.07}$&$43.8^{+0.8}_{-0.9}$&$6.35^{+1.4}_{-1.11}$\\
+		\cite{2011Prugniel}&$1.17^{+0.07}_{-0.07}$&$43.9^{+0.9}_{-0.9}$&$6.23^{+1.4}_{-1.11}$\\
+		\cite{2008Massarotti}&$1.13^{+0.07}_{-0.07}$&$43.5^{+0.9}_{-0.9}$&$5.95^{+1.37}_{-1.07}$\\
+		\cite{2009Frasca}&$1.02^{+0.04}_{-0.04}$&$44.0^{+0.8}_{-0.8}$&$10.26^{+1.54}_{-1.45}$\\
+		\hline
+		\multicolumn{4}{c}{$\log{g}$, $T_{\textrm{eff}}$, $R_\textrm{int}$, L and [FeH].} \\
+		\hline
+\cite{2012Sheffield}$_{a}$&$1.43^{+0.26}_{-0.24}$&$43.8^{+0.9}_{-0.9}$&$3.53^{+2.72}_{-1.44}$\\
+\cite{2012Sheffield}$_{b}$&$1.27^{+0.24}_{-0.2}$&$43.8^{+0.9}_{-0.9}$&$4.86^{+3.56}_{-2.04}$\\
+\cite{2011Prugniel}&$1.25^{+0.22}_{-0.19}$&$43.8^{+0.9}_{-0.9}$&$5.02^{+3.47}_{-2.02}$\\
+\cite{2008Massarotti}&$0.95^{+0.11}_{-0.05}$&$43.8^{+0.9}_{-0.9}$&$10.24^{+2.36}_{-3.08}$\\
+\cite{2009Frasca}&$0.96^{+0.04}_{-0.04}$&$44.4^{+0.8}_{-0.8}$&$11.64^{+1.43}_{-1.8}$\\
+		\hline
+		\multicolumn{4}{c}{$\nu_{\textrm{max}}$, $T_{\textrm{eff}}$, $R_\textrm{int}$, L and [FeH].} \\
+		\hline
+\cite{2012Sheffield}$_{a}$&$1.17^{+0.07}_{-0.07}$&$43.9^{+0.9}_{-0.9}$&$6.51^{+1.41}_{-1.11}$\\
+\cite{2012Sheffield}$_{b}$&$1.16^{+0.07}_{-0.07}$&$43.8^{+0.8}_{-0.9}$&$6.38^{+1.42}_{-1.12}$\\
+\cite{2011Prugniel}&$1.16^{+0.07}_{-0.07}$&$43.9^{+0.9}_{-0.9}$&$6.43^{+1.48}_{-1.15}$\\
+\cite{2008Massarotti}&$1.13^{+0.07}_{-0.07}$&$43.5^{+0.9}_{-0.9}$&$5.86^{+1.35}_{-1.05}$\\
+\cite{2009Frasca}&$1.15^{+0.07}_{-0.07}$&$43.9^{+0.9}_{-0.8}$&$6.47^{+1.51}_{-1.17}$\\
+		\hline
+	\end{tabular}
+
+\end{table*}
 

### 2017-09-25 paper/aldebaran.tex
deleting duplicate bib entries

@@
 In the frequency-power spectrum of either radial velocity of photometric observations, these modes of oscillation form an often intricate pattern of sharp peaks where the properties of the pattern can be mapped back to intrinsic stellar properties \citep[e.g.,][]{2016AN....337..774D}.
-The easiest properties of the pattern to determine are the the frequency of maximum amplitude \numax and the so-called large separation \Dnu \citep{1995A&A...293...87K}.
+The easiest properties of the pattern to determine are the the frequency of maximum amplitude \numax and the so-called large separation \Dnu \citep{Kjeldsen1995}.
 The two observables are often referred to as the global average seismic parameters.

### 2017-09-26 paper/aldebaran.tex
small changes

@@
 Aldebaran, or $\alpha$~Tauri, is a well-known first-magnitude naked-eye red giant star, and has long been the subject of astronomical investigations.
-Following the proposal of \citet{struverv} several groups began to search for planets by the radial velocity (RV) method, looking for Doppler shifts from the star's reflex motion around the common centre of mass, culminating in the discovery of 51~Peg~b \citep{51peg}, the first exoplanet to be recognized as such.
-Before this, \citet{hatzes1993} noted RV variations in Pollux \citep[$\beta$~Gem; subsequently confirmed as a planet:][]{betgemconf,betgemconf2}, Arcturus ($\alpha$~Boo, unconfirmed), and Aldebaran.
+By looking for Doppler shifts from the star's reflex motion around the common centre of mass with a companion planet \citep[the radial velocity or RV method;][]{struverv}, .
+While the hot Jupiter 51~Peg~b \citep{51peg} was the first exoplanet to be recognized as such, before this, \citet{hatzes1993} had noted RV variations in Pollux \citep[$\beta$~Gem; subsequently confirmed as a planet:][]{betgemconf,betgemconf2}, Arcturus ($\alpha$~Boo, unconfirmed), and Aldebaran.
 After further investigation by \citet{Hatzes1998}, \citet{Hatzes2015} now claim a firm RV detection of a planetary-mass companion Aldebaran~b, with a period of $628.96 \pm 0.90$~d.
@@
 In order to verify the results of the novel analysis presented above, we sought to obtain an independent detection of the oscillations of Aldebaran and compare the frequencies determined with the two methods.
-
-Aldebaran was therefore observed with the \ktwo Mission \citep{howell14}, a two-wheeled revival of the \kepler Mission \citep{2010sci...327..977b}, under Guest Observer Program 130471 in Campaign~13, from 2017-03-08 to 2017-05-27.
+Aldebaran was therefore observed with \ktwo \citep{howell14}, a two-wheeled revival of the \kepler Mission \citep{2010sci...327..977b}, under Guest Observer Program 130471 in Campaign~13, from 2017-03-08 to 2017-05-27.
 
@@
 %Using the scaling relations of \citep{eker2015}, we find that Aldebaran would have had a luminosity $L_* = 2.0 \pm 0.37~\lsun$ on the main sequence.
-Using the Mesa Isochrones and Stellar Tracks \citep[MIST:][]{mist0,mist1} models, we compute the luminosity and radius evolution of the planet along the main sequence.
+Using the Mesa Isochrones and Stellar Tracks \citep[MIST:][]{mist0,mist1} models via the \textsc{isochrones} package \citep{isochrones}, we compute the luminosity and radius evolution of the planet along the main sequence.
 $L$ evolves from $2.0 \pm 0.7$~\lsun at 0.5~Gyr to $3.5 \pm 2.3$~\lsun at 4.5~Gyr.

### 2017-09-29 paper/aldebaran.tex
no duplicate kjeldsen

@@
 In the frequency-power spectrum of either radial velocity of photometric observations, these modes of oscillation form an often intricate pattern of sharp peaks where the properties of the pattern can be mapped back to intrinsic stellar properties \citep[e.g.,][]{2016AN....337..774D}.
-The easiest properties of the pattern to determine are the the frequency of maximum amplitude \numax and the so-called large separation \Dnu \citep{Kjeldsen1995}.
+The easiest properties of the pattern to determine are the the frequency of maximum amplitude \numax and the so-called large separation \Dnu \citep{Kjeldsen95}.
 The two observables are often referred to as the global average seismic parameters.

### 2017-09-29 paper/aldebaran.tex
no duplicate kjeldsen

@@
 Aldebaran, or $\alpha$~Tauri, is a well-known first-magnitude naked-eye red giant star, and has long been the subject of astronomical investigations.
-By looking for Doppler shifts from the star's reflex motion around the common centre of mass with a companion planet \citep[the radial velocity or RV method;][]{struverv}, .
+It was one of the first stars around which an extrasolar planet candidate was identified, by looking for Doppler shifts from the star's reflex motion around the common centre of mass with its companion \citep[the radial velocity or RV method;][]{struverv}.
 While the hot Jupiter 51~Peg~b \citep{51peg} was the first exoplanet to be recognized as such, before this, \citet{hatzes1993} had noted RV variations in Pollux \citep[$\beta$~Gem; subsequently confirmed as a planet:][]{betgemconf,betgemconf2}, Arcturus ($\alpha$~Boo, unconfirmed), and Aldebaran.

### 2017-09-29 paper/aldebaran.tex
tidy introduction

@@
 While the hot Jupiter 51~Peg~b \citep{51peg} was the first exoplanet to be recognized as such, before this, \citet{hatzes1993} had noted RV variations in Pollux \citep[$\beta$~Gem; subsequently confirmed as a planet:][]{betgemconf,betgemconf2}, Arcturus ($\alpha$~Boo, unconfirmed), and Aldebaran.
-After further investigation by \citet{Hatzes1998}, \citet{Hatzes2015} now claim a firm RV detection of a planetary-mass companion Aldebaran~b, with a period of $628.96 \pm 0.90$~d.
+After further investigation by \citet{Hatzes1998}, \citet{Hatzes2015} have now claimed a firm RV detection of a planetary-mass companion Aldebaran~b, with a period of $628.96 \pm 0.90$~d.
 
 In this paper, we present a re-analysis of these original RV data in which we not only confirm this signal, but detect acoustic oscillations in Aldebaran for the first time.
-We validate this method and its result with new RV observations with the SONG~Telescope, and photometry from the \ktwo Mission.
-By measuring the frequency of maximum power ,\numax, of these \emph{p}-mode oscillations we asteroseismically determine the mass of Aldebaran to be $1.16 \pm 0.07$~\msun.
+We validate this method and its result with new independent RV observations with the SONG~Telescope, and photometry from the \ktwo Mission.
+By measuring the frequency of maximum power, \numax, of these \emph{p}-mode oscillations we asteroseismically determine the mass of Aldebaran to be $1.16 \pm 0.07$~\msun.
 This measured stellar mass allows us to calculate that any moons of this giant planet, although they are now likely to be very hot, may have had equilibrium temperatures comparable to that of the Earth when Aldebaran was on the main sequence, raising the possibility that they may have once been habitable.

### 2017-09-29 paper/aldebaran.tex
tidy introduction

@@
 In practise, for the higher luminosity red giants it is more straight forward to measure \numax than \Dnu.
-Typical values for \numax might range from $0.1$ to $20 \rm \mu Hz$ for the luminous giants and $20$ to $50 \rm \mu Hz$ for red giants with luminosities similar to the red clump .
+Typical values for \numax might range from $0.1$ to $20 \rm\,\mu Hz$ for the luminous giants and $20$ to $50 \rm\,\mu Hz$ for red giants with luminosities similar to the red clump .
 For the same sets of stars, \Dnu would take ranges from $0.02$ to $3 \rm \mu Hz$ and $3$ to $7 \rm \mu Hz$ depending on stellar mass and \teff (or metallicity which is considered more fundamental) \citep[e.g.,][]{2011A&A...525L...9M, 2013A&A...559A.137M}.

### 2017-09-29 paper/aldebaran.tex
tidy introduction

@@
 Typical values for \numax might range from $0.1$ to $20 \rm\,\mu Hz$ for the luminous giants and $20$ to $50 \rm\,\mu Hz$ for red giants with luminosities similar to the red clump .
-For the same sets of stars, \Dnu would take ranges from $0.02$ to $3 \rm \mu Hz$ and $3$ to $7 \rm \mu Hz$ depending on stellar mass and \teff (or metallicity which is considered more fundamental) \citep[e.g.,][]{2011A&A...525L...9M, 2013A&A...559A.137M}.
+For the same sets of stars, \Dnu would take ranges from $0.02$ to $3 \rm\,\mu Hz$ and $3$ to $7 \rm\,\mu Hz$ depending on stellar mass and \teff (or metallicity which is considered more fundamental) \citep[e.g.,][]{2011A&A...525L...9M, 2013A&A...559A.137M}.
 Hence, \numax requires less frequency resolution than \Dnu to establish a good measurement.

### 2017-09-29 paper/aldebaran.tex
tidy introduction
added an abstract

@@
 \begin{abstract}
-Stuff.
+The nearby red giant Aldebaran is suspected to host a gas giant planetary companion from decades of ground-based spectroscopic measurements of its Doppler shift.
+Using Gaussian Process based Continuous Auto-Regressive Moving Average (CARMA) models, we confirm this suspected planet and show that this historic data also contains evidence of acoustic oscillations in the star itself, and verify this result with further dedicated ground-based spectroscopy and space-based photometry with the \kepler Space Telescope.
+From the frequency of these oscillations we determine the mass of Aldebaran to be $1.16 \pm 0.07$~\msun, and note that this implies its planet will have been subject to insolation comparable to the Earth for some of the star's main sequence lifetime.
+Our approach to sparse, irregularly sampled time series astronomical data has the potential to unlock similar measurements for hundreds of stars in archival data, and permits more flexible observing schedules in future.
 \end{abstract}
@@
 Aldebaran was therefore observed with \ktwo \citep{howell14}, a two-wheeled revival of the \kepler Mission \citep{2010sci...327..977b}, under Guest Observer Program 130471 in Campaign~13, from 2017-03-08 to 2017-05-27.
+Aldebaran and similar K giants are known to be photometrically variable from Hipparcos photometry \citep{bedding2000}, and we sought to obtain higher-precision, continuous light curves to compare to our CARMA results.
 

### 2017-09-29 paper/aldebaran.tex
added an abstract

@@
 The nearby red giant Aldebaran is suspected to host a gas giant planetary companion from decades of ground-based spectroscopic measurements of its Doppler shift.
-Using Gaussian Process based Continuous Auto-Regressive Moving Average (CARMA) models, we confirm this suspected planet and show that this historic data also contains evidence of acoustic oscillations in the star itself, and verify this result with further dedicated ground-based spectroscopy and space-based photometry with the \kepler Space Telescope.
+Using Gaussian Process-based Continuous Auto-Regressive Moving Average (CARMA) models, we confirm this suspected planet and show that this historic data also contains evidence of acoustic oscillations in the star itself, and verify this result with further dedicated ground-based spectroscopy and space-based photometry with the \kepler Space Telescope.
 From the frequency of these oscillations we determine the mass of Aldebaran to be $1.16 \pm 0.07$~\msun, and note that this implies its planet will have been subject to insolation comparable to the Earth for some of the star's main sequence lifetime.

### 2017-09-29 paper/aldebaran.tex
periods to frequencies

@@
 This introduces a window-function effect: the power spectrum as constructed for example by a Fourier transform, or a Lomb-Scargle Periodogram \citep{lomb,scargle}, is convolved with the Fourier transform of the window function, introducing strong sidelobes adjacent to real frequency peaks and causing crosstalk between adjacent frequency channels.
-This imposes significant limitations both on the signal-to-noise and frequency resolution of power spectra derived from linear methods such as the Lomb-Scargle periodogram, and in practice makes asteroseismology with these conventional approaches difficult or impossible from the ground for stars with oscillation frequencies ranging from $\sim 12$~h to $\sim$~a few days.
+This imposes significant limitations both on the signal-to-noise and frequency resolution of power spectra derived from linear methods such as the Lomb-Scargle periodogram, and in practice makes asteroseismology with these conventional approaches difficult or impossible from the ground for stars with oscillation periods ranging from $\sim 12$~h to $\sim$~a few days.
 

### 2017-10-18 paper/aldebaran.tex
added details on seismic analysis of K2 data

@@
 
-In order to verify the results of the novel analysis presented above, we sought to obtain an independent detection of the oscillations of Aldebaran and compare the frequencies determined with the two methods.
-Aldebaran was therefore observed with \ktwo \citep{howell14}, a two-wheeled revival of the \kepler Mission \citep{2010sci...327..977b}, under Guest Observer Program 130471 in Campaign~13, from 2017-03-08 to 2017-05-27.
-Aldebaran and similar K giants are known to be photometrically variable from Hipparcos photometry \citep{bedding2000}, and we sought to obtain higher-precision, continuous light curves to compare to our CARMA results.
+%In order to verify the results of the novel analysis presented above, we sought to obtain an independent detection of the oscillations of Aldebaran and compare the frequencies determined with the two methods.
+Aldebaran was observed with \ktwo \citep{howell14}, a two-wheeled revival of the \kepler Mission \citep{2010sci...327..977b}, under Guest Observer Program 130471 in Campaign~13, from 2017-03-08 to 2017-05-27.
+Aldebaran and similar high-luminosity K giants are well known to show photometric variability due to solar-like oscillations \citep{bedding2000}, and hence the continuous, high-precision K2 light curves allow an independent confirmation of our CARMA results.
 
@@
 We instead use halo photometry \citep[as originally implemented in][]{White2017}, whereby unsaturated pixels from the outer part of the large and complicated halo of scattered light around bright stars are used to reconstruct a light curve as a weighted linear combination of individual pixel-level time series, as described in Appendix~\ref{halo}.
+
+Figure \ref{k2_lightcurve} shows the light curve and power spectrum of the K2 observations of Aldebaran.
+We detect clear variability on a average timescale of $\sim$\,5\,days, consistent with the expected timescale for solar-like oscillations for a high-luminosity red giant.
+To measure global asteroseismic parameters we model the background variability in the Fourier domain using the methodology described in \citet{huber09}, yielding a frequency of maximum power of $\numax = 2.2 \pm 0.25\,\muHz$ and amplitude per radial mode of $1850 \pm 500$\,ppm.
+Due to the limited frequency resolution of the 70-day K2 timeseries we were unable to measure the large frequency separation \Dnu, which is expected to be on the order of $\approx$\,0.5\muHz.
 
@@
 % \includegraphics[width=0.75\textwidth]{Aldebaran_lc.png}
-\includegraphics[width=0.75\textwidth]{aldebaran_lc.eps}
-\caption{\ktwo lightcurve of Aldebaran.}
+%\includegraphics[width=0.75\textwidth]{aldebaran_lc.eps}
+\includegraphics[width=\textwidth]{k2obs.png}
+\caption{\ktwo lightcurve (left) and power spectrum (right) of Aldebaran.
+The red dashed line shows the background model, and the orange line is a heavily smoothed version of the power spectrum used to measure the frequency of maximum power.
+}
 \label{k2_lightcurve}

### 2017-10-19 paper/aldebaran.tex
conclusion

@@
 
+With CARMA models we have confirmed the previously-suspected planet around Aldebaran using archival data, and additionally detected acoustic oscillations that permit a $\sim 6\%$ mass determination.
+We have confirmed both of these results with new ground-based observations, and space-based photometry with \ktwo.
+
+From this pilot study with Aldebaran, have shown that with limited quantities of data of limited quality, we can nevertheless do precision asteroseismology from the ground and obtain very precise estimates of stellar parameters.
+Sufficient data either already exist, or are trivial to obtain, in order to do this for essentially all bright giants; while for solar-like stars, these new methods allow significantly relaxed observing requirements.
+Furthermore, CARMA models and related methods will enhance deep, all-sky photometric surveys such as \emph{Gaia} in space, or \emph{LSST} from the ground, from which many thousands of new asteroseismic determinations will be possible.
+
 \acknowledgments

### 2017-10-27 paper/aldebaran.tex
suggesting surveys

@@
 From this pilot study with Aldebaran, have shown that with limited quantities of data of limited quality, we can nevertheless do precision asteroseismology from the ground and obtain very precise estimates of stellar parameters.
-Sufficient data either already exist, or are trivial to obtain, in order to do this for essentially all bright giants; while for solar-like stars, these new methods allow significantly relaxed observing requirements.
-Furthermore, CARMA models and related methods will enhance deep, all-sky photometric surveys such as \emph{Gaia} in space, or \emph{LSST} from the ground, from which many thousands of new asteroseismic determinations will be possible.
+Sufficient radial velocity data either already exist, or are trivial to obtain, in order to do this for essentially all bright giants; while for solar-like stars, these new methods allow significantly relaxed observing requirements.
+One such rich archive is the series of observations with the Hamilton \'{E}chelle Spectrograph at Lick Observatory, studying 373 G~and K~giants \citep[e.g.
+][]{frink2001,hekker2006,ortiz2016}.
+The method may also be useful in detecting rotational or other periodic variability in the LCES HIRES/Keck Precision Radial Velocity Exoplanet Survey \citep{butler2017} of 1,624 FGKM dwarfs, although the time sampling is by design too coarse to detect $p$-mode oscillations in these stars.
+Furthermore, CARMA models and related methods will enhance deep, all-sky photometric surveys such as \emph{Gaia} in space \citep{gaia}, or \emph{LSST} from the ground \citep{dmt,lsst,lsstbook}, from which many thousands of new asteroseismic determinations will be possible.
 

### 2017-10-30 paper/aldebaran.tex
comment out bit; do later
clarity

@@
 The ability to measure \numax and \Dnu, therefore, depends on the length and sampling rate of a data set.
-For regularly spaced data, typically analysed using an estimate of the frequency-power spectrum, the Nyquist limit and frequency resolution are trivial to estimate.
-For the historic data sets proposed here the typical length is \guy{blah} and typical sampling is \guy{blah}.
+For regularly spaced data, typically analysed using an estimate of the frequency-power spectrum, the Nyquist limit (half the sampling rate) and frequency resolution ($\sim 1/$~duration) are trivial to estimate.
+%For the historic data sets proposed here the typical length is \guy{blah} and typical sampling is \guy{blah}.
 This in principle gives us access to measures of \numax for \guy{these types of stars} and \Dnu $\,$ \guy{for those types of stars}.

### 2017-10-30 paper/aldebaran.tex
misc

@@
 Sufficient radial velocity data either already exist, or are trivial to obtain, in order to do this for essentially all bright giants; while for solar-like stars, these new methods allow significantly relaxed observing requirements.
-One such rich archive is the series of observations with the Hamilton \'{E}chelle Spectrograph at Lick Observatory, studying 373 G~and K~giants \citep[e.g.
+One rich archive is the series of observations with the Hamilton \'{E}chelle Spectrograph at Lick Observatory, studying 373 G~and K~giants \citep[e.g.
 ][]{frink2001,hekker2006,ortiz2016}.

### 2017-11-10 paper/aldebaran.tex
kallinger confirmation
typo

@@
 We note that with the determination of the stellar mass it is possible to infer the parameters Aldebaran had while it was on the main sequence.
-We conduct a Monte Carlo simulation, drawing masses randomly from a Gaussian distribution $M \sim \mathcal{N}(1.16,0.07)$ and metallicitiies $[\text{Fe}/\text{H}] \sim \mathcal{N}(-0.15,0.2)$ \citep{decin2003}  to predict the luminosity of the main-sequence progenitor, and the semi-major axis of the planet's orbit.
+We conduct a Monte Carlo simulation, drawing masses randomly from a Gaussian distribution $M \sim \mathcal{N}(1.16,0.07)$ and metallicities $[\text{Fe}/\text{H}] \sim \mathcal{N}(-0.15,0.2)$ \citep{decin2003}  to predict the luminosity of the main-sequence progenitor, and the semi-major axis of the planet's orbit.
 %Using the scaling relations of \citep{eker2015}, we find that Aldebaran would have had a luminosity $L_* = 2.0 \pm 0.37~\lsun$ on the main sequence.
@@
 \noindent and likewise second-order TV the equivalent expression in second-order finite differences.
+The efficacy of this method was recently confirmed by \citet{kallinger2017}, comparing BRITE-constellation observations of Atlas to \ktwo halo photometry and finding excellent agreement in frequency and amplitude of the reported oscillations.
 

### 2017-11-15 paper/aldebaran.tex
hipparcos

@@
 The method may also be useful in detecting rotational or other periodic variability in the LCES HIRES/Keck Precision Radial Velocity Exoplanet Survey \citep{butler2017} of 1,624 FGKM dwarfs, although the time sampling is by design too coarse to detect $p$-mode oscillations in these stars.
-Furthermore, CARMA models and related methods will enhance deep, all-sky photometric surveys such as \emph{Gaia} in space \citep{gaia}, or \emph{LSST} from the ground \citep{dmt,lsst,lsstbook}, from which many thousands of new asteroseismic determinations will be possible.
+Furthermore, CARMA models and related methods will enhance deep, all-sky, sparse photometric surveys such as \emph{Gaia} in space \citep{gaia}, or \emph{LSST} from the ground \citep{dmt,lsst,lsstbook}, from which many thousands of new asteroseismic determinations will be possible.
+An immediate future test of this will be from \hipparcos; while only~58 epochs of photometry are available for Aldebaran, a number and sampling insufficient for our purposes \textcolor{red}{check this is true}, stars at higher latitudes may often have 150--200 epochs and more even sampling \citep{hipparcos_phot}, and the~14 bright~K giants in \hipparcos noted by \citet{bedding2000} are an ideal first test case.
 

### 2017-11-18 code/Aldebaran.jl
Lots of random work, including tons of fits of various (p,q) orders that aren't quite working yet.

@@ -1,0 +1,42 @@
+module Aldebaran
+
+using Kepler
+using CARMAKepler
+
+function load_timeseries(before_song=true, song=true)
+    ts = []
+    ys = []
+    dys = []
+
+    if before_song 
+        for i in 3:9
+            data = readdlm("/Users/farr/Documents/Research/Aldebaran/data/$(i)/table$(i).dat")
+            push!(ts, data[:,1])
+            push!(ys, data[:,2])
+            push!(dys, data[:,3])
+        end
+    end
+
+    if song
+        data = readdlm("/Users/farr/Documents/Research/Aldebaran/data/song/tablesong.dat")
+        push!(ts, data[:,1])
+        push!(ys, data[:,2])
+        push!(dys, data[:,3])
+    end
+
+    (ts, ys, dys)
+end
+
+const P_min = 300.0
+const P_max = 1200.0
+
+const K_min = 70.0
+const K_max = 280.0
+
+const mu_Hz = 1e-6*3600.0*24.0 # per day
+
+# We think nu_max is ~2 mu_Hz
+const f_min = 1.0*mu_Hz
+const f_max = 5.0*mu_Hz
+
+end

### 2017-11-18 code/CARMAKepler.jl

@@ -1,0 +1,351 @@
+module CARMAKepler
+
+using CARMA
+using Distributions
+using Ensemble
+using Kepler
+
+type MultiEpochPosterior
+    ts::Array{Array{Float64, 1}, 1}
+    ys::Array{Array{Float64, 1}, 1}
+    dys::Array{Array{Float64, 1}, 1}
+
+    inds::Array{Array{Int, 1}, 1}
+    
+    allts::Array{Float64, 1}
+    allys::Array{Float64, 1}
+    alldys::Array{Float64, 1}
+    
+    P_min::Float64
+    P_max::Float64
+
+    K_min::Float64
+    K_max::Float64
+
+    ndrw::Int
+    nosc::Int
+
+    rate_min::Float64
+    rate_max::Float64
+
+    f_min::Float64
+    f_max::Float64
+
+    rms_min::Float64
+    rms_max::Float64
+
+    Q_min::Float64
+    Q_max::Float64
+end
+
+function MultiEpochPosterior(ts, ys, dys, per_min, per_max, ndrw, nosc, f_min, f_max)
+    allts = vcat(ts...)
+    allys = vcat(ys...)
+    alldys = vcat(dys...)
+    
+    allinds = sortperm(allts)
+    inds = []
+    i = 1
+    for t in ts
+        push!(inds, allinds[i:i+size(t,1)-1])
+        i = i+size(t,1)
+    end
+
+    allts = allts[allinds]
+    allys = allys[allinds]
+    alldys = alldys[allinds]
+
+    rms_max = maximum([std(y) for y in ys])
+    rms_min = minimum([std(y) for y in ys])
+
+    T = allts[end]-allts[1]
+    dtmin = minimum(diff(allts))
+    
+    Q_max = 10.0*f_max*T
+
+    rate_min = 1.0/(10.0*T)
+    rate_max = 10.0/dtmin
+    
+    MultiEpochPosterior(ts, ys, dys, inds, allts, allys, alldys, per_min, per_max, rms_min/100.0, 10.0*rms_max, ndrw, nosc, rate_min, rate_max, f_min, f_max, rms_min/100.0, rms_max*10.0, 0.1, Q_max)
+end
+
+type MultiEpochParams
+    mu::Array{Float64, 1}
+    nu::Array{Float64, 1}
+    
+    K::Float64
+    P::Float64
+    e::Float64
+    omega::Float64
+    chi::Float64
+
+    drw_rms::Array{Float64,1}
[…]

### 2017-11-18 code/Kepler.jl

@@ -1,0 +1,71 @@
+module Kepler
+
+function ecc_anom(M, e)
+    M = mod2pi(M)
+    if M < zero(M)
+        M = M + 2*pi
+    end
+
+    Emin = zero(M)
+    fmin = nothing
+    Emax = 2*pi*one(M)
+    fmax = nothing
+    Eguess = M
+    while true
+        f = M - Eguess + e*sin(Eguess)
+
+        if abs(f) < convert(typeof(f), 1e-10)
+            return Eguess
+        end
+        df = -one(Eguess) + e*cos(Eguess)
+
+        Enewton = Eguess - f/df
+
+        if Enewton < Emin || Enewton > Emax
+            fmin = M - Emin - e*sin(Emin)
+            fmax = M - Emax - e*sin(Emax)
+
+            if fmin == zero(fmin)
+                return Emin
+            elseif fmax == zero(fmin)
+                return Emax
+            end
+
+            if fmin*f < zero(fmin)
+                Emax = Eguess
+            elseif fmax*f < zero(fmax)
+                Emin = Eguess
+            else
+                error("Not bracketing a root in bisection.")
+            end
+            Eguess = (Emax-Emin)/2
+        else
+            Eguess = Enewton
+        end
+    end
+end
+
+function true_anom(ea, e)
+    ta = 2*atan(sqrt((one(e)+e)/(one(e)-e))*tan(ea/2))
+    if ta < zero(ta)
+        ta + 2*pi
+    else
+        ta
+    end
+end
+
+function rv(t, K, P, e, omega, chi)
+    fp, ip = modf(t/P - chi)
+
+    if fp < 0
+        fp = fp + 1.0
+    end
+
+    m = 2*pi*fp
+    
+    ta = true_anom(ecc_anom(m, e), e)
+
+    K*(cos(ta + omega) + e*cos(omega))
+end
+
+end

### 2017-11-19 code/Aldebaran.jl
Revert "Lots of random work, including tons of fits of various (p,q) orders that aren't quite working yet."
This reverts commit e0179fc68d511dbbc05fc0261d85319d1ed165b0.

@@ -1,42 +1,0 @@
-module Aldebaran
-
-using Kepler
-using CARMAKepler
-
-function load_timeseries(before_song=true, song=true)
-    ts = []
-    ys = []
-    dys = []
-
-    if before_song 
-        for i in 3:9
-            data = readdlm("/Users/farr/Documents/Research/Aldebaran/data/$(i)/table$(i).dat")
-            push!(ts, data[:,1])
-            push!(ys, data[:,2])
-            push!(dys, data[:,3])
-        end
-    end
-
-    if song
-        data = readdlm("/Users/farr/Documents/Research/Aldebaran/data/song/tablesong.dat")
-        push!(ts, data[:,1])
-        push!(ys, data[:,2])
-        push!(dys, data[:,3])
-    end
-
-    (ts, ys, dys)
-end
-
-const P_min = 300.0
-const P_max = 1200.0
-
-const K_min = 70.0
-const K_max = 280.0
-
-const mu_Hz = 1e-6*3600.0*24.0 # per day
-
-# We think nu_max is ~2 mu_Hz
-const f_min = 1.0*mu_Hz
-const f_max = 5.0*mu_Hz
-
-end

### 2017-11-19 code/CARMAKepler.jl

@@ -1,351 +1,0 @@
-module CARMAKepler
-
-using CARMA
-using Distributions
-using Ensemble
-using Kepler
-
-type MultiEpochPosterior
-    ts::Array{Array{Float64, 1}, 1}
-    ys::Array{Array{Float64, 1}, 1}
-    dys::Array{Array{Float64, 1}, 1}
-
-    inds::Array{Array{Int, 1}, 1}
-    
-    allts::Array{Float64, 1}
-    allys::Array{Float64, 1}
-    alldys::Array{Float64, 1}
-    
-    P_min::Float64
-    P_max::Float64
-
-    K_min::Float64
-    K_max::Float64
-
-    ndrw::Int
-    nosc::Int
-
-    rate_min::Float64
-    rate_max::Float64
-
-    f_min::Float64
-    f_max::Float64
-
-    rms_min::Float64
-    rms_max::Float64
-
-    Q_min::Float64
-    Q_max::Float64
-end
-
-function MultiEpochPosterior(ts, ys, dys, per_min, per_max, ndrw, nosc, f_min, f_max)
-    allts = vcat(ts...)
-    allys = vcat(ys...)
-    alldys = vcat(dys...)
-    
-    allinds = sortperm(allts)
-    inds = []
-    i = 1
-    for t in ts
-        push!(inds, allinds[i:i+size(t,1)-1])
-        i = i+size(t,1)
-    end
-
-    allts = allts[allinds]
-    allys = allys[allinds]
-    alldys = alldys[allinds]
-
-    rms_max = maximum([std(y) for y in ys])
-    rms_min = minimum([std(y) for y in ys])
-
-    T = allts[end]-allts[1]
-    dtmin = minimum(diff(allts))
-    
-    Q_max = 10.0*f_max*T
-
-    rate_min = 1.0/(10.0*T)
-    rate_max = 10.0/dtmin
-    
-    MultiEpochPosterior(ts, ys, dys, inds, allts, allys, alldys, per_min, per_max, rms_min/100.0, 10.0*rms_max, ndrw, nosc, rate_min, rate_max, f_min, f_max, rms_min/100.0, rms_max*10.0, 0.1, Q_max)
-end
-
-type MultiEpochParams
-    mu::Array{Float64, 1}
-    nu::Array{Float64, 1}
-    
-    K::Float64
-    P::Float64
-    e::Float64
-    omega::Float64
-    chi::Float64
-
-    drw_rms::Array{Float64,1}
[…]

### 2017-11-19 code/Kepler.jl

@@ -1,71 +1,0 @@
-module Kepler
-
-function ecc_anom(M, e)
-    M = mod2pi(M)
-    if M < zero(M)
-        M = M + 2*pi
-    end
-
-    Emin = zero(M)
-    fmin = nothing
-    Emax = 2*pi*one(M)
-    fmax = nothing
-    Eguess = M
-    while true
-        f = M - Eguess + e*sin(Eguess)
-
-        if abs(f) < convert(typeof(f), 1e-10)
-            return Eguess
-        end
-        df = -one(Eguess) + e*cos(Eguess)
-
-        Enewton = Eguess - f/df
-
-        if Enewton < Emin || Enewton > Emax
-            fmin = M - Emin - e*sin(Emin)
-            fmax = M - Emax - e*sin(Emax)
-
-            if fmin == zero(fmin)
-                return Emin
-            elseif fmax == zero(fmin)
-                return Emax
-            end
-
-            if fmin*f < zero(fmin)
-                Emax = Eguess
-            elseif fmax*f < zero(fmax)
-                Emin = Eguess
-            else
-                error("Not bracketing a root in bisection.")
-            end
-            Eguess = (Emax-Emin)/2
-        else
-            Eguess = Enewton
-        end
-    end
-end
-
-function true_anom(ea, e)
-    ta = 2*atan(sqrt((one(e)+e)/(one(e)-e))*tan(ea/2))
-    if ta < zero(ta)
-        ta + 2*pi
-    else
-        ta
-    end
-end
-
-function rv(t, K, P, e, omega, chi)
-    fp, ip = modf(t/P - chi)
-
-    if fp < 0
-        fp = fp + 1.0
-    end
-
-    m = 2*pi*fp
-    
-    ta = true_anom(ecc_anom(m, e), e)
-
-    K*(cos(ta + omega) + e*cos(omega))
-end
-
-end

### 2017-11-19 code/Periodogram.jl

@@ -1,82 +1,0 @@
-module Periodogram
-
-function generate_mean_vectors(ts)
-    nv = size(ts, 1)
-
-    ntot = sum([size(t,1) for t in ts])
-
-    mvs = zeros(ntot, nv)
-
-    i = 1
-    for j in eachindex(ts)
-        t = ts[j]
-
-        k = i + size(t, 1) - 1
-        mvs[i:k, j] = 1.0
-
-        i = k+1
-    end
-
-    mvs
-end
-
-function response_matrix(ts, ys, dys, f)
-    mvs = generate_mean_vectors(ts)
-
-    all_ts = vcat(ts...)
-    all_ys = vcat(ys...)
-    all_dys = vcat(dys...)
-
-    cv = cos.(all_ts.*(2.0*pi*f))
-    sv = sin.(all_ts.*(2.0*pi*f))
-
-    all_ts, all_ys, all_dys, hcat(mvs, cv, sv)
-end
-
-function response_matrix(ts, ys, dys, fs::Array{Float64, 1})
-    mvs = generate_mean_vectors(ts)
-
-    all_ts = vcat(ts...)
-    all_ys = vcat(ys...)
-    all_dys = vcat(dys...)
-
-    csvs = zeros(size(all_ts, 1), 2*size(fs, 1))
-
-    for i in eachindex(all_ts)
-        for j in eachindex(fs)
-            csvs[i,2*j-1] = cos(2.0*pi*fs[j]*all_ts[i])
-            csvs[i,2*j] = sin(2.0*pi*fs[j]*all_ts[i])
-        end
-    end
-
-    all_ts, all_ys, all_dys, hcat(mvs, csvs)
-end
-
-function pgram(ts, ys, dys, f)
-    all_ts, all_ys, all_dys, A = response_matrix(ts, ys, dys, f)
-
-    for i in eachindex(all_ys)
-        all_ys[i] *= 1.0./all_dys[i]
-        A[i,1:end] *= 1.0./all_dys[i]
-    end
-
-    A\all_ys
-end
-
-function residual(ts, ys, dys, f, pgram)
-    all_ts, all_ys, all_dys, A = response_matrix(ts, ys, dys, f)
-
-    all_rs = all_ys - A*pgram
-
-    rs = Array{Float64, 1}[]
-    i = 1
-    for j in eachindex(ts)
-        k = i + size(ys[j],1) - 1
-        push!(rs, all_rs[i:k])
-        i = k+1
-    end
-
-    rs
-end
-
-end

### 2017-11-19 code/CARMAKepler.jl
Initial work on LargeSpacing estimation.

@@ -1,0 +1,351 @@
+module CARMAKepler
+
+using CARMA
+using Distributions
+using Ensemble
+using Kepler
+
+type MultiEpochPosterior
+    ts::Array{Array{Float64, 1}, 1}
+    ys::Array{Array{Float64, 1}, 1}
+    dys::Array{Array{Float64, 1}, 1}
+
+    inds::Array{Array{Int, 1}, 1}
+    
+    allts::Array{Float64, 1}
+    allys::Array{Float64, 1}
+    alldys::Array{Float64, 1}
+    
+    P_min::Float64
+    P_max::Float64
+
+    K_min::Float64
+    K_max::Float64
+
+    ndrw::Int
+    nosc::Int
+
+    rate_min::Float64
+    rate_max::Float64
+
+    f_min::Float64
+    f_max::Float64
+
+    rms_min::Float64
+    rms_max::Float64
+
+    Q_min::Float64
+    Q_max::Float64
+end
+
+function MultiEpochPosterior(ts, ys, dys, per_min, per_max, ndrw, nosc, f_min, f_max)
+    allts = vcat(ts...)
+    allys = vcat(ys...)
+    alldys = vcat(dys...)
+    
+    allinds = sortperm(allts)
+    inds = []
+    i = 1
+    for t in ts
+        push!(inds, allinds[i:i+size(t,1)-1])
+        i = i+size(t,1)
+    end
+
+    allts = allts[allinds]
+    allys = allys[allinds]
+    alldys = alldys[allinds]
+
+    rms_max = maximum([std(y) for y in ys])
+    rms_min = minimum([std(y) for y in ys])
+
+    T = allts[end]-allts[1]
+    dtmin = minimum(diff(allts))
+    
+    Q_max = 10.0*f_max*T
+
+    rate_min = 1.0/(10.0*T)
+    rate_max = 10.0/dtmin
+    
+    MultiEpochPosterior(ts, ys, dys, inds, allts, allys, alldys, per_min, per_max, rms_min/100.0, 10.0*rms_max, ndrw, nosc, rate_min, rate_max, f_min, f_max, rms_min/100.0, rms_max*10.0, 0.1, Q_max)
+end
+
+type MultiEpochParams
+    mu::Array{Float64, 1}
+    nu::Array{Float64, 1}
+    
+    K::Float64
+    P::Float64
+    e::Float64
+    omega::Float64
+    chi::Float64
+
+    drw_rms::Array{Float64,1}
[…]

### 2017-11-19 paper/aldebaran.tex

--- 2017-11-10 paper/aldebaran.tex
@@
 From this pilot study with Aldebaran, have shown that with limited quantities of data of limited quality, we can nevertheless do precision asteroseismology from the ground and obtain very precise estimates of stellar parameters.
+Furthermore, improvements in the algorithm may also permit the detection of not only~\numax but also~\Dnu from similar measurements, permitting more sophisticated asteroseismic analysis.
 Sufficient radial velocity data either already exist, or are trivial to obtain, in order to do this for essentially all bright giants; while for solar-like stars, these new methods allow significantly relaxed observing requirements.
@@
 The method may also be useful in detecting rotational or other periodic variability in the LCES HIRES/Keck Precision Radial Velocity Exoplanet Survey \citep{butler2017} of 1,624 FGKM dwarfs, although the time sampling is by design too coarse to detect $p$-mode oscillations in these stars.
-Furthermore, CARMA models and related methods will enhance deep, all-sky photometric surveys such as \emph{Gaia} in space \citep{gaia}, or \emph{LSST} from the ground \citep{dmt,lsst,lsstbook}, from which many thousands of new asteroseismic determinations will be possible.
+Furthermore, CARMA models and related methods will enhance deep, all-sky, sparse photometric surveys: an immediate future test of this will be from \hipparcos; while only~58 epochs of photometry are available for Aldebaran, a number and sampling insufficient for our purposes \textcolor{red}{check this is true}, stars at higher latitudes may often have 150--200 epochs and more even sampling \citep{hipparcos_phot}, and the~14 bright~K giants in \hipparcos noted by \citet{bedding2000} are an ideal first test case.
+This can naturally be extended to \emph{Gaia} in space \citep{gaia}, or \emph{LSST} from the ground \citep{dmt,lsst,lsstbook}, from which many thousands of new asteroseismic determinations will be possible.
 
@@
 We thank the academy....
+
+We would like to thank David Hogg and Dan Foreman-Mackey for their helpful discussions.
 

### 2017-11-20 code/Periodogram.jl
Files brought over from reverted commit.

@@ -1,0 +1,82 @@
+module Periodogram
+
+function generate_mean_vectors(ts)
+    nv = size(ts, 1)
+
+    ntot = sum([size(t,1) for t in ts])
+
+    mvs = zeros(ntot, nv)
+
+    i = 1
+    for j in eachindex(ts)
+        t = ts[j]
+
+        k = i + size(t, 1) - 1
+        mvs[i:k, j] = 1.0
+
+        i = k+1
+    end
+
+    mvs
+end
+
+function response_matrix(ts, ys, dys, f)
+    mvs = generate_mean_vectors(ts)
+
+    all_ts = vcat(ts...)
+    all_ys = vcat(ys...)
+    all_dys = vcat(dys...)
+
+    cv = cos.(all_ts.*(2.0*pi*f))
+    sv = sin.(all_ts.*(2.0*pi*f))
+
+    all_ts, all_ys, all_dys, hcat(mvs, cv, sv)
+end
+
+function response_matrix(ts, ys, dys, fs::Array{Float64, 1})
+    mvs = generate_mean_vectors(ts)
+
+    all_ts = vcat(ts...)
+    all_ys = vcat(ys...)
+    all_dys = vcat(dys...)
+
+    csvs = zeros(size(all_ts, 1), 2*size(fs, 1))
+
+    for i in eachindex(all_ts)
+        for j in eachindex(fs)
+            csvs[i,2*j-1] = cos(2.0*pi*fs[j]*all_ts[i])
+            csvs[i,2*j] = sin(2.0*pi*fs[j]*all_ts[i])
+        end
+    end
+
+    all_ts, all_ys, all_dys, hcat(mvs, csvs)
+end
+
+function pgram(ts, ys, dys, f)
+    all_ts, all_ys, all_dys, A = response_matrix(ts, ys, dys, f)
+
+    for i in eachindex(all_ys)
+        all_ys[i] *= 1.0./all_dys[i]
+        A[i,1:end] *= 1.0./all_dys[i]
+    end
+
+    A\all_ys
+end
+
+function residual(ts, ys, dys, f, pgram)
+    all_ts, all_ys, all_dys, A = response_matrix(ts, ys, dys, f)
+
+    all_rs = all_ys - A*pgram
+
+    rs = Array{Float64, 1}[]
+    i = 1
+    for j in eachindex(ts)
+        k = i + size(ys[j],1) - 1
+        push!(rs, all_rs[i:k])
+        i = k+1
+    end
+
+    rs
+end
+
+end

### 2017-11-21 paper/aldebaran.tex
wording

@@
 The nearby red giant Aldebaran is suspected to host a gas giant planetary companion from decades of ground-based spectroscopic measurements of its Doppler shift.
-Using Gaussian Process-based Continuous Auto-Regressive Moving Average (CARMA) models, we confirm this suspected planet and show that this historic data also contains evidence of acoustic oscillations in the star itself, and verify this result with further dedicated ground-based spectroscopy and space-based photometry with the \kepler Space Telescope.
+Using Gaussian Process-based Continuous Auto-Regressive Moving Average (CARMA) models, we confirm this suspected planet and show that these historic data also contain evidence of acoustic oscillations in the star itself, and verify this result with further dedicated ground-based spectroscopy and space-based photometry with the \kepler Space Telescope.
 From the frequency of these oscillations we determine the mass of Aldebaran to be $1.16 \pm 0.07$~\msun, and note that this implies its planet will have been subject to insolation comparable to the Earth for some of the star's main sequence lifetime.
-Our approach to sparse, irregularly sampled time series astronomical data has the potential to unlock similar measurements for hundreds of stars in archival data, and permits more flexible observing schedules in future.
+Our approach to sparse, irregularly sampled time series astronomical observations has the potential to unlock similar measurements for hundreds of stars in archival data, and permits more flexible observing schedules in future.
 \end{abstract}

### 2017-12-11 paper/aldebaran.tex
bedding
new address

@@
 \affiliation{Sydney Institute for Astronomy, School of Physics, University of Sydney, Sydney NSW 2006, Australia} 
-\affiliation{Center for Cosmology and Particle Physics, Department of Physics, New York University, 4 Washington Place, New York, NY 10003, USA}
+\affiliation{Center for Cosmology and Particle Physics, Department of Physics, New York University, 726 Broadway, New York, NY 10003, USA}
 \affiliation{NASA Sagan Fellow}
@@
 
-We would like to thank David Hogg and Dan Foreman-Mackey for their helpful discussions.
+We would like to thank Tim Bedding, David Hogg and Dan Foreman-Mackey for their helpful discussions.
 

### 2017-12-13 paper/aldebaran.tex
changes

@@
 
-\subsection{Stellar Parameters}
+\subsection{Planetary and Stellar Parameters}
 
-We have determined the stellar properties of Aldebaran with the additional constraint of \numax determined from our analysis here.
+By jointly fitting a Keplerian for the planet with a CARMA model for the stellar oscillations, we obtain a precise estimate of both \numax and the planetary orbital parameters.
+The planetary parameters we obtain are similar to \citet{Hatzes2015}: period $P = 629.0^{+ 1.7}_{-2.1}$~d, eccentricity $e = 0.168^{+ 0.069}_{- 0.071}$; and radial velocity semi-amplitude $K = 127^{+ 15}_{- 14}$~m/s.
+
+The stellar oscillations are best fit with a single CARMA component with $\numax = 2.238^{+ 0.086}_{- 0.080}$ \muHz, and including additional CARMA components does not resolve any further structure in the power spectrum.
+We have determined the stellar properties of Aldebaran with the additional constraint of \numax determined from this analysis.
 The details of the stellar modelling are presented in detail in \ref{sec:mod}.
@@
 It is clear that the addition of \numax provides substantially more precise estimates of mass and age.
+With this stellar mass, the radial velocity signal translates to an inclination-dependent planetary mass of $M\,\sin\,{i} = 5.81 \pm 0.72~M_J$ \citep[Equation~1]{Torres2008}, a massive giant planet; for sufficiently high inclinations ($\gtrsim 45^\circ$), the mass of Aldebaran~b may exceed $13~M_J$, making it a possible low-mass brown dwarf candidate.
 % re sheffield - I have substituted the citation from the .bib but you said 'b'; afaik she only published one paper in 2012?
@@
 
-We note that with the determination of the stellar mass it is possible to infer the parameters Aldebaran had while it was on the main sequence.
+With the determination of the stellar mass it is possible to infer the parameters Aldebaran had while it was on the main sequence.
 We conduct a Monte Carlo simulation, drawing masses randomly from a Gaussian distribution $M \sim \mathcal{N}(1.16,0.07)$ and metallicities $[\text{Fe}/\text{H}] \sim \mathcal{N}(-0.15,0.2)$ \citep{decin2003} to predict the luminosity of the main-sequence progenitor, and the semi-major axis of the planet's orbit.
 %Using the scaling relations of \citep{eker2015}, we find that Aldebaran would have had a luminosity $L_* = 2.0 \pm 0.37~\lsun$ on the main sequence.
-Using the Mesa Isochrones and Stellar Tracks \citep[MIST:][]{mist0,mist1} models via the \textsc{isochrones} package \citep{isochrones}, we compute the luminosity and radius evolution of the planet along the main sequence.
-$L$ evolves from $2.0 \pm 0.7$~\lsun at 0.5~Gyr to $3.5 \pm 2.3$~\lsun at 4.5~Gyr.
-We note that this implies that the planet at a semi-major axis of $1.507 \pm 0.03$~AU would have been subject to an insolation comparable to that of the Earth, evolving from 0.5~Gyr to 4.5~Gyr from $0.86 \pm 0.27$ to $1.55 \pm 1.0$ times that of the Earth today.
-Subject as well as this to the great uncertainties of the planet's orbital evolution and albedo, this planet and any of its moons may well have had temperate environments for some of their previous evolution, now long-since destroyed by their host star's evolution away from the main sequence.
+Using the Mesa Isochrones and Stellar Tracks \citep[MIST:][]{mist0,mist1} models via the \textsc{isochrones} package \citep{isochrones}, we compute the evolution of the luminosity $L$ of the star along the main sequence; $L$ evolves from $2.0 \pm 0.7$~\lsun at 0.5~Gyr to $3.5 \pm 2.3$~\lsun at 4.5~Gyr.
+We note that this implies that the planet at a semi-major axis of $1.50\, \pm\, 0.03$~AU would have been subject to an insolation comparable to that of the Earth, evolving from 0.5~Gyr to 4.5~Gyr from $0.86\, \pm\, 0.27$ to $1.55 \,\pm\,1.0$ times that of the Earth today.
+Subject as well as this to the great uncertainties of the planet's orbital evolution and albedo, Aldebaran~b and any of its moons may well have hosted temperate environments for some of their history, now long-since destroyed by their star's evolution away from the main sequence.
 
@@
 ][]{frink2001,hekker2006,ortiz2016}.
-The method may also be useful in detecting rotational or other periodic variability in the LCES HIRES/Keck Precision Radial Velocity Exoplanet Survey \citep{butler2017} of 1,624 FGKM dwarfs, although the time sampling is by design too coarse to detect $p$-mode oscillations in these stars.
+The method may also be useful in detecting rotational or other periodic variability in the LCES HIRES/Keck Precision Radial Velocity Exoplanet Survey \citep{butler17} of 1,624 FGKM dwarfs, although the time sampling is by design too coarse to detect $p$-mode oscillations in these stars.
 Furthermore, CARMA models and related methods will enhance deep, all-sky, sparse photometric surveys: an immediate future test of this will be from \hipparcos; while only~58 epochs of photometry are available for Aldebaran, a number and sampling insufficient for our purposes \textcolor{red}{check this is true}, stars at higher latitudes may often have 150--200 epochs and more even sampling \citep{hipparcos_phot}, and the~14 bright~K giants in \hipparcos noted by \citet{bedding2000} are an ideal first test case.

### 2017-12-15 code/CARMAKepler.jl
First stab at predicted state with Kepler RV and stochastic process.

@@ -350,6 +350,24 @@ function psd(post::MultiEpochPosterior, p::MultiEpochParams, fs::Array{Float64, 
     Celerite.psd(filt, fs)
 end
 
+function predict(post::MultiEpochPosterior, x::Array{Float64, 1}, ts::Array{Float64, 1})
+    predict(post, to_params(post, x), ts)
+end
+
+function predict(post::MultiEpochPosterior, p::MultiEpochParams, ts::Array{Float64, 1})
+    rvs = [Kepler.rv(t, p.K, p.P, p.e, p.omega, p.chi) for t in ts]
+
+    ys, dys = produce_ys_dys(post, p)
+
+    filt = Celerite.CeleriteKalmanFilter(0.0, p.drw_rms, p.drw_rate, p.osc_rms, p.osc_freq, p.osc_Q)
+
+    ysp, vysp = Celerite.predict(filt, post.allts, ys, dys, ts)
+
+    ysp = ysp .+ rvs
+
+    ysp, vysp
+end
+
 function posterior_predictive(post::MultiEpochPosterior, x::Array{Float64, 1})
     posterior_predictive(post, to_params(post, x))
 end

### 2017-12-15 paper/aldebaran.tex
gadigal
space
orbital

@@
 By jointly fitting a Keplerian for the planet with a CARMA model for the stellar oscillations, we obtain a precise estimate of both \numax and the planetary orbital parameters.
-The planetary parameters we obtain are similar to \citet{Hatzes2015}: period $P = 629.0^{+ 1.7}_{-2.1}$~d, eccentricity $e = 0.168^{+ 0.069}_{- 0.071}$; and radial velocity semi-amplitude $K = 127^{+ 15}_{- 14}$~m/s.
+The planetary orbital parameters we obtain are similar to \citet{Hatzes2015}: period $P = 629.0^{+ 1.7}_{-2.1}$~d, eccentricity $e = 0.168^{+ 0.069}_{- 0.071}$; and radial velocity semi-amplitude $K = 127^{+ 15}_{- 14}$~m/s.
 
@@
 This work was performed in part under contract with the Jet Propulsion Laboratory (JPL) funded by NASA through the Sagan Fellowship Program executed by the NASA Exoplanet Science Institute.
+
+The authors would like to acknowledge the Gadigal people of the Eora nation on whose ancestral lands the University of Sydney is built.
+
 Funding for the Stellar Astrophysics Centre is provided by The Danish National Research Foundation. TRW acknowledges the support of the Villum Foundation (research grant 10118).

### 2017-12-18 paper/aldebaran.tex
citation

@@
 As Aldebaran is extremely bright, it saturates the \kepler detector and it is therefore not possible to use standard photometry pipelines to extract a K2 lightcurve.
-We instead use halo photometry \citep[as originally implemented in][]{White2017}, whereby unsaturated pixels from the outer part of the large and complicated halo of scattered light around bright stars are used to reconstruct a light curve as a weighted linear combination of individual pixel-level time series, as described in Appendix~\ref{halo}.
+We instead use halo photometry \citep{White2017}, whereby unsaturated pixels from the outer part of the large and complicated halo of scattered light around bright stars are used to reconstruct a light curve as a weighted linear combination of individual pixel-level time series, as described in Appendix~\ref{halo}.
 

### 2017-12-18 paper/aldebaran.tex
hipparcos

@@
 The method may also be useful in detecting rotational or other periodic variability in the LCES HIRES/Keck Precision Radial Velocity Exoplanet Survey \citep{butler17} of 1,624 FGKM dwarfs, although the time sampling is by design too coarse to detect $p$-mode oscillations in these stars.
-Furthermore, CARMA models and related methods will enhance deep, all-sky, sparse photometric surveys: an immediate future test of this will be from \hipparcos; while only~58 epochs of photometry are available for Aldebaran, a number and sampling insufficient for our purposes \textcolor{red}{check this is true}, stars at higher latitudes may often have 150--200 epochs and more even sampling \citep{hipparcos_phot}, and the~14 bright~K giants in \hipparcos noted by \citet{bedding2000} are an ideal first test case.
+Furthermore, CARMA models and related methods will enhance deep, all-sky, sparse photometric surveys: an immediate future test of this will be from \hipparcos; while only~58 epochs of photometry are available for Aldebaran, at a sampling we find to be insufficient for our purposes, stars at higher latitudes may often have 150--200 epochs and more even sampling \citep{hipparcos_phot}, and the~14 bright~K giants in \hipparcos noted by \citet{bedding2000} are an ideal first test case.
 This can naturally be extended to \emph{Gaia} in space \citep{gaia}, or \emph{LSST} from the ground \citep{dmt,lsst,lsstbook}, from which many thousands of new asteroseismic determinations will be possible.

### 2017-12-21 paper/aldebaran.tex
Added input from Frank/SONG

@@
 
+\author{Victoria Antoci}
+\affiliation{ Stellar Astrophysics Centre (SAC), Department of Physics and Astronomy, Aarhus University, Ny Munkegade 120, DK-8000, Denmark}
+
+\author{Mads Fredslund Andersen} 
+\affiliation{ Stellar Astrophysics Centre (SAC), Department of Physics and Astronomy, Aarhus University, Ny Munkegade 120, DK-8000, Denmark}
+\author{Frank Grundahl}
+\affiliation{ Stellar Astrophysics Centre (SAC), Department of Physics and Astronomy, Aarhus University, Ny Munkegade 120, DK-8000, Denmark}
+
 \email{w.farr@bham.ac.uk, G.R.Davies@bham.ac.uk, b.pope@sydney.edu.au, txn016@student.bham.ac.uk, jimbarrett27@gmail.com}
@@
 Our new approach to asteroseismic data analysis, based on Continuous Auto-Regressive Moving Average (CARMA) models, can extract exoplanet signals together with measures of \numax from sparse and irregularly-sampled time series.
-An all-sky survey to find planetary companions and to precisely measure the masses of all nearby red giant stars is feasible with this new approach, and the required data either already exist in large exoplanet surveys, or are easy to obtain with ground-based telescopes.
+An all-sky survey to find planetary companions and to precisely measure the masses of all nearby red giant stars is feasible with this new approach, and the required data either already exist in large radial velocity exoplanet surveys, or are easy to obtain with ground-based telescopes.
 
@@
 
-\subsection{SONG Observations}
+\subsection{SONG Observations}    % this goes into section 4.1
+
+In order to detect the oscillations of Aldebaran via the radial velocity 
+method we used the Hertzsprung SONG telescope \cite{2017ApJ...836..142G}. All 
+observations were carried out in the highest resolution mode (R$\approx\,110.000$)
+with an integration time of 30s. The observations were carried out between 27 September 
+2016 and  12 January 2017, during this time we attempted to obtain at least one 
+visit per available night. A total of 254 spectra was obtained. The spectra were
+extracted using the SONG pipeline (see \cite{2017ApJ...836..142G}). SONG employs an iodine cell for
+precise wavelength calibration and the velocities were determined using the $\it iSONG$
+software following the same procedures as described in \cite{2017ApJ...836..142G}.   The typical
+velocity precision per visit is 2.5$m\,s^{-1}$ allowing us to easily detect the oscillations
+which displays a peak-to-peak amplitude of $\approx \, 170 \, \rm m \, s^{-1}$. $\bf Figure XX$ show
+the radial velocities obtained -- oscillations are easily visible as well as a long-term
+trend. This is probably due to the a manifestation of the planetary companion reported by
+\cite{Hatzes2015}.
 
@@
 Support for MAST for non-HST data is provided by the NASA Office of Space Science via grant NNX13AC07G and by other grants and contracts.
+
+ The Villum Foundation, The Danish Council for Independent Research | Natural Sciences and 
+ the Carlsberg Foundation are thanked for their support on building the SONG prototype on 
+ Tenerife. 
+ We thank Jens Jessen-Hansen, Rasmus Handberg and René Tronsgaard Rasmussen for their contribution
+ to the development of the SONG data reduction pipeline and archive system.
+
+ Based on observations made with the Hertzsprung SONG telescope operated on the Spanish Observatorio
+  del Teide on the island of Tenerife by the Aarhus and Copenhagen Universities and by the Instituto de 
+  Astrofisica de Canarias.
+
 This work made use of the \textsc{IPython} package \citep{PER-GRA:2007}; SciPy \citep{scipy};  \textsc{matplotlib}, a \textsc{Python} library for publication quality graphics \citep{Hunter:2007}; \textsc{Astropy}, a community-developed core \textsc{Python} package for astronomy \citep{2013A&A...558A..33A}; and \textsc{NumPy} \citep{van2011numpy}.

### 2017-12-21 paper/aldebaran.tex
affiliation

@@
 
-\email{w.farr@bham.ac.uk, G.R.Davies@bham.ac.uk, b.pope@sydney.edu.au, txn016@student.bham.ac.uk, jimbarrett27@gmail.com}
+\email{w.farr@bham.ac.uk, G.R.Davies@bham.ac.uk, benjamin.pope@nyu.edu, txn016@student.bham.ac.uk, jimbarrett27@gmail.com}
 
@@
 
- The Villum Foundation, The Danish Council for Independent Research | Natural Sciences and 
+The Villum Foundation, The Danish Council for Independent Research~|~Natural Sciences and 
  the Carlsberg Foundation are thanked for their support on building the SONG prototype on 

### 2017-12-22 presub/natastron_presub_abstract.tex
presubmission letter and abstract draft

--- 2017-12-18 paper/aldebaran.tex
@@
-% !TEX TS-program = pdflatexmk
-
-\documentclass[modern]{aastex61}
-\usepackage[utf8]{inputenc}
+\documentclass{natureprintstyle}
+%\documentclass{nature}
+\bibliographystyle{naturemag}
+\usepackage{epsfig,caption}
+\usepackage{color}
+\usepackage{bm}
+\usepackage{graphicx}
+\usepackage{longtable}
+\usepackage{amssymb}
+\usepackage{rotating}
+\usepackage{latexsym}
 \usepackage{hyperref}
-\usepackage{natbib}
-\usepackage{graphicx}
+\usepackage{hyperref}
+% \usepackage{natbib}
 \usepackage{xspace}
 \usepackage{amsmath}
-\usepackage{amssymb}
 \usepackage{mathrsfs}
 
-\include{macros}
-% Journals
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%journal commands
+\newcommand{\apj}{Astrophys. J.}
+\newcommand{\spie}{Proc. SPIE}
+\newcommand{\pasp}{Publ. Astron. Soc. Pac.}
+\newcommand{\apjs}{Astrophys. J. Supp.}
+\newcommand{\araa}{Annu. Rev. Astron. Astrophys.}
+\newcommand{\mnras}{Mon. Not. R. Astron. Soc.}
+\newcommand{\apjl}{Astrophys. J. Let.}
+\newcommand{\aap}{Astron. Astrophys.}
+\newcommand{\aj}{Astron. J.}
+\newcommand{\nat}{Nature}
+\newcommand{\na}{New Astron. Rev.}
+\newcommand{\aaps}{A\&AS}
+\newcommand{\procspie}{Proc. SPIE}
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % Astronomical abbreviations (thanks to Dan Huber)
@@
 \newcommand{\guy}[1]{{\bf \color{blue} #1}}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% title brainstorm:
+% According to nature astronomy style guide: "a title of, at most, 90 characters (including spaces)"
+%formal document says 100 characters
+\title{Detection of Oscillations in Aldebaran with Ground-Based Observations}
+
+\author{Will M. Farr$^{1,2}$\thanks{w.farr@bham.ac.uk}, Guy R. Davies$^{2,3}$, Benjamin J. S. Pope$^{4,5,6}$, et al.}
 
@@
 
-\title{Detection of Oscillations in Aldebaran with Ground-Based Observations}
+\maketitle
 
-\author[0000-0003-1540-8562]{Will M. Farr}
+\let\thefootnote\relax\footnote{
+\begin{affiliations}
+\item {Birmingham Institute for Gravitational Wave Astronomy,
+  University of Birmingham, Birmingham, B15 2TT, United Kingdom}
+\item {School of Physics and Astronomy, University of
+  Birmingham, Birmingham, B15 2TT, United Kingdom}
+\item {Stellar Astrophysics Centre, Department of Physics 
+and Astronomy, Aarhus University, Ny Munkegade 120, DK-8000 Aarhus C, Denmark}
+\item {Sydney Institute for Astronomy, School of Physics, University of Sydney, Sydney NSW 2006, Australia} 
+\item {Center for Cosmology and Particle Physics, Department of Physics, New York University, 726 Broadway, New York, NY 10003, USA}
+\item {NASA Sagan Fellow}
 
-\affiliation{Birmingham Institute for Gravitational Wave Astronomy,
-  University of Birmingham, Birmingham, B15 2TT, United Kingdom}
-
-\affiliation{School of Physics and Astronomy, University of
-  Birmingham, Birmingham, B15 2TT, United Kingdom}
-
-\author[0000-0002-4290-7351]{Guy R. Davies}
-
-\affiliation{School of Physics and Astronomy, University of
-  Birmingham, Birmingham, B15 2TT, United Kingdom}
-\affiliation{Stellar Astrophysics Centre, Department of Physics 
-and Astronomy, Aarhus University, Ny Munkegade 120, DK-8000 Aarhus C, Denmark}
-
-\author[0000-0003-2595-9114]{Benjamin J. S. Pope}
-\affiliation{Sydney Institute for Astronomy, School of Physics, University of Sydney, Sydney NSW 2006, Australia} 
-\affiliation{Center for Cosmology and Particle Physics, Department of Physics, New York University, 726 Broadway, New York, NY 10003, USA}
-\affiliation{NASA Sagan Fellow}
-
-\author{Thomas North}
-\affiliation{School of Physics and Astronomy, University of Birmingham, Birmingham, B15 2TT, United Kingdom}
-\affiliation{Stellar Astrophysics Centre, Department of Physics 
-and Astronomy, Aarhus University, Ny Munkegade 120, DK-8000 Aarhus C, Denmark}
-
-\author[0000-0002-6980-3392]{Timothy R. White}
-\affiliation{Stellar Astrophysics Centre, Department of Physics 
-and Astronomy, Aarhus University, Ny Munkegade 120, DK-8000 Aarhus C, Denmark}
-
-\author{James Barrett}
-\affiliation{Birmingham Institute for Gravitational Wave Astronomy,
-  University of Birmingham, Birmingham, B15 2TT, United Kingdom}
-
-\affiliation{School of Physics and Astronomy, University of
-  Birmingham, Birmingham, B15 2TT, United Kingdom}
-
-\author{Andrea Miglio}
-
-\affiliation{School of Physics and Astronomy, University of
-  Birmingham, Birmingham, B15 2TT, United Kingdom}
-\affiliation{Stellar Astrophysics Centre, Department of Physics 
-and Astronomy, Aarhus University, Ny Munkegade 120, DK-8000 Aarhus C, Denmark}
-
-\author{Mikkel N. Lund}
-
-\affiliation{Stellar Astrophysics Centre, Department of Physics 
-and Astronomy, Aarhus University, Ny Munkegade 120, DK-8000 Aarhus C, Denmark}
-\affiliation{School of Physics and Astronomy, University of
-  Birmingham, Birmingham, B15 2TT, United Kingdom}
-
-\email{w.farr@bham.ac.uk, G.R.Davies@bham.ac.uk, b.pope@sydney.edu.au, txn016@student.bham.ac.uk, jimbarrett27@gmail.com}
+\end{affiliations}
+}
 
 \begin{abstract}
+
 The nearby red giant Aldebaran is suspected to host a gas giant planetary companion from decades of ground-based spectroscopic measurements of its Doppler shift.
@@
 Our approach to sparse, irregularly sampled time series astronomical observations has the potential to unlock similar measurements for hundreds of stars in archival data, and permits more flexible observing schedules in future.
+
+Aldebaran, or $\alpha$~Tauri, is a well-known first-magnitude naked-eye red giant star, and has long been the subject of astronomical investigations.
+It was one of the first stars around which an extrasolar planet candidate was identified, by looking for Doppler shifts from the star's reflex motion around the common centre of mass with its companion\cite{struverv}.
+While the hot Jupiter 51~Peg~b was the first exoplanet to be recognized as such in 1995\cite{51peg}, as early as 1993\cite{hatzes1993} RV variations in Pollux \cite{betgemconf,betgemconf2}, Arcturus (as yet unconfirmed), and Aldebaran suggested these giant stars may themselves host massive planets.
+After further investigation\cite{Hatzes1998}, Hatzes et al.\cite{Hatzes2015} have now claimed a firm RV detection of a planetary-mass companion Aldebaran~b, with a period of $628.96 \pm 0.90$~d.
+
+In this paper, we present a re-analysis of these original RV data in which we not only confirm this signal, but detect acoustic oscillations in Aldebaran for the first time.
+We validate this method and its result with new independent RV observations with the SONG~Telescope, and photometry from the \ktwo Mission (displayed in Figure~\ref{fig:k2obs}.
+By measuring the frequency of maximum power of these \emph{p}-mode oscillations we asteroseismically determine the mass of Aldebaran to be $1.16 \pm 0.07$~\msun.
+This measured stellar mass allows us to calculate that any moons of this giant planet, although they are now likely to be very hot, may have had equilibrium temperatures comparable to that of the Earth when Aldebaran was on the main sequence, raising the possibility that they may have once been habitable.
+
+Our new approach to asteroseismic data analysis, based on Continuous Auto-Regressive Moving Average (CARMA) models, can extract exoplanet signals together with measures of the frequency of maximum power from sparse and irregularly-sampled time series.
+An all-sky survey to find planetary companions and to precisely measure the masses of all nearby red giant stars is feasible with this new approach, and the required data either already exist in large radial velocity exoplanet surveys, or are easy to obtain with ground-based telescopes.
+
 \end{abstract}
 
-\section{Introduction}
-
-Aldebaran, or $\alpha$~Tauri, is a well-known first-magnitude naked-eye red giant star, and has long been the subject of astronomical investigations.
-It was one of the first stars around which an extrasolar planet candidate was identified, by looking for Doppler shifts from the star's reflex motion around the common centre of mass with its companion \citep[the radial velocity or RV method;][]{struverv}.
-While the hot Jupiter 51~Peg~b \citep{51peg} was the first exoplanet to be recognized as such, before this, \citet{hatzes1993} had noted RV variations in Pollux \citep[$\beta$~Gem; subsequently confirmed as a planet:][]{betgemconf,betgemconf2}, Arcturus ($\alpha$~Boo, unconfirmed), and Aldebaran.
-After further investigation by \citet{Hatzes1998}, \citet{Hatzes2015} have now claimed a firm RV detection of a planetary-mass companion Aldebaran~b, with a period of $628.96 \pm 0.90$~d.
-
-In this paper, we present a re-analysis of these original RV data in which we not only confirm this signal, but detect acoustic oscillations in Aldebaran for the first time.
-We validate this method and its result with new independent RV observations with the SONG~Telescope, and photometry from the \ktwo Mission.
-By measuring the frequency of maximum power, \numax, of these \emph{p}-mode oscillations we asteroseismically determine the mass of Aldebaran to be $1.16 \pm 0.07$~\msun.
-This measured stellar mass allows us to calculate that any moons of this giant planet, although they are now likely to be very hot, may have had equilibrium temperatures comparable to that of the Earth when Aldebaran was on the main sequence, raising the possibility that they may have once been habitable.
-
-Our new approach to asteroseismic data analysis, based on Continuous Auto-Regressive Moving Average (CARMA) models, can extract exoplanet signals together with measures of \numax from sparse and irregularly-sampled time series.
-An all-sky survey to find planetary companions and to precisely measure the masses of all nearby red giant stars is feasible with this new approach, and the required data either already exist in large exoplanet surveys, or are easy to obtain with ground-based telescopes.
-
-\section{Asteroseismology of Red Giants}
-Asteroseismology is a powerful tool for the charaterisation of red giant stars \citep[see][for a detailed review]{2013ARA&A..51..353C}.
-Red giants are solar-type pulsators, they exhibit acoustic modes of oscillation that are both excited and damped by stellar convection.
-In the frequency-power spectrum of either radial velocity of photometric observations, these modes of oscillation form an often intricate pattern of sharp peaks where the properties of the pattern can be mapped back to intrinsic stellar properties \citep[e.g.,][]{2016AN....337..774D}.
-The easiest properties of the pattern to determine are the the frequency of maximum amplitude \numax and the so-called large separation \Dnu \citep{Kjeldsen95}.
-The two observables are often referred to as the global average seismic parameters.
-
-Stellar properties can be determined with the average seismic measures and effective temperature \teff.
-A set of scaling relations exist that illustrate the constraint the global asteroseismic measures provide.
-These scaling relations are understood to be approximate but never the less are commonly used for radius $R$, mass $M$, and surface gravity $g$:
-\begin{eqnarray}
-\left( \frac{R}{\mathrm{R_{\odot}}} \right) \simeq & \left( \frac{\nu_{\rm max}}{\nu_{\rm max, \odot}} \right) \, 
-\left( \frac{\Delta \nu}{\Delta \nu_{\odot}} \right)^{-2} \, \left( \frac{T_{\rm eff}}{T_{\rm eff, \odot}} \right)^{0.5}, \\
-\left( \frac{M}{\mathrm{M_{\odot}}} \right) \simeq & \left( \frac{\nu_{\rm max}}{\nu_{\rm max, \odot}} \right)^{3} \, 
-\left( \frac{\Delta \nu}{\Delta \nu_{\odot}} \right)^{-4} \, \left( \frac{T_{\rm eff}}{T_{\rm eff, \odot}} \right)^{1.5}, \\
-\left( \frac{g}{\mathrm{g_{\odot}}} \right) \simeq & \left( \frac{\nu_{\rm max}}{\nu_{\rm max, \odot}} \right) \, 
- \left( \frac{T_{\rm eff}}{T_{\rm eff, \odot}} \right)^{0.5}. \\
-\end{eqnarray}
-While the accuracy of these scaling relations is still a matter of ongoing work \citep[e.g.,][]{2017ApJ...844..102H}, stellar parameters can be estimated by comparing observables to parameter from models of stellar evolution.
-
-As can be seen from the scaling relations, asteroseismology can provide constraint on Mass, Radius, and surface gravity if \Dnu, \numax, and \teff can be measured.
-In practise, for the higher luminosity red giants it is more straight forward to measure \numax than \Dnu.
-Typical values for \numax might range from $0.1$ to $20 \rm\,\mu Hz$ for the luminous giants and $20$ to $50 \rm\,\mu Hz$ for red giants with luminosities similar to the red clump .
-For the same sets of stars, \Dnu would take ranges from $0.02$ to $3 \rm\,\mu Hz$ and $3$ to $7 \rm\,\mu Hz$ depending on stellar mass and \teff (or metallicity which is considered more fundamental) \citep[e.g.,][]{2011A&A...525L...9M, 2013A&A...559A.137M}.
-Hence, \numax requires less frequency resolution than \Dnu to establish a good measurement.
-
-The ability to measure \numax and \Dnu, therefore, depends on the length and sampling rate of a data set.
-For regularly spaced data, typically analysed using an estimate of the frequency-power spectrum, the Nyquist limit (half the sampling rate) and frequency resolution ($\sim 1/$~duration) are trivial to estimate.
-%For the historic data sets proposed here the typical length is \guy{blah} and typical sampling is \guy{blah}.
-This in principle gives us access to measures of \numax for \guy{these types of stars} and \Dnu $\,$ \guy{for those types of stars}.
-
-\section{Time-Domain Models}
-
-It is easy to observe Aldebaran and similarly-bright stars with ground-based spectroscopic instruments, typically only requiring short exposures that can be obtained even under adverse observing conditions.
-There is indeed a considerable archive of such observations already, as a legacy of radial velocity (RV) surveys conducted to find exoplanets.
-In most cases, however, these have not so far been useful for asteroseismology, because these RV data are sparsely and irregularly-sampled.
-Because we have to pause observations during the day, during poor weather conditions, or simply when targets of higher priority are being observed, we get time series of only a few points and which may have significant and uneven gaps.
-This introduces a window-function effect: the power spectrum as constructed for example by a Fourier transform, or a Lomb-Scargle Periodogram \citep{lomb,scargle}, is convolved with the Fourier transform of the window function, introducing strong sidelobes adjacent to real frequency peaks and causing crosstalk between adjacent frequency channels.
-This imposes significant limitations both on the signal-to-noise and frequency resolution of power spectra derived from linear methods such as the Lomb-Scargle periodogram, and in practice makes asteroseismology with these conventional approaches difficult or impossible from the ground for stars with oscillation periods ranging from $\sim 12$~h to $\sim$~a few days.
-
-If we apply nonlinear statistical inference methods this situation can be improved.
-\citet{brewer2009} show that a system of driven, damped harmonic oscillators such as we encounter in asteroseismology can be statistically modelled as a Gaussian Process, with a covariance kernel consisting of a sum of damped sinusoids.
-The hyperparameters of this process can be used to retrieve very accurate power spectral density distributions.
-Unfortunately, this is impractical for long time series, as the computational cost of evaluating this likelihood function scales as $\mathcal{O}(n^3)$.
-
-Reference \citet{Kelly2014}, compare to \citet{Foreman-Mackey2017}.
-
-\section{Aldebaran}
-Aldebaran is a red giant star with spectral type K5, one of the nearest such stars at a distance of only $19.96 \pm 0.38$~pc \citep{hipparcos}.
-Its position near the Ecliptic permits the determination of its angular diameter by lunar occultations and by interferometry \citep[$20.58 \pm 0.03$ mas;][]{richichi2005,1979ApJ...228L.111B,brown1979,panek1980}.
-
-\subsection{SONG Observations}
-
-\subsection{K2 Observations}
-
-%In order to verify the results of the novel analysis presented above, we sought to obtain an independent detection of the oscillations of Aldebaran and compare the frequencies determined with the two methods.
-Aldebaran was observed with \ktwo \citep{howell14}, a two-wheeled revival of the \kepler Mission \citep{2010sci...327..977b}, under Guest Observer Program 130471 in Campaign~13, from 2017-03-08 to 2017-05-27.
-Aldebaran and similar high-luminosity K giants are well known to show photometric variability due to solar-like oscillations \citep{bedding2000}, and hence the continuous, high-precision K2 light curves allow an independent confirmation of our CARMA results.
-
-As Aldebaran is extremely bright, it saturates the \kepler detector and it is therefore not possible to use standard photometry pipelines to extract a K2 lightcurve.
-We instead use halo photometry \citep{White2017}, whereby unsaturated pixels from the outer part of the large and complicated halo of scattered light around bright stars are used to reconstruct a light curve as a weighted linear combination of individual pixel-level time series, as described in Appendix~\ref{halo}.
-
-Figure \ref{k2_lightcurve} shows the light curve and power spectrum of the K2 observations of Aldebaran.
-We detect clear variability on a average timescale of $\sim$\,5\,days, consistent with the expected timescale for solar-like oscillations for a high-luminosity red giant.
-To measure global asteroseismic parameters we model the background variability in the Fourier domain using the methodology described in \citet{huber09}, yielding a frequency of maximum power of $\numax = 2.2 \pm 0.25\,\muHz$ and amplitude per radial mode of $1850 \pm 500$\,ppm.
-Due to the limited frequency resolution of the 70-day K2 timeseries we were unable to measure the large frequency separation \Dnu, which is expected to be on the order of $\approx$\,0.5\muHz.
-
-Using the \textsc{k2ps} planet-search code \citep{k2ps,Pope2016} to examine this light curve, we search for transits across a wide range of periods, and find no evidence either of short-period planetary transits or an eclipsing stellar companion.
-
-\begin{figure}
-\centering
-% \includegraphics[width=0.75\textwidth]{Aldebaran_lc.png}
-%\includegraphics[width=0.75\textwidth]{aldebaran_lc.eps}
-\includegraphics[width=\textwidth]{k2obs.png}
+\begin{figure}[!t]
+\begin{center}
+\includegraphics[width=\linewidth]{k2obs.png}
 \caption{\ktwo lightcurve (left) and power spectrum (right) of Aldebaran.
@@
 }
-\label{k2_lightcurve}
+\label{fig:k2obs}
+\end{center}
 \end{figure}
 
-\subsection{Planetary and Stellar Parameters}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-By jointly fitting a Keplerian for the planet with a CARMA model for the stellar oscillations, we obtain a precise estimate of both \numax and the planetary orbital parameters.
-The planetary orbital parameters we obtain are similar to \citet{Hatzes2015}: period $P = 629.0^{+ 1.7}_{-2.1}$~d, eccentricity $e = 0.168^{+ 0.069}_{- 0.071}$; and radial velocity semi-amplitude $K = 127^{+ 15}_{- 14}$~m/s.
+%\bibliographystyle{mnras}
+\bibliography{aldebaran.bib}
 
-The stellar oscillations are best fit with a single CARMA component with $\numax = 2.238^{+ 0.086}_{- 0.080}$ \muHz, and including additional CARMA components does not resolve any further structure in the power spectrum.
-We have determined the stellar properties of Aldebaran with the additional constraint of \numax determined from this analysis.
-The details of the stellar modelling are presented in detail in \ref{sec:mod}.
-We have considered the impact of our additional \numax constraint and run our analysis for different sets of spectroscopic estimates.
-For final values we adopt the \citet{2012Sheffield} spectroscopic solution as being `middle of the road' estimates.
-For the analysis without constraint on \numax we find estimates of the stellar properties as $M = 1.27^{+0.24}_{-0.2} \, \mathrm{M_{\odot}}$ and age $4.86^{+3.56}_{-2.04} \, \rm Gyr$.
-With the inclusion of \numax we find $M = 1.16^{+0.07}_{-0.07} \, \mathrm{M_{\odot}}$ and age $6.38^{+1.42}_{-1.12} \, \rm Gyr$.
-It is clear that the addition of \numax provides substantially more precise estimates of mass and age.
-With this stellar mass, the radial velocity signal translates to an inclination-dependent planetary mass of $M\,\sin\,{i} = 5.81 \pm 0.72~M_J$ \citep[Equation~1]{Torres2008}, a massive giant planet; for sufficiently high inclinations ($\gtrsim 45^\circ$), the mass of Aldebaran~b may exceed $13~M_J$, making it a possible low-mass brown dwarf candidate.
-% re sheffield - I have substituted the citation from the .bib but you said 'b'; afaik she only published one paper in 2012?
-- Ben
+%\begin{thebibliography}{10}
+%\expandafter\ifx\csname url\endcsname\relax
+%  \def\url#1{\texttt{#1}}\fi
+%\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi
+%\providecommand{\bibinfo}[2]{#2}
+%\providecommand{\eprint}[2][]{\url{#2}}
+%
+%\bibitem{TNaab2009}
+%\bibinfo{author}{{Naab}, T.}, \bibinfo{author}{{Johansson}, P.~H.} \&
+%  \bibinfo{author}{{Ostriker}, J.~P.}
+%\newblock \bibinfo{title}{{Minor Mergers and the Size Evolution of Elliptical
+%  Galaxies}}.
+%\newblock \emph{\bibinfo{journal}{\apjl}} \textbf{\bibinfo{volume}{699}},
+%  \bibinfo{pages}{L178--L182} (\bibinfo{year}{2009}).
+%
+%\end{thebibliography}
 
-With the determination of the stellar mass it is possible to infer the parameters Aldebaran had while it was on the main sequence.
-We conduct a Monte Carlo simulation, drawing masses randomly from a Gaussian distribution $M \sim \mathcal{N}(1.16,0.07)$ and metallicities $[\text{Fe}/\text{H}] \sim \mathcal{N}(-0.15,0.2)$ \citep{decin2003} to predict the luminosity of the main-sequence progenitor, and the semi-major axis of the planet's orbit.
-%Using the scaling relations of \citep{eker2015}, we find that Aldebaran would have had a luminosity $L_* = 2.0 \pm 0.37~\lsun$ on the main sequence.
-Using the Mesa Isochrones and Stellar Tracks \citep[MIST:][]{mist0,mist1} models via the \textsc{isochrones} package \citep{isochrones}, we compute the evolution of the luminosity $L$ of the star along the main sequence; $L$ evolves from $2.0 \pm 0.7$~\lsun at 0.5~Gyr to $3.5 \pm 2.3$~\lsun at 4.5~Gyr.
-We note that this implies that the planet at a semi-major axis of $1.50\, \pm\, 0.03$~AU would have been subject to an insolation comparable to that of the Earth, evolving from 0.5~Gyr to 4.5~Gyr from $0.86\, \pm\, 0.27$ to $1.55 \,\pm\,1.0$ times that of the Earth today.
-Subject as well as this to the great uncertainties of the planet's orbital evolution and albedo, Aldebaran~b and any of its moons may well have hosted temperate environments for some of their history, now long-since destroyed by their star's evolution away from the main sequence.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\section{Conclusions}
-
-With CARMA models we have confirmed the previously-suspected planet around Aldebaran using archival data, and additionally detected acoustic oscillations that permit a $\sim 6\%$ mass determination.
-We have confirmed both of these results with new ground-based observations, and space-based photometry with \ktwo.
-
-From this pilot study with Aldebaran, have shown that with limited quantities of data of limited quality, we can nevertheless do precision asteroseismology from the ground and obtain very precise estimates of stellar parameters.
-Furthermore, improvements in the algorithm may also permit the detection of not only~\numax but also~\Dnu from similar measurements, permitting more sophisticated asteroseismic analysis.
-Sufficient radial velocity data either already exist, or are trivial to obtain, in order to do this for essentially all bright giants; while for solar-like stars, these new methods allow significantly relaxed observing requirements.
-One rich archive is the series of observations with the Hamilton \'{E}chelle Spectrograph at Lick Observatory, studying 373 G~and K~giants \citep[e.g.
-][]{frink2001,hekker2006,ortiz2016}.
-The method may also be useful in detecting rotational or other periodic variability in the LCES HIRES/Keck Precision Radial Velocity Exoplanet Survey \citep{butler17} of 1,624 FGKM dwarfs, although the time sampling is by design too coarse to detect $p$-mode oscillations in these stars.
-Furthermore, CARMA models and related methods will enhance deep, all-sky, sparse photometric surveys: an immediate future test of this will be from \hipparcos; while only~58 epochs of photometry are available for Aldebaran, at a sampling we find to be insufficient for our purposes, stars at higher latitudes may often have 150--200 epochs and more even sampling \citep{hipparcos_phot}, and the~14 bright~K giants in \hipparcos noted by \citet{bedding2000} are an ideal first test case.
-This can naturally be extended to \emph{Gaia} in space \citep{gaia}, or \emph{LSST} from the ground \citep{dmt,lsst,lsstbook}, from which many thousands of new asteroseismic determinations will be possible.
-
-\acknowledgments
-
-We thank the academy....
-
-We would like to thank Tim Bedding, David Hogg and Dan Foreman-Mackey for their helpful discussions.
-
-BP is grateful for funding from the Breakthrough Prize Foundation.
-This work was performed in part under contract with the Jet Propulsion Laboratory (JPL) funded by NASA through the Sagan Fellowship Program executed by the NASA Exoplanet Science Institute.
-
-The authors would like to acknowledge the Gadigal people of the Eora nation on whose ancestral lands the University of Sydney is built.
-
-Funding for the Stellar Astrophysics Centre is provided by The Danish National Research Foundation. TRW acknowledges the support of the Villum Foundation (research grant 10118).
-This research has made use of the SIMBAD database, operated at CDS, Strasbourg, France, and NASA's Astrophysics Data System.
-The \kepler data presented in this paper were obtained from the Mikulski Archive for Space Telescopes (MAST).
-STScI is operated by the Association of Universities for Research in Astronomy, Inc., under NASA contract NAS5-26555.
-Support for MAST for non-HST data is provided by the NASA Office of Space Science via grant NNX13AC07G and by other grants and contracts.
-This work made use of the \textsc{IPython} package \citep{PER-GRA:2007}; SciPy \citep{scipy};  \textsc{matplotlib}, a \textsc{Python} library for publication quality graphics \citep{Hunter:2007}; \textsc{Astropy}, a community-developed core \textsc{Python} package for astronomy \citep{2013A&A...558A..33A}; and \textsc{NumPy} \citep{van2011numpy}.
-Some of the acknowledgements were compiled using the Astronomy Acknowledgement Generator.
-
-\appendix
-
-\section{Halo Photometry}
-\label{halo}
-
-The \kepler Space Telescope \citep{2010sci...327..977b}  
-suffered a critical reaction wheel failure in May 2013, which made it impossible to maintain a stable pointing and therefore continue its nominal mission.
-It was revived as \ktwo \citep{howell14}, balanced by orienting perpendicular to the Sun.
-This requires that \ktwo observes fields in the Ecliptic in $\sim~80$~d Campaigns.
-
-The \kepler detector saturates for stars brighter than the $\sim~11^\text{th}$ magnitude.
-Nevertheless, the excess flux deposited in a saturated pixel spills conservatively up and down the pixel column, such that it is possible to sum this `bleed column' for bright stars and still obtain precise photometry, such as was done for the brightest star in the nominal \kepler mission, $\theta$~Cyg \citep[$V = 4.48$;][]{guzik2011,thetacygwhite,guzik2016}.
-There are two main reasons why this is not possible for all bright stars in general.
-First, because the downlink bandwidth from \kepler is very limited, it is often not desirable to download the large pixel stamps as are required for such bright stars; these must be even larger in \ktwo to account for the reduced image stability.
-Second, if the bleed column for a sufficiently bright star reaches the edge of the chip, flux spills over and is not conserved, imposing a hard brightness limit that depends on the distance to the detector edge.
-Collateral `smear' data, which are collected to help calibrate the photometric bias from stars sharing the same column as a target, can be used to reconstruct light curves for un-downloaded bright stars and thereby avoid bandwidth constraints \citep{k2smear}, but these data are still rendered unusable if the bleed column falls off the edge of the chip and contaminates the smear rows.
-
-Bright stars have a wide, complicated, position-dependent point spread function (PSF) arising from diffraction and scattering from secondary and higher-order reflections inside the instrument, with the result that they may contaminate thousands of nearby pixels with significant flux.
-We can therefore use this `halo' of unsaturated pixels for photometry.
-The brightness of this halo varies in the same way as that of the primary star, and we therefore obtain data in a region of 20~pixel radius around the mean position of Aldebaran, and discard saturated pixels.
-In this paper we proceed as in \citet{White2017}, in which the method was demonstrated on the seven bright Pleiades, with only minor changes.
-
-The flux $f_i$ at each cadence $i$ is constructed as a weighted sum of pixel values $p_{ij}$:
-
-\begin{equation}
-	f_i = \sum_{j=1}^{M} w_j p_{ij}.
-\end{equation}
-
-\noindent We choose the weights $w_j$ such that they lie between 0~and~1, add to unity, and minimize the Total Variation (TV) of the weighted light curve.
-In the continuous case, $n$-th order TV is defined as the integral of the absolute value of the $n$-th derivative of a function; in the discrete case, replacing the derivative with finite differences, first-order TV becomes
-
-\begin{equation}
-\text{TV} = \dfrac{\sum_{i=1}^{N} |{f}_i - {f}_{i-1}|}{\sum_{j=1}^{N} {f}_j}
-\end{equation}
-
-\noindent and likewise second-order TV the equivalent expression in second-order finite differences.
-The efficacy of this method was recently confirmed by \citet{kallinger2017}, comparing BRITE-constellation observations of Atlas to \ktwo halo photometry and finding excellent agreement in frequency and amplitude of the reported oscillations.
-
-In an improvement since \citet{White2017}, we use the \textsc{Theano} library \citep{theano} to calculate analytic derivatives for this objective function, which reduces the computational time for a single halo light curve on a commercial laptop from tens of minutes to tens of seconds.
-
-As a final step to reduce residual uncorrected systematics, we apply the \textsc{k2sc} \citep{k2sc} Gaussian Process-based systematics correction code to the initial halo light curve, but the effect of this in the present case for Aldebaran is minimal.
-There is somewhat higher than usual residual noise at multiples of $4 c/d$ (the satellite thruster firing frequency), but this is nevertheless very small in comparison to the signal from Aldebaran, and may be ascribed to the large fraction of the pixel mask occupied by the bleed column from this extremely bright star.
-
-\section{Stellar Modelling}\label{sec:mod}
-\subsection{Stellar Models}\label{sec:stell_mod}
-We used our determination of \numax and several combinations of the asteroseismic and spectroscopic parameters, along with luminosity, to estimate the fundamental stellar parameters, via fitting to stellar models.
-We used \textsc{MESA} models \citep{2011Paxton,2013Paxton} in conjunction with the Bayesian code \textsc{PARAM} \citep{2006dasilva, 2017Rod}.
-A summary of our selected ``benchmark'' options is as follows;
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{itemize}%GUY THIS BIT COULD PROBABLY BE DROPPED AND REFERENCED OUT TO \citep{2017Rod}
-\item Heavy element partitioning from \cite{1993Grevesse}.
-\item OPAL equation of state \citep{2002Rogers} along with OPAL opacities \citep{1996Iglesias}, with complementary values at low temperatures from \cite{2005Ferguson}.
-\item Nuclear reaction rates from NACRE \citep{1999Angulo}.
-\item The atmosphere model is taken according to \cite{1966Kris}.
-\item The mixing length theory was used to describe convection (a solar-calibrated parameter $\alpha_{\textrm{MLT}} =1.9657$ was adopted).
-\item Convective overshooting on the main sequence is set to $\alpha_{\textrm{ov}}=0.2H_{p}$, with $H_{p}$ the pressure scale height at the border of the convective core.
-Overshooting was applied according to the \cite{1975Maeder} step function
-scheme.
-\item No rotational mixing or diffusion is included.
-\end{itemize}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-Note that the precision on our estimate of \numax is orders of magnitude less that the point at which we would have needed to correct for the line-of-sight doppler shift \citep{2014MNRAS.445L..94D}.
-\subsection{Additional modelling inputs}\label{sec:addmod}
-In addition to the asteroseismic parameters, temperature and metallicity values are needed.
-There exist multiple literature values for Aldebaran.
-We chose to use a selection of sources as a separate modelling variation, to investigate what variation this would produce in final stellar properties.
-
-To ensure the values are self-consistent, when a literature value was chosen for temperature, we took the stellar metallicity from the same source i.e.
-matched pairs of temperature and metallicity.
-The final constraint is the stellar luminosity, which may be estimated as follows (e.g.
-see \citealt{pijpers2003}):
-\begin{multline}
-\log_{10} \frac{L}{L_{\odot}} = 4.0+
-0.4 M_{{\textrm{bol}},\odot} -2.0 \log_{10} {\pi [{\textrm{mas}]}} \\-0.4(V-A_V + BC(V)).
-\label{eqn:lum}
-\end{multline}
-The solar bolometric magnitude $M_{\textrm{bol},\odot}=4.73$ is taken from \cite{Torres2010}, from which we also take the polynomial expression for the bolometric correction $BC(V)$.
-Extinction $A_V$ was assumed to be zero.
-
-The final constraint available for Aldebaran is the angular diameter of the star, measured by long baseline interferometry, combined with the parallax to produce a physical radius constraint of $R_{\textrm{int}}=44.2\pm0.9\textrm{R}_{\odot}$ \citep{richichi2005}.
-
-% Example table
-\begin{table*}
-	\centering
-	\caption{Spectroscopic from each literature source, along with calculated luminosity.
-All temperature uncertainties assumed to be 50K, 0.2 dex in $\log{g}$, and 0.1 dex in [FeH].
-For the two \protect{\cite{2012Sheffield}} results, the reason for discrepancy between the two sets of metallicity results is not discussed.}
-	\label{tab:spec}
-	\begin{tabular}{lllll} % 5 columns, alignment for each
-		\hline
-		Spectrosocpy Source & $T_{\textrm{eff}}$ (K) & $\log{g}$ (dex) & [FeH] (dex) & Luminosity (L$_{\odot}$)\\
-		\hline
-		\cite{2012Sheffield}$_{a}$	&	3900	&	1.3	&	0.17	&	480\\
-		\cite{2012Sheffield}$_{b}$	&	3900	&	1.3	&	0.05	&	480\\
-		\cite{2011Prugniel}	&	3870 & 1.66 & -0.04 & 507\\
-		\cite{2008Massarotti} & 3936 & 1 & -0.34 & 456\\
-		\cite{2009Frasca} & 3850 & 0.55 & -0.1 & 526\\
-		\hline
-	\end{tabular}
-\end{table*}
-
-As Table \ref{tab:spec} shows, the spectroscopic parameters of Aldebaran are somewhat unclear, particularly $\log{g}$ and [FeH], which may have an impact on the recovered stellar properties when fitting to models.
-To explore what impact each parameter is having on the final stellar properties, multiple \textsc{PARAM} runs were performed, using different constraints.
-Two constraints potentially in tension were $\nu_{\textrm{max}}$ and $\log{g}$.
-$\nu_{\textrm{max}}$ has been shown to scale with $\log{g}$ \citep{Kjeldsen95, 2011A&A...530A.142B},
-\begin{equation}
-\frac{\nu_{\textrm{max}}}{\nu_{\textrm{max},\odot}}=\frac{\log{g}}{\log{g},\odot}\left(\frac{T_{\textrm{eff}}}{T_{\textrm{eff},\odot}}\right)^{-1/2}.
-\label{eqn:numax}
-\end{equation}
-
-Using Eq \ref{eqn:numax} with the values in Table \ref{tab:spec} predicts $\nu_{\textrm{max}}$ in the range $0.5-6\mu$Hz.
-Reversing the equation to produce a predicted $\log{g}$ from the observed $\nu_{\textrm{max,obs}}=2.23\pm0.1\mu$Hz results in a predicted $\log{g}\sim1.2$ dex, using an assumed temperature of 3900K.
-The solar calibration values used here are $g_{\odot}=274$ms$^{-2}$, $\nu_{\textrm{max},\odot}=3150\mu$Hz and $T_{\textrm{eff},\odot}=5777$K.
-
-Table \ref{tab:res_mass} shows the results, for all modelling variations, both different inputs and different constraints.
-It shows that results with the addition of $\nu_{\textrm{max}}$ as a constraint exhibit in general smaller uncertainties, with or without the addition of $\log{g}$ as a constraint.
-
-Recovering the mass without the use of asteroseismic constraints produces considerable scatter on the results ($0.96-1.5\textrm{M}_{\odot}$), whilst the use of asteroseismology brings the mass estimates into closer agreement, with the exception of the very low metallicity solution of \cite{2008Massarotti}.
-Any systematic offset between asteroseismic masses and spectroscopic masses is sensitive to chosen reference mass, in agreement with \cite{2017North}.
-
-Within the results of Table \ref{tab:res_mass}, the parallax of Aldebaran is contained within the angular diameter and the luminosity of the star.
-To avoid this, \textsc{PARAM} was also ran without luminosity constraint, the average absolute mass offset between the two sets of results was $<0.04\textrm{M}_{\odot}$ for the runs without $\nu_{\textrm{max}}$ constraint, and $<0.02\textrm{M}_{\odot}$ for the runs with $\nu_{\textrm{max}}$.
-
-% Example table
-\begin{table*}
-	\centering
-
-	\caption{Recovered stellar properties from \textsc{PARAM} using various constraints. Uncertainties quoted are the 68\% credible interval.}
-	\label{tab:res_mass}
-	\begin{tabular}{llll} % 5 columns, alignment for each
-		\hline
-		Spectroscopy Source & Mass ($\textrm{M}_{\odot}$) & Radius ($\textrm{R}_{\odot}$) & Age (Gyr)\\
-		\hline
-		\multicolumn{4}{c}{$\nu_{\textrm{max}}$, $\log{g}$, $T_{\textrm{eff}}$, $R_\textrm{int}$, L and [FeH].} \\
-		\hline
-		\cite{2012Sheffield}$_{a}$ &$1.17^{+0.07}_{-0.07}$&$43.9^{+0.9}_{-0.9}$&$6.47^{+1.39}_{-1.1}$\\
-		\cite{2012Sheffield}$_{b}$&$1.17^{+0.07}_{-0.07}$&$43.8^{+0.8}_{-0.9}$&$6.35^{+1.4}_{-1.11}$\\
-		\cite{2011Prugniel}&$1.17^{+0.07}_{-0.07}$&$43.9^{+0.9}_{-0.9}$&$6.23^{+1.4}_{-1.11}$\\
-		\cite{2008Massarotti}&$1.13^{+0.07}_{-0.07}$&$43.5^{+0.9}_{-0.9}$&$5.95^{+1.37}_{-1.07}$\\
-		\cite{2009Frasca}&$1.02^{+0.04}_{-0.04}$&$44.0^{+0.8}_{-0.8}$&$10.26^{+1.54}_{-1.45}$\\
-		\hline
-		\multicolumn{4}{c}{$\log{g}$, $T_{\textrm{eff}}$, $R_\textrm{int}$, L and [FeH].} \\
-		\hline
-\cite{2012Sheffield}$_{a}$&$1.43^{+0.26}_{-0.24}$&$43.8^{+0.9}_{-0.9}$&$3.53^{+2.72}_{-1.44}$\\
-\cite{2012Sheffield}$_{b}$&$1.27^{+0.24}_{-0.2}$&$43.8^{+0.9}_{-0.9}$&$4.86^{+3.56}_{-2.04}$\\
-\cite{2011Prugniel}&$1.25^{+0.22}_{-0.19}$&$43.8^{+0.9}_{-0.9}$&$5.02^{+3.47}_{-2.02}$\\
-\cite{2008Massarotti}&$0.95^{+0.11}_{-0.05}$&$43.8^{+0.9}_{-0.9}$&$10.24^{+2.36}_{-3.08}$\\
-\cite{2009Frasca}&$0.96^{+0.04}_{-0.04}$&$44.4^{+0.8}_{-0.8}$&$11.64^{+1.43}_{-1.8}$\\
-		\hline
-		\multicolumn{4}{c}{$\nu_{\textrm{max}}$, $T_{\textrm{eff}}$, $R_\textrm{int}$, L and [FeH].} \\
-		\hline
-\cite{2012Sheffield}$_{a}$&$1.17^{+0.07}_{-0.07}$&$43.9^{+0.9}_{-0.9}$&$6.51^{+1.41}_{-1.11}$\\
-\cite{2012Sheffield}$_{b}$&$1.16^{+0.07}_{-0.07}$&$43.8^{+0.8}_{-0.9}$&$6.38^{+1.42}_{-1.12}$\\
-\cite{2011Prugniel}&$1.16^{+0.07}_{-0.07}$&$43.9^{+0.9}_{-0.9}$&$6.43^{+1.48}_{-1.15}$\\
-\cite{2008Massarotti}&$1.13^{+0.07}_{-0.07}$&$43.5^{+0.9}_{-0.9}$&$5.86^{+1.35}_{-1.05}$\\
-\cite{2009Frasca}&$1.15^{+0.07}_{-0.07}$&$43.9^{+0.9}_{-0.8}$&$6.47^{+1.51}_{-1.17}$\\
-		\hline
-	\end{tabular}
-
-\end{table*}
-
-\bibliography{aldebaran}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2017-12-22 presub/natastron_presub_coverletter.tex
presubmission letter and abstract draft

% Cover letter using letter.cls
\documentclass[11pt]{letter} % Uses 10pt
%\usepackage{helvetica} % uses helvetica postscript font (download helvetica.sty)
%\usepackage{newcent}   % uses new century schoolbook postscript font 
% the following commands control the margins:

\topmargin=-1.25in    % Make letterhead start about 1 inch from top of page 
\textheight=10.75in    % text height can be bigger for a longer letter
\oddsidemargin=0pt   % leftmargin is 1 inch
\textwidth=6.5in     % textwidth of 6.5in leaves 1 inch for right margin
\usepackage{url}
\usepackage{hyperref}
\hypersetup{
    colorlinks=false,
    pdfborder={0 0 0},
}
\usepackage{graphicx}
\usepackage[strict]{changepage}
\usepackage{xcolor}

%--------------------------------------------------------------
% A few colors for hyperlinks.
%--------------------------------------------------------------
\definecolor{plum}{rgb}{0.36078, 0.20784, 0.4}
\definecolor{chameleon}{rgb}{0.30588, 0.60392, 0.023529}
\definecolor{cornflower}{rgb}{0.12549, 0.29020, 0.52941}
\definecolor{scarlet}{rgb}{0.8, 0, 0}
\definecolor{brick}{rgb}{0.64314, 0, 0}

\newcommand{\email}[1]{\href{mailto:#1}{\tt \textcolor{cornflower}{#1}}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

%
\longindentation=0pt                       % needed to get closing flush left
\let\raggedleft\raggedright                % needed to get date flush left

\nopagebreak  
\begin{letter}
{Nature Astronomy \\
natureastro@nature.com \\}

{
\leftskip = 0pt plus 1 fill
\rightskip = 0pt
\parindent 0pt
%\obeylines
%
%
{{\bf Dr Will M. Farr}} \\
School of Physics and Astronomy\\
University of Birmingham\\
Birmingham, B15 2TT\\
United Kingdom\\
 \url{w.farr@bham.ac.uk} \\

}

\begin{minipage}[c]{5in}\vskip-4.25cm
\begin{flushleft}
	\begin{minipage}[c]{3cm}
		\begin{flushleft}
			\includegraphics*[width=8cm]{birmingham_logo.png}%
		\end{flushleft}
	\end{minipage}
\end{flushleft}
\end{minipage}

\opening{Dear Editor,}

\noindent I would like to express my interest in submitting a Letter to \emph{Nature Astronomy}, on the ``Detection of Oscillations in Aldebaran with Ground-Based Observations''. The principal authors are myself, Prof. Guy Davies (Birmingham), and Dr Benjamin Pope (NYU).

Aldebaran ($\alpha$~Tauri, the Eye of the Bull) is a nearby naked-eye red giant star. Since the identification of the planet candidate Aldebaran~b in~1993 in the first modern radial velocity (RV) surveys, a further two decades of observations have firmly established the existence of a massive planet around the star.

With radical new data processing techniques, we have re-analysed these historical RV data and augmented them with new observations of our own. Using our Gaussian Process-based model we infer much more accurate orbital parameters for the planet, but also detect the signal of stellar oscillations, missed by previous authors but uncovered by our powerful new algorithms. With asteroseismology we constrain the star's age for the first time, measure its mass with a factor of two better precision than had previously been possible (to 6\%), and consequently obtain the mass of the giant planet to a precision of 12\%.

We have also independently confirmed our results with new photometry from the \emph{Kepler}-2 (K2) mission, again using a highly novel data processing technique to extract a precise light curve of the brightest ever observed by \emph{Kepler}, ten thousand times above its saturation limit. The K2 light curve shows extremely clear oscillations that are in perfect agreement with the frequency we infer from RV.

A neat and surprising result of this work is that while the planet has a hellish temperature today well upwards of a thousand Kelvin, when Aldebaran was on the main sequence, the planet (and more importantly any of its moons) would have been subject to similar incident starlight as the Earth receives from the Sun. We posit that this is the first clear example of a formerly-habitable world destroyed by stellar evolution.

Thank you for considering our submission. Please do not hesitate to contact me with any questions you may have.\\

\closing{Yours sincerely, \\
\vspace{1.75cm}
Will Farr}

%\encl{One paragraph Letter outline}					% Enclosures

\end{letter}

\end{document}

### 2018-01-03 paper/aldebaran.tex
added graph of insolation vs time

@@
 Using the Mesa Isochrones and Stellar Tracks \citep[MIST:][]{mist0,mist1} models via the \textsc{isochrones} package \citep{isochrones}, we compute the evolution of the luminosity $L$ of the star along the main sequence; $L$ evolves from $2.0 \pm 0.7$~\lsun at 0.5~Gyr to $3.5 \pm 2.3$~\lsun at 4.5~Gyr.
-We note that this implies that the planet at a semi-major axis of $1.50\, \pm\, 0.03$~AU would have been subject to an insolation comparable to that of the Earth, evolving from 0.5~Gyr to 4.5~Gyr from $0.86\, \pm\, 0.27$ to $1.55 \,\pm\,1.0$ times that of the Earth today.
+We note that this implies that the planet at a semi-major axis of $1.50\, \pm\, 0.03$~AU would have been subject to an insolation comparable to that of the Earth, evolving from 0.5~Gyr to 4.5~Gyr from $0.86\, \pm\, 0.27$ to $1.55 \,\pm\,1.0$ times that of the Earth today (Figure~\ref{insolation}).
 Subject as well as this to the great uncertainties of the planet's orbital evolution and albedo, Aldebaran~b and any of its moons may well have hosted temperate environments for some of their history, now long-since destroyed by their star's evolution away from the main sequence.
+
+\begin{figure}
+\centering
+% \includegraphics[width=0.75\textwidth]{Aldebaran_lc.png}
+%\includegraphics[width=0.75\textwidth]{aldebaran_lc.eps}
+\includegraphics[width=\textwidth]{insolation.png}
+\caption{Evolution of the insolation of Aldebaran~b relative to the incident sunlight on Earth, as a function of the age of Aldebaran on the main sequence.
+The solid line is the mean of simulations, and the shaded region extends from the 5th to the 95th percentile.
+It is likely that Aldebaran~b spent some time on the main sequence with an equilibrium temperature similar to Earth.}
+\label{insolation}
+\end{figure}
 

### 2018-01-05 paper/aldebaran.tex
1 sigma
s type planet

@@
 By measuring the frequency of maximum power, \numax, of these \emph{p}-mode oscillations we asteroseismically determine the mass of Aldebaran to be $1.16 \pm 0.07$~\msun.
-This measured stellar mass allows us to calculate that any moons of this giant planet, although they are now likely to be very hot, may have had equilibrium temperatures comparable to that of the Earth when Aldebaran was on the main sequence, raising the possibility that they may have once been habitable.
+This measured stellar mass allows us to calculate that any moons of this giant planet (S-type planets if Aldebaran~b is a brown dwarf), although they are now likely to be very hot, may have had equilibrium temperatures comparable to that of the Earth when Aldebaran was on the main sequence, raising the possibility that they may have once been habitable.
 
@@
 \caption{Evolution of the insolation of Aldebaran~b relative to the incident sunlight on Earth, as a function of the age of Aldebaran on the main sequence.
-The solid line is the mean of simulations, and the shaded region extends from the 5th to the 95th percentile.
+The solid line is the mean of simulations, and the shaded region represents $1\,\sigma$ equivalent deviations.
 It is likely that Aldebaran~b spent some time on the main sequence with an equilibrium temperature similar to Earth.}

### 2018-01-05 paper/aldebaran.tex
satellites

@@
 By measuring the frequency of maximum power, \numax, of these \emph{p}-mode oscillations we asteroseismically determine the mass of Aldebaran to be $1.16 \pm 0.07$~\msun.
-This measured stellar mass allows us to calculate that any moons of this giant planet (S-type planets if Aldebaran~b is a brown dwarf), although they are now likely to be very hot, may have had equilibrium temperatures comparable to that of the Earth when Aldebaran was on the main sequence, raising the possibility that they may have once been habitable.
+This measured stellar mass allows us to calculate that any satellites of Aldebaran~b, although they are now likely to be very hot, may have had equilibrium temperatures comparable to that of the Earth when Aldebaran was on the main sequence, raising the possibility that they may have once been habitable.
 
@@
 We note that this implies that the planet at a semi-major axis of $1.50\, \pm\, 0.03$~AU would have been subject to an insolation comparable to that of the Earth, evolving from 0.5~Gyr to 4.5~Gyr from $0.86\, \pm\, 0.27$ to $1.55 \,\pm\,1.0$ times that of the Earth today (Figure~\ref{insolation}).
-Subject as well as this to the great uncertainties of the planet's orbital evolution and albedo, Aldebaran~b and any of its moons may well have hosted temperate environments for some of their history, now long-since destroyed by their star's evolution away from the main sequence.
+Subject as well as this to the great uncertainties of the planet's orbital evolution and albedo, Aldebaran~b and any of its moons (or its S-type planets if it is a brown dwarf) may well have hosted temperate environments for some of their history, now long-since destroyed by their star's evolution away from the main sequence.
 

### 2018-01-05 paper/aldebaran.tex
precision

@@
 
-From this pilot study with Aldebaran, have shown that with limited quantities of data of limited quality, we can nevertheless do precision asteroseismology from the ground and obtain very precise estimates of stellar parameters.
+From this pilot study with Aldebaran, have shown that with limited quantities of data of limited quality, we can nevertheless do asteroseismology from the ground and obtain very precise estimates of stellar parameters.
 Furthermore, improvements in the algorithm may also permit the detection of not only~\numax but also~\Dnu from similar measurements, permitting more sophisticated asteroseismic analysis.

### 2018-01-05 paper/aldebaran.tex
will allow

@@
 Furthermore, improvements in the algorithm may also permit the detection of not only~\numax but also~\Dnu from similar measurements, permitting more sophisticated asteroseismic analysis.
-Sufficient radial velocity data either already exist, or are trivial to obtain, in order to do this for essentially all bright giants; while for solar-like stars, these new methods allow significantly relaxed observing requirements.
+Sufficient radial velocity data either already exist, or are trivial to obtain, in order to do this for essentially all bright giants; while for solar-like stars, these new methods will allow significantly relaxed observing requirements.
 One rich archive is the series of observations with the Hamilton \'{E}chelle Spectrograph at Lick Observatory, studying 373 G~and K~giants \citep[e.g.

### 2018-01-05 paper/aldebaran.tex
remove placeholder
new section; phrasing

@@
 
+\subsection{Archival Observations}
+
 \subsection{SONG Observations}    % this goes into section 4.1
 
-In order to detect the oscillations of Aldebaran via the radial velocity 
-method we used the Hertzsprung SONG telescope \cite{2017ApJ...836..142G}. All 
-observations were carried out in the highest resolution mode (R$\approx\,110.000$)
-with an integration time of 30s. The observations were carried out between 27 September 
-2016 and  12 January 2017, during this time we attempted to obtain at least one 
-visit per available night. A total of 254 spectra was obtained. The spectra were
+In order to confirm the oscillations detected in the archival data, we used the Hertzsprung SONG telescope \cite{2017ApJ...836..142G} to conduct high time cadence follow-up observations.
+These were carried out in the highest resolution mode (R$\approx\,110.000$)
+with an integration time of 30s, between 27 September 
+2016 and  12 January 2017. During this time we attempted to obtain at least one 
+visit per available night, for a total of 254 epochs over the campaign. The spectra were
 extracted using the SONG pipeline (see \cite{2017ApJ...836..142G}). SONG employs an iodine cell for
@@
 We conduct a Monte Carlo simulation, drawing masses randomly from a Gaussian distribution $M \sim \mathcal{N}(1.16,0.07)$ and metallicities $[\text{Fe}/\text{H}] \sim \mathcal{N}(-0.15,0.2)$ \citep{decin2003} to predict the luminosity of the main-sequence progenitor, and the semi-major axis of the planet's orbit.
-%Using the scaling relations of \citep{eker2015}, we find that Aldebaran would have had a luminosity $L_* = 2.0 \pm 0.37~\lsun$ on the main sequence.
 Using the Mesa Isochrones and Stellar Tracks \citep[MIST:][]{mist0,mist1} models via the \textsc{isochrones} package \citep{isochrones}, we compute the evolution of the luminosity $L$ of the star along the main sequence; $L$ evolves from $2.0 \pm 0.7$~\lsun at 0.5~Gyr to $3.5 \pm 2.3$~\lsun at 4.5~Gyr.
@@
 \acknowledgments
-
-We thank the academy....
 

### 2018-01-05 paper/aldebaran.tex
phrasing

@@
 the radial velocities obtained -- oscillations are easily visible as well as a long-term
-trend. This is probably due to the a manifestation of the planetary companion reported by
-\cite{Hatzes2015}.
+trend, probably due to the planetary companion.
 

### 2018-01-05 code/CARMAKepler.jl
Pre-SONG and SONG alone runs with 1 oscillatory mode.

@@ -3,7 +3,11 @@ module CARMAKepler
 using CARMA
 using Distributions
 using Ensemble
+using HDF5
 using Kepler
+
+import Base:
+    write, read
 
 type MultiEpochPosterior
     ts::Array{Array{Float64, 1}, 1}
@@ -389,6 +393,55 @@ function posterior_predictive(post::MultiEpochPosterior, p::MultiEpochParams)
     end
 
     obs_ys
+end
+
+function write(f::Union{HDF5File, HDF5Group}, post::MultiEpochPosterior, ns::EnsembleNest.NestState)
+    ns_group = g_create(f, "nest_state")
+    write(ns_group, ns)
+
+    xs, lnlikes = EnsembleNest.postsample(ns)
+
+    ps = [to_params(post, xs[:,j]) for j in 1:size(xs,2)]
+    
+    f["lnlike", "compress", 3, "shuffle", ()] = lnlikes
+    f["mu", "compress", 3, "shuffle", ()] = hcat([p.mu for p in ps]...)
+    f["nu", "compress", 3, "shuffle", ()] = hcat([p.nu for p in ps]...)
+    f["K", "compress", 3, "shuffle", ()] = [p.K for p in ps]
+    f["P", "compress", 3, "shuffle", ()] = [p.P for p in ps]
+    f["e", "compress", 3, "shuffle", ()] = [p.e for p in ps]
+    f["omega", "compress", 3, "shuffle", ()] = [p.omega for p in ps]
+    f["chi", "compress", 3, "shuffle", ()] = [p.chi for p in ps]
+
+    f["drw_rms", "compress", 3, "shuffle", ()] = hcat([p.drw_rms for p in ps]...)
+    f["drw_rate", "compress", 3, "shuffle", ()] = hcat([p.drw_rate for p in ps]...)
+
+    f["osc_rms", "compress", 3, "shuffle", ()] = hcat([p.osc_rms for p in ps]...)
+    f["osc_freq", "compress", 3, "shuffle", ()] = hcat([p.osc_freq for p in ps]...)
+    f["osc_Q", "compress", 3, "shuffle", ()] = hcat([p.osc_Q for p in ps]...)
+end
+
+function read(f::Union{HDF5File, HDF5Group}, post::MultiEpochPosterior)
+    mus = read(f, "mu")
+    nus = read(f, "nu")
+    Ks = read(f, "K")
[…]

### 2018-01-05 paper/aldebaran.tex
add comment back in
comments
Lots of technical CARMA crap, should eventually move most of it to Methods section.

@@
 \newcommand{\bibtex}{\textsc{Bib}\!\TeX} % bibtex. Not quite the correct typesetting, but close enough
+
+\newcommand{\dd}{\mathrm{d}}
+\newcommand{\diff}[2]{\frac{\dd #1}{\dd #2}}
 
@@
 The hyperparameters of this process can be used to retrieve very accurate power spectral density distributions.
-Unfortunately, this is impractical for long time series, as the computational cost of evaluating this likelihood function scales as $\mathcal{O}(n^3)$.
+Unfortunately, this is impractical for long time series, as the computational cost of evaluating the standard Gaussian process likelihood function scales as $\mathcal{O}(n^3)$.
 
-Reference \citet{Kelly2014}, compare to \citet{Foreman-Mackey2017}.
+A Gaussian process such as that used by \citet{brewer2009} also
+corresponds to the solution, $y(t)$, of the stochastic ODE
+\begin{equation}
+  \label{eq:GP-definition}
+  \prod_{i=1}^{p} \left[ \diff{}{t} - r_i \right] y(t) = \prod_{j = 1}^{q} \left[ \diff{}{t} - b_i \right] \eta(t),
+\end{equation}
+where $p, q \in \mathbb{Z}$, $p > q \geq 0$ define the order of the
+process; $r_i \in \mathbb{C}$, $\Re r_i < 0$ are the roots of the
+characteristic equation of the ODE; $b_j \in \mathbb{C}$,
+$\Re b_j < 0$ are corresponding roots of the inverse process; and
+$\eta(t)$ is a white-noise Gaussian process with
+$\left\langle \eta(t) \right\rangle = 0$ and
+$\left\langle \eta(t) \eta(t') \right\rangle = \sigma^2 \delta\left(t
+  - t' \right)$.  Such a solution is called a CARMA (Continuous
+AutoRegressive Moving Average) process.  The constraint that
+$\Re r_i < 0$ and $\Re b_j < 0$ ensures that the linear operator
+defining the process and its inverse are invertable.  Under the
+assumption that the solution, $y(t)$, is real, roots $r_i$ and $b_j$
+must either be real or occur in complex conjugate pairs.  The power
+spectrum of the process is
+\begin{equation}
+  \label{eq:CARMA-PSD}
+  P_y(f) = \sigma^2 \left| \frac{\prod_{j=1}^q \left[ 2 \pi i f - b_j \right]}{\prod_{i=1}^p \left[ 2 \pi i f - r_i \right]} \right|^2.
+\end{equation}
+The power spectrum is a rational function of frequency with poles at
+$2\pi i f = r_i$ and zeros at $2\pi i f = b_i$; because of the
+invertable constraints, these poles and zeros all occur when
+$\Im f \neq 0$.  The autocorrelation function corresponding to the
+power spectrum in Eq.\ \eqref{eq:CARMA-PSD} is 
+\begin{equation}
+  \label{eq:CARMA-ACF}
+  \rho_y (\tau) = \left\langle y\left( t \right) y \left( t + \tau \right) \right\rangle = \sum_{i=1}^p A_i^2 e^{r_i \tau},
+\end{equation}
+where the RMS amplitudes of the $p$ different modes, $A_i$, are
+functions of $\sigma$ and the values of the roots $r_i$ and $b_j$.  If
+$q = p-1$, then the $A_i$ are independent, while if $q < p-1$ then
+there are interdependencies among the $A_i$.  From either the power
+spectrum in Eq.\ \eqref{eq:CARMA-PSD} or the autocorrelation function
+in Eq.\ \eqref{eq:CARMA-ACF}, it is apparent that a real root $r_i$
+represents an exponentially-decaying mode with $r_i$ the rate
+constant, while a complex conjugate pair of $r_i$ represent an
+oscillatory mode with decay rate $\Re r_i$ and angular frequency
+$\left|\Im r_i\right|$.  The former is called an CAR(1) process, an
+Ornstein-Uhlenbeck process, or a damped random walk.
+\citet{Kelly2014} gives more details on CARMA processes, including
+formulae for $A_i$ as a function of $\sigma$, $r_i$, and $b_j$, and
+shows how a state-space model of the solution $y(t)$ can be tracked
+through a time-series of uncertain observations
+$y_k = y\left( t_k \right) + \epsilon_k$, with Gaussian noise
+$\epsilon_k$ of zero mean and known (heteroskedastic) variance, using
+a Kalman filter to produce a likelihood function
+$p\left( y_k \mid \sigma, r_i, b_j \right)$ in
+$\mathcal{O}\left( n \right)$ computational time.  Additionally,
+\citet{Foreman-Mackey2017} show how a dimensional expansion can be
+used to reduce the standard Gaussian process covariance matrix for the
+process with autocorrelation given by Eq.\ \eqref{eq:CARMA-ACF} under
+observations at times $t_k$ to a banded form that can then be
+Cholesky-decomposed in the standard Gaussian process likelihood
+function in $\mathcal{O}\left( n \right)$ time.
 
@@
 
-%In order to verify the results of the novel analysis presented above, we sought to obtain an independent detection of the oscillations of Aldebaran and compare the frequencies determined with the two methods.
+In order to verify the results of the novel analysis presented above, we sought to obtain an independent detection of the oscillations of Aldebaran and compare the frequencies determined with the two methods.
 Aldebaran was observed with \ktwo \citep{howell14}, a two-wheeled revival of the \kepler Mission \citep{2010sci...327..977b}, under Guest Observer Program 130471 in Campaign~13, from 2017-03-08 to 2017-05-27.
@@
 \centering
-% \includegraphics[width=0.75\textwidth]{Aldebaran_lc.png}
-%\includegraphics[width=0.75\textwidth]{aldebaran_lc.eps}
 \includegraphics[width=\textwidth]{insolation.png}

### 2018-01-05 paper/aldebaran.tex
minor text edits, mostly Sec 2

@@
 \affiliation{ Stellar Astrophysics Centre (SAC), Department of Physics and Astronomy, Aarhus University, Ny Munkegade 120, DK-8000, Denmark}
+
 \author{Frank Grundahl}
+\affiliation{ Stellar Astrophysics Centre (SAC), Department of Physics and Astronomy, Aarhus University, Ny Munkegade 120, DK-8000, Denmark}
+
+\author[0000-0001-8832-4488]{Daniel Huber}
+\affiliation{Institute for Astronomy, University of Hawai`i, 2680 Woodlawn Drive, Honolulu, HI 96822, USA}
+\affiliation{Sydney Institute for Astronomy (SIfA), School of Physics, University of Sydney, NSW 2006, Australia}
+\affiliation{SETI Institute, 189 Bernardo Avenue, Mountain View, CA 94043, USA}
 \affiliation{ Stellar Astrophysics Centre (SAC), Department of Physics and Astronomy, Aarhus University, Ny Munkegade 120, DK-8000, Denmark}
@@
 \section{Asteroseismology of Red Giants}
-Asteroseismology is a powerful tool for the charaterisation of red giant stars \citep[see][for a detailed review]{2013ARA&A..51..353C}.
-Red giants are solar-type pulsators, they exhibit acoustic modes of oscillation that are both excited and damped by stellar convection.
-In the frequency-power spectrum of either radial velocity of photometric observations, these modes of oscillation form an often intricate pattern of sharp peaks where the properties of the pattern can be mapped back to intrinsic stellar properties \citep[e.g.,][]{2016AN....337..774D}.
-The easiest properties of the pattern to determine are the the frequency of maximum amplitude \numax and the so-called large separation \Dnu \citep{Kjeldsen95}.
-The two observables are often referred to as the global average seismic parameters.
+Asteroseismology is a powerful tool for the charaterisation of red giant stars \citep[see][for detailed reviews]{2013ARA&A..51..353C, hekker17}.
+Red giants exhibit oscillations that are excited and damped by stellar convection.
+In the power spectrum of either radial velocity of photometric observations, these modes of oscillation form a characteristic pattern of peaks which can be used to infer intrinsic stellar properties \citep[e.g.,][]{2016AN....337..774D}.
+The easiest properties of the pattern to determine are the the frequency of maximum amplitude \numax and the so-called large separation \Dnu \citep{Kjeldsen95}, which are often referred to as global asteroseismic parameters.
 
-Stellar properties can be determined with the average seismic measures and effective temperature \teff.
+%Stellar properties can be determined with the average seismic measures and effective temperature \teff.
 A set of scaling relations exist that illustrate the constraint the global asteroseismic measures provide.
 These scaling relations are understood to be approximate but never the less are commonly used for radius $R$, mass $M$, and surface gravity $g$:
+Global asteroseismic parameters can be used to estimate the radius $R$, mass $M$, and surface gravity $g$ of stars when combined with an estimate of the effective temperature \teff\ through so-called scaling relations:
+
 \begin{eqnarray}
@@
 \end{eqnarray}
+
 While the accuracy of these scaling relations is still a matter of ongoing work \citep[e.g.,][]{2017ApJ...844..102H}, stellar parameters can be estimated by comparing observables to parameter from models of stellar evolution.
 
-As can be seen from the scaling relations, asteroseismology can provide constraint on Mass, Radius, and surface gravity if \Dnu, \numax, and \teff can be measured.
+%As can be seen from the scaling relations, asteroseismology can provide constraint on Mass, Radius, and surface gravity if \Dnu, \numax, and \teff can be measured.  
+The ability to measure \numax and \Dnu depends on the length and sampling rate of a data set.
 In practise, for the higher luminosity red giants it is more straight forward to measure \numax than \Dnu.
-Typical values for \numax might range from $0.1$ to $20 \rm\,\mu Hz$ for the luminous giants and $20$ to $50 \rm\,\mu Hz$ for red giants with luminosities similar to the red clump .
-For the same sets of stars, \Dnu would take ranges from $0.02$ to $3 \rm\,\mu Hz$ and $3$ to $7 \rm\,\mu Hz$ depending on stellar mass and \teff (or metallicity which is considered more fundamental) \citep[e.g.,][]{2011A&A...525L...9M, 2013A&A...559A.137M}.
-Hence, \numax requires less frequency resolution than \Dnu to establish a good measurement.
+Typical values for \numax\ range from $0.1-20 \rm\,\mu Hz$ for luminous giants and $20-50 \rm\,\mu Hz$ for stars near the red clump.
+For the evolutionary stages, \Dnu typically ranges from $0.02-3 \rm\,\mu Hz$ and $3-7 \rm\,\mu Hz$ depending on stellar mass and \teff
+%(or metallicity which is considered more fundamental)
+\citep[e.g.,][]{2011A&A...525L...9M, 2013A&A...559A.137M}.  Hence, \numax requires less frequency resolution than \Dnu to establish a good measurement.
 
-The ability to measure \numax and \Dnu, therefore, depends on the length and sampling rate of a data set.
+%The ability to measure \numax and \Dnu, therefore, depends on the length and sampling rate of a data set.
 For regularly spaced data, typically analysed using an estimate of the frequency-power spectrum, the Nyquist limit (half the sampling rate) and frequency resolution ($\sim 1/$~duration) are trivial to estimate.
@@
 
-With CARMA models we have confirmed the previously-suspected planet around Aldebaran using archival data, and additionally detected acoustic oscillations that permit a $\sim 6\%$ mass determination.
+With CARMA models we have confirmed the previously-suspected planet around Aldebaran using archival data, and additionally detected acoustic oscillations that permit a mass determination with a precision of $\sim 6\%$.
 We have confirmed both of these results with new ground-based observations, and space-based photometry with \ktwo.
@@
   Astrofisica de Canarias.
+
+D.H.
+acknowledges support by the Australian Research Council's Discovery Projects funding scheme (project number DE140101364) and support by the National Aeronautics and Space Administration under Grants NNX14AB92G and NNX16AH45G issued through the Kepler Participating Scientist Program and the K2 Guest Observer Program.
 

### 2018-01-09 paper/aldebaran.tex
minor text edits

@@
 \author[0000-0003-2595-9114]{Benjamin J. S. Pope}
-\affiliation{Sydney Institute for Astronomy, School of Physics, University of Sydney, Sydney NSW 2006, Australia} 
+\affiliation{Sydney Institute for Astronomy (SIfA), School of Physics, University of Sydney, NSW 2006, Australia}
 \affiliation{Center for Cosmology and Particle Physics, Department of Physics, New York University, 726 Broadway, New York, NY 10003, USA}
@@
 \author{Victoria Antoci}
-\affiliation{ Stellar Astrophysics Centre (SAC), Department of Physics and Astronomy, Aarhus University, Ny Munkegade 120, DK-8000, Denmark}
+\affiliation{Stellar Astrophysics Centre, Department of Physics and Astronomy, Aarhus University, Ny Munkegade 120, DK-8000 Aarhus C, Denmark}
 
 \author{Mads Fredslund Andersen} 
-\affiliation{ Stellar Astrophysics Centre (SAC), Department of Physics and Astronomy, Aarhus University, Ny Munkegade 120, DK-8000, Denmark}
+\affiliation{Stellar Astrophysics Centre, Department of Physics and Astronomy, Aarhus University, Ny Munkegade 120, DK-8000 Aarhus C, Denmark}
 
 \author{Frank Grundahl}
-\affiliation{ Stellar Astrophysics Centre (SAC), Department of Physics and Astronomy, Aarhus University, Ny Munkegade 120, DK-8000, Denmark}
+\affiliation{Stellar Astrophysics Centre, Department of Physics and Astronomy, Aarhus University, Ny Munkegade 120, DK-8000 Aarhus C, Denmark}
 
@@
 \affiliation{SETI Institute, 189 Bernardo Avenue, Mountain View, CA 94043, USA}
-\affiliation{ Stellar Astrophysics Centre (SAC), Department of Physics and Astronomy, Aarhus University, Ny Munkegade 120, DK-8000, Denmark}
+\affiliation{Stellar Astrophysics Centre, Department of Physics and Astronomy, Aarhus University, Ny Munkegade 120, DK-8000 Aarhus C, Denmark}
 
@@
 \left( \frac{g}{\mathrm{g_{\odot}}} \right) \simeq & \left( \frac{\nu_{\rm max}}{\nu_{\rm max, \odot}} \right) \, 
- \left( \frac{T_{\rm eff}}{T_{\rm eff, \odot}} \right)^{0.5}. \\
+ \left( \frac{T_{\rm eff}}{T_{\rm eff, \odot}} \right)^{0.5}.
 \end{eqnarray}
@@
 
-In order to confirm the oscillations detected in the archival data, we used the Hertzsprung SONG telescope \cite{2017ApJ...836..142G} to conduct high time cadence follow-up observations.
-These were carried out in the highest resolution mode (R$\approx\,110.000$)
-with an integration time of 30s, between 27 September 
-2016 and  12 January 2017. During this time we attempted to obtain at least one 
+In order to confirm the oscillations detected in the archival data, we used the Hertzsprung SONG telescope \citep{2017ApJ...836..142G} to conduct high time cadence follow-up observations.
+These were carried out in the highest resolution mode ($R\approx110000$)
+with an integration time of 30\,s, between 2016 September 27 and 2017 January 12. During this time we attempted to obtain at least one 
 visit per available night, for a total of 254 epochs over the campaign. The spectra were
-extracted using the SONG pipeline (see \cite{2017ApJ...836..142G}). SONG employs an iodine cell for
-precise wavelength calibration and the velocities were determined using the $\it iSONG$
+extracted using the SONG pipeline \citep[see][]{2017ApJ...836..142G}. SONG employs an iodine cell for
+precise wavelength calibration and the velocities were determined using the \textit{iSONG}
 software following the same procedures as described in \cite{2017ApJ...836..142G}. The typical
-velocity precision per visit is 2.5$m\,s^{-1}$ allowing us to easily detect the oscillations
+velocity precision per visit is 2.5$\,\rm m\,s^{-1}$ allowing us to easily detect the oscillations
 which displays a peak-to-peak amplitude of $\approx \, 170 \, \rm m \, s^{-1}$. $\bf Figure XX$ show
@@
 In order to verify the results of the novel analysis presented above, we sought to obtain an independent detection of the oscillations of Aldebaran and compare the frequencies determined with the two methods.
-Aldebaran was observed with \ktwo \citep{howell14}, a two-wheeled revival of the \kepler Mission \citep{2010sci...327..977b}, under Guest Observer Program 130471 in Campaign~13, from 2017-03-08 to 2017-05-27.
+Aldebaran was observed with \ktwo \citep{howell14}, a two-wheeled revival of the \kepler Mission \citep{2010sci...327..977b}, under Guest Observer Program 130471 in Campaign~13, from 2017 March 8 to 2017 May 27.
 Aldebaran and similar high-luminosity K giants are well known to show photometric variability due to solar-like oscillations \citep{bedding2000}, and hence the continuous, high-precision K2 light curves allow an independent confirmation of our CARMA results.
@@
 We have determined the stellar properties of Aldebaran with the additional constraint of \numax determined from this analysis.
-The details of the stellar modelling are presented in detail in \ref{sec:mod}.
+The details of the stellar modelling are presented in detail in Appendix \ref{sec:mod}.
 We have considered the impact of our additional \numax constraint and run our analysis for different sets of spectroscopic estimates.
@@
 There are two main reasons why this is not possible for all bright stars in general.
-First, because the downlink bandwidth from \kepler is very limited, it is often not desirable to download the large pixel stamps as are required for such bright stars; these must be even larger in \ktwo to account for the reduced image stability.
+First, because the on-board data storage and downlink bandwidth from \kepler are limited, it is often not desirable to download the large number of pixels that are required for such bright stars.
 Second, if the bleed column for a sufficiently bright star reaches the edge of the chip, flux spills over and is not conserved, imposing a hard brightness limit that depends on the distance to the detector edge.

### 2018-01-11 code/Aldebaran.jl
some fixes

@@ -10,7 +10,7 @@ function load_timeseries(;before_song=true, song=true)
 
     if before_song 
         for i in 3:9
-            data = readdlm("/Users/farr/Documents/Research/Aldebaran/data/$(i)/table$(i).dat")
+            data = readdlm("..//data/$(i)/table$(i).dat")
             push!(ts, data[:,1])
             push!(ys, data[:,2])
             push!(dys, data[:,3])
@@ -18,7 +18,7 @@ function load_timeseries(;before_song=true, song=true)
     end
 
     if song
-        data = readdlm("/Users/farr/Documents/Research/Aldebaran/data/song/tablesong.dat")
+        data = readdlm("../data/song/tablesong.dat")
         push!(ts, data[:,1])
         push!(ys, data[:,2])
         push!(dys, data[:,3])

### 2018-01-12 paper/aldebaran.tex
Some work on time-domain models (more to follow tonight after Katherine sleeps).

@@
 
+Here we have adopted the \citet{Kelly2014} approach.
+The \texttt{CARMA.jl} package\footnote{\url{https://github.com/farr/CARMA.jl}} implements Kalman filters that can compute the likelihood for a set of observations parameterised by either the roots $r_i$ and $b_j$ or the RMS amplitudes $A_i$ and roots $r_i$.
+Because we have a set of observations taken with different instruments at different sites (see Section \ref{sec:archival-obs}), we include RV offset (i.e.\ mean value) parameters that are site-dependent and also a site-dependent scaling parameter.
+Thus, the $i$th RV measured at site $j$ is assumed to be
+\begin{equation}
+\label{eq:RV-model}
+y_{ij} = v_r\left(t_i\right) + \mu_j + y\left( t_i \right) + \nu_j \epsilon_{ij},
+\end{equation}
+$t_i$ is the time of the measurement, $v_r$ is a planetary RV signal \citep{Hatzes2015}, $\mu_j$ is the site-dependent mean RV offset, $y(t)$ is the CARMA process representing stellar activity, $\nu_j$ is the site-dependent uncertainty scaling parameter, and $\epsilon_{ij}$ is an independent Gaussian random variable with mean zero and standard deviation equal to the quoted uncertainty of the observation.
+We impose broad priors (see Table \ref{tab:parameters}) on the parameters of the planetary RV signal, the CARMA process, the $\mu_j$ offsets, and the $\nu_j$ scale factors.
+
+\begin{deluxetable}{lll}
+\tablecolumns{3}
+\tablecaption{\label{tab:parameters} The parameters in our model (see Eq.\ \eqref{eq:RV-model}).}
+\tablehead{\colhead{Parameter} & \colhead{Prior} & \colhead{Description}}
+\startdata
+$\mu_j$ & Normal\tablenotemark{a} & The per-site RV offset. \\
+\enddata
+\tablenotetext{a}{The prior for $\mu_j$ is centered on the mean of the data from site $j$ and has standard deviation 10 times the standard deviation of the data.}
+\end{deluxetable}
+
 \section{Aldebaran}
@@
 \subsection{Archival Observations}
+\label{sec:archival-obs}
 

### 2018-01-12 paper/aldebaran.tex
abstract

@@
 \begin{abstract}
-The nearby red giant Aldebaran is suspected to host a gas giant planetary companion from decades of ground-based spectroscopic measurements of its Doppler shift.
-Using Gaussian Process-based Continuous Auto-Regressive Moving Average (CARMA) models, we confirm this suspected planet and show that these historic data also contain evidence of acoustic oscillations in the star itself, and verify this result with further dedicated ground-based spectroscopy and space-based photometry with the \kepler Space Telescope.
+The nearby red giant Aldebaran is known to host a gas giant planetary companion from decades of ground-based spectroscopic measurements of its Doppler shift.
+Using Gaussian Process-based Continuous Auto-Regressive Moving Average (CARMA) models, we show that these historic data also contain evidence of acoustic oscillations in the star itself, and verify this result with further dedicated ground-based spectroscopy and space-based photometry with the \kepler Space Telescope.
 From the frequency of these oscillations we determine the mass of Aldebaran to be $1.16 \pm 0.07$~\msun, and note that this implies its planet will have been subject to insolation comparable to the Earth for some of the star's main sequence lifetime.
-Our approach to sparse, irregularly sampled time series astronomical observations has the potential to unlock similar measurements for hundreds of stars in archival data, and permits more flexible observing schedules in future.
+Our approach to sparse, irregularly sampled time series astronomical observations has the potential to unlock similar measurements for hundreds of stars in archival data.
+%, and permits more flexible observing schedules in future.
 \end{abstract}

### 2018-01-12 paper/aldebaran.tex
first paragraph changes

@@
 
-Aldebaran, or $\alpha$~Tauri, is a well-known first-magnitude naked-eye red giant star, and has long been the subject of astronomical investigations.
+Aldebaran ($\alpha$~Tauri) is a well-known first-magnitude naked-eye red giant star, and has long been the subject of astronomical investigations.
 It was one of the first stars around which an extrasolar planet candidate was identified, by looking for Doppler shifts from the star's reflex motion around the common centre of mass with its companion \citep[the radial velocity or RV method;][]{struverv}.
@@
 In this paper, we present a re-analysis of these original RV data in which we not only confirm this signal, but detect acoustic oscillations in Aldebaran for the first time.
-We validate this method and its result with new independent RV observations with the SONG~Telescope, and photometry from the \ktwo Mission.
-By measuring the frequency of maximum power, \numax, of these \emph{p}-mode oscillations we asteroseismically determine the mass of Aldebaran to be $1.16 \pm 0.07$~\msun.
-This measured stellar mass allows us to calculate that any satellites of Aldebaran~b, although they are now likely to be very hot, may have had equilibrium temperatures comparable to that of the Earth when Aldebaran was on the main sequence, raising the possibility that they may have once been habitable.
+We validate this method and its result with new independent RV observations with the Hertzsprung SONG~Telescope, and photometry from the \ktwo Mission.
+By measuring the frequency of maximum power of these \emph{p}-mode oscillations, \numax, we asteroseismically determine the mass of Aldebaran to be $1.16 \pm 0.07$~\msun.
+This precise stellar mass allows us to calculate that Aldebaran~b and any satellites it may have, although they are now likely to be very hot, would have had equilibrium temperatures comparable to that of the Earth when Aldebaran was on the main sequence.
+It is possible that they may have once been habitable, billions of years ago.
 

### 2018-01-12 paper/aldebaran.tex
second paragraph

@@
 These scaling relations are understood to be approximate but never the less are commonly used for radius $R$, mass $M$, and surface gravity $g$:
-Global asteroseismic parameters can be used to estimate the radius $R$, mass $M$, and surface gravity $g$ of stars when combined with an estimate of the effective temperature \teff\ through so-called scaling relations:
+These parameters can be used to estimate the radius $R$, mass $M$, and surface gravity $g$ of stars when combined with an estimate of the effective temperature \teff\ through so-called scaling relations:
 

### 2018-01-12 paper/aldebaran.tex
appendix ref

@@
 There is indeed a considerable archive of such observations already, as a legacy of radial velocity (RV) surveys conducted to find exoplanets.
-In most cases, however, these have not so far been useful for asteroseismology, because these RV data are sparsely and irregularly-sampled.
-Because we have to pause observations during the day, during poor weather conditions, or simply when targets of higher priority are being observed, we get time series of only a few points and which may have significant and uneven gaps.
+In most cases, however, these have not so far been useful for asteroseismology, because these RV data are sparsely and irregularly sampled.
+Because we have to pause observations during the day, during poor weather conditions, or simply when targets of higher priority are being observed, we get time series which may have significant and uneven gaps.
 This introduces a window-function effect: the power spectrum as constructed for example by a Fourier transform, or a Lomb-Scargle Periodogram \citep{lomb,scargle}, is convolved with the Fourier transform of the window function, introducing strong sidelobes adjacent to real frequency peaks and causing crosstalk between adjacent frequency channels.
@@
 \citet{brewer2009} show that a system of driven, damped harmonic oscillators such as we encounter in asteroseismology can be statistically modelled as a Gaussian Process, with a covariance kernel consisting of a sum of damped sinusoids.
-The hyperparameters of this process can be used to retrieve very accurate power spectral density distributions.
-Unfortunately, this is impractical for long time series, as the computational cost of evaluating the standard Gaussian process likelihood function scales as $\mathcal{O}(n^3)$.
+The hyperparameters of this process encode features of the power spectral density distributions and are insensitive to the window function.
+Unfortunately, this is impractical for long time series, as the computational cost of evaluating the standard Gaussian Process likelihood function scales as $\mathcal{O}(n^3)$.
 
-A Gaussian process such as that used by \citet{brewer2009} also
+A Gaussian Process such as that used by \citet{brewer2009} also
 corresponds to the solution, $y(t)$, of the stochastic ODE
@@
 $\Re b_j < 0$ are corresponding roots of the inverse process; and
-$\eta(t)$ is a white-noise Gaussian process with
+$\eta(t)$ is a white-noise Gaussian Process with
 $\left\langle \eta(t) \right\rangle = 0$ and
@@
 \citet{Foreman-Mackey2017} show how a dimensional expansion can be
-used to reduce the standard Gaussian process covariance matrix for the
+used to reduce the standard Gaussian Process covariance matrix for the
 process with autocorrelation given by Eq.\ \eqref{eq:CARMA-ACF} under
 observations at times $t_k$ to a banded form that can then be
-Cholesky-decomposed in the standard Gaussian process likelihood
+Cholesky-decomposed in the standard Gaussian Process likelihood
 function in $\mathcal{O}\left( n \right)$ time.
@@
 Its position near the Ecliptic permits the determination of its angular diameter by lunar occultations and by interferometry \citep[$20.58 \pm 0.03$ mas;][]{richichi2005,1979ApJ...228L.111B,brown1979,panek1980}.
+These tight constraints are valuable in breaking degeneracies in stellar modelling and make this an ideal star for asteroseismic characterization (Appendix~/ref{sec:mod}).
 

### 2018-01-12 paper/aldebaran.tex
acknowledgements

@@
 Sufficient radial velocity data either already exist, or are trivial to obtain, in order to do this for essentially all bright giants; while for solar-like stars, these new methods will allow significantly relaxed observing requirements.
-One rich archive is the series of observations with the Hamilton \'{E}chelle Spectrograph at Lick Observatory, studying 373 G~and K~giants \citep[e.g.
-][]{frink2001,hekker2006,ortiz2016}.
+One rich archive is the series of observations with the Hamilton \'{E}chelle Spectrograph at Lick Observatory, studying 373 G~and K~giants \citep[e.g.][]{frink2001,hekker2006,ortiz2016}.
 The method may also be useful in detecting rotational or other periodic variability in the LCES HIRES/Keck Precision Radial Velocity Exoplanet Survey \citep{butler17} of 1,624 FGKM dwarfs, although the time sampling is by design too coarse to detect $p$-mode oscillations in these stars.
@@
 
+The authors would like to acknowledge the Gadigal people of the Eora nation on whose ancestral lands the University of Sydney is built.
+
 BP is grateful for funding from the Breakthrough Prize Foundation.
 This work was performed in part under contract with the Jet Propulsion Laboratory (JPL) funded by NASA through the Sagan Fellowship Program executed by the NASA Exoplanet Science Institute.
+TRW acknowledges the support of the Villum Foundation (research grant 10118).
+D.H.
+acknowledges support by the Australian Research Council's Discovery Projects funding scheme (project number DE140101364) and support by the National Aeronautics and Space Administration under Grants NNX14AB92G and NNX16AH45G issued through the Kepler Participating Scientist Program and the K2 Guest Observer Program.
 
-The authors would like to acknowledge the Gadigal people of the Eora nation on whose ancestral lands the University of Sydney is built.
-
-Funding for the Stellar Astrophysics Centre is provided by The Danish National Research Foundation. TRW acknowledges the support of the Villum Foundation (research grant 10118).
+Funding for the Stellar Astrophysics Centre is provided by the Danish National Research Foundation. 
 This research has made use of the SIMBAD database, operated at CDS, Strasbourg, France, and NASA's Astrophysics Data System.
@@
 
-The Villum Foundation, The Danish Council for Independent Research~|~Natural Sciences and
+This work has made use of observations made with the Hertzsprung SONG telescope operated on the Spanish Observatorio
+  del Teide on the island of Tenerife by Aarhus and Copenhagen Universities and by the Instituto de
+  Astrofisica de Canarias.
+
+The Villum Foundation, the Danish Council for Independent Research~|~Natural Sciences and
  the Carlsberg Foundation are thanked for their support on building the SONG prototype on
@@
  to the development of the SONG data reduction pipeline and archive system.
-
-Based on observations made with the Hertzsprung SONG telescope operated on the Spanish Observatorio
-  del Teide on the island of Tenerife by the Aarhus and Copenhagen Universities and by the Instituto de
-  Astrofisica de Canarias.
-
-D.H.
-acknowledges support by the Australian Research Council's Discovery Projects funding scheme (project number DE140101364) and support by the National Aeronautics and Space Administration under Grants NNX14AB92G and NNX16AH45G issued through the Kepler Participating Scientist Program and the K2 Guest Observer Program.
 

### 2018-01-12 paper/aldebaran.tex
software section

@@
 We conduct a Monte Carlo simulation, drawing masses randomly from a Gaussian distribution $M \sim \mathcal{N}(1.16,0.07)$ and metallicities $[\text{Fe}/\text{H}] \sim \mathcal{N}(-0.15,0.2)$ \citep{decin2003} to predict the luminosity of the main-sequence progenitor, and the semi-major axis of the planet's orbit.
-Using the Mesa Isochrones and Stellar Tracks \citep[MIST:][]{mist0,mist1} models via the \textsc{isochrones} package \citep{isochrones}, we compute the evolution of the luminosity $L$ of the star along the main sequence; $L$ evolves from $2.0 \pm 0.7$~\lsun at 0.5~Gyr to $3.5 \pm 2.3$~\lsun at 4.5~Gyr.
+Using the Mesa Isochrones and Stellar Tracks \citep[MIST:][]{mist0,mist1} models, we compute the evolution of the luminosity $L$ of the star along the main sequence; $L$ evolves from $2.0 \pm 0.7$~\lsun at 0.5~Gyr to $3.5 \pm 2.3$~\lsun at 4.5~Gyr.
 We note that this implies that the planet at a semi-major axis of $1.50\, \pm\, 0.03$~AU would have been subject to an insolation comparable to that of the Earth, evolving from 0.5~Gyr to 4.5~Gyr from $0.86\, \pm\, 0.27$ to $1.55 \,\pm\,1.0$ times that of the Earth today (Figure~\ref{insolation}).
@@
 
-This work made use of the \textsc{IPython} package \citep{PER-GRA:2007}; SciPy \citep{scipy};  \textsc{matplotlib}, a \textsc{Python} library for publication quality graphics \citep{Hunter:2007}; \textsc{Astropy}, a community-developed core \textsc{Python} package for astronomy \citep{2013A&A...558A..33A}; and \textsc{NumPy} \citep{van2011numpy}.
+\software{\textsc{IPython} \citep{PER-GRA:2007}, SciPy \citep{scipy}, \textsc{matplotlib} \citep{Hunter:2007}, \textsc{Astropy} \citep{2013A&A...558A..33A}, \textsc{NumPy} \citep{van2011numpy}, \textsc{isochrones} \citep{isochrones}.}
+
+% This work made use of the \textsc{IPython} package \citep{PER-GRA:2007}; SciPy \citep{scipy};  \textsc{matplotlib}, a \textsc{Python} library for publication quality graphics \citep{Hunter:2007}; \textsc{Astropy}, a community-developed core \textsc{Python} package for astronomy \citep{2013A&A...558A..33A}; and \textsc{NumPy} \citep{van2011numpy}.
 Some of the acknowledgements were compiled using the Astronomy Acknowledgement Generator.

### 2018-01-12 paper/aldebaran.tex
campaign
capital Nation

@@
 
-The authors would like to acknowledge the Gadigal people of the Eora nation on whose ancestral lands the University of Sydney is built.
+The authors would like to acknowledge the Gadigal people of the Eora Nation on whose ancestral lands the University of Sydney is built.
 
@@
 It was revived as \ktwo \citep{howell14}, balanced by orienting perpendicular to the Sun.
-This requires that \ktwo observes fields in the Ecliptic in $\sim~80$~d Campaigns.
+This requires that \ktwo observes fields in the Ecliptic in $\sim~80$~d Campaigns, of which Aldebaran was observed in Campaign~13.
 

### 2018-01-12 paper/aldebaran.tex
phrasing

@@
 \item Nuclear reaction rates from NACRE \citep{1999Angulo}.
-\item The atmosphere model is taken according to \cite{1966Kris}.
-\item The mixing length theory was used to describe convection (a solar-calibrated parameter $\alpha_{\textrm{MLT}} =1.9657$ was adopted).
+\item The atmosphere model according to \cite{1966Kris}.
+\item The mixing length theory to describe convection (we adopt a solar-calibrated parameter $\alpha_{\textrm{MLT}} =1.9657$).
 \item Convective overshooting on the main sequence is set to $\alpha_{\textrm{ov}}=0.2H_{p}$, with $H_{p}$ the pressure scale height at the border of the convective core.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-Note that the precision on our estimate of \numax is orders of magnitude less that the point at which we would have needed to correct for the line-of-sight doppler shift \citep{2014MNRAS.445L..94D}.
+We do not need to correct for the line-of-sight Doppler shift at the frequency precision available in our data \citep{2014MNRAS.445L..94D}.
 \subsection{Additional modelling inputs}\label{sec:addmod}
-In addition to the asteroseismic parameters, temperature and metallicity values are needed.
+In addition to the asteroseismic parameters, spectroscopically-determined temperature and metallicity values are needed.
 There exist multiple literature values for Aldebaran.
-We chose to use a selection of sources as a separate modelling variation, to investigate what variation this would produce in final stellar properties.
+We chose to compare a range of literature values to investigate what uncertainty these systematically-differing models produce in inferred stellar properties.
 
@@
 	\centering
-	\caption{Spectroscopic from each literature source, along with calculated luminosity.
+	\caption{Spectroscopic stellar parameters from each literature source, along with calculated luminosity.
 All temperature uncertainties assumed to be 50K, 0.2 dex in $\log{g}$, and 0.1 dex in [FeH].

### 2018-01-12 paper/aldebaran.tex
specify both parallax and angular diameter explicitly

@@
 \section{Aldebaran}
-Aldebaran is a red giant star with spectral type K5, one of the nearest such stars at a distance of only $19.96 \pm 0.38$~pc \citep{hipparcos}.
+Aldebaran is a red giant star with spectral type K5, one of the nearest such stars at a distance of only $19.96 \pm 0.38$~pc as determined by \hipparcos \citep{hipparcos}.
 Its position near the Ecliptic permits the determination of its angular diameter by lunar occultations and by interferometry \citep[$20.58 \pm 0.03$ mas;][]{richichi2005,1979ApJ...228L.111B,brown1979,panek1980}.
@@
 
-The final constraint available for Aldebaran is the angular diameter of the star, measured by long baseline interferometry, combined with the parallax to produce a physical radius constraint of $R_{\textrm{int}}=44.2\pm0.9\textrm{R}_{\odot}$ \citep{richichi2005}.
+The final constraint available for Aldebaran is the angular diameter of the star as measured by long baseline interferometry and lunar occultations \citep[$20.58 \pm 0.03$ mas;][]{richichi2005,1979ApJ...228L.111B,brown1979,panek1980}, combined with the \hipparcos parallax of $19.96 \pm 0.38$~pc to produce a physical radius constraint of $R_{\textrm{int}}=44.2\pm0.9\textrm{R}_{\odot}$.
 

### 2018-01-12 paper/aldebaran.tex
a bit of a garbled final paragraph

@@
 The solar bolometric magnitude $M_{\textrm{bol},\odot}=4.73$ is taken from \cite{Torres2010}, from which we also take the polynomial expression for the bolometric correction $BC(V)$.
-Extinction $A_V$ was assumed to be zero.
+We assume extinction $A_V$ to be zero.
 
@@
 To explore what impact each parameter is having on the final stellar properties, multiple \textsc{PARAM} runs were performed, using different constraints.
-Two constraints potentially in tension were $\nu_{\textrm{max}}$ and $\log{g}$.
-$\nu_{\textrm{max}}$ has been shown to scale with $\log{g}$ \citep{Kjeldsen95, 2011A&A...530A.142B},
+Two constraints potentially in tension were $\nu_{\textrm{max}}$ and the spectroscopically-determined $\log{g}$.
+$\nu_{\textrm{max}}$ has been shown to scale with the stellar $\log{g}$ \citep{Kjeldsen95, 2011A&A...530A.142B},
 \begin{equation}
@@
 
-Recovering the mass without the use of asteroseismic constraints produces considerable scatter on the results ($0.96-1.5\textrm{M}_{\odot}$), whilst the use of asteroseismology brings the mass estimates into closer agreement, with the exception of the very low metallicity solution of \cite{2008Massarotti}.
+Recovering the mass without the use of asteroseismic constraints produces considerable scatter on the results ($0.96-1.5\textrm{M}_{\odot}$), whilst the use of asteroseismology brings the mass estimates into closer agreement with one another, with the exception of the very low metallicity solution of \cite{2008Massarotti}.
 Any systematic offset between asteroseismic masses and spectroscopic masses is sensitive to chosen reference mass, in agreement with \cite{2017North}.
 
-Within the results of Table \ref{tab:res_mass}, the parallax of Aldebaran is contained within the angular diameter and the luminosity of the star.
-To avoid this, \textsc{PARAM} was also ran without luminosity constraint, the average absolute mass offset between the two sets of results was $<0.04\textrm{M}_{\odot}$ for the runs without $\nu_{\textrm{max}}$ constraint, and $<0.02\textrm{M}_{\odot}$ for the runs with $\nu_{\textrm{max}}$.
+We have also investigated running \textsc{PARAM} and relaxing one or more constraints: without the luminosity constraint, the average absolute mass offset between the two sets of results was $<0.04\textrm{M}_{\odot}$ for the runs without $\nu_{\textrm{max}}$ constraint, and $<0.02\textrm{M}_{\odot}$ for the runs with $\nu_{\textrm{max}}$.
 

### 2018-01-13 paper/aldebaran.tex
More edits to time-series analysis section.

@@
 Because we have a set of observations taken with different instruments at different sites (see Section \ref{sec:archival-obs}), we include RV offset (i.e.\ mean value) parameters that are site-dependent and also a site-dependent scaling parameter.
-Thus, the $i$th RV measured at site $j$ is assumed to be
+Thus, the $i$th RV measured at site $j$, $y_{ij}$, is assumed to be
 \begin{equation}
@@
 \end{equation}
-$t_i$ is the time of the measurement, $v_r$ is a planetary RV signal \citep{Hatzes2015}, $\mu_j$ is the site-dependent mean RV offset, $y(t)$ is the CARMA process representing stellar activity, $\nu_j$ is the site-dependent uncertainty scaling parameter, and $\epsilon_{ij}$ is an independent Gaussian random variable with mean zero and standard deviation equal to the quoted uncertainty of the observation.
+$t_i$ is the time of the measurement, $v_r$ is a planetary RV signal \citep{Hatzes2015,Gregory2005}, $\mu_j$ is the site-dependent mean RV offset, $y(t)$ is the CARMA process representing stellar activity, $\nu_j$ is the site-dependent uncertainty scaling parameter, and $\epsilon_{ij}$ is an independent Gaussian random variable with mean zero and standard deviation equal to the quoted uncertainty of the observation.
 We impose broad priors (see Table \ref{tab:parameters}) on the parameters of the planetary RV signal, the CARMA process, the $\mu_j$ offsets, and the $\nu_j$ scale factors.
@@
 \startdata
-$\mu_j$ & Normal\tablenotemark{a} & The per-site RV offset. \\
+\cutinhead{Per-Site RV Offset and Uncertainty Scaling}
+$\mu_j$ & $N\left[ \left\langle y_{j} \right\rangle, 10 \sigma_{y_{j}} \right] \left( \mu _j \right)\tablenotemark{a}$ & The per-site RV offset. \\
+$\nu_j$ & $1/\nu_j, \quad \nu_j \in [1/2, 2]$ & The per-site uncertainty scale factor. \\
+\cutinhead{Keplerian Parameters}
+$K$ & $1/K, \quad K \in \left[ \min_j \sigma_{y_j}/100, \max_j 10 \sigma_{y_j} \right]$ & RV semi-amplitude. \\
+$P$ & $1/P, \quad P \in [600, 700] \, \mathrm{d}$ & Keplerian period \\
+$e$ & $2e$\tablenotemark{b} & Keplerian eccentricity \\
+$\omega$ & $\mathrm{const}$ & Longitude of pericentre \\
+$\chi$ & $\mathrm{const}$ & Pericentre passage is at $t = \chi P$\\
 \enddata
-\tablenotetext{a}{The prior for $\mu_j$ is centered on the mean of the data from site $j$ and has standard deviation 10 times the standard deviation of the data.}
+\tablenotetext{a}{$\left\langle y_j \right\rangle$ and $\sigma_{y_j}$ are the mean and standard deviation of the RV timeseries from site $j$.}
+\tablenotetext{b}{We actually sample in and impose a (flat) prior on $x \equiv e \cos \omega$ and $y \equiv e \sin \omega$.}
 \end{deluxetable}

### 2018-01-15 paper/aldebaran.tex
WIP on data analysis section.

@@
 $P$ & $1/P, \quad P \in [600, 700] \, \mathrm{d}$ & Keplerian period \\
-$e$ & $2e$\tablenotemark{b} & Keplerian eccentricity \\
-$\omega$ & $\mathrm{const}$ & Longitude of pericentre \\
-$\chi$ & $\mathrm{const}$ & Pericentre passage is at $t = \chi P$\\
+$e$ & $2e,\quad 0 \leq e < 1$\tablenotemark{b} & Keplerian eccentricity \\
+$\omega$ & $\mathrm{const}, \quad 0 \leq \omega < 2\pi$ & Longitude of pericentre \\
+$\chi$ & $\mathrm{const}, \quad 0 \leq \chi < 1$ & Pericentre passage is at $t = \chi P$\\
 \enddata

### 2018-01-17 paper/aldebaran.tex
WIP on archival data section, time series analysis section.

@@
 
-A Gaussian Process such as that used by \citet{brewer2009} also
-corresponds to the solution, $y(t)$, of the stochastic ODE
-\begin{equation}
-  \label{eq:GP-definition}
-  \prod_{i=1}^{p} \left[ \diff{}{t} - r_i \right] y(t) = \prod_{j = 1}^{q} \left[ \diff{}{t} - b_i \right] \eta(t),
-\end{equation}
-where $p, q \in \mathbb{Z}$, $p > q \geq 0$ define the order of the
-process; $r_i \in \mathbb{C}$, $\Re r_i < 0$ are the roots of the
-characteristic equation of the ODE; $b_j \in \mathbb{C}$,
-$\Re b_j < 0$ are corresponding roots of the inverse process; and
-$\eta(t)$ is a white-noise Gaussian Process with
-$\left\langle \eta(t) \right\rangle = 0$ and
-$\left\langle \eta(t) \eta(t') \right\rangle = \sigma^2 \delta\left(t
-  - t' \right)$.  Such a solution is called a CARMA (Continuous
-AutoRegressive Moving Average) process.  The constraint that
-$\Re r_i < 0$ and $\Re b_j < 0$ ensures that the linear operator
-defining the process and its inverse are invertable.  Under the
-assumption that the solution, $y(t)$, is real, roots $r_i$ and $b_j$
-must either be real or occur in complex conjugate pairs.  The power
-spectrum of the process is
-\begin{equation}
-  \label{eq:CARMA-PSD}
-  P_y(f) = \sigma^2 \left| \frac{\prod_{j=1}^q \left[ 2 \pi i f - b_j \right]}{\prod_{i=1}^p \left[ 2 \pi i f - r_i \right]} \right|^2.
-\end{equation}
-The power spectrum is a rational function of frequency with poles at
-$2\pi i f = r_i$ and zeros at $2\pi i f = b_i$; because of the
-invertable constraints, these poles and zeros all occur when
-$\Im f \neq 0$.  The autocorrelation function corresponding to the
-power spectrum in Eq.\ \eqref{eq:CARMA-PSD} is
-\begin{equation}
-  \label{eq:CARMA-ACF}
-  \rho_y (\tau) = \left\langle y\left( t \right) y \left( t + \tau \right) \right\rangle = \sum_{i=1}^p A_i^2 e^{r_i \tau},
-\end{equation}
-where the RMS amplitudes of the $p$ different modes, $A_i$, are
-functions of $\sigma$ and the values of the roots $r_i$ and $b_j$.  If
-$q = p-1$, then the $A_i$ are independent, while if $q < p-1$ then
-there are interdependencies among the $A_i$.  From either the power
-spectrum in Eq.\ \eqref{eq:CARMA-PSD} or the autocorrelation function
-in Eq.\ \eqref{eq:CARMA-ACF}, it is apparent that a real root $r_i$
-represents an exponentially-decaying mode with $r_i$ the rate
-constant, while a complex conjugate pair of $r_i$ represent an
-oscillatory mode with decay rate $\Re r_i$ and angular frequency
-$\left|\Im r_i\right|$.  The former is called an CAR(1) process, an
-Ornstein-Uhlenbeck process, or a damped random walk.
-\citet{Kelly2014} gives more details on CARMA processes, including
-formulae for $A_i$ as a function of $\sigma$, $r_i$, and $b_j$, and
-shows how a state-space model of the solution $y(t)$ can be tracked
-through a time-series of uncertain observations
-$y_k = y\left( t_k \right) + \epsilon_k$, with Gaussian noise
-$\epsilon_k$ of zero mean and known (heteroskedastic) variance, using
-a Kalman filter to produce a likelihood function
-$p\left( y_k \mid \sigma, r_i, b_j \right)$ in
+A Gaussian Process such as that used by \citet{brewer2009} also corresponds to
+the solution, $y(t)$, of the stochastic ODE \begin{equation}
+\label{eq:GP-definition} \prod_{i=1}^{p} \left[ \diff{}{t} - r_i \right] y(t) =
+\prod_{j = 1}^{q} \left[ \diff{}{t} - b_i \right] \eta(t), \end{equation} where
+$p, q \in \mathbb{Z}$, $p > q \geq 0$ define the order of the process; $r_i \in
+\mathbb{C}$, $\Re r_i < 0$ are the roots of the characteristic equation of the
+ODE; $b_j \in \mathbb{C}$, $\Re b_j < 0$ are corresponding roots of the inverse
+process; and $\eta(t)$ is a white-noise Gaussian Process with $\left\langle
+\eta(t) \right\rangle = 0$ and $\left\langle \eta(t) \eta(t') \right\rangle =
+\sigma^2 \delta\left(t - t' \right)$.  Such a solution is called a CARMA
+(Continuous AutoRegressive Moving Average) process.  The constraint that $\Re
+r_i < 0$ and $\Re b_j < 0$ ensures that the linear operator defining the process
+and its inverse are invertable.  Under the assumption that the solution, $y(t)$,
+is real, roots $r_i$ and $b_j$ must either be real or occur in complex conjugate
+pairs.  The power spectrum of the process is \begin{equation}
+\label{eq:CARMA-PSD} P_y(f) = \sigma^2 \left| \frac{\prod_{j=1}^q \left[ 2 \pi i
+f - b_j \right]}{\prod_{i=1}^p \left[ 2 \pi i f - r_i \right]} \right|^2.
+\end{equation} The power spectrum is a rational function of frequency with poles
+at $2\pi i f = r_i$ and zeros at $2\pi i f = b_i$; because of the invertable
+constraints, these poles and zeros all occur when $\Im f \neq 0$.  The
+autocorrelation function corresponding to the power spectrum in Eq.\
+\eqref{eq:CARMA-PSD} is \begin{equation} \label{eq:CARMA-ACF} \rho_y (\tau) =
+\left\langle y\left( t \right) y \left( t + \tau \right) \right\rangle =
+\sum_{i=1}^p A_i^2 e^{r_i \tau}, \end{equation} where the RMS amplitudes of the
+$p$ different modes, $A_i$, are functions of $\sigma$ and the values of the
+roots $r_i$ and $b_j$.  If $q = p-1$, then the $A_i$ are independent, while if
+$q < p-1$ then there are interdependencies among the $A_i$.  From either the
+power spectrum in Eq.\ \eqref{eq:CARMA-PSD} or the autocorrelation function in
+Eq.\ \eqref{eq:CARMA-ACF}, it is apparent that a real root $r_i$ represents an
+exponentially-decaying mode with $r_i$ the rate constant, while a complex
+conjugate pair of $r_i$ represent an oscillatory mode with decay rate $\Re r_i$
+and angular frequency $\left|\Im r_i\right|$.  Oscillatory modes can also be
+described by their frequency $f$ and quality factor $Q$ via $r = f/(2Q) \pm 2\pi
+i f$.  Decaying modes generate a damped random walk, also known as a
+Ornstein-Uhlenbeck process.  \citet{Kelly2014} gives more details on CARMA
+processes in an astronomical context, including formulae for $A_i$ as a function
+of $\sigma$, $r_i$, and $b_j$, and shows how a state-space model of the solution
+$y(t)$ can be tracked through a time-series of uncertain observations $y_k =
+y\left( t_k \right) + \epsilon_k$, with Gaussian noise $\epsilon_k$ of zero mean
+and known (heteroskedastic) variance, using a Kalman filter to produce a
+likelihood function $p\left( y_k \mid \sigma, r_i, b_j \right)$ in
 $\mathcal{O}\left( n \right)$ computational time.  Additionally,
-\citet{Foreman-Mackey2017} show how a dimensional expansion can be
-used to reduce the standard Gaussian Process covariance matrix for the
-process with autocorrelation given by Eq.\ \eqref{eq:CARMA-ACF} under
-observations at times $t_k$ to a banded form that can then be
-Cholesky-decomposed in the standard Gaussian Process likelihood
-function in $\mathcal{O}\left( n \right)$ time.
+\citet{Foreman-Mackey2017} show how a dimensional expansion can be used to
+reduce the standard Gaussian Process covariance matrix for the process with
+autocorrelation given by Eq.\ \eqref{eq:CARMA-ACF} under observations at times
+$t_k$ to a banded form that can then be Cholesky-decomposed in the standard
+Gaussian Process likelihood function in $\mathcal{O}\left( n \right)$ time.
 
@@
 The \texttt{CARMA.jl} package\footnote{\url{https://github.com/farr/CARMA.jl}} implements Kalman filters that can compute the likelihood for a set of observations parameterised by either the roots $r_i$ and $b_j$ or the RMS amplitudes $A_i$ and roots $r_i$.
-Because we have a set of observations taken with different instruments at different sites (see Section \ref{sec:archival-obs}), we include RV offset (i.e.\ mean value) parameters that are site-dependent and also a site-dependent scaling parameter.
-Thus, the $i$th RV measured at site $j$, $y_{ij}$, is assumed to be
+We also include a Keplerian RV signal from a planetary companion to Aldebaran \citep{Hatzes2015,Gregory2005}.
+Because we have a set of observations taken with different instruments at different sites (see Section \ref{sec:archival-obs}), we include RV offset (i.e.\ mean value) parameters that are instrument- and site-dependent and also an instrument- and site-dependent scaling parameter.
+Thus, the $i$th RV measured at site and intstrument combination $j$, $y_{ij}$, is assumed to be
 \begin{equation}
@@
 \end{equation}
-$t_i$ is the time of the measurement, $v_r$ is a planetary RV signal \citep{Hatzes2015,Gregory2005}, $\mu_j$ is the site-dependent mean RV offset, $y(t)$ is the CARMA process representing stellar activity, $\nu_j$ is the site-dependent uncertainty scaling parameter, and $\epsilon_{ij}$ is an independent Gaussian random variable with mean zero and standard deviation equal to the quoted uncertainty of the observation.
+$t_i$ is the time of the measurement, $v_r$ is the planetary RV signal, $\mu_j$ is the mean RV offset for site and instrument combination $j$, $y(t)$ is the CARMA process representing stellar activity, $\nu_j$ is the uncertainty scaling parameter for site and instrument combination $j$, and $\epsilon_{ij}$ is an independent Gaussian random variable with mean zero and standard deviation equal to the quoted uncertainty of the observation.
 We impose broad priors (see Table \ref{tab:parameters}) on the parameters of the planetary RV signal, the CARMA process, the $\mu_j$ offsets, and the $\nu_j$ scale factors.
+We treat the frequency of maximum asteroseismic power,  $\nu_\mathrm{max}$, as the frequency of an appropriate oscillatory mode in the CARMA process representing the stellar activity; this is equivalent to fitting a Lorentzian profile to a set of unresolved asteroseismic modes.
+We used the nested sampling algorithm of the \texttt{Ensemble.jl} package\footnote{\url{https://github.com/farr/Ensemble.jl}; this package implements several stochastic sampling algorithms based on the ``stretch move'' proposal used in \texttt{emcee} \citep{Foreman-Mackey2013} and introduced in \citet{Goodman2010}} to calculate the marginal likelihood (evidences) and draw samples from the posterior distribution over the parameters of our various models.
 
@@
 \tablecolumns{3}
-\tablecaption{\label{tab:parameters} The parameters in our model (see Eq.\ \eqref{eq:RV-model}).}
+\tablecaption{\label{tab:parameters} The parameters in our model Eq.\ \eqref{eq:RV-model} and their priors.
+Site and instrument combinations are indexed by $j$.
+Modes in the stochastic process are indexed by $k \in [1,p]$ and are either real (parameterised by a rate parameter $r_k$) or complex (parameterised by a frequency $f_k$ and quality factor $Q_k$).}
 \tablehead{\colhead{Parameter} & \colhead{Prior} & \colhead{Description}}
 \startdata
-\cutinhead{Per-Site RV Offset and Uncertainty Scaling}
-$\mu_j$ & $N\left[ \left\langle y_{j} \right\rangle, 10 \sigma_{y_{j}} \right] \left( \mu _j \right)\tablenotemark{a}$ & The per-site RV offset. \\
-$\nu_j$ & $1/\nu_j, \quad \nu_j \in [1/2, 2]$ & The per-site uncertainty scale factor. \\
+\cutinhead{Site/Instrument RV Offset and Uncertainty Scaling}
+$\mu_j$ & $N\left[ \left\langle y_{j} \right\rangle, 10 \sigma_{y_{j}} \right] \left( \mu _j \right)\tablenotemark{a}$ & The site/instrument RV offset. \\
+$\nu_j$ & $1/\nu_j, \quad \nu_j \in [1/2, 2]$ & The site/instrument uncertainty scale factor. \\
 \cutinhead{Keplerian Parameters}
@@
 $\chi$ & $\mathrm{const}, \quad 0 \leq \chi < 1$ & Pericentre passage is at $t = \chi P$\\
+\cutinhead{CARMA Parameters}
+$A_k$ & $1/A_k, \quad A_k \in \left[\min_j \sigma_{y_j}/100, \max_j 10 \sigma_{y_j} \right]$ & RMS amplitude of stochastic mode \\
+$r_k$ & $1/r_k, \quad \Re r_k \in \left[ 1/ (2T), 0.05 \, \mu\mathrm{Hz}\right], \Im r_k = 0$\tablenotemark{c} & The damping rates for real modes \\
+$f_k$ & $1/f_k, \quad f_k \in \left[ 0.1, 10 \right] \mu \mathrm{Hz}$ & The frequency of oscillatory modes \\
+$Q_k$ & $1/Q_k, \quad Q_k \in \left[ 1, 1000 \right]$ & The quality factor of oscillatory modes \\
 \enddata
-\tablenotetext{a}{$\left\langle y_j \right\rangle$ and $\sigma_{y_j}$ are the mean and standard deviation of the RV timeseries from site $j$.}
+\tablenotetext{a}{$\left\langle y_j \right\rangle$ and $\sigma_{y_j}$ are the mean and standard deviation of the RV timeseries from site and instrument combination $j$.}
 \tablenotetext{b}{We actually sample in and impose a (flat) prior on $x \equiv e \cos \omega$ and $y \equiv e \sin \omega$.}
+\tablenotetext{c}{Here $T$ is the time span of the measurements over all sites.}
 \end{deluxetable}
@@
 Its position near the Ecliptic permits the determination of its angular diameter by lunar occultations and by interferometry \citep[$20.58 \pm 0.03$ mas;][]{richichi2005,1979ApJ...228L.111B,brown1979,panek1980}.
-These tight constraints are valuable in breaking degeneracies in stellar modelling and make this an ideal star for asteroseismic characterization (Appendix~/ref{sec:mod}).
+These tight constraints are valuable in breaking degeneracies in stellar modelling and make this an ideal star for asteroseismic characterization (Appendix~\ref{sec:mod}).
 
@@
 \label{sec:archival-obs}
+
+\citet{Hatzes2015} reported on RV observations of Aldebaran from the coude
+échelle spectrograph of the Thüringer Landessternwarte Tautenburg (TLS), the
+Tull Spectrograph of the McDonald 2.7 m telescope, and the Bohyunsan Observatory
+Échelle Spectrograph (BOES) spectrograph of the Bohyunsan Optical Astronomy
+Observatory (BOAO) spanning the period from 2000.01 to 2013.92.  Combined with earlier observations \citep{hatzes1993}
 

### 2018-01-17 paper/aldebaran.tex
accents

@@
 
-\citet{Hatzes2015} reported on RV observations of Aldebaran from the coude
-échelle spectrograph of the Thüringer Landessternwarte Tautenburg (TLS), the
+\citet{Hatzes2015} reported on RV observations of Aldebaran from the coud\'{e}
+\'{e}chelle spectrograph of the Th\"{u}ringer Landessternwarte Tautenburg (TLS), the
 Tull Spectrograph of the McDonald 2.7 m telescope, and the Bohyunsan Observatory
-Échelle Spectrograph (BOES) spectrograph of the Bohyunsan Optical Astronomy
+\'{E}chelle Spectrograph (BOES) spectrograph of the Bohyunsan Optical Astronomy
 Observatory (BOAO) spanning the period from 2000.01 to 2013.92.  Combined with earlier observations \citep{hatzes1993}

### 2018-01-18 paper/aldebaran.tex
carma psd; fixed citation
text

@@
 \'{E}chelle Spectrograph (BOES) spectrograph of the Bohyunsan Optical Astronomy
-Observatory (BOAO) spanning the period from 2000.01 to 2013.92.  Combined with earlier observations \citep{hatzes1993}
+Observatory (BOAO) spanning the period from 2000.01 to 2013.92.
+Combined with earlier observations \citep{hatzes1993}, we construct a full time series spanning more than two decades and attempt to fit this with a Keplerian model and a CARMA noise process.
 
@@
 trend, probably due to the planetary companion.
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{psd-zoom.pdf}
+\caption{CARMA-derived power spectral density from archival and SONG RV observations of Aldebaran, showing the clear peak at around 2\,\muHz.
+The dark and light shaded regions represent $1\,\sigma$~and $2\,\sigma$ posterior probability contours.
+}
+\label{psd}
+\end{figure}
 

### 2018-01-18 paper/aldebaran.tex
titles
got that accursed bounding box
caption
figure labels

@@
 Observatory (BOAO) spanning the period from 2000.01 to 2013.92.
-Combined with earlier observations \citep{hatzes1993}, we construct a full time series spanning more than two decades and attempt to fit this with a Keplerian model and a CARMA noise process.
+Combined with earlier observations \citep{hatzes1993}, we construct a full time series spanning more than two decades (Figure~\ref{alldata}) and attempt to fit this with a Keplerian model and a CARMA noise process.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.8\textwidth]{all-data.pdf}
+\caption{All archival and new radial velocities used for inference in this work.}
+\label{alldata}
+\end{figure}
 
@@
 velocity precision per visit is 2.5$\,\rm m\,s^{-1}$ allowing us to easily detect the oscillations
-which displays a peak-to-peak amplitude of $\approx \, 170 \, \rm m \, s^{-1}$. $\bf Figure XX$ show
+which displays a peak-to-peak amplitude of $\approx \, 170 \, \rm m \, s^{-1}$. Figure~\ref{songdata} displays
 the radial velocities obtained -- oscillations are easily visible as well as a long-term
 trend, probably due to the planetary companion.
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{song-data.pdf}
+\caption{New radial velocities of Aldebaran obtained by the Hertzsprung~SONG Telescope. Clear oscillations can be seen, plus a trend representing part of the planetary orbit.}
+\label{songdata}
+\end{figure}
 

### 2018-01-18 paper/aldebaran.tex
carma appendix; placeholder text in body

@@
 
-A Gaussian Process such as that used by \citet{brewer2009} also corresponds to
-the solution, $y(t)$, of the stochastic ODE \begin{equation}
-\label{eq:GP-definition} \prod_{i=1}^{p} \left[ \diff{}{t} - r_i \right] y(t) =
-\prod_{j = 1}^{q} \left[ \diff{}{t} - b_i \right] \eta(t), \end{equation} where
-$p, q \in \mathbb{Z}$, $p > q \geq 0$ define the order of the process; $r_i \in
-\mathbb{C}$, $\Re r_i < 0$ are the roots of the characteristic equation of the
-ODE; $b_j \in \mathbb{C}$, $\Re b_j < 0$ are corresponding roots of the inverse
-process; and $\eta(t)$ is a white-noise Gaussian Process with $\left\langle
-\eta(t) \right\rangle = 0$ and $\left\langle \eta(t) \eta(t') \right\rangle =
-\sigma^2 \delta\left(t - t' \right)$.  Such a solution is called a CARMA
-(Continuous AutoRegressive Moving Average) process.  The constraint that $\Re
-r_i < 0$ and $\Re b_j < 0$ ensures that the linear operator defining the process
-and its inverse are invertable.  Under the assumption that the solution, $y(t)$,
-is real, roots $r_i$ and $b_j$ must either be real or occur in complex conjugate
-pairs.  The power spectrum of the process is \begin{equation}
-\label{eq:CARMA-PSD} P_y(f) = \sigma^2 \left| \frac{\prod_{j=1}^q \left[ 2 \pi i
-f - b_j \right]}{\prod_{i=1}^p \left[ 2 \pi i f - r_i \right]} \right|^2.
-\end{equation} The power spectrum is a rational function of frequency with poles
-at $2\pi i f = r_i$ and zeros at $2\pi i f = b_i$; because of the invertable
-constraints, these poles and zeros all occur when $\Im f \neq 0$.  The
-autocorrelation function corresponding to the power spectrum in Eq.\
-\eqref{eq:CARMA-PSD} is \begin{equation} \label{eq:CARMA-ACF} \rho_y (\tau) =
-\left\langle y\left( t \right) y \left( t + \tau \right) \right\rangle =
-\sum_{i=1}^p A_i^2 e^{r_i \tau}, \end{equation} where the RMS amplitudes of the
-$p$ different modes, $A_i$, are functions of $\sigma$ and the values of the
-roots $r_i$ and $b_j$.  If $q = p-1$, then the $A_i$ are independent, while if
-$q < p-1$ then there are interdependencies among the $A_i$.  From either the
-power spectrum in Eq.\ \eqref{eq:CARMA-PSD} or the autocorrelation function in
-Eq.\ \eqref{eq:CARMA-ACF}, it is apparent that a real root $r_i$ represents an
-exponentially-decaying mode with $r_i$ the rate constant, while a complex
-conjugate pair of $r_i$ represent an oscillatory mode with decay rate $\Re r_i$
-and angular frequency $\left|\Im r_i\right|$.  Oscillatory modes can also be
-described by their frequency $f$ and quality factor $Q$ via $r = f/(2Q) \pm 2\pi
-i f$.  Decaying modes generate a damped random walk, also known as a
-Ornstein-Uhlenbeck process.  \citet{Kelly2014} gives more details on CARMA
-processes in an astronomical context, including formulae for $A_i$ as a function
-of $\sigma$, $r_i$, and $b_j$, and shows how a state-space model of the solution
+Fortunately, this GP can be written as the solution of a class of stochastic ordinary differential equations, the Continuous Auto-Regressive Moving Average (CARMA) models, whose likelihoods can be evaluated in linear time.
+For problems such as the damped driven harmonic oscillator, we therefore have access to these powerful computational tools for inferring their power spectral densities.
+\citet{Kelly2014} gives more details on CARMA
+processes in an astronomical context, and shows how a state-space model of the solution
 $y(t)$ can be tracked through a time-series of uncertain observations $y_k =
@@
 
+\section{CARMA Models}
+\label{carmatheory}
+
+A Gaussian Process such as that used by \citet{brewer2009} also corresponds to
+the solution, $y(t)$, of the stochastic ODE \begin{equation}
+\label{eq:GP-definition} \prod_{i=1}^{p} \left[ \diff{}{t} - r_i \right] y(t) =
+\prod_{j = 1}^{q} \left[ \diff{}{t} - b_i \right] \eta(t), \end{equation} where
+$p, q \in \mathbb{Z}$, $p > q \geq 0$ define the order of the process; $r_i \in
+\mathbb{C}$, $\Re r_i < 0$ are the roots of the characteristic equation of the
+ODE; $b_j \in \mathbb{C}$, $\Re b_j < 0$ are corresponding roots of the inverse
+process; and $\eta(t)$ is a white-noise Gaussian Process with $\left\langle
+\eta(t) \right\rangle = 0$ and $\left\langle \eta(t) \eta(t') \right\rangle =
+\sigma^2 \delta\left(t - t' \right)$.  Such a solution is called a CARMA
+(Continuous AutoRegressive Moving Average) process.  The constraint that $\Re
+r_i < 0$ and $\Re b_j < 0$ ensures that the linear operator defining the process
+and its inverse are invertable.  Under the assumption that the solution, $y(t)$,
+is real, roots $r_i$ and $b_j$ must either be real or occur in complex conjugate
+pairs.  The power spectrum of the process is \begin{equation}
+\label{eq:CARMA-PSD} P_y(f) = \sigma^2 \left| \frac{\prod_{j=1}^q \left[ 2 \pi i
+f - b_j \right]}{\prod_{i=1}^p \left[ 2 \pi i f - r_i \right]} \right|^2.
+\end{equation} The power spectrum is a rational function of frequency with poles
+at $2\pi i f = r_i$ and zeros at $2\pi i f = b_i$; because of the invertable
+constraints, these poles and zeros all occur when $\Im f \neq 0$.  The
+autocorrelation function corresponding to the power spectrum in Eq.\
+\eqref{eq:CARMA-PSD} is \begin{equation} \label{eq:CARMA-ACF} \rho_y (\tau) =
+\left\langle y\left( t \right) y \left( t + \tau \right) \right\rangle =
+\sum_{i=1}^p A_i^2 e^{r_i \tau}, \end{equation} where the RMS amplitudes of the
+$p$ different modes, $A_i$, are functions of $\sigma$ and the values of the
+roots $r_i$ and $b_j$.  If $q = p-1$, then the $A_i$ are independent, while if
+$q < p-1$ then there are interdependencies among the $A_i$.  From either the
+power spectrum in Eq.\ \eqref{eq:CARMA-PSD} or the autocorrelation function in
+Eq.\ \eqref{eq:CARMA-ACF}, it is apparent that a real root $r_i$ represents an
+exponentially-decaying mode with $r_i$ the rate constant, while a complex
+conjugate pair of $r_i$ represent an oscillatory mode with decay rate $\Re r_i$
+and angular frequency $\left|\Im r_i\right|$.  Oscillatory modes can also be
+described by their frequency $f$ and quality factor $Q$ via $r = f/(2Q) \pm 2\pi
+i f$.  Decaying modes generate a damped random walk, also known as a
+Ornstein-Uhlenbeck process.
+
 \section{Halo Photometry}

### 2018-01-18 presub/natastron_presub_coverletter.tex
remove principal authors
cover letter wording

@@
 \noindent I would like to express my interest in submitting a Letter to \emph{Nature Astronomy}, on the ``Detection of Oscillations in Aldebaran with Ground-Based Observations''.
-The principal authors are myself, Prof.
-Guy Davies (Birmingham), and Dr Benjamin Pope (NYU).
 
-Aldebaran ($\alpha$~Tauri, the Eye of the Bull) is a nearby naked-eye red giant star.
-Since the identification of the planet candidate Aldebaran~b in~1993 in the first modern radial velocity (RV) surveys, a further two decades of observations have firmly established the existence of a massive planet around the star.
+Aldebaran ($\alpha$~Tauri, the Eye of the Bull) is a nearby first-magnitude red giant star.
+Since the identification of the planet candidate Aldebaran~b in~1993 in the first modern radial velocity (RV) surveys, the existence of a massive planet orbiting the star has been firmly established with a further two decades of observations.
 
-With radical new data processing techniques, we have re-analysed these historical RV data and augmented them with new observations of our own.
-Using our Gaussian Process-based model we infer much more accurate orbital parameters for the planet, but also detect the signal of stellar oscillations, missed by previous authors but uncovered by our powerful new algorithms.
+One astronomer's noise is another astronomer's data: we have found that the noise in these archival observations is actually the signal of stellar oscillations.
+With radical new data processing techniques, we have re-analysed historical RV data and confirmed our result with new observations.
+Using our Gaussian Process-based model we detect the signal of stellar oscillations, missed by previous authors but uncovered by our powerful new algorithms, and consequently infer much more accurate orbital parameters for the planet.
 With asteroseismology we constrain the star's age for the first time, measure its mass with a factor of two better precision than had previously been possible (to 6\%), and consequently obtain the mass of the giant planet to a precision of 12\%.
 
-We have also independently confirmed our results with new photometry from the \emph{Kepler}-2 (K2) mission, again using a highly novel data processing technique to extract a precise light curve of the brightest ever observed by \emph{Kepler}, ten thousand times above its saturation limit.
+We have also independently confirmed our results with new photometry from the \emph{Kepler}-2 (K2) mission, again using a highly novel data processing technique to extract a precise light curve of the brightest star ever observed by \emph{Kepler}.
 The K2 light curve shows extremely clear oscillations that are in perfect agreement with the frequency we infer from RV.
@@
 \closing{Yours sincerely, \\
-\vspace{1.75cm}
+\vspace{1.5cm}
 Will Farr}

### 2018-01-19 paper/aldebaran.tex
Added Julia to software stack in ack's.

@@
 
-\software{\textsc{IPython} \citep{PER-GRA:2007}, SciPy \citep{scipy}, \textsc{matplotlib} \citep{Hunter:2007}, \textsc{Astropy} \citep{2013A&A...558A..33A}, \textsc{NumPy} \citep{van2011numpy}, \textsc{isochrones} \citep{isochrones}.}
+\software{\textsc{IPython} \citep{PER-GRA:2007}, SciPy \citep{scipy}, \textsc{matplotlib} \citep{Hunter:2007}, \textsc{Astropy} \citep{2013A&A...558A..33A}, \textsc{NumPy} \citep{van2011numpy}, \textsc{isochrones} \citep{isochrones}, \textsc{Julia} \citep{Bezanson2017}.}
 

### 2018-01-19 paper/aldebaran.tex
Several updates around results and archival data section.
* Some wordsmithing and clarity in time-series and CARMA appendix.

* Plot of nu_max posteriors from archival, SONG, and combined data sets.

* More numbers (nu_max and Q posterior ranges) in archival, SONG, and
combined sections.

@@
 \newcommand{\bibtex}{\textsc{Bib}\!\TeX} % bibtex. Not quite the correct typesetting, but close enough
+
+% Numbers in the document
+\newcommand{\twosidedrange}[3]{\ensuremath{#1^{+#2}_{-#3}}}
+
+%% Q
+\newcommand{\HatzesQRange}{\twosidedrange{2.12}{0.69}{0.56}}
+\newcommand{\SONGQRange}{\twosidedrange{3.06}{1.20}{0.81}}
+\newcommand{\CombinedQRange}{\twosidedrange{2.83}{0.60}{0.51}}
+
+%% NuMax
+\newcommand{\HatzesNuMaxRange}{\twosidedrange{2.21}{0.13}{0.12}}
+\newcommand{\SONGNuMaxRange}{\twosidedrange{2.33}{0.13}{0.14}}
+\newcommand{\combinedNuMaxRange}{\twosidedrange{2.238}{0.086}{0.080}}
+
+%% P
+\newcommand{\PeriodRange}{\twosidedrange{629.0}{1.7}{2.1}}
+
+%% e
+\newcommand{\EccentricityRange}{\twosidedrange{0.168}{0.069}{0.071}}
+
+%% K
+\newcommand{\PlanetRVAmpRange}{\twosidedrange{127}{15}{14}}
 
@@
 
-If we apply nonlinear statistical inference methods this situation can be improved.
-\citet{brewer2009} show that a system of driven, damped harmonic oscillators such as we encounter in asteroseismology can be statistically modelled as a Gaussian Process, with a covariance kernel consisting of a sum of damped sinusoids.
-The hyperparameters of this process encode features of the power spectral density distributions and are insensitive to the window function.
-Unfortunately, this is impractical for long time series, as the computational cost of evaluating the standard Gaussian Process likelihood function scales as $\mathcal{O}(n^3)$.
+If we apply nonlinear statistical inference methods this situation can be
+improved. \citet{brewer2009} show that a system of driven, damped harmonic
+oscillators such as we encounter in asteroseismology can be statistically
+modelled as a Gaussian Process, with a covariance kernel consisting of a sum of
+damped sinusoids. The hyperparameters of this process encode features of the
+power spectral density distributions and are insensitive to the window function.
+Unfortunately, this is impractical for long time series, as the computational
+cost of evaluating the standard Gaussian Process likelihood function scales as
+$\mathcal{O}(n^3)$.
 
-Fortunately, this GP can be written as the solution of a class of stochastic ordinary differential equations, the Continuous Auto-Regressive Moving Average (CARMA) models, whose likelihoods can be evaluated in linear time.
-For problems such as the damped driven harmonic oscillator, we therefore have access to these powerful computational tools for inferring their power spectral densities.
-\citet{Kelly2014} gives more details on CARMA
-processes in an astronomical context, and shows how a state-space model of the solution
-$y(t)$ can be tracked through a time-series of uncertain observations $y_k =
-y\left( t_k \right) + \epsilon_k$, with Gaussian noise $\epsilon_k$ of zero mean
-and known (heteroskedastic) variance, using a Kalman filter to produce a
-likelihood function $p\left( y_k \mid \sigma, r_i, b_j \right)$ in
-$\mathcal{O}\left( n \right)$ computational time.  Additionally,
-\citet{Foreman-Mackey2017} show how a dimensional expansion can be used to
-reduce the standard Gaussian Process covariance matrix for the process with
-autocorrelation given by Eq.\ \eqref{eq:CARMA-ACF} under observations at times
-$t_k$ to a banded form that can then be Cholesky-decomposed in the standard
-Gaussian Process likelihood function in $\mathcal{O}\left( n \right)$ time.
+Fortunately, this GP can be written as the solution of a class of stochastic
+ordinary differential equations, the Continuous Auto-Regressive Moving Average
+(CARMA) models, whose likelihoods can be evaluated in linear time. For problems
+such as the damped driven harmonic oscillator, we therefore have access to these
+powerful computational tools for inferring their power spectral densities.
+\citet{Kelly2014} gives more details on CARMA processes in an astronomical
+context, and shows how a state-space model of the solution $y(t)$ can be tracked
+through a time-series of uncertain observations $y_k = y\left( t_k \right) +
+\epsilon_k$, with Gaussian noise $\epsilon_k$ of zero mean and known
+(heteroskedastic) variance, using a Kalman filter to produce a likelihood
+function $p\left( y_k \mid \sigma, r_i, b_j \right)$ depending on the amplitude
+of the stochastic forcing, $\sigma$, the (possibly complex) eigenfrequencies of
+the ODE, $r_i$, and parameters describing the power spectrum of the stochastic
+forcing, $b_j$ in $\mathcal{O}\left( n \right)$ computational time.
+Additionally, \citet{Foreman-Mackey2017} show how a dimensional expansion can be
+used to reduce the standard Gaussian Process covariance matrix for this same
+process,  with autocorrelation given by Eq.\ \eqref{eq:CARMA-ACF} under
+observations at times $t_k$, to a banded form that can then be
+Cholesky-decomposed in the standard Gaussian Process likelihood function in
+$\mathcal{O}\left( n \right)$ time.
 
-Here we have adopted the \citet{Kelly2014} approach.
-The \texttt{CARMA.jl} package\footnote{\url{https://github.com/farr/CARMA.jl}} implements Kalman filters that can compute the likelihood for a set of observations parameterised by either the roots $r_i$ and $b_j$ or the RMS amplitudes $A_i$ and roots $r_i$.
-We also include a Keplerian RV signal from a planetary companion to Aldebaran \citep{Hatzes2015,Gregory2005}.
-Because we have a set of observations taken with different instruments at different sites (see Section \ref{sec:archival-obs}), we include RV offset (i.e.\ mean value) parameters that are instrument- and site-dependent and also an instrument- and site-dependent scaling parameter.
-Thus, the $i$th RV measured at site and intstrument combination $j$, $y_{ij}$, is assumed to be
+Here we have adopted the \citet{Kelly2014} approach.  The \texttt{CARMA.jl}
+package\footnote{\url{https://github.com/farr/CARMA.jl}} implements Kalman
+filters that can compute the likelihood for a set of observations parameterised
+by either the $r_i$ and $b_j$ or the RMS amplitudes $A_i$ of each eigenmode of
+the ODE and the corresponding roots $r_i$.  We also include a Keplerian RV
+signal from a planetary companion to Aldebaran \citep{Hatzes2015,Gregory2005}.
+Because we have a set of observations taken with different instruments at
+different sites (see Section \ref{sec:archival-obs}), we include RV offset
+(i.e.\ mean value) parameters that are instrument- and site-dependent and also
+an instrument- and site-dependent scaling parameter.  Thus, the $i$th RV
+measured at site and intstrument combination $j$, $y_{ij}$, is assumed to be
+%
 \begin{equation}
   \label{eq:RV-model}
-  y_{ij} = v_r\left(t_i\right) + \mu_j + y\left( t_i \right) + \nu_j \epsilon_{ij},
+  y_{ij} = v_r\left(t_i\right) + \mu_j + y\left( t_i \right) + \nu_j
+\epsilon_{ij},
 \end{equation}
-$t_i$ is the time of the measurement, $v_r$ is the planetary RV signal, $\mu_j$ is the mean RV offset for site and instrument combination $j$, $y(t)$ is the CARMA process representing stellar activity, $\nu_j$ is the uncertainty scaling parameter for site and instrument combination $j$, and $\epsilon_{ij}$ is an independent Gaussian random variable with mean zero and standard deviation equal to the quoted uncertainty of the observation.
-We impose broad priors (see Table \ref{tab:parameters}) on the parameters of the planetary RV signal, the CARMA process, the $\mu_j$ offsets, and the $\nu_j$ scale factors.
-We treat the frequency of maximum asteroseismic power,  $\nu_\mathrm{max}$, as the frequency of an appropriate oscillatory mode in the CARMA process representing the stellar activity; this is equivalent to fitting a Lorentzian profile to a set of unresolved asteroseismic modes.
-We used the nested sampling algorithm of the \texttt{Ensemble.jl} package\footnote{\url{https://github.com/farr/Ensemble.jl}; this package implements several stochastic sampling algorithms based on the ``stretch move'' proposal used in \texttt{emcee} \citep{Foreman-Mackey2013} and introduced in \citet{Goodman2010}} to calculate the marginal likelihood (evidences) and draw samples from the posterior distribution over the parameters of our various models.
-
-\begin{deluxetable}{lll}
-\tablecolumns{3}
-\tablecaption{\label{tab:parameters} The parameters in our model Eq.\ \eqref{eq:RV-model} and their priors.
-Site and instrument combinations are indexed by $j$.
-Modes in the stochastic process are indexed by $k \in [1,p]$ and are either real (parameterised by a rate parameter $r_k$) or complex (parameterised by a frequency $f_k$ and quality factor $Q_k$).}
-\tablehead{\colhead{Parameter} & \colhead{Prior} & \colhead{Description}}
-\startdata
-\cutinhead{Site/Instrument RV Offset and Uncertainty Scaling}
-$\mu_j$ & $N\left[ \left\langle y_{j} \right\rangle, 10 \sigma_{y_{j}} \right] \left( \mu _j \right)\tablenotemark{a}$ & The site/instrument RV offset. \\
-$\nu_j$ & $1/\nu_j, \quad \nu_j \in [1/2, 2]$ & The site/instrument uncertainty scale factor. \\
-\cutinhead{Keplerian Parameters}
-$K$ & $1/K, \quad K \in \left[ \min_j \sigma_{y_j}/100, \max_j 10 \sigma_{y_j} \right]$ & RV semi-amplitude. \\
-$P$ & $1/P, \quad P \in [600, 700] \, \mathrm{d}$ & Keplerian period \\
-$e$ & $2e,\quad 0 \leq e < 1$\tablenotemark{b} & Keplerian eccentricity \\
-$\omega$ & $\mathrm{const}, \quad 0 \leq \omega < 2\pi$ & Longitude of pericentre \\
-$\chi$ & $\mathrm{const}, \quad 0 \leq \chi < 1$ & Pericentre passage is at $t = \chi P$\\
-\cutinhead{CARMA Parameters}
-$A_k$ & $1/A_k, \quad A_k \in \left[\min_j \sigma_{y_j}/100, \max_j 10 \sigma_{y_j} \right]$ & RMS amplitude of stochastic mode \\
-$r_k$ & $1/r_k, \quad \Re r_k \in \left[ 1/ (2T), 0.05 \, \mu\mathrm{Hz}\right], \Im r_k = 0$\tablenotemark{c} & The damping rates for real modes \\
-$f_k$ & $1/f_k, \quad f_k \in \left[ 0.1, 10 \right] \mu \mathrm{Hz}$ & The frequency of oscillatory modes \\
-$Q_k$ & $1/Q_k, \quad Q_k \in \left[ 1, 1000 \right]$ & The quality factor of oscillatory modes \\
-\enddata
-\tablenotetext{a}{$\left\langle y_j \right\rangle$ and $\sigma_{y_j}$ are the mean and standard deviation of the RV timeseries from site and instrument combination $j$.}
-\tablenotetext{b}{We actually sample in and impose a (flat) prior on $x \equiv e \cos \omega$ and $y \equiv e \sin \omega$.}
-\tablenotetext{c}{Here $T$ is the time span of the measurements over all sites.}
-\end{deluxetable}
+%
+where $t_i$ is the time of the measurement, $v_r$ is the planetary RV signal,
+$\mu_j$ is the mean RV offset for site and instrument combination $j$, $y(t)$ is
+the CARMA process representing stellar activity, $\nu_j$ is the uncertainty
+scaling parameter for site and instrument combination $j$, and $\epsilon_{ij}$
+is an independent Gaussian random variable with mean zero and standard deviation
+equal to the quoted uncertainty of the observation.  We impose broad priors (see
+Table \ref{tab:parameters}) on the parameters of the planetary RV signal, the
+CARMA process, the $\mu_j$ offsets, and the $\nu_j$ scale factors.  We treat the
+frequency of maximum asteroseismic power, $\nu_\mathrm{max}$, as the frequency
+of an appropriate eigenmode in the CARMA process representing the stellar
+activity; this is equivalent to fitting a Lorentzian profile to a set of
+unresolved asteroseismic modes.  We used the nested sampling algorithm of the
+\texttt{Ensemble.jl} package\footnote{\url{https://github.com/farr/Ensemble.jl};
+this package implements several stochastic sampling algorithms based on the
+``stretch move'' proposal used in \texttt{emcee} \citep{Foreman-Mackey2013} and
+introduced in \citet{Goodman2010}} to calculate the marginal likelihood
+(evidences) and draw samples from the posterior distribution over the parameters
+of our various models.  The priors we impose on the parameters of our model are given in Table \ref{tab:parameters}.
 
@@
 \citet{Hatzes2015} reported on RV observations of Aldebaran from the coud\'{e}
-\'{e}chelle spectrograph of the Th\"{u}ringer Landessternwarte Tautenburg (TLS), the
-Tull Spectrograph of the McDonald 2.7 m telescope, and the Bohyunsan Observatory
-\'{E}chelle Spectrograph (BOES) spectrograph of the Bohyunsan Optical Astronomy
-Observatory (BOAO) spanning the period from 2000.01 to 2013.92.
-Combined with earlier observations \citep{hatzes1993}, we construct a full time series spanning more than two decades (Figure~\ref{alldata}) and attempt to fit this with a Keplerian model and a CARMA noise process.
+\'{e}chelle spectrograph of the Th\"{u}ringer Landessternwarte Tautenburg (TLS),
+the Tull Spectrograph of the McDonald 2.7 m telescope, and the Bohyunsan
+Observatory \'{E}chelle Spectrograph (BOES) spectrograph of the Bohyunsan
+Optical Astronomy Observatory (BOAO) spanning the period from 2000.01 to
+2013.92.  Combined with earlier observations \citep{hatzes1993}, we construct a
+full time series spanning more than two decades from 1980.80 to 2013.71
+(Figure~\ref{alldata}) and fit this with a Keplerian model and a CARMA process.
+We find evidence for a low-quality ($Q = \HatzesQRange{}$) oscillatory mode with
+$\nu_\mathrm{max} = \HatzesNuMaxRange{} \, \mu\mathrm{Hz}$.  This is consistent
+with the presence of a number of un-resolved asteroseismic modes in the RV data.
+Our posterior for \numax{} from this data set is shown in Figure
+\ref{fig:numax-datasets}.
 
@@
 
+\begin{figure}
+  \includegraphics[width=\columnwidth]{../plots/numax}
+  \caption{Posterior for \numax{} from the complete, combined RV data set described in this paper (blue), the archival data from \citet{Hatzes2015} (green), and the SONG data described in Section \ref{sec:SONG-observations} (red).
+The inference on the archival data is consistent with the independent inference on the SONG data set alone.}
+  \label{fig:numax-datasets}
+\end{figure}
+
 \subsection{SONG Observations}    % this goes into section 4.1
-
-In order to confirm the oscillations detected in the archival data, we used the Hertzsprung SONG telescope \citep{2017ApJ...836..142G} to conduct high time cadence follow-up observations.
-These were carried out in the highest resolution mode ($R\approx110000$)
-with an integration time of 30\,s, between 2016 September 27 and 2017 January 12. During this time we attempted to obtain at least one
-visit per available night, for a total of 254 epochs over the campaign. The spectra were
-extracted using the SONG pipeline \citep[see][]{2017ApJ...836..142G}. SONG employs an iodine cell for
-precise wavelength calibration and the velocities were determined using the \textit{iSONG}
-software following the same procedures as described in \cite{2017ApJ...836..142G}. The typical
-velocity precision per visit is 2.5$\,\rm m\,s^{-1}$ allowing us to easily detect the oscillations
-which displays a peak-to-peak amplitude of $\approx \, 170 \, \rm m \, s^{-1}$. Figure~\ref{songdata} displays
-the radial velocities obtained -- oscillations are easily visible as well as a long-term
-trend, probably due to the planetary companion.
+\label{sec:SONG-observations} In order to confirm the oscillations detected in
+the archival data, we used the Hertzsprung SONG telescope
+\citep{2017ApJ...836..142G} to conduct high time cadence follow-up observations.
+These were carried out in the highest resolution mode ($R\approx110000$) with an
+integration time of 30\,s, between 2016 September 27 and 2017 January 12. During
+this time we attempted to obtain at least one visit per available night, for a
+total of 254 epochs over the campaign. The spectra were extracted using the SONG
+pipeline \citep[see][]{2017ApJ...836..142G}. SONG employs an iodine cell for
+precise wavelength calibration and the velocities were determined using the
+\textit{iSONG} software following the same procedures as described in
+\cite{2017ApJ...836..142G}. The typical velocity precision per visit is
+2.5$\,\rm m\,s^{-1}$ allowing us to easily detect the oscillations which
+displays a peak-to-peak amplitude of $\approx \, 170 \, \rm m \, s^{-1}$.
+Figure~\ref{songdata} displays the radial velocities obtained -- oscillations
+are easily visible as well as a long-term trend, probably due to the planetary
+companion.  From this data set alone we also find evidence for a low-quality ($Q
+= \SONGQRange{}$) oscillatory mode with $\nu_\mathrm{max} = \SONGNuMaxRange{} \,
+\mu\mathrm{Hz}$ which is consistent with the inference from the archival data.
+This posterior is shown in Figure \ref{fig:numax-datasets}.
 
@@
 
-By jointly fitting a Keplerian for the planet with a CARMA model for the stellar oscillations, we obtain a precise estimate of both \numax and the planetary orbital parameters.
-The planetary orbital parameters we obtain are similar to \citet{Hatzes2015}: period $P = 629.0^{+ 1.7}_{-2.1}$~d, eccentricity $e = 0.168^{+ 0.069}_{- 0.071}$; and radial velocity semi-amplitude $K = 127^{+ 15}_{- 14}$~m/s.
+By jointly fitting a Keplerian for the planet with a CARMA model for the stellar
+oscillations to the combined archival and SONG data sets (Figure \ref{alldata}),
+we obtain a precise estimate of both \numax and the planetary orbital
+parameters. The planetary orbital parameters we obtain are similar to
+\citet{Hatzes2015}, but with larger uncertainty that is likely due to our
+more-flexible and correlated model for the stellar component of the RV signal:
+period $P = \PeriodRange \, \mathrm{d}$, eccentricity $e =
+\EccentricityRange{}$; and radial velocity semi-amplitude $K =
+\PlanetRVAmpRange{} \, \mathrm{m} \, \mathrm{s}^{-1}$.
 
-The stellar oscillations are best fit with a single CARMA component with $\numax = 2.238^{+ 0.086}_{- 0.080}$ \muHz, and including additional CARMA components does not resolve any further structure in the power spectrum.
-We have determined the stellar properties of Aldebaran with the additional constraint of \numax determined from this analysis.
-The details of the stellar modelling are presented in detail in Appendix \ref{sec:mod}.
-We have considered the impact of our additional \numax constraint and run our analysis for different sets of spectroscopic estimates.
-For final values we adopt the \citet{2012Sheffield} spectroscopic solution as being `middle of the road' estimates.
-For the analysis without constraint on \numax we find estimates of the stellar properties as $M = 1.27^{+0.24}_{-0.2} \, \mathrm{M_{\odot}}$ and age $4.86^{+3.56}_{-2.04} \, \rm Gyr$.
-With the inclusion of \numax we find $M = 1.16^{+0.07}_{-0.07} \, \mathrm{M_{\odot}}$ and age $6.38^{+1.42}_{-1.12} \, \rm Gyr$.
-It is clear that the addition of \numax provides substantially more precise estimates of mass and age.
-With this stellar mass, the radial velocity signal translates to an inclination-dependent planetary mass of $M\,\sin\,{i} = 5.81 \pm 0.72~M_J$ \citep[Equation~1]{Torres2008}, a massive giant planet; for sufficiently high inclinations ($\gtrsim 45^\circ$), the mass of Aldebaran~b may exceed $13~M_J$, making it a possible low-mass brown dwarf candidate.
-% re sheffield - I have substituted the citation from the .bib but you said 'b'; afaik she only published one paper in 2012?
-- Ben
+The stellar oscillations are best fit with a single, low-quality ($Q =
+\CombinedQRange{}$) oscillatory mode with $\numax = 2.238^{+ 0.086}_{- 0.080}$
+\muHz.  We find no improvement in the marginal likelihood (evidence) or other
+model-selection information criteria \citep{Gelman2013} from models with
+additional oscillatory modes, so we conclude that the large spacing $\Delta\nu$
+is not observable in this data set.  We have determined the stellar properties
+of Aldebaran with the additional constraint of \numax determined from this
+analysis.  The details of the stellar modelling are presented in Appendix
+\ref{sec:mod}.  We have considered the impact of our additional \numax
+constraint and run our analysis for different sets of spectroscopic estimates.
+For final values we adopt the \citet{2012Sheffield} spectroscopic solution as
+being `middle of the road' estimates.  For the analysis without constraint on
+\numax we find estimates of the stellar properties as $M = 1.27^{+0.24}_{-0.2}
+\, \mathrm{M_{\odot}}$ and age $4.86^{+3.56}_{-2.04} \, \rm Gyr$.  With the
+inclusion of \numax we find $M = 1.16^{+0.07}_{-0.07} \, \mathrm{M_{\odot}}$ and
+age $6.38^{+1.42}_{-1.12} \, \rm Gyr$.  It is clear that the addition of \numax
+provides substantially more precise estimates of mass and age. With this stellar
+mass, the radial velocity signal translates to an inclination-dependent
+planetary mass of $M\,\sin\,{i} = 5.81 \pm 0.72~M_J$
+\citep[Equation~1]{Torres2008}, a massive giant planet; for sufficiently high
+inclinations ($\gtrsim 45^\circ$), the mass of Aldebaran~b may exceed $13~M_J$,
+making it a possible low-mass brown dwarf candidate.
 
@@
 
+The data, code, and \LaTeX{} source used to produce this document can be found
+at \url{https://github.com/farr/Aldebaran} under an open-source ``MIT'' license.
+
 % This work made use of the \textsc{IPython} package \citep{PER-GRA:2007}; SciPy \citep{scipy};  \textsc{matplotlib}, a \textsc{Python} library for publication quality graphics \citep{Hunter:2007}; \textsc{Astropy}, a community-developed core \textsc{Python} package for astronomy \citep{2013A&A...558A..33A}; and \textsc{NumPy} \citep{van2011numpy}.
@@
 
-A Gaussian Process such as that used by \citet{brewer2009} also corresponds to
-the solution, $y(t)$, of the stochastic ODE \begin{equation}
-\label{eq:GP-definition} \prod_{i=1}^{p} \left[ \diff{}{t} - r_i \right] y(t) =
-\prod_{j = 1}^{q} \left[ \diff{}{t} - b_i \right] \eta(t), \end{equation} where
-$p, q \in \mathbb{Z}$, $p > q \geq 0$ define the order of the process; $r_i \in
-\mathbb{C}$, $\Re r_i < 0$ are the roots of the characteristic equation of the
-ODE; $b_j \in \mathbb{C}$, $\Re b_j < 0$ are corresponding roots of the inverse
-process; and $\eta(t)$ is a white-noise Gaussian Process with $\left\langle
-\eta(t) \right\rangle = 0$ and $\left\langle \eta(t) \eta(t') \right\rangle =
-\sigma^2 \delta\left(t - t' \right)$.  Such a solution is called a CARMA
-(Continuous AutoRegressive Moving Average) process.  The constraint that $\Re
+A CARMA process corresponds to the solution, $y(t)$, of the stochastic ODE
+%
+\begin{equation}
+  \label{eq:GP-definition}
+  \prod_{i=1}^{p} \left[ \diff{}{t} - r_i \right] y(t) =
+  \prod_{j = 1}^{q} \left[ \diff{}{t} - b_i \right] \eta(t),
+\end{equation}
+%
+where $p, q \in \mathbb{Z}$, $p > q \geq 0$ define the order of the process;
+$r_i \in \mathbb{C}$, $\Re r_i < 0$ are the roots of the characteristic equation
+of the ODE; $b_j \in \mathbb{C}$, $\Re b_j < 0$ are corresponding roots of the
+inverse process; and $\eta(t)$ is a white-noise Gaussian Process with
+$\left\langle \eta(t) \right\rangle = 0$ and $\left\langle \eta(t) \eta(t')
+\right\rangle = \sigma^2 \delta\left(t - t' \right)$.  The constraint that $\Re
 r_i < 0$ and $\Re b_j < 0$ ensures that the linear operator defining the process
@@
 is real, roots $r_i$ and $b_j$ must either be real or occur in complex conjugate
-pairs.  The power spectrum of the process is \begin{equation}
-\label{eq:CARMA-PSD} P_y(f) = \sigma^2 \left| \frac{\prod_{j=1}^q \left[ 2 \pi i
-f - b_j \right]}{\prod_{i=1}^p \left[ 2 \pi i f - r_i \right]} \right|^2.
-\end{equation} The power spectrum is a rational function of frequency with poles
-at $2\pi i f = r_i$ and zeros at $2\pi i f = b_i$; because of the invertable
-constraints, these poles and zeros all occur when $\Im f \neq 0$.  The
-autocorrelation function corresponding to the power spectrum in Eq.\
-\eqref{eq:CARMA-PSD} is \begin{equation} \label{eq:CARMA-ACF} \rho_y (\tau) =
-\left\langle y\left( t \right) y \left( t + \tau \right) \right\rangle =
-\sum_{i=1}^p A_i^2 e^{r_i \tau}, \end{equation} where the RMS amplitudes of the
-$p$ different modes, $A_i$, are functions of $\sigma$ and the values of the
-roots $r_i$ and $b_j$.  If $q = p-1$, then the $A_i$ are independent, while if
-$q < p-1$ then there are interdependencies among the $A_i$.  From either the
-power spectrum in Eq.\ \eqref{eq:CARMA-PSD} or the autocorrelation function in
-Eq.\ \eqref{eq:CARMA-ACF}, it is apparent that a real root $r_i$ represents an
-exponentially-decaying mode with $r_i$ the rate constant, while a complex
-conjugate pair of $r_i$ represent an oscillatory mode with decay rate $\Re r_i$
-and angular frequency $\left|\Im r_i\right|$.  Oscillatory modes can also be
-described by their frequency $f$ and quality factor $Q$ via $r = f/(2Q) \pm 2\pi
-i f$.  Decaying modes generate a damped random walk, also known as a
-Ornstein-Uhlenbeck process.
+pairs.  The power spectrum of the process is
+%
+\begin{equation}
+  \label{eq:CARMA-PSD}
+  P_y(f) = \sigma^2 \left| \frac{\prod_{j=1}^q \left[ 2 \pi i f - b_j \right]}{\prod_{i=1}^p \left[ 2 \pi i f - r_i \right]} \right|^2.
+\end{equation}
+%
+The power spectrum is a rational function of frequency with poles at $2\pi i f =
+r_i$ and zeros at $2\pi i f = b_i$; because of the invertable constraints, these
+poles and zeros all occur when $\Im f \neq 0$.  The autocorrelation function
+corresponding to the power spectrum in Eq.\ \eqref{eq:CARMA-PSD} is
+%
+\begin{equation}
+  \label{eq:CARMA-ACF}
+  \rho_y (\tau) = \left\langle y\left( t \right) y \left( t + \tau \right) \right\rangle = \sum_{i=1}^p A_i^2 e^{r_i \tau},
+\end{equation}
+%
+where the RMS amplitudes of the $p$ different modes, $A_i$, are functions of
+$\sigma$ and the values of the roots $r_i$ and $b_j$.  If $q = p-1$, then the
+$A_i$ are independent, while if $q < p-1$ then there are interdependencies among
+the $A_i$.  From either the power spectrum in Eq.\ \eqref{eq:CARMA-PSD} or the
+autocorrelation function in Eq.\ \eqref{eq:CARMA-ACF}, it is apparent that a
+real root $r_i$ represents an exponentially-decaying mode with $r_i$ the rate
+constant, while a complex conjugate pair of $r_i$ represent an oscillatory mode
+with decay rate $\Re r_i$ and angular frequency $\left|\Im r_i\right|$.
+Oscillatory modes can also be described by their frequency $f$ and quality
+factor $Q$ via $r = f/(2Q) \pm 2\pi i f$.  Decaying modes generate a damped
+random walk, also known as a Ornstein-Uhlenbeck process.
+More discussion of CARMA processes, including formulae for $A_i$ in terms of $\sigma$, $r_i$, and $b_j$, can be found in \citet{Kelly2014} and references therein.
+
+The priors we impose on the parameters of the CARMA model, planetary RV signal,
+and instrumental and site effects are given in Table \ref{tab:parameters}.
+
+\begin{deluxetable}{lll}
+\tablecolumns{3}
+\tablecaption{\label{tab:parameters} The parameters in our model Eq.\ \eqref{eq:RV-model} and their priors.
+Site and instrument combinations are indexed by $j$.
+Modes in the stochastic process are indexed by $k \in [1,p]$ and are either real (parameterised by a rate parameter $r_k$) or complex (parameterised by a frequency $f_k$ and quality factor $Q_k$).}
+\tablehead{\colhead{Parameter} & \colhead{Prior} & \colhead{Description}}
+\startdata
+\cutinhead{Site/Instrument RV Offset and Uncertainty Scaling}
+$\mu_j$ & $N\left[ \left\langle y_{j} \right\rangle, 10 \sigma_{y_{j}} \right] \left( \mu _j \right)\tablenotemark{a}$ & The site/instrument RV offset. \\
+$\nu_j$ & $1/\nu_j, \quad \nu_j \in [1/2, 2]$ & The site/instrument uncertainty scale factor. \\
+\cutinhead{Keplerian Parameters}
+$K$ & $1/K, \quad K \in \left[ \min_j \sigma_{y_j}/100, \max_j 10 \sigma_{y_j} \right]$ & RV semi-amplitude. \\
+$P$ & $1/P, \quad P \in [600, 700] \, \mathrm{d}$ & Keplerian period \\
+$e$ & $2e,\quad 0 \leq e < 1$\tablenotemark{b} & Keplerian eccentricity \\
+$\omega$ & $\mathrm{const}, \quad 0 \leq \omega < 2\pi$ & Longitude of pericentre \\
+$\chi$ & $\mathrm{const}, \quad 0 \leq \chi < 1$ & Pericentre passage is at $t = \chi P$\\
+\cutinhead{CARMA Parameters}
+$A_k$ & $1/A_k, \quad A_k \in \left[\min_j \sigma_{y_j}/100, \max_j 10 \sigma_{y_j} \right]$ & RMS amplitude of stochastic mode \\
+$r_k$ & $1/r_k, \quad \Re r_k \in \left[ 1/ (2T), 0.05 \, \mu\mathrm{Hz}\right], \Im r_k = 0$\tablenotemark{c} & The damping rates for real modes \\
+$f_k$ & $1/f_k, \quad f_k \in \left[ 0.1, 10 \right] \mu \mathrm{Hz}$ & The frequency of oscillatory modes \\
+$Q_k$ & $1/Q_k, \quad Q_k \in \left[ 1, 1000 \right]$ & The quality factor of oscillatory modes \\
+\enddata
+\tablenotetext{a}{$\left\langle y_j \right\rangle$ and $\sigma_{y_j}$ are the mean and standard deviation of the RV timeseries from site and instrument combination $j$.}
+\tablenotetext{b}{We actually sample in and impose a (flat) prior on $x \equiv e \cos \omega$ and $y \equiv e \sin \omega$.}
+\tablenotetext{c}{Here $T$ is the time span of the measurements over all sites.}
+\end{deluxetable}
 

### 2018-01-20 code/Aldebaran.jl
CARMAKepler has moved to CARMA package; code updated.

@@ -1,14 +1,13 @@
 module Aldebaran
 
-using Kepler
-using CARMAKepler
+using CARMA
 
 function load_timeseries(;before_song=true, song=true)
     ts = []
     ys = []
     dys = []
 
-    if before_song 
+    if before_song
         for i in 3:9
             data = readdlm("..//data/$(i)/table$(i).dat")
             push!(ts, data[:,1])

### 2018-01-20 code/CARMAKepler.jl

@@ -1,447 +1,0 @@
-module CARMAKepler
-
-using CARMA
-using Distributions
-using Ensemble
-using HDF5
-using Kepler
-
-import Base:
-    write, read
-
-type MultiEpochPosterior
-    ts::Array{Array{Float64, 1}, 1}
-    ys::Array{Array{Float64, 1}, 1}
-    dys::Array{Array{Float64, 1}, 1}
-
-    inds::Array{Array{Int, 1}, 1}
-    
-    allts::Array{Float64, 1}
-    allys::Array{Float64, 1}
-    alldys::Array{Float64, 1}
-    
-    P_min::Float64
-    P_max::Float64
-
-    K_min::Float64
-    K_max::Float64
-
-    ndrw::Int
-    nosc::Int
-
-    rate_min::Float64
-    rate_max::Float64
-
-    f_min::Float64
-    f_max::Float64
-
-    rms_min::Float64
-    rms_max::Float64
-
-    Q_min::Float64
-    Q_max::Float64
-end
-
-function MultiEpochPosterior(ts, ys, dys, per_min, per_max, ndrw, nosc, rate_min, rate_max, f_min, f_max, Q_max)
-    allts = vcat(ts...)
-    allys = vcat(ys...)
-    alldys = vcat(dys...)
-    
-    allinds = sortperm(allts)
-
-    iallinds = invperm(allinds)
-    
-    inds = []
-    i = 1
-    for t in ts
-        push!(inds, iallinds[i:i+size(t,1)-1])
-        i = i+size(t,1)
-    end
-
-    allts = allts[allinds]
-    allys = allys[allinds]
-    alldys = alldys[allinds]
-
-    rms_max = maximum([std(y) for y in ys])
-    rms_min = minimum([std(y) for y in ys])
-
-    T = allts[end]-allts[1]
-    dtmin = minimum(diff(allts))
-    
-    MultiEpochPosterior(ts, ys, dys, inds, allts, allys, alldys, per_min, per_max, rms_min/100.0, 10.0*rms_max, ndrw, nosc, rate_min, rate_max, f_min, f_max, rms_min/100.0, rms_max*10.0, 1.0, Q_max)
-end
-
-type MultiEpochParams
-    mu::Array{Float64, 1}
-    nu::Array{Float64, 1}
-    
-    K::Float64
-    P::Float64
-    e::Float64
-    omega::Float64
-    chi::Float64
-
[…]

### 2018-01-20 code/Kepler.jl

@@ -1,71 +1,0 @@
-module Kepler
-
-function ecc_anom(M, e)
-    M = mod2pi(M)
-    if M < zero(M)
-        M = M + 2*pi
-    end
-
-    Emin = zero(M)
-    fmin = nothing
-    Emax = 2*pi*one(M)
-    fmax = nothing
-    Eguess = M
-    while true
-        f = M - Eguess + e*sin(Eguess)
-
-        if abs(f) < convert(typeof(f), 1e-10)
-            return Eguess
-        end
-        df = -one(Eguess) + e*cos(Eguess)
-
-        Enewton = Eguess - f/df
-
-        if Enewton < Emin || Enewton > Emax
-            fmin = M - Emin - e*sin(Emin)
-            fmax = M - Emax - e*sin(Emax)
-
-            if fmin == zero(fmin)
-                return Emin
-            elseif fmax == zero(fmin)
-                return Emax
-            end
-
-            if fmin*f < zero(fmin)
-                Emax = Eguess
-            elseif fmax*f < zero(fmax)
-                Emin = Eguess
-            else
-                error("Not bracketing a root in bisection.")
-            end
-            Eguess = (Emax-Emin)/2
-        else
-            Eguess = Enewton
-        end
-    end
-end
-
-function true_anom(ea, e)
-    ta = 2*atan(sqrt((one(e)+e)/(one(e)-e))*tan(ea/2))
-    if ta < zero(ta)
-        ta + 2*pi
-    else
-        ta
-    end
-end
-
-function rv(t, K, P, e, omega, chi)
-    fp, ip = modf(t/P - chi)
-
-    if fp < zero(fp)
-        fp = fp + one(fp)
-    end
-
-    m = 2*pi*fp
-    
-    ta = true_anom(ecc_anom(m, e), e)
-
-    K*(cos(ta + omega) + e*cos(omega))
-end
-
-end

### 2018-01-25 paper/aldebaran.tex
Ack Hatzes+(2015) for making data public.
haywood and rajpaul citations

@@
 
+Radial velocity variations intrinsic to the star, such as those caused by stellar oscillations, have previously limited the precision with which red giant planets have been studied \citep{2005PASJ...57...97S}.
+We aim to model out these noise processes and use them recover asteroseismic information, as well as to improve our estimates of the orbital parameters of the planet.
 It is easy to observe Aldebaran and similarly-bright stars with ground-based spectroscopic instruments, typically only requiring short exposures that can be obtained even under adverse observing conditions.
@@
 power spectral density distributions and are insensitive to the window function.
+Gaussian processes with quasi-periodic kernels have previously been used to re-analyse RV data for main sequence stars \citep{2014MNRAS.443.2517H,2015MNRAS.452.2269R}, where the noise is from stellar activity, but have not previously been applied to red giant asteroseismology.
 Unfortunately, this is impractical for long time series, as the computational
@@
 We would like to thank Tim Bedding, David Hogg and Dan Foreman-Mackey for their helpful discussions.
+
+We would like to commend the authors of \citet{Hatzes2015} for making their data
+publicly available on SIMBAD.
 

### 2018-01-25 paper/aldebaran.tex
phrasing

@@
 
-Fortunately, this GP can be written as the solution of a class of stochastic
-ordinary differential equations, the Continuous Auto-Regressive Moving Average
+Fortunately, the damped and driven harmonic oscillator GP can be written as the solution of a class of stochastic ordinary differential equations, the Continuous Auto-Regressive Moving Average
 (CARMA) models, whose likelihoods can be evaluated in linear time. For problems
-such as the damped driven harmonic oscillator, we therefore have access to these
+such as this, we therefore have access to these
 powerful computational tools for inferring their power spectral densities.

### 2018-01-25 paper/aldebaran.tex
text about gamma dra

@@
 making it a possible low-mass brown dwarf candidate.
+A recent study by \citet{2018arXiv180105239H} has cast doubt on the validity of the detection of the massive planet $\gamma$~Draconis~b, and by extension similar planets around similar K~giants such as Aldebaran.
+We remain convinced that Aldebaran~b is a bona fide planet for reasons outlined in Appendix~\ref{gamma_dra}.
 
@@
 
+\section{The $\gamma$~Draconis Problem is Not a Problem}
+\label{gamma_dra}
+
+$\gamma$~Draconis (also known as Eltanin, or $\gamma$~Dra) is a second-magnitude K5III giant which had been observed by \citet{hatzes1993} as part of the same campaign that led to the discovery of Aldebaran~b and Pollux~b.
+\citet{2018arXiv180105239H} have recently shown that the 702~d period RV variations of the putative Eltanin~b disappeared from 2011-2013, returning in 2014 with a different amplitude and phase.
+They ascribe this to the beating between two modes of a previously-unidentified kind of stellar variability, and warn that ``Given that the periods found in α Tau are comparable to those in $\gamma$~Dra and both stars are evolved with large radii, a closer scrutiny of the RV variability of $\alpha$~Tau \[Aldebaran\] is warranted.''
+
+For several reasons we are not convinced that this poses an issue for Aldebaran or for K~giant RV planets more generally.
+
 \bibliography{aldebaran}

### 2018-01-25 paper/aldebaran.tex
some reasons why gamma dra is probably fine

@@
 For several reasons we are not convinced that this poses an issue for Aldebaran or for K~giant RV planets more generally.
+\citet{2018arXiv180105239H} suggest that the new type of stellar variability may be oscillatory convective modes, but both Aldebaran and Eltanin have much lower luminosities and longer periods than would seem to be allowed by the period-luminosity relation for these otherwise unobserved modes \citep[Figure~\9]{2015MNRAS.452.3863S,2018arXiv180105239H}.
+If these are identified with the long secondary periods (LSPs) observed in some bright red giants ($L \approx 1000 \lsun$), then Aldebaran and Eltanin are both too faint, and lack the mid-IR excess typical of LSP stars \citep{2009ApJ...707..573W}.
+It would also be surprising if the shape of the RV curve could reproduce the harmonic structure of an eccentric Keplerian such as in Aldebaran~b ($e = \EccentricityRange{}$), let alone the much higher eccentricities observed in other giants \citep[e.g.
+$\iota$~Dra, $e=0.7$:][]{2002ApJ...576..478F}.
 

### 2018-01-25 paper/aldebaran.tex
transits

@@
 
+It would be a cruel conspiracy of nature if red giants support a type of oscillation which is common and closely resembles a planetary signal.
+We believe this cannot be the case: the populations of planets around subgiants and giants evolved from intermediate mass stars are similar \citep{2014A&A...566A.113J}, as expected if subgiants evolve into giants and retain their planetary systems in a similar configuration, and with very different stellar structures subgiants and giants are unlikely to share similar modes of long-period pulsation.
+The similarity of distributions of systems hosted by subgiant and giants could not be reproduced if a large fraction of the giants' planets were false positives.
+Moreover giant planets are expected to be common around intermediate-mass stars \citep{2008ApJ...673..502K}, and some are definitely known to be bona fide planets because they transit their star \citep[e.g.][]{2014A&A...562A.109L,2015A&A...573L...6O,2016AJ....152..185G,2017AJ....154..254G}.
+
 \bibliography{aldebaran}

### 2018-01-25 paper/aldebaran.tex
coda

@@
 
+We therefore believe that either Eltanin~b is not a false positive, or if it is, that it is not a common type of false positive, and is unlikely to affect our certainty that Aldebaran~b is real.
+\citet{2018arXiv180105239H} offer an alternative to the pulsation hypothesis, that one `mode' is the stellar rotation and the other is the planetary signal; this does not seem implausible.
+
 \bibliography{aldebaran}

### 2018-01-25 paper/aldebaran.tex
fixes

@@
 \citet{2018arXiv180105239H} have recently shown that the 702~d period RV variations of the putative Eltanin~b disappeared from 2011-2013, returning in 2014 with a different amplitude and phase.
-They ascribe this to the beating between two modes of a previously-unidentified kind of stellar variability, and warn that ``Given that the periods found in α Tau are comparable to those in $\gamma$~Dra and both stars are evolved with large radii, a closer scrutiny of the RV variability of $\alpha$~Tau \[Aldebaran\] is warranted.''
+They ascribe this to the beating between two modes of a previously-unidentified kind of stellar variability, and warn that ``Given that the periods found in $\alpha$~Tau [Aldebaran] are comparable to those in $\gamma$~Dra and both stars are evolved with large radii, a closer scrutiny of the RV variability of $\alpha$~Tau is warranted.''
 
 For several reasons we are not convinced that this poses an issue for Aldebaran or for K~giant RV planets more generally.
-\citet{2018arXiv180105239H} suggest that the new type of stellar variability may be oscillatory convective modes, but both Aldebaran and Eltanin have much lower luminosities and longer periods than would seem to be allowed by the period-luminosity relation for these otherwise unobserved modes \citep[Figure~\9]{2015MNRAS.452.3863S,2018arXiv180105239H}.
+\citet{2018arXiv180105239H} suggest that the new type of stellar variability may be oscillatory convective modes, but both Aldebaran and Eltanin have much lower luminosities and longer periods than would seem to be allowed by the period-luminosity relation for these otherwise unobserved modes \citep[Figure~9]{2015MNRAS.452.3863S,2018arXiv180105239H}.
 If these are identified with the long secondary periods (LSPs) observed in some bright red giants ($L \approx 1000 \lsun$), then Aldebaran and Eltanin are both too faint, and lack the mid-IR excess typical of LSP stars \citep{2009ApJ...707..573W}.
 It would also be surprising if the shape of the RV curve could reproduce the harmonic structure of an eccentric Keplerian such as in Aldebaran~b ($e = \EccentricityRange{}$), let alone the much higher eccentricities observed in other giants \citep[e.g.
-$\iota$~Dra, $e=0.7$:][]{2002ApJ...576..478F}.
+$\iota$~Dra~b, $e=0.7$:][]{2002ApJ...576..478F}.
 

### 2018-01-25 paper/aldebaran.tex
phrasing
reference to sato

@@
 A recent study by \citet{2018arXiv180105239H} has cast doubt on the validity of the detection of the massive planet $\gamma$~Draconis~b, and by extension similar planets around similar K~giants such as Aldebaran.
-We remain convinced that Aldebaran~b is a bona fide planet for reasons outlined in Appendix~\ref{gamma_dra}.
+We remain convinced that the $\gamma$~Draconis$ problem is not an issue in this case, and Aldebaran~b is a bona fide planet, for reasons outlined in Appendix~\ref{gamma_dra}.
 
@@
 This can naturally be extended to \emph{Gaia} in space \citep{gaia}, or \emph{LSST} from the ground \citep{dmt,lsst,lsstbook}, from which many thousands of new asteroseismic determinations will be possible.
+In future RV surveys, it may even be possible to beat the \citet{2005PASJ...57...97S} $\sim 30$~m/s RV precision limit for red giants by taking many closely spaced observations and modelling-out the effects of stellar oscillations: this will allow us to dig deeper into this intrinsic stellar noise to detect less-massive planets around these stars.
 

### 2018-01-25 paper/aldebaran.tex
fixes

@@
 A recent study by \citet{2018arXiv180105239H} has cast doubt on the validity of the detection of the massive planet $\gamma$~Draconis~b, and by extension similar planets around similar K~giants such as Aldebaran.
-We remain convinced that the $\gamma$~Draconis$ problem is not an issue in this case, and Aldebaran~b is a bona fide planet, for reasons outlined in Appendix~\ref{gamma_dra}.
+We remain convinced that the $\gamma$~Draconis problem is not an issue in this case, and Aldebaran~b is a bona fide planet, for reasons outlined in Appendix~\ref{gamma_dra}.
 

### 2018-01-26 paper/aldebaran.tex
Minor edits to wording, some references back and forth to Appendix 1.

@@
 Gaussian processes with quasi-periodic kernels have previously been used to re-analyse RV data for main sequence stars \citep{2014MNRAS.443.2517H,2015MNRAS.452.2269R}, where the noise is from stellar activity, but have not previously been applied to red giant asteroseismology.
-Unfortunately, this is impractical for long time series, as the computational
+Unfortunately, these methods are impractical for long time series, as the computational
 cost of evaluating the standard Gaussian Process likelihood function scales as
@@
 
-Fortunately, the damped and driven harmonic oscillator GP can be written as the solution of a class of stochastic ordinary differential equations, the Continuous Auto-Regressive Moving Average
-(CARMA) models, whose likelihoods can be evaluated in linear time. For problems
-such as this, we therefore have access to these
-powerful computational tools for inferring their power spectral densities.
-\citet{Kelly2014} gives more details on CARMA processes in an astronomical
-context, and shows how a state-space model of the solution $y(t)$ can be tracked
-through a time-series of uncertain observations $y_k = y\left( t_k \right) +
-\epsilon_k$, with Gaussian noise $\epsilon_k$ of zero mean and known
-(heteroskedastic) variance, using a Kalman filter to produce a likelihood
-function $p\left( y_k \mid \sigma, r_i, b_j \right)$ depending on the amplitude
-of the stochastic forcing, $\sigma$, the (possibly complex) eigenfrequencies of
-the ODE, $r_i$, and parameters describing the power spectrum of the stochastic
-forcing, $b_j$ in $\mathcal{O}\left( n \right)$ computational time.
-Additionally, \citet{Foreman-Mackey2017} show how a dimensional expansion can be
-used to reduce the standard Gaussian Process covariance matrix for this same
-process,  with autocorrelation given by Eq.\ \eqref{eq:CARMA-ACF} under
+Fortunately, the damped and driven harmonic oscillator GP can be written as the
+solution of a class of stochastic ordinary differential equations, the
+Continuous Auto-Regressive Moving Average (CARMA) models, whose likelihoods can
+be evaluated in linear time. For problems such as this, we therefore have access
+to these powerful computational tools for inferring their power spectral
+densities.  Our treatment of CARMA models is described more fully in Appendix
+\ref{carmatheory}.   \citet{Kelly2014} also gives more details on CARMA
+processes in an astronomical context, and shows how a state-space model of the
+process $y(t)$ can be tracked through a time-series of uncertain observations
+$y_k = y\left( t_k \right) + \epsilon_k$, with Gaussian noise $\epsilon_k$ of
+zero mean and known (heteroskedastic) variance, using a Kalman filter to produce
+a likelihood function $p\left( y_k \mid \sigma, r_i, b_j \right)$ depending on
+the amplitude of the stochastic forcing, $\sigma$, the (possibly complex)
+eigenfrequencies of the ODE, $r_i$, and parameters describing the power spectrum
+of the stochastic forcing, $b_j$ in $\mathcal{O}\left( n \right)$ computational
+time.  Additionally, \citet{Foreman-Mackey2017} show how a dimensional expansion
+can be used to reduce the standard Gaussian Process covariance matrix for this
+same process,  with autocorrelation given by Eq.\ \eqref{eq:CARMA-ACF} under
 observations at times $t_k$, to a banded form that can then be
@@
 (evidences) and draw samples from the posterior distribution over the parameters
-of our various models.  The priors we impose on the parameters of our model are given in Table \ref{tab:parameters}.
+of our various models.
 
@@
 \subsection{Planetary and Stellar Parameters}
+\label{sec:planet-star-parameters}
 
@@
 
-The priors we impose on the parameters of the CARMA model, planetary RV signal,
+Our preferred model for the data uses one decaying mode and one oscillatory mode
+in the CARMA process to capture the un-resolved superposition of many
+asteroseismic modes (see Section \ref{sec:planet-star-parameters}); we have fit
+models with more than one mode of each type, but find no improvement in
+model-selection criteria from the expanded parameter space.  The priors we
+impose in all cases on the parameters of the CARMA model, planetary RV signal,
 and instrumental and site effects are given in Table \ref{tab:parameters}.

### 2018-01-26 presub/coverletter.tex
updated coverletter

--- 2018-01-18 presub/natastron_presub_coverletter.tex
@@
 
-\topmargin=-1.25in    % Make letterhead start about 1 inch from top of page 
-\textheight=10.75in    % text height can be bigger for a longer letter
+\topmargin=-1.5in    % Make letterhead start about 1 inch from top of page 
+\textheight=11.75in    % text height can be bigger for a longer letter
 \oddsidemargin=0pt   % leftmargin is 1 inch
@@
 \begin{letter}
-{Nature Astronomy \\
-natureastro@nature.com \\}
+{Leslie Sage \\
+l.sage@us.nature.com \\}
 
@@
 
-\opening{Dear Editor,}
+\opening{Dear Leslie,}
 
-\noindent I would like to express my interest in submitting a Letter to \emph{Nature Astronomy}, on the ``Detection of Oscillations in Aldebaran with Ground-Based Observations''.
+\noindent I submit for your consideration a Letter on the ``Detection of Oscillations in Aldebaran with Ground-Based Observations''.
 
@@
 With radical new data processing techniques, we have re-analysed historical RV data and confirmed our result with new observations.
-Using our Gaussian Process-based model we detect the signal of stellar oscillations, missed by previous authors but uncovered by our powerful new algorithms, and consequently infer much more accurate orbital parameters for the planet.
+Using our Gaussian Process model we detect the signal of stellar oscillations, missed by previous authors, and consequently infer much more accurate orbital parameters for the planet.
+This technique will be widely applicable to legacy and future datasets, allowing both asteroseismology with RV, and also searches for lower-mass companions to giants than was previously possible.
+
+We have also independently confirmed our results with new photometry from the \emph{Kepler}-2 (K2) mission, again using a highly novel algorithm to extract a precise light curve of the brightest star ever observed by \emph{Kepler}.
+The K2 light curve shows very clear oscillations that are in perfect agreement with the frequency we infer from RV.
+
 With asteroseismology we constrain the star's age for the first time, measure its mass with a factor of two better precision than had previously been possible (to 6\%), and consequently obtain the mass of the giant planet to a precision of 12\%.
 
-We have also independently confirmed our results with new photometry from the \emph{Kepler}-2 (K2) mission, again using a highly novel data processing technique to extract a precise light curve of the brightest star ever observed by \emph{Kepler}.
-The K2 light curve shows extremely clear oscillations that are in perfect agreement with the frequency we infer from RV.
+A neat and surprising result of this work is that while the planet has a hellish temperature today well upwards of a thousand Kelvin, when Aldebaran was on the main sequence, the planet (and more importantly any of its moons) would have been subject to similar incident starlight as the Earth receives from the Sun.
+We posit that this is the first clear example of a formerly-habitable world rendered uninhabitable by stellar evolution.
 
-A neat and surprising result of this work is that while the planet has a hellish temperature today well upwards of a thousand Kelvin, when Aldebaran was on the main sequence, the planet (and more importantly any of its moons) would have been subject to similar incident starlight as the Earth receives from the Sun.
-We posit that this is the first clear example of a formerly-habitable world destroyed by stellar evolution.
-
-Thank you for considering our submission. Please do not hesitate to contact me with any questions you may have.\\
+Thank you for considering our submission. Please do not hesitate to contact me with any questions.\\
 
 \closing{Yours sincerely, \\
-\vspace{1.5cm}
+\vspace{1.0cm}
 Will Farr}

### 2018-01-30 paper/aldebaran.tex
Pope, Davies swap in author order.
Only w.farr@bham.ac.uk on email.
wording

@@
 
+\email{w.farr@bham.ac.uk}
+
 \affiliation{Birmingham Institute for Gravitational Wave Astronomy,
@@
   Birmingham, Birmingham, B15 2TT, United Kingdom}
+
+\author[0000-0003-2595-9114]{Benjamin J. S. Pope}
+\affiliation{Sydney Institute for Astronomy (SIfA), School of Physics, University of Sydney, NSW 2006, Australia}
+\affiliation{Center for Cosmology and Particle Physics, Department of Physics, New York University, 726 Broadway, New York, NY 10003, USA}
+\affiliation{NASA Sagan Fellow}
 
@@
 and Astronomy, Aarhus University, Ny Munkegade 120, DK-8000 Aarhus C, Denmark}
-
-\author[0000-0003-2595-9114]{Benjamin J. S. Pope}
-\affiliation{Sydney Institute for Astronomy (SIfA), School of Physics, University of Sydney, NSW 2006, Australia}
-\affiliation{Center for Cosmology and Particle Physics, Department of Physics, New York University, 726 Broadway, New York, NY 10003, USA}
-\affiliation{NASA Sagan Fellow}
 
@@
 \affiliation{Stellar Astrophysics Centre, Department of Physics and Astronomy, Aarhus University, Ny Munkegade 120, DK-8000 Aarhus C, Denmark}
-
-\email{w.farr@bham.ac.uk, G.R.Davies@bham.ac.uk, benjamin.pope@nyu.edu, txn016@student.bham.ac.uk, jimbarrett27@gmail.com}
 
@@
 \citet{2018arXiv180105239H} suggest that the new type of stellar variability may be oscillatory convective modes, but both Aldebaran and Eltanin have much lower luminosities and longer periods than would seem to be allowed by the period-luminosity relation for these otherwise unobserved modes \citep[Figure~9]{2015MNRAS.452.3863S,2018arXiv180105239H}.
-If these are identified with the long secondary periods (LSPs) observed in some bright red giants ($L \approx 1000 \lsun$), then Aldebaran and Eltanin are both too faint, and lack the mid-IR excess typical of LSP stars \citep{2009ApJ...707..573W}.
+If these are identified with the long secondary periods (LSPs) observed in some bright red giants ($L \approx 1000\, \lsun$), then Aldebaran and Eltanin are both too faint, and lack the mid-IR excess typical of LSP stars \citep{2009ApJ...707..573W}.
 It would also be surprising if the shape of the RV curve could reproduce the harmonic structure of an eccentric Keplerian such as in Aldebaran~b ($e = \EccentricityRange{}$), let alone the much higher eccentricities observed in other giants \citep[e.g.
@@
 It would be a cruel conspiracy of nature if red giants support a type of oscillation which is common and closely resembles a planetary signal.
-We believe this cannot be the case: the populations of planets around subgiants and giants evolved from intermediate mass stars are similar \citep{2014A&A...566A.113J}, as expected if subgiants evolve into giants and retain their planetary systems in a similar configuration, and with very different stellar structures subgiants and giants are unlikely to share similar modes of long-period pulsation.
+We believe this cannot be the case: the populations of planets around subgiants and giants evolved from intermediate mass stars are similar \citep{2014A&A...566A.113J}, as expected if subgiants evolve into giants and retain their planetary systems, and with very different stellar structures subgiants and giants are unlikely to share modes of long-period pulsation.
 The similarity of distributions of systems hosted by subgiant and giants could not be reproduced if a large fraction of the giants' planets were false positives.

### 2018-01-30 paper/aldebaran.tex
wording

@@
 \citet{2018arXiv180105239H} have recently shown that the 702~d period RV variations of the putative Eltanin~b disappeared from 2011-2013, returning in 2014 with a different amplitude and phase.
-They ascribe this to the beating between two modes of a previously-unidentified kind of stellar variability, and warn that ``Given that the periods found in $\alpha$~Tau [Aldebaran] are comparable to those in $\gamma$~Dra and both stars are evolved with large radii, a closer scrutiny of the RV variability of $\alpha$~Tau is warranted.''
+They ascribe this to a previously-unidentified kind of stellar variability, and warn that ``Given that the periods found in $\alpha$~Tau [Aldebaran] are comparable to those in $\gamma$~Dra and both stars are evolved with large radii, a closer scrutiny of the RV variability of $\alpha$~Tau is warranted.''
 

### 2018-01-30 paper/aldebaran.tex
wording

@@
 For several reasons we are not convinced that this poses an issue for Aldebaran or for K~giant RV planets more generally.
-\citet{2018arXiv180105239H} suggest that the new type of stellar variability may be oscillatory convective modes, but both Aldebaran and Eltanin have much lower luminosities and longer periods than would seem to be allowed by the period-luminosity relation for these otherwise unobserved modes \citep[Figure~9]{2015MNRAS.452.3863S,2018arXiv180105239H}.
+\citet{2018arXiv180105239H} suggest that the new type of stellar variability may be oscillatory convective modes, but both Aldebaran and Eltanin have much lower luminosities and longer periods than would seem to be allowed by the period-luminosity relation predicted for these otherwise unobserved modes \citep[Figure~9]{2015MNRAS.452.3863S,2018arXiv180105239H}.
 If these are identified with the long secondary periods (LSPs) observed in some bright red giants ($L \approx 1000\, \lsun$), then Aldebaran and Eltanin are both too faint, and lack the mid-IR excess typical of LSP stars \citep{2009ApJ...707..573W}.

### 2018-01-30 paper/aldebaran.tex
wording

@@
 In the power spectrum of either radial velocity of photometric observations, these modes of oscillation form a characteristic pattern of peaks which can be used to infer intrinsic stellar properties \citep[e.g.,][]{2016AN....337..774D}.
-The easiest properties of the pattern to determine are the the frequency of maximum amplitude \numax and the so-called large separation \Dnu \citep{Kjeldsen95}, which are often referred to as global asteroseismic parameters.
+The easiest properties of the pattern to determine are the frequency of maximum amplitude \numax and the so-called large separation \Dnu \citep{Kjeldsen95}, which are often referred to as global asteroseismic parameters.
 
@@
 We therefore believe that either Eltanin~b is not a false positive, or if it is, that it is not a common type of false positive, and is unlikely to affect our certainty that Aldebaran~b is real.
-\citet{2018arXiv180105239H} offer an alternative to the pulsation hypothesis, that one `mode' is the stellar rotation and the other is the planetary signal; this does not seem implausible.
+\citet{2018arXiv180105239H} offer an alternative to the pulsation hypothesis, namely beating between the stellar rotation and the planetary signal; this does not seem implausible.
 

### 2018-01-30 paper/aldebaran.tex
wording

@@
 Reversing the equation to produce a predicted $\log{g}$ from the observed $\nu_{\textrm{max,obs}}=2.23\pm0.1\mu$Hz results in a predicted $\log{g}\sim1.2$ dex, using an assumed temperature of 3900K.
-The solar calibration values used here are $g_{\odot}=274$ms$^{-2}$, $\nu_{\textrm{max},\odot}=3150\mu$Hz and $T_{\textrm{eff},\odot}=5777$K.
+The solar calibration values used here are $g_{\odot}=274\,$ms$^{-2}$, $\nu_{\textrm{max},\odot}=3150\,\mu$Hz and $T_{\textrm{eff},\odot}=5777\,$K.
 
@@
 
-We have also investigated running \textsc{PARAM} and relaxing one or more constraints: without the luminosity constraint, the average absolute mass offset between the two sets of results was $<0.04\textrm{M}_{\odot}$ for the runs without $\nu_{\textrm{max}}$ constraint, and $<0.02\textrm{M}_{\odot}$ for the runs with $\nu_{\textrm{max}}$.
+We have also investigated running \textsc{PARAM} and relaxing one or more constraints: between runs with and without the luminosity constraint, the average absolute mass offset between the two sets of results was $<0.04\textrm{M}_{\odot}$ if we also ignore the $\nu_{\textrm{max}}$ constraint, and $<0.02\textrm{M}_{\odot}$ including the $\nu_{\textrm{max}}$.
 

### 2018-01-30 paper/aldebaran.tex
wording

@@
 As a final step to reduce residual uncorrected systematics, we apply the \textsc{k2sc} \citep{k2sc} Gaussian Process-based systematics correction code to the initial halo light curve, but the effect of this in the present case for Aldebaran is minimal.
-There is somewhat higher than usual residual noise at multiples of $4 c/d$ (the satellite thruster firing frequency), but this is nevertheless very small in comparison to the signal from Aldebaran, and may be ascribed to the large fraction of the pixel mask occupied by the bleed column from this extremely bright star.
+There is somewhat higher than usual residual noise at harmonics of $4 c/d$ (the satellite thruster firing frequency), but this is nevertheless very small in comparison to the signal from Aldebaran, and may be ascribed to the large fraction of the pixel mask occupied by the bleed column from this extremely bright star.
 

### 2018-01-30 paper/aldebaran.tex
formatting

@@
 It was revived as \ktwo \citep{howell14}, balanced by orienting perpendicular to the Sun.
-This requires that \ktwo observes fields in the Ecliptic in $\sim~80$~d Campaigns, of which Aldebaran was observed in Campaign~13.
+This requires that \ktwo observes fields in the Ecliptic in $\sim80$~d Campaigns, of which Aldebaran was observed in Campaign~13.
 
-The \kepler detector saturates for stars brighter than the $\sim~11^\text{th}$ magnitude.
+The \kepler detector saturates for stars brighter than the $\sim11^\text{th}$ magnitude.
 Nevertheless, the excess flux deposited in a saturated pixel spills conservatively up and down the pixel column, such that it is possible to sum this `bleed column' for bright stars and still obtain precise photometry, such as was done for the brightest star in the nominal \kepler mission, $\theta$~Cyg \citep[$V = 4.48$;][]{guzik2011,thetacygwhite,guzik2016}.
@@
 \noindent and likewise second-order TV the equivalent expression in second-order finite differences.
-The efficacy of this method was recently confirmed by \citet{kallinger2017}, comparing BRITE-constellation observations of Atlas to \ktwo halo photometry and finding excellent agreement in frequency and amplitude of the reported oscillations.
+The efficacy of this method was recently confirmed by \citet{kallinger2017}, comparing BRITE-Constellation observations of Atlas to \ktwo halo photometry and finding excellent agreement in the frequency and amplitude of the reported oscillations.
 
-In an improvement since \citet{White2017}, we use the \textsc{Theano} library \citep{theano} to calculate analytic derivatives for this objective function, which reduces the computational time for a single halo light curve on a commercial laptop from tens of minutes to tens of seconds.
+In an improvement since \citet{White2017}, we use the \textsc{Theano} library \citep{theano} to calculate analytic derivatives for the TV objective function, which reduces the computational time for a single halo light curve on a commercial laptop from tens of minutes to tens of seconds.
 

### 2018-01-30 paper/aldebaran.tex
flashy first sentence

@@
 
-In this paper, we present a re-analysis of these original RV data in which we not only confirm this signal, but detect acoustic oscillations in Aldebaran for the first time.
+In this paper, we present a re-analysis of these original RV data in which we not only confirm the planetary signal, but detect acoustic oscillations in Aldebaran for the first time.
+\citet{hatzes1993} noted night-to-night RV variability: in this planet-hunting noise we find an asteroseismic signal.
 We validate this method and its result with new independent RV observations with the Hertzsprung SONG~Telescope, and photometry from the \ktwo Mission.

### 2018-02-02 paper/aldebaran.tex
Fix some plotting issues and correct Jim's name in author list.
even

@@
 
-\author{James Barrett}
+\author{Jim W. Barrett}
 \affiliation{Birmingham Institute for Gravitational Wave Astronomy,
@@
 This can naturally be extended to \emph{Gaia} in space \citep{gaia}, or \emph{LSST} from the ground \citep{dmt,lsst,lsstbook}, from which many thousands of new asteroseismic determinations will be possible.
-In future RV surveys, it may even be possible to beat the \citet{2005PASJ...57...97S} $\sim 30$~m/s RV precision limit for red giants by taking many closely spaced observations and modelling-out the effects of stellar oscillations: this will allow us to dig deeper into this intrinsic stellar noise to detect less-massive planets around these stars.
+In future RV surveys, it may be possible to beat the \citet{2005PASJ...57...97S} $\sim 30$~m/s RV precision limit for red giants by taking many closely spaced observations and modelling-out the effects of stellar oscillations: this will allow us to dig deeper into this intrinsic stellar noise to detect less-massive planets around these stars.
 

### 2018-02-02 paper/aldebaran.tex
extra abstract clause

@@
 From the frequency of these oscillations we determine the mass of Aldebaran to be $1.16 \pm 0.07$~\msun, and note that this implies its planet will have been subject to insolation comparable to the Earth for some of the star's main sequence lifetime.
-Our approach to sparse, irregularly sampled time series astronomical observations has the potential to unlock similar measurements for hundreds of stars in archival data.
-%, and permits more flexible observing schedules in future.
+Our approach to sparse, irregularly sampled time series astronomical observations has the potential to unlock asteroseismic measurements for hundreds of stars in archival data, and push to lower-mass planets around red giant stars.
 \end{abstract}

### 2018-02-02 paper/aldebaran.tex
wording

@@
 In this paper, we present a re-analysis of these original RV data in which we not only confirm the planetary signal, but detect acoustic oscillations in Aldebaran for the first time.
-\citet{hatzes1993} noted night-to-night RV variability: in this planet-hunting noise we find an asteroseismic signal.
+\citet{hatzes1993} noted night-to-night RV variability: this is the noise floor limiting the sensitivity to sub-$M_J$ planets around giants in RV surveys \citep{2005PASJ...57...97S,2014A&A...566A.113J}.
+We pull out the asteroseismic signal in this planet hunting noise.
 We validate this method and its result with new independent RV observations with the Hertzsprung SONG~Telescope, and photometry from the \ktwo Mission.

### 2018-02-02 paper/aldebaran.tex
Minor changes with a noteable note on RGB only treatment.

--- 2018-02-02 paper/aldebaran.tex
@@
 From the frequency of these oscillations we determine the mass of Aldebaran to be $1.16 \pm 0.07$~\msun, and note that this implies its planet will have been subject to insolation comparable to the Earth for some of the star's main sequence lifetime.
-Our approach to sparse, irregularly sampled time series astronomical observations has the potential to unlock similar measurements for hundreds of stars in archival data.
-%, and permits more flexible observing schedules in future.
+Our approach to sparse, irregularly sampled time series astronomical observations has the potential to unlock asteroseismic measurements for thousands of stars in archival data, and push to lower-mass planets around red giant stars.
 \end{abstract}
@@
 We validate this method and its result with new independent RV observations with the Hertzsprung SONG~Telescope, and photometry from the \ktwo Mission.
-By measuring the frequency of maximum power of these \emph{p}-mode oscillations, \numax, we asteroseismically determine the mass of Aldebaran to be $1.16 \pm 0.07$~\msun.
+By measuring the frequency of maximum power of these \emph{p}-mode oscillations, \numax, we can determine the mass of Aldebaran to be $1.16 \pm 0.07$~\msun.
 This precise stellar mass allows us to calculate that Aldebaran~b and any satellites it may have, although they are now likely to be very hot, would have had equilibrium temperatures comparable to that of the Earth when Aldebaran was on the main sequence.
@@
 These scaling relations are understood to be approximate but never the less are commonly used for radius $R$, mass $M$, and surface gravity $g$:
-These parameters can be used to estimate the radius $R$, mass $M$, and surface gravity $g$ of stars when combined with an estimate of the effective temperature \teff\ through so-called scaling relations:
+These parameters can be used to estimate the radius $R$, mass $M$, and surface gravity $g$ of stars when combined with an estimate of the effective temperature \teff\ through scaling relations:
 
@@
 
-While the accuracy of these scaling relations is still a matter of ongoing work \citep[e.g.,][]{2017ApJ...844..102H}, stellar parameters can be estimated by comparing observables to parameter from models of stellar evolution.
+While the accuracy of these scaling relations is still a matter of ongoing work \citep[e.g.,][]{2017ApJ...844..102H}, stellar parameters can be estimated by comparing observables to parameters from models of stellar evolution.
 
@@
 2.5$\,\rm m\,s^{-1}$ allowing us to easily detect the oscillations which
-displays a peak-to-peak amplitude of $\approx \, 170 \, \rm m \, s^{-1}$.
+display a peak-to-peak amplitude of $\approx \, 170 \, \rm m \, s^{-1}$.
 Figure~\ref{songdata} displays the radial velocities obtained -- oscillations
@@
 Figure \ref{k2_lightcurve} shows the light curve and power spectrum of the K2 observations of Aldebaran.
-We detect clear variability on a average timescale of $\sim$\,5\,days, consistent with the expected timescale for solar-like oscillations for a high-luminosity red giant.
+We detect clear variability on an average timescale of $\sim$\,5\,days, consistent with the expected timescale for solar-like oscillations for a high-luminosity red giant.
 To measure global asteroseismic parameters we model the background variability in the Fourier domain using the methodology described in \citet{huber09}, yielding a frequency of maximum power of $\numax = 2.2 \pm 0.25\,\muHz$ and amplitude per radial mode of $1850 \pm 500$\,ppm.
@@
 additional oscillatory modes, so we conclude that the large spacing $\Delta\nu$
-is not observable in this data set.  We have determined the stellar properties
+is not observable in this data set.
+
+We have determined the stellar properties
 of Aldebaran with the additional constraint of \numax determined from this
@@
 scheme.
+\item We consider only the red giant branch phase to avoid the treatment of poorly understood mass loss at the tip of the red giant branch.
 \item No rotational mixing or diffusion is included.

### 2018-02-02 paper/aldebaran.tex

@@
 In this paper, we present a re-analysis of these original RV data in which we not only confirm the planetary signal, but detect acoustic oscillations in Aldebaran for the first time.
-\citet{hatzes1993} noted night-to-night RV variability: in this planet-hunting noise we find an asteroseismic signal.
+\citet{hatzes1993} noted night-to-night RV variability: this is the noise floor limiting the sensitivity to sub-$M_J$ planets around giants in RV surveys \citep{2005PASJ...57...97S,2014A&A...566A.113J}.
+We pull out the asteroseismic signal in this planet hunting noise.
 We validate this method and its result with new independent RV observations with the Hertzsprung SONG~Telescope, and photometry from the \ktwo Mission.
-By measuring the frequency of maximum power of these \emph{p}-mode oscillations, \numax, we can determine the mass of Aldebaran to be $1.16 \pm 0.07$~\msun.
+By measuring the frequency of maximum power of these \emph{p}-mode oscillations, \numax, we asteroseismically determine the mass of Aldebaran to be $1.16 \pm 0.07$~\msun.
 This precise stellar mass allows us to calculate that Aldebaran~b and any satellites it may have, although they are now likely to be very hot, would have had equilibrium temperatures comparable to that of the Earth when Aldebaran was on the main sequence.

### 2018-02-05 paper/aldebaran.tex
title
planet not star
Fixed typos

@@
 
-\title{Detection of Oscillations in Aldebaran with Ground-Based Observations}
+\title{Aldebaran~b's temperate past uncovered in planet search data}
 
@@
 The ability to measure \numax and \Dnu depends on the length and sampling rate of a data set.
-In practise, for the higher luminosity red giants it is more straight forward to measure \numax than \Dnu.
+In practice, for the higher luminosity red giants it is more straightforward to measure \numax than \Dnu.
 Typical values for \numax\ range from $0.1-20 \rm\,\mu Hz$ for luminous giants and $20-50 \rm\,\mu Hz$ for stars near the red clump.
@@
 Radial velocity variations intrinsic to the star, such as those caused by stellar oscillations, have previously limited the precision with which red giant planets have been studied \citep{2005PASJ...57...97S}.
-We aim to model out these noise processes and use them recover asteroseismic information, as well as to improve our estimates of the orbital parameters of the planet.
+We aim to model out these noise processes and use them to recover asteroseismic information, as well as to improve our estimates of the orbital parameters of the planet.
 It is easy to observe Aldebaran and similarly-bright stars with ground-based spectroscopic instruments, typically only requiring short exposures that can be obtained even under adverse observing conditions.
 There is indeed a considerable archive of such observations already, as a legacy of radial velocity (RV) surveys conducted to find exoplanets.
-In most cases, however, these have not so far been useful for asteroseismology, because these RV data are sparsely and irregularly sampled.
+In most cases, however, these have not so far been useful for asteroseismology because these RV data are sparsely and irregularly sampled.
 Because we have to pause observations during the day, during poor weather conditions, or simply when targets of higher priority are being observed, we get time series which may have significant and uneven gaps.
@@
 oscillators such as we encounter in asteroseismology can be statistically
-modelled as a Gaussian Process, with a covariance kernel consisting of a sum of
+modelled as a Gaussian Process (GP), with a covariance kernel consisting of a sum of
 damped sinusoids. The hyperparameters of this process encode features of the
@@
 densities.  Our treatment of CARMA models is described more fully in Appendix
-\ref{carmatheory}.   \citet{Kelly2014} also gives more details on CARMA
-processes in an astronomical context, and shows how a state-space model of the
+\ref{carmatheory}.   \citet{Kelly2014} also give more details on CARMA
+processes in an astronomical context, and show how a state-space model of the
 process $y(t)$ can be tracked through a time-series of uncertain observations
@@
 
-From this pilot study with Aldebaran, have shown that with limited quantities of data of limited quality, we can nevertheless do asteroseismology from the ground and obtain very precise estimates of stellar parameters.
+From this pilot study with Aldebaran, we have shown that with limited quantities of data of limited quality, we can nevertheless do asteroseismology from the ground and obtain very precise estimates of stellar parameters.
 Furthermore, improvements in the algorithm may also permit the detection of not only~\numax but also~\Dnu from similar measurements, permitting more sophisticated asteroseismic analysis.
@@
 There are two main reasons why this is not possible for all bright stars in general.
-First, because the on-board data storage and downlink bandwidth from \kepler are limited, it is often not desirable to download the large number of pixels that are required for such bright stars.
+First, because the on-board data storage and downlink bandwidth from \kepler are limited, it is often not desirable to store and download the large number of pixels that are required for such bright stars.
 Second, if the bleed column for a sufficiently bright star reaches the edge of the chip, flux spills over and is not conserved, imposing a hard brightness limit that depends on the distance to the detector edge.
@@
 As a final step to reduce residual uncorrected systematics, we apply the \textsc{k2sc} \citep{k2sc} Gaussian Process-based systematics correction code to the initial halo light curve, but the effect of this in the present case for Aldebaran is minimal.
-There is somewhat higher than usual residual noise at harmonics of $4 c/d$ (the satellite thruster firing frequency), but this is nevertheless very small in comparison to the signal from Aldebaran, and may be ascribed to the large fraction of the pixel mask occupied by the bleed column from this extremely bright star.
+There is somewhat higher than usual residual noise at harmonics of $46\,\mu\mathrm{Hz}$ ($4\,\mathrm{d}^{-1}$, the satellite thruster firing frequency), but this is nevertheless very small in comparison to the signal from Aldebaran, and may be ascribed to the large fraction of the pixel mask occupied by the bleed column from this extremely bright star.
 
@@
 \log_{10} \frac{L}{L_{\odot}} = 4.0+
-0.4 M_{{\textrm{bol}},\odot} -2.0 \log_{10} {\pi [{\textrm{mas}]}} \\-0.4(V-A_V + BC(V)).
+0.4 M_{{\textrm{bol}},\odot} -2.0 \log_{10} {\pi [{\textrm{mas}]}} -0.4(V-A_V + BC(V)).
 \label{eqn:lum}

### 2018-02-05 paper/aldebaran.tex
Fixed uncertainty significant figures

@@
 %% Q
-\newcommand{\HatzesQRange}{\twosidedrange{2.12}{0.69}{0.56}}
-\newcommand{\SONGQRange}{\twosidedrange{3.06}{1.20}{0.81}}
-\newcommand{\CombinedQRange}{\twosidedrange{2.83}{0.60}{0.51}}
+\newcommand{\HatzesQRange}{\twosidedrange{2.1}{0.7}{0.6}}
+\newcommand{\SONGQRange}{\twosidedrange{3.1}{1.2}{0.8}}
+\newcommand{\CombinedQRange}{\twosidedrange{2.8}{0.6}{0.5}}
 
@@
 \newcommand{\SONGNuMaxRange}{\twosidedrange{2.33}{0.13}{0.14}}
-\newcommand{\combinedNuMaxRange}{\twosidedrange{2.238}{0.086}{0.080}}
+\newcommand{\combinedNuMaxRange}{\twosidedrange{2.24}{0.09}{0.08}}
 
@@
 %% e
-\newcommand{\EccentricityRange}{\twosidedrange{0.168}{0.069}{0.071}}
+\newcommand{\EccentricityRange}{0.17$\pm$0.07} %\twosidedrange{0.168}{0.069}{0.071}}
 
@@
 While the hot Jupiter 51~Peg~b \citep{51peg} was the first exoplanet to be recognized as such, before this, \citet{hatzes1993} had noted RV variations in Pollux \citep[$\beta$~Gem; subsequently confirmed as a planet:][]{betgemconf,betgemconf2}, Arcturus ($\alpha$~Boo, unconfirmed), and Aldebaran.
-After further investigation by \citet{Hatzes1998}, \citet{Hatzes2015} have now claimed a firm RV detection of a planetary-mass companion Aldebaran~b, with a period of $628.96 \pm 0.90$~d.
+After further investigation by \citet{Hatzes1998}, \citet{Hatzes2015} have now claimed a firm RV detection of a planetary-mass companion Aldebaran~b, with a period of $629.96 \pm 0.90$~d.
 
@@
 The stellar oscillations are best fit with a single, low-quality ($Q =
-\CombinedQRange{}$) oscillatory mode with $\numax = 2.238^{+ 0.086}_{- 0.080}$
+\CombinedQRange{}$) oscillatory mode with $\numax = 2.24^{+ 0.09}_{- 0.08}$
 \muHz.  We find no improvement in the marginal likelihood (evidence) or other
@@
 \numax we find estimates of the stellar properties as $M = 1.27^{+0.24}_{-0.2}
-\, \mathrm{M_{\odot}}$ and age $4.86^{+3.56}_{-2.04} \, \rm Gyr$.  With the
+\, \mathrm{M_{\odot}}$ and age $4.9^{+3.6}_{-2.0} \, \rm Gyr$.  With the
 inclusion of \numax we find $M = 1.16^{+0.07}_{-0.07} \, \mathrm{M_{\odot}}$ and
-age $6.38^{+1.42}_{-1.12} \, \rm Gyr$.  It is clear that the addition of \numax
+age $6.4^{+1.4}_{-1.1} \, \rm Gyr$.  It is clear that the addition of \numax
 provides substantially more precise estimates of mass and age. With this stellar
 mass, the radial velocity signal translates to an inclination-dependent
-planetary mass of $M\,\sin\,{i} = 5.81 \pm 0.72~M_J$
+planetary mass of $M\,\sin\,{i} = 5.8 \pm 0.7~M_J$
 \citep[Equation~1]{Torres2008}, a massive giant planet; for sufficiently high
@@
 		\hline
-		\cite{2012Sheffield}$_{a}$ &$1.17^{+0.07}_{-0.07}$&$43.9^{+0.9}_{-0.9}$&$6.47^{+1.39}_{-1.1}$\\
-		\cite{2012Sheffield}$_{b}$&$1.17^{+0.07}_{-0.07}$&$43.8^{+0.8}_{-0.9}$&$6.35^{+1.4}_{-1.11}$\\
-		\cite{2011Prugniel}&$1.17^{+0.07}_{-0.07}$&$43.9^{+0.9}_{-0.9}$&$6.23^{+1.4}_{-1.11}$\\
-		\cite{2008Massarotti}&$1.13^{+0.07}_{-0.07}$&$43.5^{+0.9}_{-0.9}$&$5.95^{+1.37}_{-1.07}$\\
-		\cite{2009Frasca}&$1.02^{+0.04}_{-0.04}$&$44.0^{+0.8}_{-0.8}$&$10.26^{+1.54}_{-1.45}$\\
+		\cite{2012Sheffield}$_{a}$ &$1.17^{+0.07}_{-0.07}$&$43.9^{+0.9}_{-0.9}$&$6.5^{+1.4}_{-1.1}$\\
+		\cite{2012Sheffield}$_{b}$&$1.17^{+0.07}_{-0.07}$&$43.8^{+0.8}_{-0.9}$&$6.4^{+1.4}_{-1.1}$\\
+		\cite{2011Prugniel}&$1.17^{+0.07}_{-0.07}$&$43.9^{+0.9}_{-0.9}$&$6.2^{+1.4}_{-1.1}$\\
+		\cite{2008Massarotti}&$1.13^{+0.07}_{-0.07}$&$43.5^{+0.9}_{-0.9}$&$6.0^{+1.4}_{-1.1}$\\
+		\cite{2009Frasca}&$1.02^{+0.04}_{-0.04}$&$44.0^{+0.8}_{-0.8}$&$10.3^{+1.5}_{-1.4}$\\
 		\hline
@@
 		\hline
-\cite{2012Sheffield}$_{a}$&$1.43^{+0.26}_{-0.24}$&$43.8^{+0.9}_{-0.9}$&$3.53^{+2.72}_{-1.44}$\\
-\cite{2012Sheffield}$_{b}$&$1.27^{+0.24}_{-0.2}$&$43.8^{+0.9}_{-0.9}$&$4.86^{+3.56}_{-2.04}$\\
-\cite{2011Prugniel}&$1.25^{+0.22}_{-0.19}$&$43.8^{+0.9}_{-0.9}$&$5.02^{+3.47}_{-2.02}$\\
-\cite{2008Massarotti}&$0.95^{+0.11}_{-0.05}$&$43.8^{+0.9}_{-0.9}$&$10.24^{+2.36}_{-3.08}$\\
-\cite{2009Frasca}&$0.96^{+0.04}_{-0.04}$&$44.4^{+0.8}_{-0.8}$&$11.64^{+1.43}_{-1.8}$\\
+\cite{2012Sheffield}$_{a}$&$1.43^{+0.26}_{-0.24}$&$43.8^{+0.9}_{-0.9}$&$3.5^{+2.7}_{-1.4}$\\
+\cite{2012Sheffield}$_{b}$&$1.27^{+0.24}_{-0.2}$&$43.8^{+0.9}_{-0.9}$&$4.9^{+3.6}_{-2.0}$\\
+\cite{2011Prugniel}&$1.25^{+0.22}_{-0.19}$&$43.8^{+0.9}_{-0.9}$&$5.0^{+3.5}_{-2.0}$\\
+\cite{2008Massarotti}&$0.95^{+0.11}_{-0.05}$&$43.8^{+0.9}_{-0.9}$&$10.2^{+2.4}_{-3.1}$\\
+\cite{2009Frasca}&$0.96^{+0.04}_{-0.04}$&$44.4^{+0.8}_{-0.8}$&$11.6^{+1.4}_{-1.8}$\\
 		\hline
@@
 		\hline
-\cite{2012Sheffield}$_{a}$&$1.17^{+0.07}_{-0.07}$&$43.9^{+0.9}_{-0.9}$&$6.51^{+1.41}_{-1.11}$\\
-\cite{2012Sheffield}$_{b}$&$1.16^{+0.07}_{-0.07}$&$43.8^{+0.8}_{-0.9}$&$6.38^{+1.42}_{-1.12}$\\
-\cite{2011Prugniel}&$1.16^{+0.07}_{-0.07}$&$43.9^{+0.9}_{-0.9}$&$6.43^{+1.48}_{-1.15}$\\
-\cite{2008Massarotti}&$1.13^{+0.07}_{-0.07}$&$43.5^{+0.9}_{-0.9}$&$5.86^{+1.35}_{-1.05}$\\
-\cite{2009Frasca}&$1.15^{+0.07}_{-0.07}$&$43.9^{+0.9}_{-0.8}$&$6.47^{+1.51}_{-1.17}$\\
+\cite{2012Sheffield}$_{a}$&$1.17^{+0.07}_{-0.07}$&$43.9^{+0.9}_{-0.9}$&$6.5^{+1.4}_{-1.1}$\\
+\cite{2012Sheffield}$_{b}$&$1.16^{+0.07}_{-0.07}$&$43.8^{+0.8}_{-0.9}$&$6.4^{+1.4}_{-1.1}$\\
+\cite{2011Prugniel}&$1.16^{+0.07}_{-0.07}$&$43.9^{+0.9}_{-0.9}$&$6.4^{+1.5}_{-1.2}$\\
+\cite{2008Massarotti}&$1.13^{+0.07}_{-0.07}$&$43.5^{+0.9}_{-0.9}$&$5.9^{+1.4}_{-1.0}$\\
+\cite{2009Frasca}&$1.15^{+0.07}_{-0.07}$&$43.9^{+0.9}_{-0.8}$&$6.5^{+1.5}_{-1.2}$\\
 		\hline

### 2018-02-05 paper/aldebaran.tex
Remove comment on mass loss treatment
misc

@@
 %% e
-\newcommand{\EccentricityRange}{0.17$\pm$0.07} %\twosidedrange{0.168}{0.069}{0.071}}
+\newcommand{\EccentricityRange}{0.17\pm0.07} %\twosidedrange{0.168}{0.069}{0.071}}
 
@@
 more-flexible and correlated model for the stellar component of the RV signal:
-period $P = \PeriodRange \, \mathrm{d}$, eccentricity $e =
+period $P = \PeriodRange{} \, \mathrm{d}$,  eccentricity $e =
 \EccentricityRange{}$; and radial velocity semi-amplitude $K =
-\PlanetRVAmpRange{} \, \mathrm{m} \, \mathrm{s}^{-1}$.
+\PlanetRVAmpRange{} \, \mathrm{m}\, \mathrm{s}^{-1}$.
 

### 2018-02-05 paper/aldebaran.tex
gaussian process -> GP

@@
 \begin{abstract}
-The nearby red giant Aldebaran is known to host a gas giant planetary companion from decades of ground-based spectroscopic measurements of its Doppler shift.
+The nearby red giant Aldebaran is known to host a gas giant planetary companion from decades of ground-based spectroscopic radial velocity measurements.
 Using Gaussian Process-based Continuous Auto-Regressive Moving Average (CARMA) models, we show that these historic data also contain evidence of acoustic oscillations in the star itself, and verify this result with further dedicated ground-based spectroscopy and space-based photometry with the \kepler Space Telescope.
@@
 
-Aldebaran ($\alpha$~Tauri) is a well-known first-magnitude naked-eye red giant star, and has long been the subject of astronomical investigations.
+Aldebaran ($\alpha$~Tauri) is a well-known first-magnitude red giant star, and has long been the subject of astronomical investigations.
 It was one of the first stars around which an extrasolar planet candidate was identified, by looking for Doppler shifts from the star's reflex motion around the common centre of mass with its companion \citep[the radial velocity or RV method;][]{struverv}.
@@
 power spectral density distributions and are insensitive to the window function.
-Gaussian processes with quasi-periodic kernels have previously been used to re-analyse RV data for main sequence stars \citep{2014MNRAS.443.2517H,2015MNRAS.452.2269R}, where the noise is from stellar activity, but have not previously been applied to red giant asteroseismology.
+GPs with quasi-periodic kernels have previously been used to re-analyse RV data for main sequence stars \citep{2014MNRAS.443.2517H,2015MNRAS.452.2269R}, where the noise is from stellar activity, but have not previously been applied to red giant asteroseismology.
 Unfortunately, these methods are impractical for long time series, as the computational
-cost of evaluating the standard Gaussian Process likelihood function scales as
+cost of evaluating the standard GP likelihood function scales as
 $\mathcal{O}(n^3)$.
@@
 time.  Additionally, \citet{Foreman-Mackey2017} show how a dimensional expansion
-can be used to reduce the standard Gaussian Process covariance matrix for this
+can be used to reduce the standard GP covariance matrix for this
 same process,  with autocorrelation given by Eq.\ \eqref{eq:CARMA-ACF} under
 observations at times $t_k$, to a banded form that can then be
-Cholesky-decomposed in the standard Gaussian Process likelihood function in
+Cholesky-decomposed in the standard GP likelihood function in
 $\mathcal{O}\left( n \right)$ time.
@@
 of the ODE; $b_j \in \mathbb{C}$, $\Re b_j < 0$ are corresponding roots of the
-inverse process; and $\eta(t)$ is a white-noise Gaussian Process with
+inverse process; and $\eta(t)$ is a white-noise GP with
 $\left\langle \eta(t) \right\rangle = 0$ and $\left\langle \eta(t) \eta(t')
@@
 
-As a final step to reduce residual uncorrected systematics, we apply the \textsc{k2sc} \citep{k2sc} Gaussian Process-based systematics correction code to the initial halo light curve, but the effect of this in the present case for Aldebaran is minimal.
+As a final step to reduce residual uncorrected systematics, we apply the \textsc{k2sc} \citep{k2sc} GP-based systematics correction code to the initial halo light curve, but the effect of this in the present case for Aldebaran is minimal.
 There is somewhat higher than usual residual noise at harmonics of $46\,\mu\mathrm{Hz}$ ($4\,\mathrm{d}^{-1}$, the satellite thruster firing frequency), but this is nevertheless very small in comparison to the signal from Aldebaran, and may be ascribed to the large fraction of the pixel mask occupied by the bleed column from this extremely bright star.
@@
 scheme.
-\item We consider only the red giant branch phase to avoid the treatment of poorly understood mass loss at the tip of the red giant branch.
 \item No rotational mixing or diffusion is included.

### 2018-02-05 paper/aldebaran.tex
celerite to appendix

@@
 of the stochastic forcing, $b_j$ in $\mathcal{O}\left( n \right)$ computational
-time.  Additionally, \citet{Foreman-Mackey2017} show how a dimensional expansion
-can be used to reduce the standard GP covariance matrix for this
-same process,  with autocorrelation given by Eq.\ \eqref{eq:CARMA-ACF} under
-observations at times $t_k$, to a banded form that can then be
-Cholesky-decomposed in the standard GP likelihood function in
-$\mathcal{O}\left( n \right)$ time.
+time.
 
@@
 
+The Kalman filter implementation of the CARMA likelihood is not unique. 
+\citet{Foreman-Mackey2017} show how a dimensional expansion
+can be used to reduce the standard GP covariance matrix for this
+same process,  with autocorrelation given by Eq.\ \eqref{eq:CARMA-ACF} under
+observations at times $t_k$, to a banded form that can then be
+Cholesky-decomposed in the standard GP likelihood function in
+$\mathcal{O}\left( n \right)$ time. These `\emph{celerite}' models are formally equivalent to the Kalman CARMA model and differ only in their computational implementation.
+
 \begin{deluxetable}{lll}

### 2018-02-05 paper/aldebaran.tex
celerite in main body
clearer language

@@
 time.
+\citet{Foreman-Mackey2017} has demonstrated that the same model can also be implemented via a novel matrix factorization, and demonstrated its asteroseismic potential modelling the light curve of KIC~11615890.
 
@@
 by either the $r_i$ and $b_j$ or the RMS amplitudes $A_i$ of each eigenmode of
-the ODE and the corresponding roots $r_i$.  We also include a Keplerian RV
-signal from a planetary companion to Aldebaran \citep{Hatzes2015,Gregory2005}.
+the ODE and the corresponding roots $r_i$. This likelihood is computed for the 
+residuals of a deterministic Keplerian RV mean model for the planetary companion.
 Because we have a set of observations taken with different instruments at

### 2018-02-05 paper/aldebaran.tex
uncertainties; phrasing

@@
 We find evidence for a low-quality ($Q = \HatzesQRange{}$) oscillatory mode with
-$\nu_\mathrm{max} = \HatzesNuMaxRange{} \, \mu\mathrm{Hz}$.  This is consistent
-with the presence of a number of un-resolved asteroseismic modes in the RV data.
+$\nu_\mathrm{max} = \HatzesNuMaxRange{} \, \mu\mathrm{Hz}$. 
+This is consistent with the presence of a number of un-resolved asteroseismic modes in the RV data. 
+The low Q factor in this case refers to the inverse fractional width of the band over which the modes are excited, not the quality of any individual mode.
 Our posterior for \numax{} from this data set is shown in Figure
@@
 To measure global asteroseismic parameters we model the background variability in the Fourier domain using the methodology described in \citet{huber09}, yielding a frequency of maximum power of $\numax = 2.2 \pm 0.25\,\muHz$ and amplitude per radial mode of $1850 \pm 500$\,ppm.
-Due to the limited frequency resolution of the 70-day K2 timeseries we were unable to measure the large frequency separation \Dnu, which is expected to be on the order of $\approx$\,0.5\muHz.
+Due to the limited frequency resolution of the 70-day K2 time series we were unable to measure the large frequency separation \Dnu, which is expected to be on the order of $\approx$\,0.5\muHz.
 
@@
 being `middle of the road' estimates.  For the analysis without constraint on
-\numax we find estimates of the stellar properties as $M = 1.27^{+0.24}_{-0.2}
+\numax we find estimates of the stellar properties as $M = 1.27^{+0.24}_{-0.20}
 \, \mathrm{M_{\odot}}$ and age $4.9^{+3.6}_{-2.0} \, \rm Gyr$.  With the
-inclusion of \numax we find $M = 1.16^{+0.07}_{-0.07} \, \mathrm{M_{\odot}}$ and
+inclusion of \numax we find $M = 1.16\, \pm \, {0.07} \, \mathrm{M_{\odot}}$ and
 age $6.4^{+1.4}_{-1.1} \, \rm Gyr$.  It is clear that the addition of \numax
@@
 Furthermore, improvements in the algorithm may also permit the detection of not only~\numax but also~\Dnu from similar measurements, permitting more sophisticated asteroseismic analysis.
-Sufficient radial velocity data either already exist, or are trivial to obtain, in order to do this for essentially all bright giants; while for solar-like stars, these new methods will allow significantly relaxed observing requirements.
+Sufficient radial velocity data either already exist, or are trivial to obtain, in order to do this for essentially all bright giants; while for asteroseismology of solar-like stars, these new methods will allow significantly relaxed observing requirements.
 One rich archive is the series of observations with the Hamilton \'{E}chelle Spectrograph at Lick Observatory, studying 373 G~and K~giants \citep[e.g.][]{frink2001,hekker2006,ortiz2016}.
@@
 \enddata
-\tablenotetext{a}{$\left\langle y_j \right\rangle$ and $\sigma_{y_j}$ are the mean and standard deviation of the RV timeseries from site and instrument combination $j$.}
+\tablenotetext{a}{$\left\langle y_j \right\rangle$ and $\sigma_{y_j}$ are the mean and standard deviation of the RV time series from site and instrument combination $j$.}
 \tablenotetext{b}{We actually sample in and impose a (flat) prior on $x \equiv e \cos \omega$ and $y \equiv e \sin \omega$.}

### 2018-02-05 paper/aldebaran.tex
Comments from Frank.

--- 2018-02-05 paper/aldebaran.tex
@@
 %% Q
-\newcommand{\HatzesQRange}{\twosidedrange{2.1}{0.7}{0.6}}
-\newcommand{\SONGQRange}{\twosidedrange{3.1}{1.2}{0.8}}
-\newcommand{\CombinedQRange}{\twosidedrange{2.8}{0.6}{0.5}}
+\newcommand{\HatzesQRange}{\twosidedrange{2.12}{0.69}{0.56}}
+\newcommand{\SONGQRange}{\twosidedrange{3.06}{1.20}{0.81}}
+\newcommand{\CombinedQRange}{\twosidedrange{2.83}{0.60}{0.51}}
 
@@
 \newcommand{\SONGNuMaxRange}{\twosidedrange{2.33}{0.13}{0.14}}
-\newcommand{\combinedNuMaxRange}{\twosidedrange{2.24}{0.09}{0.08}}
+\newcommand{\combinedNuMaxRange}{\twosidedrange{2.238}{0.086}{0.080}}
 
@@
 %% e
-\newcommand{\EccentricityRange}{0.17\pm0.07} %\twosidedrange{0.168}{0.069}{0.071}}
+\newcommand{\EccentricityRange}{\twosidedrange{0.168}{0.069}{0.071}}
 
@@
 
-\title{Aldebaran~b's temperate past uncovered in planet search data}
+\title{Detection of Oscillations in Aldebaran with Ground-Based Observations}
 
@@
 \begin{abstract}
-The nearby red giant Aldebaran is known to host a gas giant planetary companion from decades of ground-based spectroscopic radial velocity measurements.
+The nearby red giant Aldebaran is known to host a gas giant planetary companion from decades of ground-based spectroscopic measurements of its Doppler shift.
 Using Gaussian Process-based Continuous Auto-Regressive Moving Average (CARMA) models, we show that these historic data also contain evidence of acoustic oscillations in the star itself, and verify this result with further dedicated ground-based spectroscopy and space-based photometry with the \kepler Space Telescope.
 From the frequency of these oscillations we determine the mass of Aldebaran to be $1.16 \pm 0.07$~\msun, and note that this implies its planet will have been subject to insolation comparable to the Earth for some of the star's main sequence lifetime.
-Our approach to sparse, irregularly sampled time series astronomical observations has the potential to unlock asteroseismic measurements for thousands of stars in archival data, and push to lower-mass planets around red giant stars.
+Our approach to sparse, irregularly sampled time series astronomical observations has the potential to unlock similar measurements for hundreds of stars in archival data.
+%, and permits more flexible observing schedules in future.
 \end{abstract}
@@
 
-Aldebaran ($\alpha$~Tauri) is a well-known first-magnitude red giant star, and has long been the subject of astronomical investigations.
+Aldebaran ($\alpha$~Tauri) is a well-known first-magnitude naked-eye red giant star, and has long been the subject of astronomical investigations.
 It was one of the first stars around which an extrasolar planet candidate was identified, by looking for Doppler shifts from the star's reflex motion around the common centre of mass with its companion \citep[the radial velocity or RV method;][]{struverv}.
 While the hot Jupiter 51~Peg~b \citep{51peg} was the first exoplanet to be recognized as such, before this, \citet{hatzes1993} had noted RV variations in Pollux \citep[$\beta$~Gem; subsequently confirmed as a planet:][]{betgemconf,betgemconf2}, Arcturus ($\alpha$~Boo, unconfirmed), and Aldebaran.
-After further investigation by \citet{Hatzes1998}, \citet{Hatzes2015} have now claimed a firm RV detection of a planetary-mass companion Aldebaran~b, with a period of $629.96 \pm 0.90$~d.
+After further investigation by \citet{Hatzes1998}, \citet{Hatzes2015} have now claimed a firm RV detection of a planetary-mass companion Aldebaran~b, with a period of $628.96 \pm 0.90$~d.
 
 In this paper, we present a re-analysis of these original RV data in which we not only confirm the planetary signal, but detect acoustic oscillations in Aldebaran for the first time.
-\citet{hatzes1993} noted night-to-night RV variability: this is the noise floor limiting the sensitivity to sub-$M_J$ planets around giants in RV surveys \citep{2005PASJ...57...97S,2014A&A...566A.113J}.
-We pull out the asteroseismic signal in this planet hunting noise.
+\citet{hatzes1993} noted night-to-night RV variability: in this planet-hunting noise we find an asteroseismic signal.
 We validate this method and its result with new independent RV observations with the Hertzsprung SONG~Telescope, and photometry from the \ktwo Mission.
@@
 These scaling relations are understood to be approximate but never the less are commonly used for radius $R$, mass $M$, and surface gravity $g$:
-These parameters can be used to estimate the radius $R$, mass $M$, and surface gravity $g$ of stars when combined with an estimate of the effective temperature \teff\ through scaling relations:
+These parameters can be used to estimate the radius $R$, mass $M$, and surface gravity $g$ of stars when combined with an estimate of the effective temperature \teff\ through so-called scaling relations:
 
@@
 
-While the accuracy of these scaling relations is still a matter of ongoing work \citep[e.g.,][]{2017ApJ...844..102H}, stellar parameters can be estimated by comparing observables to parameters from models of stellar evolution.
+While the accuracy of these scaling relations is still a matter of ongoing work \citep[e.g.,][]{2017ApJ...844..102H}, stellar parameters can be estimated by comparing observables to parameter from models of stellar evolution.
 
@@
 The ability to measure \numax and \Dnu depends on the length and sampling rate of a data set.
-In practice, for the higher luminosity red giants it is more straightforward to measure \numax than \Dnu.
+In practise, for the higher luminosity red giants it is more straight forward to measure \numax than \Dnu.
 Typical values for \numax\ range from $0.1-20 \rm\,\mu Hz$ for luminous giants and $20-50 \rm\,\mu Hz$ for stars near the red clump.
@@
 Radial velocity variations intrinsic to the star, such as those caused by stellar oscillations, have previously limited the precision with which red giant planets have been studied \citep{2005PASJ...57...97S}.
-We aim to model out these noise processes and use them to recover asteroseismic information, as well as to improve our estimates of the orbital parameters of the planet.
+We aim to model out these noise processes and use them recover asteroseismic information, as well as to improve our estimates of the orbital parameters of the planet.
 It is easy to observe Aldebaran and similarly-bright stars with ground-based spectroscopic instruments, typically only requiring short exposures that can be obtained even under adverse observing conditions.
 There is indeed a considerable archive of such observations already, as a legacy of radial velocity (RV) surveys conducted to find exoplanets.
-In most cases, however, these have not so far been useful for asteroseismology because these RV data are sparsely and irregularly sampled.
+In most cases, however, these have not so far been useful for asteroseismology, because these RV data are sparsely and irregularly sampled.
 Because we have to pause observations during the day, during poor weather conditions, or simply when targets of higher priority are being observed, we get time series which may have significant and uneven gaps.
@@
 oscillators such as we encounter in asteroseismology can be statistically
-modelled as a Gaussian Process (GP), with a covariance kernel consisting of a sum of
+modelled as a Gaussian Process, with a covariance kernel consisting of a sum of
 damped sinusoids. The hyperparameters of this process encode features of the
 power spectral density distributions and are insensitive to the window function.
-GPs with quasi-periodic kernels have previously been used to re-analyse RV data for main sequence stars \citep{2014MNRAS.443.2517H,2015MNRAS.452.2269R}, where the noise is from stellar activity, but have not previously been applied to red giant asteroseismology.
+Gaussian processes with quasi-periodic kernels have previously been used to re-analyse RV data for main sequence stars \citep{2014MNRAS.443.2517H,2015MNRAS.452.2269R}, where the noise is from stellar activity, but have not previously been applied to red giant asteroseismology.
 Unfortunately, these methods are impractical for long time series, as the computational
-cost of evaluating the standard GP likelihood function scales as
+cost of evaluating the standard Gaussian Process likelihood function scales as
 $\mathcal{O}(n^3)$.
@@
 densities.  Our treatment of CARMA models is described more fully in Appendix
-\ref{carmatheory}.   \citet{Kelly2014} also give more details on CARMA
-processes in an astronomical context, and show how a state-space model of the
+\ref{carmatheory}.   \citet{Kelly2014} also gives more details on CARMA
+processes in an astronomical context, and shows how a state-space model of the
 process $y(t)$ can be tracked through a time-series of uncertain observations
@@
 of the stochastic forcing, $b_j$ in $\mathcal{O}\left( n \right)$ computational
-time.
+time.  Additionally, \citet{Foreman-Mackey2017} show how a dimensional expansion
+can be used to reduce the standard Gaussian Process covariance matrix for this
+same process,  with autocorrelation given by Eq.\ \eqref{eq:CARMA-ACF} under
+observations at times $t_k$, to a banded form that can then be
+Cholesky-decomposed in the standard Gaussian Process likelihood function in
+$\mathcal{O}\left( n \right)$ time.
 
@@
 2.5$\,\rm m\,s^{-1}$ allowing us to easily detect the oscillations which
-display a peak-to-peak amplitude of $\approx \, 170 \, \rm m \, s^{-1}$.
+displays a peak-to-peak amplitude of $\approx \, 170 \, \rm m \, s^{-1}$.
 Figure~\ref{songdata} displays the radial velocities obtained -- oscillations
@@
 Figure \ref{k2_lightcurve} shows the light curve and power spectrum of the K2 observations of Aldebaran.
-We detect clear variability on an average timescale of $\sim$\,5\,days, consistent with the expected timescale for solar-like oscillations for a high-luminosity red giant.
+We detect clear variability on a average timescale of $\sim$\,5\,days, consistent with the expected timescale for solar-like oscillations for a high-luminosity red giant.
 To measure global asteroseismic parameters we model the background variability in the Fourier domain using the methodology described in \citet{huber09}, yielding a frequency of maximum power of $\numax = 2.2 \pm 0.25\,\muHz$ and amplitude per radial mode of $1850 \pm 500$\,ppm.
@@
 more-flexible and correlated model for the stellar component of the RV signal:
-period $P = \PeriodRange{} \, \mathrm{d}$,  eccentricity $e =
+period $P = \PeriodRange \, \mathrm{d}$, eccentricity $e =
 \EccentricityRange{}$; and radial velocity semi-amplitude $K =
-\PlanetRVAmpRange{} \, \mathrm{m}\, \mathrm{s}^{-1}$.
+\PlanetRVAmpRange{} \, \mathrm{m} \, \mathrm{s}^{-1}$.
 
 The stellar oscillations are best fit with a single, low-quality ($Q =
-\CombinedQRange{}$) oscillatory mode with $\numax = 2.24^{+ 0.09}_{- 0.08}$
+\CombinedQRange{}$) oscillatory mode with $\numax = 2.238^{+ 0.086}_{- 0.080}$
 \muHz.  We find no improvement in the marginal likelihood (evidence) or other
@@
 additional oscillatory modes, so we conclude that the large spacing $\Delta\nu$
-is not observable in this data set.
-
-We have determined the stellar properties
+is not observable in this data set.  We have determined the stellar properties
 of Aldebaran with the additional constraint of \numax determined from this
@@
 \numax we find estimates of the stellar properties as $M = 1.27^{+0.24}_{-0.2}
-\, \mathrm{M_{\odot}}$ and age $4.9^{+3.6}_{-2.0} \, \rm Gyr$.  With the
+\, \mathrm{M_{\odot}}$ and age $4.86^{+3.56}_{-2.04} \, \rm Gyr$.  With the
 inclusion of \numax we find $M = 1.16^{+0.07}_{-0.07} \, \mathrm{M_{\odot}}$ and
-age $6.4^{+1.4}_{-1.1} \, \rm Gyr$.  It is clear that the addition of \numax
-provides substantially more precise estimates of mass and age. With this stellar
-mass, the radial velocity signal translates to an inclination-dependent
-planetary mass of $M\,\sin\,{i} = 5.8 \pm 0.7~M_J$
-\citep[Equation~1]{Torres2008}, a massive giant planet; for sufficiently high
-inclinations ($\gtrsim 45^\circ$), the mass of Aldebaran~b may exceed $13~M_J$,
-making it a possible low-mass brown dwarf candidate.
-A recent study by \citet{2018arXiv180105239H} has cast doubt on the validity of the detection of the massive planet $\gamma$~Draconis~b, and by extension similar planets around similar K~giants such as Aldebaran.
-We remain convinced that the $\gamma$~Draconis problem is not an issue in this case, and Aldebaran~b is a bona fide planet, for reasons outlined in Appendix~\ref{gamma_dra}.
+age $6.38^{+1.42}_{-1.12} \, \rm Gyr$.  Our measurement of \numax provides
+substantially more precise estimates of mass and age, and the uncertainties are
+now dominated by the uncertainty in our measurement. With this stellar mass, the
+radial velocity signal translates to an inclination-dependent planetary mass of
+$M\,\sin\,{i} = 5.81 \pm 0.72~M_J$ \citep[Equation~1]{Torres2008}, a massive
+giant planet; for sufficiently high inclinations ($\gtrsim 45^\circ$), the mass
+of Aldebaran~b may exceed $13~M_J$, making it a possible low-mass brown dwarf
+candidate. A recent study by \citet{2018arXiv180105239H} has cast doubt on the
+validity of the detection of the massive planet $\gamma$~Draconis~b, and by
+extension similar planets around similar K~giants such as Aldebaran. We remain
+convinced that the $\gamma$~Draconis problem is not an issue in this case, and
+Aldebaran~b is a bona fide planet, for reasons outlined in
+Appendix~\ref{gamma_dra}.
 
@@
 
-From this pilot study with Aldebaran, we have shown that with limited quantities of data of limited quality, we can nevertheless do asteroseismology from the ground and obtain very precise estimates of stellar parameters.
-Furthermore, improvements in the algorithm may also permit the detection of not only~\numax but also~\Dnu from similar measurements, permitting more sophisticated asteroseismic analysis.
-Sufficient radial velocity data either already exist, or are trivial to obtain, in order to do this for essentially all bright giants; while for solar-like stars, these new methods will allow significantly relaxed observing requirements.
-One rich archive is the series of observations with the Hamilton \'{E}chelle Spectrograph at Lick Observatory, studying 373 G~and K~giants \citep[e.g.][]{frink2001,hekker2006,ortiz2016}.
-The method may also be useful in detecting rotational or other periodic variability in the LCES HIRES/Keck Precision Radial Velocity Exoplanet Survey \citep{butler17} of 1,624 FGKM dwarfs, although the time sampling is by design too coarse to detect $p$-mode oscillations in these stars.
-Furthermore, CARMA models and related methods will enhance deep, all-sky, sparse photometric surveys: an immediate future test of this will be from \hipparcos; while only~58 epochs of photometry are available for Aldebaran, at a sampling we find to be insufficient for our purposes, stars at higher latitudes may often have 150--200 epochs and more even sampling \citep{hipparcos_phot}, and the~14 bright~K giants in \hipparcos noted by \citet{bedding2000} are an ideal first test case.
-This can naturally be extended to \emph{Gaia} in space \citep{gaia}, or \emph{LSST} from the ground \citep{dmt,lsst,lsstbook}, from which many thousands of new asteroseismic determinations will be possible.
-In future RV surveys, it may be possible to beat the \citet{2005PASJ...57...97S} $\sim 30$~m/s RV precision limit for red giants by taking many closely spaced observations and modelling-out the effects of stellar oscillations: this will allow us to dig deeper into this intrinsic stellar noise to detect less-massive planets around these stars.
+From this pilot study with Aldebaran, have shown that with limited quantities of
+data of limited quality, we can nevertheless do asteroseismology from the ground
+and obtain very precise estimates of stellar parameters. Furthermore,
+improvements in the algorithm may also permit the detection of not only~\numax
+but also~\Dnu from similar measurements, permitting more sophisticated
+asteroseismic analysis. Sufficient radial velocity data either already exist, or
+are trivial to obtain, in order to do this for essentially all bright giants;
+while for solar-like stars, these new methods will allow significantly relaxed
+observing requirements. One rich archive is the series of observations with the
+Hamilton \'{E}chelle Spectrograph at Lick Observatory, studying 373 G~and
+K~giants \citep[e.g.][]{frink2001,hekker2006,ortiz2016}. The method may also be
+useful in detecting rotational or other periodic variability in the LCES
+HIRES/Keck Precision Radial Velocity Exoplanet Survey \citep{butler17} of 1,624
+FGKM dwarfs, although the time sampling is by design too coarse to detect
+$p$-mode oscillations in these stars. Furthermore, CARMA models and related
+methods will enhance deep, all-sky, sparse photometric surveys: an immediate
+future test of this will be from \hipparcos; while only~58 epochs of photometry
+are available for Aldebaran, at a sampling we find to be insufficient for our
+purposes, stars at higher latitudes may often have 150--200 epochs and more even
+sampling \citep{hipparcos_phot}, and the~14 bright~K giants in \hipparcos noted
+by \citet{bedding2000} are an ideal first test case. This can naturally be
+extended to \emph{Gaia} in space \citep{gaia}, or \emph{LSST} from the ground
+\citep{dmt,lsst,lsstbook}, from which many thousands of new asteroseismic
+determinations will be possible. In future RV surveys, it may even be possible
+to beat the \citet{2005PASJ...57...97S} $\sim 30$~m/s RV precision limit for red
+giants by taking many closely spaced observations and modelling the effects of
+stellar oscillations: this will allow us to dig deeper into this intrinsic
+stellar noise to detect less-massive planets around these stars.
 
@@
   del Teide on the island of Tenerife by Aarhus and Copenhagen Universities and by the Instituto de
-  Astrofisica de Canarias.
+  Astrof\`{i}sica de Canarias.
 
@@
 of the ODE; $b_j \in \mathbb{C}$, $\Re b_j < 0$ are corresponding roots of the
-inverse process; and $\eta(t)$ is a white-noise GP with
+inverse process; and $\eta(t)$ is a white-noise Gaussian Process with
 $\left\langle \eta(t) \right\rangle = 0$ and $\left\langle \eta(t) \eta(t')
@@
 
-The Kalman filter implementation of the CARMA likelihood is not unique. 
-\citet{Foreman-Mackey2017} show how a dimensional expansion
-can be used to reduce the standard GP covariance matrix for this
-same process,  with autocorrelation given by Eq.\ \eqref{eq:CARMA-ACF} under
-observations at times $t_k$, to a banded form that can then be
-Cholesky-decomposed in the standard GP likelihood function in
-$\mathcal{O}\left( n \right)$ time. These `\emph{celerite}' models are formally equivalent to the Kalman CARMA model and differ only in their computational implementation.
-
 \begin{deluxetable}{lll}
@@
 There are two main reasons why this is not possible for all bright stars in general.
-First, because the on-board data storage and downlink bandwidth from \kepler are limited, it is often not desirable to store and download the large number of pixels that are required for such bright stars.
+First, because the on-board data storage and downlink bandwidth from \kepler are limited, it is often not desirable to download the large number of pixels that are required for such bright stars.
 Second, if the bleed column for a sufficiently bright star reaches the edge of the chip, flux spills over and is not conserved, imposing a hard brightness limit that depends on the distance to the detector edge.
@@
 
-As a final step to reduce residual uncorrected systematics, we apply the \textsc{k2sc} \citep{k2sc} GP-based systematics correction code to the initial halo light curve, but the effect of this in the present case for Aldebaran is minimal.
-There is somewhat higher than usual residual noise at harmonics of $46\,\mu\mathrm{Hz}$ ($4\,\mathrm{d}^{-1}$, the satellite thruster firing frequency), but this is nevertheless very small in comparison to the signal from Aldebaran, and may be ascribed to the large fraction of the pixel mask occupied by the bleed column from this extremely bright star.
+As a final step to reduce residual uncorrected systematics, we apply the \textsc{k2sc} \citep{k2sc} Gaussian Process-based systematics correction code to the initial halo light curve, but the effect of this in the present case for Aldebaran is minimal.
+There is somewhat higher than usual residual noise at harmonics of $4 c/d$ (the satellite thruster firing frequency), but this is nevertheless very small in comparison to the signal from Aldebaran, and may be ascribed to the large fraction of the pixel mask occupied by the bleed column from this extremely bright star.
 
@@
 \log_{10} \frac{L}{L_{\odot}} = 4.0+
-0.4 M_{{\textrm{bol}},\odot} -2.0 \log_{10} {\pi [{\textrm{mas}]}} -0.4(V-A_V + BC(V)).
+0.4 M_{{\textrm{bol}},\odot} -2.0 \log_{10} {\pi [{\textrm{mas}]}} \\-0.4(V-A_V + BC(V)).
 \label{eqn:lum}
@@
 		\hline
-		\cite{2012Sheffield}$_{a}$ &$1.17^{+0.07}_{-0.07}$&$43.9^{+0.9}_{-0.9}$&$6.5^{+1.4}_{-1.1}$\\
-		\cite{2012Sheffield}$_{b}$&$1.17^{+0.07}_{-0.07}$&$43.8^{+0.8}_{-0.9}$&$6.4^{+1.4}_{-1.1}$\\
-		\cite{2011Prugniel}&$1.17^{+0.07}_{-0.07}$&$43.9^{+0.9}_{-0.9}$&$6.2^{+1.4}_{-1.1}$\\
-		\cite{2008Massarotti}&$1.13^{+0.07}_{-0.07}$&$43.5^{+0.9}_{-0.9}$&$6.0^{+1.4}_{-1.1}$\\
-		\cite{2009Frasca}&$1.02^{+0.04}_{-0.04}$&$44.0^{+0.8}_{-0.8}$&$10.3^{+1.5}_{-1.4}$\\
+		\cite{2012Sheffield}$_{a}$ &$1.17^{+0.07}_{-0.07}$&$43.9^{+0.9}_{-0.9}$&$6.47^{+1.39}_{-1.1}$\\
+		\cite{2012Sheffield}$_{b}$&$1.17^{+0.07}_{-0.07}$&$43.8^{+0.8}_{-0.9}$&$6.35^{+1.4}_{-1.11}$\\
+		\cite{2011Prugniel}&$1.17^{+0.07}_{-0.07}$&$43.9^{+0.9}_{-0.9}$&$6.23^{+1.4}_{-1.11}$\\
+		\cite{2008Massarotti}&$1.13^{+0.07}_{-0.07}$&$43.5^{+0.9}_{-0.9}$&$5.95^{+1.37}_{-1.07}$\\
+		\cite{2009Frasca}&$1.02^{+0.04}_{-0.04}$&$44.0^{+0.8}_{-0.8}$&$10.26^{+1.54}_{-1.45}$\\
 		\hline
@@
 		\hline
-\cite{2012Sheffield}$_{a}$&$1.43^{+0.26}_{-0.24}$&$43.8^{+0.9}_{-0.9}$&$3.5^{+2.7}_{-1.4}$\\
-\cite{2012Sheffield}$_{b}$&$1.27^{+0.24}_{-0.2}$&$43.8^{+0.9}_{-0.9}$&$4.9^{+3.6}_{-2.0}$\\
-\cite{2011Prugniel}&$1.25^{+0.22}_{-0.19}$&$43.8^{+0.9}_{-0.9}$&$5.0^{+3.5}_{-2.0}$\\
-\cite{2008Massarotti}&$0.95^{+0.11}_{-0.05}$&$43.8^{+0.9}_{-0.9}$&$10.2^{+2.4}_{-3.1}$\\
-\cite{2009Frasca}&$0.96^{+0.04}_{-0.04}$&$44.4^{+0.8}_{-0.8}$&$11.6^{+1.4}_{-1.8}$\\
+\cite{2012Sheffield}$_{a}$&$1.43^{+0.26}_{-0.24}$&$43.8^{+0.9}_{-0.9}$&$3.53^{+2.72}_{-1.44}$\\
+\cite{2012Sheffield}$_{b}$&$1.27^{+0.24}_{-0.2}$&$43.8^{+0.9}_{-0.9}$&$4.86^{+3.56}_{-2.04}$\\
+\cite{2011Prugniel}&$1.25^{+0.22}_{-0.19}$&$43.8^{+0.9}_{-0.9}$&$5.02^{+3.47}_{-2.02}$\\
+\cite{2008Massarotti}&$0.95^{+0.11}_{-0.05}$&$43.8^{+0.9}_{-0.9}$&$10.24^{+2.36}_{-3.08}$\\
+\cite{2009Frasca}&$0.96^{+0.04}_{-0.04}$&$44.4^{+0.8}_{-0.8}$&$11.64^{+1.43}_{-1.8}$\\
 		\hline
@@
 		\hline
-\cite{2012Sheffield}$_{a}$&$1.17^{+0.07}_{-0.07}$&$43.9^{+0.9}_{-0.9}$&$6.5^{+1.4}_{-1.1}$\\
-\cite{2012Sheffield}$_{b}$&$1.16^{+0.07}_{-0.07}$&$43.8^{+0.8}_{-0.9}$&$6.4^{+1.4}_{-1.1}$\\
-\cite{2011Prugniel}&$1.16^{+0.07}_{-0.07}$&$43.9^{+0.9}_{-0.9}$&$6.4^{+1.5}_{-1.2}$\\
-\cite{2008Massarotti}&$1.13^{+0.07}_{-0.07}$&$43.5^{+0.9}_{-0.9}$&$5.9^{+1.4}_{-1.0}$\\
-\cite{2009Frasca}&$1.15^{+0.07}_{-0.07}$&$43.9^{+0.9}_{-0.8}$&$6.5^{+1.5}_{-1.2}$\\
+\cite{2012Sheffield}$_{a}$&$1.17^{+0.07}_{-0.07}$&$43.9^{+0.9}_{-0.9}$&$6.51^{+1.41}_{-1.11}$\\
+\cite{2012Sheffield}$_{b}$&$1.16^{+0.07}_{-0.07}$&$43.8^{+0.8}_{-0.9}$&$6.38^{+1.42}_{-1.12}$\\
+\cite{2011Prugniel}&$1.16^{+0.07}_{-0.07}$&$43.9^{+0.9}_{-0.9}$&$6.43^{+1.48}_{-1.15}$\\
+\cite{2008Massarotti}&$1.13^{+0.07}_{-0.07}$&$43.5^{+0.9}_{-0.9}$&$5.86^{+1.35}_{-1.05}$\\
+\cite{2009Frasca}&$1.15^{+0.07}_{-0.07}$&$43.9^{+0.9}_{-0.8}$&$6.47^{+1.51}_{-1.17}$\\
 		\hline

### 2018-02-05 paper/aldebaran.tex
STFC in ack's, closing #1
CCA affiliation for WMF.
Fix typo.

@@
   Birmingham, Birmingham, B15 2TT, United Kingdom}
+
+\affiliation{Center for Computational Astrophysics, Flatiron Institute, 162 Fifth Avenue, New York NY 10010, United States}
 
@@
 Red giants exhibit oscillations that are excited and damped by stellar convection.
-In the power spectrum of either radial velocity of photometric observations, these modes of oscillation form a characteristic pattern of peaks which can be used to infer intrinsic stellar properties \citep[e.g.,][]{2016AN....337..774D}.
+In the power spectrum of either radial velocity or photometric observations, these modes of oscillation form a characteristic pattern of peaks which can be used to infer intrinsic stellar properties \citep[e.g.,][]{2016AN....337..774D}.
 The easiest properties of the pattern to determine are the frequency of maximum amplitude \numax and the so-called large separation \Dnu \citep{Kjeldsen95}, which are often referred to as global asteroseismic parameters.
@@
 
+ WMF, GRD, TN, JWB, AM, and MNL are supported in part by the STFC.
+
 \software{\textsc{IPython} \citep{PER-GRA:2007}, SciPy \citep{scipy}, \textsc{matplotlib} \citep{Hunter:2007}, \textsc{Astropy} \citep{2013A&A...558A..33A}, \textsc{NumPy} \citep{van2011numpy}, \textsc{isochrones} \citep{isochrones}, \textsc{Julia} \citep{Bezanson2017}.}

### 2018-02-05 paper/aldebaran.tex
Final edits to manuscript before submission.

@@
 Radial velocity variations intrinsic to the star, such as those caused by stellar oscillations, have previously limited the precision with which red giant planets have been studied \citep{2005PASJ...57...97S}.
-We aim to model out these noise processes and use them to recover asteroseismic information, as well as to improve our estimates of the orbital parameters of the planet.
+We aim to model these noise processes and use them to recover asteroseismic information, as well as to improve our estimates of the orbital parameters of the planet.
 It is easy to observe Aldebaran and similarly-bright stars with ground-based spectroscopic instruments, typically only requiring short exposures that can be obtained even under adverse observing conditions.
@@
 We find evidence for a low-quality ($Q = \HatzesQRange{}$) oscillatory mode with
-$\nu_\mathrm{max} = \HatzesNuMaxRange{} \, \mu\mathrm{Hz}$.
-This is consistent with the presence of a number of un-resolved asteroseismic modes in the RV data.
-The low Q factor in this case refers to the inverse fractional width of the band over which the modes are excited, not the quality of any individual mode.
-Our posterior for \numax{} from this data set is shown in Figure
-\ref{fig:numax-datasets}.
+$\nu_\mathrm{max} = \HatzesNuMaxRange{} \, \mu\mathrm{Hz}$ (here and throughout
+we quote the posterior median and the range from the 0.16 to the 0.84 posterior
+quantile).  This is consistent with the presence of a number of un-resolved
+asteroseismic modes in the RV data.  The low Q factor in this case refers to the
+inverse fractional width of the band over which the modes are excited, not the
+quality of any individual mode.  Our posterior for \numax{} from this data set
+is shown in Figure \ref{fig:numax-datasets}.
 
@@
 \cite{2017ApJ...836..142G}. The typical velocity precision per visit is
-2.5$\,\rm m\,s^{-1}$ allowing us to easily detect the oscillations which
-display a peak-to-peak amplitude of $\approx \, 170 \, \rm m \, s^{-1}$.
-Figure~\ref{songdata} displays the radial velocities obtained -- oscillations
-are easily visible as well as a long-term trend, probably due to the planetary
-companion.  From this data set alone we also find evidence for a low-quality ($Q
-= \SONGQRange{}$) oscillatory mode with $\nu_\mathrm{max} = \SONGNuMaxRange{} \,
+2.5$\,\rm m\,s^{-1}$ allowing us to easily detect the oscillations which display
+a peak-to-peak amplitude of $\approx \, 170 \, \rm m \, s^{-1}$.
+Figure~\ref{songdata} displays the radial velocities obtained---oscillations
+are easily visible as well as a long-term trend, due to the planetary companion.
+From this data set alone we also find evidence for a low-quality ($Q =
+\SONGQRange{}$) oscillatory mode with $\nu_\mathrm{max} = \SONGNuMaxRange{} \,
 \mu\mathrm{Hz}$ which is consistent with the inference from the archival data.
@@
 
-We have determined the stellar properties
-of Aldebaran with the additional constraint of \numax determined from this
-analysis.  The details of the stellar modelling are presented in Appendix
-\ref{sec:mod}.  We have considered the impact of our additional \numax
-constraint and run our analysis for different sets of spectroscopic estimates.
-For final values we adopt the \citet{2012Sheffield} spectroscopic solution as
-being `middle of the road' estimates.  For the analysis without constraint on
-\numax we find estimates of the stellar properties as $M = 1.27^{+0.24}_{-0.20}
-\, \mathrm{M_{\odot}}$ and age $4.9^{+3.6}_{-2.0} \, \rm Gyr$.  With the
-inclusion of \numax we find $M = 1.16\, \pm \, {0.07} \, \mathrm{M_{\odot}}$ and
-age $6.4^{+1.4}_{-1.1} \, \rm Gyr$.  It is clear that the addition of \numax
-provides substantially more precise estimates of mass and age. With this stellar
-mass, the radial velocity signal translates to an inclination-dependent
-planetary mass of $M\,\sin\,{i} = 5.8 \pm 0.7~M_J$
-\citep[Equation~1]{Torres2008}, a massive giant planet; for sufficiently high
-inclinations ($\gtrsim 45^\circ$), the mass of Aldebaran~b may exceed $13~M_J$,
-making it a possible low-mass brown dwarf candidate.
-A recent study by \citet{2018arXiv180105239H} has cast doubt on the validity of the detection of the massive planet $\gamma$~Draconis~b, and by extension similar planets around similar K~giants such as Aldebaran.
-We remain convinced that the $\gamma$~Draconis problem is not an issue in this case, and Aldebaran~b is a bona fide planet, for reasons outlined in Appendix~\ref{gamma_dra}.
+We have used the additional constraint of \numax determined from this analysis
+to update the stellar properties of Aldebaran.  The details of the stellar
+modelling are presented in Appendix \ref{sec:mod}.  We have considered the
+impact of our additional \numax constraint and run our analysis for different
+sets of spectroscopic estimates.  For final values we adopt the
+\citet{2012Sheffield} spectroscopic solution as being `middle of the road'
+estimates.  For the analysis without constraint on \numax we find estimates of
+the stellar properties as $M = 1.27^{+0.24}_{-0.20} \, \mathrm{M_{\odot}}$ and
+age $4.9^{+3.6}_{-2.0} \, \rm Gyr$.  With the inclusion of \numax we find $M =
+1.16\, \pm \, {0.07} \, \mathrm{M_{\odot}}$ and age $6.4^{+1.4}_{-1.1} \, \rm
+Gyr$.  It is clear that the addition of \numax provides substantially more
+precise estimates of mass and age. With this stellar mass, the radial velocity
+signal translates to an inclination-dependent planetary mass of $M\,\sin\,{i} =
+5.8 \pm 0.7~M_J$ \citep[Equation~1]{Torres2008}, a massive giant planet; for
+sufficiently high inclinations ($\gtrsim 45^\circ$), the mass of Aldebaran~b may
+exceed $13~M_J$, making it a possible low-mass brown dwarf candidate. A recent
+study by \citet{2018arXiv180105239H} has cast doubt on the validity of the
+detection of the massive planet $\gamma$~Draconis~b, and by extension similar
+planets around similar K~giants such as Aldebaran. We remain convinced that the
+$\gamma$~Draconis problem is not an issue in this case, and Aldebaran~b is a
+bona fide planet, for reasons outlined in Appendix~\ref{gamma_dra}.
 
@@
 
-With CARMA models we have confirmed the previously-suspected planet around Aldebaran using archival data, and additionally detected acoustic oscillations that permit a mass determination with a precision of $\sim 6\%$.
-We have confirmed both of these results with new ground-based observations, and space-based photometry with \ktwo.
+Using sophisticated time-domain CARMA models for the stellar activity in
+Aldebaran, we have confirmed the previously-suspected planet and additionally
+detected acoustic oscillations that permit a mass determination with a precision
+of $\sim 6\%$. We have confirmed both of these results with new ground-based
+observations, and space-based photometry with \ktwo.
 
@@
 The Kalman filter implementation of the CARMA likelihood is not unique.
-\citet{Foreman-Mackey2017} show how a dimensional expansion
-can be used to reduce the standard GP covariance matrix for this
-same process,  with autocorrelation given by Eq.\ \eqref{eq:CARMA-ACF} under
-observations at times $t_k$, to a banded form that can then be
-Cholesky-decomposed in the standard GP likelihood function in
-$\mathcal{O}\left( n \right)$ time. These `\emph{celerite}' models are formally equivalent to the Kalman CARMA model and differ only in their computational implementation.
+\citet{Foreman-Mackey2017} show how a dimensional expansion can be used to
+reduce the standard GP covariance matrix for this same process,  with
+autocorrelation given by Eq.\ \eqref{eq:CARMA-ACF} under observations at times
+$t_k$, to a banded form that can then be Cholesky-decomposed in the standard GP
+likelihood function in $\mathcal{O}\left( n \right)$ time. These
+`\emph{celerite}' models are formally equivalent to a Kalman CARMA model with $q
+= p-1$ and differ only in their computational implementation.
 
@@
 It was revived as \ktwo \citep{howell14}, balanced by orienting perpendicular to the Sun.
-This requires that \ktwo observes fields in the Ecliptic in $\sim80$~d Campaigns, of which Aldebaran was observed in Campaign~13.
+This requires that \ktwo observes fields in the Ecliptic in $\sim80$~d Campaigns; Aldebaran was observed in Campaign~13.
 

### 2018-02-05 presub/coverletter.tex
As sumbitted.

@@
-% Cover letter using letter.cls
-\documentclass[11pt]{letter} % Uses 10pt
-%\usepackage{helvetica} % uses helvetica postscript font (download helvetica.sty)
-%\usepackage{newcent}   % uses new century schoolbook postscript font 
-% the following commands control the margins:
-
-\topmargin=-1.5in    % Make letterhead start about 1 inch from top of page 
-\textheight=11.75in    % text height can be bigger for a longer letter
-\oddsidemargin=0pt   % leftmargin is 1 inch
-\textwidth=6.5in     % textwidth of 6.5in leaves 1 inch for right margin
+\documentclass{letter}
 \usepackage{url}
 \usepackage{hyperref}
-\hypersetup{
-    colorlinks=false,
-    pdfborder={0 0 0},
-}
 \usepackage{graphicx}
-\usepackage[strict]{changepage}
 \usepackage{xcolor}
-
-%--------------------------------------------------------------
-% A few colors for hyperlinks.
-%--------------------------------------------------------------
-\definecolor{plum}{rgb}{0.36078, 0.20784, 0.4}
-\definecolor{chameleon}{rgb}{0.30588, 0.60392, 0.023529}
-\definecolor{cornflower}{rgb}{0.12549, 0.29020, 0.52941}
-\definecolor{scarlet}{rgb}{0.8, 0, 0}
-\definecolor{brick}{rgb}{0.64314, 0, 0}
-
-\newcommand{\email}[1]{\href{mailto:#1}{\tt \textcolor{cornflower}{#1}}}
 
@@
 \begin{document}
+\signature{Will Farr}
 
-%
-\longindentation=0pt                       % needed to get closing flush left
-\let\raggedleft\raggedright                % needed to get date flush left
+\address{Will M. Farr\\School of Physics and Astronomy\\University of Birmingham\\Birmingham\\B15 2TT\\United Kingdom}
 
-\nopagebreak  
 \begin{letter}
@@
 
-{
-\leftskip = 0pt plus 1 fill
-\rightskip = 0pt
-\parindent 0pt
-%\obeylines
-%
-%
-{{\bf Dr Will M. Farr}} \\
-School of Physics and Astronomy\\
-University of Birmingham\\
-Birmingham, B15 2TT\\
-United Kingdom\\
- \url{w.farr@bham.ac.uk} \\
+\opening{Dear Dr. Sage:}
 
-}
+Please find enclosed a submission for a Nature Letter titled ``Aldebaran~b's
+temperate past uncovered in planet search data''.
 
-\begin{minipage}[c]{5in}\vskip-4.25cm
-\begin{flushleft}
-	\begin{minipage}[c]{3cm}
-		\begin{flushleft}
-			\includegraphics*[width=8cm]{birmingham_logo.png}%
-		\end{flushleft}
-	\end{minipage}
-\end{flushleft}
-\end{minipage}
+Aldebaran ($\alpha$~Tauri, the Eye of the Bull) is a nearby first-magnitude red
+giant star. Since the identification of the planet candidate Aldebaran~b in~1993
+in the first modern radial velocity (RV) surveys, the existence of a massive
+planet orbiting the star has been firmly established with a further two decades
+of observations.
 
-\opening{Dear Leslie,}
+One astronomer's noise is another astronomer's data: using new data processing
+techniques, we have found that some of the ``noise'' in these archival
+observations is actually the signal of stellar oscillations. We confirmed our
+result with new observations. The asteroseismic oscillations allow us to infer
+much more accurate stellar parameters and therefore orbital parameters for the
+planet; we constrain the star's age, measure its mass with a factor of two
+better precision than had previously been possible (to 6\%), and consequently
+obtain the mass of the giant planet to a precision of 12\%.
 
-\noindent I submit for your consideration a Letter on the ``Detection of Oscillations in Aldebaran with Ground-Based Observations''.
+Our technique will be widely applicable to legacy and future datasets, allowing
+both asteroseismology with RV, and also searches for lower-mass companions to
+giants than was previously possible.  There are many targets with \emph{already
+existing} archival data amenable to this technique, and it will also be
+applicable to upcoming and ongoing large surveys (Gaia, LSST, etc).
 
-Aldebaran ($\alpha$~Tauri, the Eye of the Bull) is a nearby first-magnitude red giant star.
-Since the identification of the planet candidate Aldebaran~b in~1993 in the first modern radial velocity (RV) surveys, the existence of a massive planet orbiting the star has been firmly established with a further two decades of observations.
+We have also independently confirmed our results with new photometry from the
+\emph{Kepler}-2 (K2) mission, again using a novel algorithm to extract a precise
+light curve of the brightest star ever observed by \emph{Kepler}. The K2 light
+curve shows very clear oscillations that are in agreement with the frequency we
+infer from RV.
 
-One astronomer's noise is another astronomer's data: we have found that the noise in these archival observations is actually the signal of stellar oscillations.
-With radical new data processing techniques, we have re-analysed historical RV data and confirmed our result with new observations.
-Using our Gaussian Process model we detect the signal of stellar oscillations, missed by previous authors, and consequently infer much more accurate orbital parameters for the planet.
-This technique will be widely applicable to legacy and future datasets, allowing both asteroseismology with RV, and also searches for lower-mass companions to giants than was previously possible.
+A neat and surprising result of this work is that while the planet has a hellish
+temperature today well upwards of a thousand Kelvin, when Aldebaran was on the
+main sequence, the planet (and more importantly any of its moons) would have
+been subject to similar incident starlight as the Earth receives from the Sun.
+We posit that this is the first clear example of a formerly-habitable world
+rendered uninhabitable by stellar evolution.
 
-We have also independently confirmed our results with new photometry from the \emph{Kepler}-2 (K2) mission, again using a highly novel algorithm to extract a precise light curve of the brightest star ever observed by \emph{Kepler}.
-The K2 light curve shows very clear oscillations that are in perfect agreement with the frequency we infer from RV.
+Should you be interested in publishing the paper in Nature, it will require some
+minor reformatting.  It is currently about 3000 words in the main body, with
+five display items; however many details currently in the main body can be
+easily moved to the appendices / methods sections.  It also contains more
+references than permitted in a Letter, but many of these can also be moved to
+the Methods section and others can be eliminated if necessary.
 
-With asteroseismology we constrain the star's age for the first time, measure its mass with a factor of two better precision than had previously been possible (to 6\%), and consequently obtain the mass of the giant planet to a precision of 12\%.
+Thank you for considering our submission. Please do not hesitate to contact me
+with any questions.  Should you feel that our submission is not appropriate for
+Nature, we would be happy to have it considered for Nature Astronomy.
 
-A neat and surprising result of this work is that while the planet has a hellish temperature today well upwards of a thousand Kelvin, when Aldebaran was on the main sequence, the planet (and more importantly any of its moons) would have been subject to similar incident starlight as the Earth receives from the Sun.
-We posit that this is the first clear example of a formerly-habitable world rendered uninhabitable by stellar evolution.
-
-Thank you for considering our submission. Please do not hesitate to contact me with any questions.\\
-
-\closing{Yours sincerely, \\
-\vspace{1.0cm}
-Will Farr}
-
-%\encl{One paragraph Letter outline}					% Enclosures
+\closing{Sincerely,}
 

### 2018-02-06 paper/aldebaran.tex
Site-dependent *uncertainty* scaling factor (thanks, Vichi!).

@@
 the ODE and the corresponding roots $r_i$. This likelihood is computed for the
-residuals of a deterministic Keplerian RV mean model for the planetary companion.
-Because we have a set of observations taken with different instruments at
-different sites (see Section \ref{sec:archival-obs}), we include RV offset
-(i.e.\ mean value) parameters that are instrument- and site-dependent and also
-an instrument- and site-dependent scaling parameter.  Thus, the $i$th RV
-measured at site and intstrument combination $j$, $y_{ij}$, is assumed to be
+residuals of a deterministic Keplerian RV mean model for the planetary
+companion.  Because we have a set of observations taken with different
+instruments at different sites (see Section \ref{sec:archival-obs}), we include
+RV offset (i.e.\ mean value) parameters that are instrument- and site-dependent
+and also an instrument- and site-dependent uncertainty scaling parameter.  Thus,
+the $i$th RV measured at site and intstrument combination $j$, $y_{ij}$, is
+assumed to be
 %

### 2018-02-15 paper/aldebaran.tex
Tom gets middle initials, some re-formatting of captions.

@@
 
-\author{Thomas North}
+\author{Thomas S. H. North}
 \affiliation{School of Physics and Astronomy, University of Birmingham, Birmingham, B15 2TT, United Kingdom}
@@
 To measure global asteroseismic parameters we model the background variability in the Fourier domain using the methodology described in \citet{huber09}, yielding a frequency of maximum power of $\numax = 2.2 \pm 0.25\,\muHz$ and amplitude per radial mode of $1850 \pm 500$\,ppm.
-Due to the limited frequency resolution of the 70-day K2 time series we were unable to measure the large frequency separation \Dnu, which is expected to be on the order of $\approx$\,0.5\muHz.
+Due to the limited frequency resolution of the 70-day K2 time series we were unable to measure the large frequency separation \Dnu, which is expected to be $\approx$\,0.5\muHz.
 
@@
 Reversing the equation to produce a predicted $\log{g}$ from the observed $\nu_{\textrm{max,obs}}=2.23\pm0.1\mu$Hz results in a predicted $\log{g}\sim1.2$ dex, using an assumed temperature of 3900K.
-The solar calibration values used here are $g_{\odot}=274\,$ms$^{-2}$, $\nu_{\textrm{max},\odot}=3150\,\mu$Hz and $T_{\textrm{eff},\odot}=5777\,$K.
+The solar calibration values used here are $\log g_{\odot} / \left( \mathrm{m} \, \mathrm{s}^{-1}\right)=2.44 \, \mathrm{dex}$, $\nu_{\textrm{max},\odot}=3150\,\mu$Hz and $T_{\textrm{eff},\odot}=5777\,$K.
 

### 2018-02-23 paper/aldebaran.tex
Close #3
Re-configured Figure 1 to fit better into the available space (the units
were cut off).
Close #6.
Show plot of model against data in appendix.

@@
 \centering
-\includegraphics[width=0.8\textwidth]{all-data.pdf}
-\caption{All archival and new radial velocities used for inference in this work.}
+\plotone{all-data.pdf}
+\caption{All archival and new radial velocities used for inference in this work.  The site and instrument for each data set is indicated in the legend (see text for a full description).}
 \label{alldata}
@@
 impose in all cases on the parameters of the CARMA model, planetary RV signal,
-and instrumental and site effects are given in Table \ref{tab:parameters}.
+and instrumental and site effects are given in Table \ref{tab:parameters}.  A
+draw from the posterior for our canonical model in shown in data space in Figure
+\ref{fig:data-space}.
+
+\begin{figure}
+\plotone{data-space}
+
+\caption{\label{fig:data-space} A posterior draw from our model plotted with the
+observed radial velocities.  Each site's radial velocity has been shifted by the
+model's site-specific RV offset and errorbars have been scaled by the model's
+site-specific errorbar scaling.  Data colors are as in Figure \ref{alldata}.
+The black line gives the sum of the model's Keplerian component and the mean of
+the inferred stochastic component; the dark and light bands give the 1- and
+2-sigma uncertainty in the inferred stochastic component.  Between observations,
+the uncertainty in the stochastic component increases; it reduces near
+observations, as each observation provides some information about the stellar
+noise that propogates to nearby times because of the temporal correlation in the
+noise.  Stellar oscillations are apparent in the short-timescale wiggles in the
+stochastic component between nearby observations; there is also a significant
+long-period correlated component (represented in our model by an
+exponentially-decaying eigenmode).}
+
+\end{figure}
 

### 2018-02-23 presub/coverletter.tex
Cover letter updated for submission.

@@
 
-\address{Will M. Farr\\School of Physics and Astronomy\\University of Birmingham\\Birmingham\\B15 2TT\\United Kingdom}
+\address{Will M. Farr\\School of Physics and Astronomy\\University of Birmingham\\Birmingham\\B15 2TT\\United Kingdom\\\href{mailto:w.farr@bham.ac.uk}{\texttt{w.farr@bham.ac.uk}}}
 
 \begin{letter}
-{Leslie Sage \\
-l.sage@us.nature.com \\}
+{Fred Rasio \\
+\href{mailto:ApJLFred@northwestern.edu}{\texttt{ApJLFred@northwestern.edu}}}
 
-\opening{Dear Dr. Sage:}
+\opening{Dear Dr. Rasio:}
 
-Please find enclosed a submission for a Nature Letter titled ``Aldebaran~b's
+Please find enclosed a submission for an ApJ Letter titled ``Aldebaran~b's
 temperate past uncovered in planet search data''.
@@
 
-Should you be interested in publishing the paper in Nature, it will require some
-minor reformatting.  It is currently about 3000 words in the main body, with
-five display items; however many details currently in the main body can be
-easily moved to the appendices / methods sections.  It also contains more
-references than permitted in a Letter, but many of these can also be moved to
-the Methods section and others can be eliminated if necessary.
+We would like to have this submission considered for a linked Letter+Supplement;
+we have formatted it as one document, but the appendices are intended for the
+Supplement.  It is not clear to us whether there are length limits still applied
+to letters; the main body is currently about 4000 words, contains six figures
+(two of which could be merged if necessary), and sixty references (this could be
+reduced by moving some references to the Supplement if necessary).
 
 Thank you for considering our submission. Please do not hesitate to contact me
-with any questions.  Should you feel that our submission is not appropriate for
-Nature, we would be happy to have it considered for Nature Astronomy.
+with any questions.
 

### 2018-02-25 paper/aldebaran.tex
As submitted.
autograd

@@
 
-In an improvement since \citet{White2017}, we use the \textsc{Theano} library \citep{theano} to calculate analytic derivatives for the TV objective function, which reduces the computational time for a single halo light curve on a commercial laptop from tens of minutes to tens of seconds.
+In an improvement since \citet{White2017}, we use the \textsc{autograd} library \citep{maclaurinautograd} to calculate analytic derivatives for the TV objective function, which reduces the computational time for a single halo light curve on a commercial laptop from tens of minutes to a few seconds.
 

### 2018-04-05 paper/aldebaran.tex
Added arXiver tag.
Removed unnecessary \include{macros}.
new models

@@
 
-\include{macros}
+%\include{macros}
 % Journals
+
+% arxiver suggestions for plots:
+%@arxiver{data-space.pdf,numax.pdf,psd-zoom.pdf}
 
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{../plots/numax}
+  \includegraphics[width=\columnwidth]{numax}
   \caption{Posterior for \numax{} from the complete, combined RV data set described in this paper (blue), the archival data from \citet{Hatzes2015} (green), and the SONG data described in Section \ref{sec:SONG-observations} (red).
@@
 We conduct a Monte Carlo simulation, drawing masses randomly from a Gaussian distribution $M \sim \mathcal{N}(1.16,0.07)$ and metallicities $[\text{Fe}/\text{H}] \sim \mathcal{N}(-0.15,0.2)$ \citep{decin2003} to predict the luminosity of the main-sequence progenitor, and the semi-major axis of the planet's orbit.
-Using the Mesa Isochrones and Stellar Tracks \citep[MIST:][]{mist0,mist1} models, we compute the evolution of the luminosity $L$ of the star along the main sequence; $L$ evolves from $2.0 \pm 0.7$~\lsun at 0.5~Gyr to $3.5 \pm 2.3$~\lsun at 4.5~Gyr.
-We note that this implies that the planet at a semi-major axis of $1.50\, \pm\, 0.03$~AU would have been subject to an insolation comparable to that of the Earth, evolving from 0.5~Gyr to 4.5~Gyr from $0.86\, \pm\, 0.27$ to $1.55 \,\pm\,1.0$ times that of the Earth today (Figure~\ref{insolation}).
+Using the same \citet{2017Rod} MESA models as were used to compute the stellar parameters, interpolating with the \textsc{isochrones} package \citep{isochrones}, we compute the evolution of the luminosity $L$ of the star along the main sequence; $L$ evolves from $2.0 \pm 0.7$~\lsun at 0.5~Gyr to $3.5 \pm 2.3$~\lsun at 4.5~Gyr.
+We note that this implies that the planet at a semi-major axis of $1.50\, \pm\, 0.03$~AU would have been subject to an insolation comparable to that of the Earth, evolving from 0.5~Gyr to 3.24~Gyr from $0.82\, \pm\, 0.27$ to $1.25^{+0.52}_{-0.40}$ times that of the Earth today (Figure~\ref{insolation}).
+After $\sim 3.2$ Gyr, the high mass tail of the simulations have evolved off the main sequence and dominate estimates for the insolation, while masses at the low end permit a substantially longer duration on the main sequence ($\sim Gyr$).
 Subject as well as this to the great uncertainties of the planet's orbital evolution and albedo, Aldebaran~b and any of its moons (or its S-type planets if it is a brown dwarf) may well have hosted temperate environments for some of their history, now long-since destroyed by their star's evolution away from the main sequence.
@@
 \caption{Evolution of the insolation of Aldebaran~b relative to the incident sunlight on Earth, as a function of the age of Aldebaran on the main sequence.
+We end the simulation plot at 3.2~Gyr, as after this, the high tail of the distribution of simulated masses has left the main sequence and disturbs estimates of the insolation.
 The solid line is the mean of simulations, and the shaded region represents $1\,\sigma$ equivalent deviations.

### 2018-04-05 paper/aldebaran.tex
new calculations

@@
 We note that this implies that the planet at a semi-major axis of $1.50\, \pm\, 0.03$~AU would have been subject to an insolation comparable to that of the Earth, evolving from 0.5~Gyr to 3.24~Gyr from $0.82\, \pm\, 0.27$ to $1.25^{+0.52}_{-0.40}$ times that of the Earth today (Figure~\ref{insolation}).
-After $\sim 3.2$ Gyr, the high mass tail of the simulations have evolved off the main sequence and dominate estimates for the insolation, while masses at the low end permit a substantially longer duration on the main sequence ($\sim Gyr$).
-Subject as well as this to the great uncertainties of the planet's orbital evolution and albedo, Aldebaran~b and any of its moons (or its S-type planets if it is a brown dwarf) may well have hosted temperate environments for some of their history, now long-since destroyed by their star's evolution away from the main sequence.
+After $\sim 3.2$ Gyr, the high mass tail of the simulations have evolved off the main sequence and dominate estimates for the insolation, while masses at the low end permit a substantially longer duration on the main sequence ($\sim$~Gyr).
+Subject as well as this to the great uncertainties of the planet's orbital evolution and albedo, Aldebaran~b and any of its moons (or its S-type planets if it is a brown dwarf, i.e.
+planets orbiting the brown dwarf and not the primary or wider binary) may well have hosted temperate environments for some of their history, now long-since destroyed by their star's evolution away from the main sequence.
 

### 2018-04-24 paper/aldebaran.tex
lenape acknowledgement
eora acknowledgement

@@
 
-The authors would like to acknowledge the Gadigal people of the Eora Nation on whose ancestral lands the University of Sydney is built.
+The authors would like to acknowledge the Gadigal people of the Eora Nation on whose ancestral lands the University of Sydney is built, and we pay our respects their knowledge, and their elders past and present.
+We also give blessings and thanks to the Lunaape people and Lunaape Nation for allowing us to do this work on their homeland at New York University.
 

### 2018-05-07 paper/aldebaran.tex
removing acknowledgement
updated acknowledgments

@@
 The authors would like to acknowledge the Gadigal people of the Eora Nation on whose ancestral lands the University of Sydney is built, and we pay our respects their knowledge, and their elders past and present.
-We also give blessings and thanks to the Lunaape people and Lunaape Nation for allowing us to do this work on their homeland at New York University.
 
@@
 D.H.
-acknowledges support by the Australian Research Council's Discovery Projects funding scheme (project number DE140101364) and support by the National Aeronautics and Space Administration under Grants NNX14AB92G and NNX16AH45G issued through the Kepler Participating Scientist Program and the K2 Guest Observer Program.
+acknowledges support by the Australian Research Council's Discovery Projects funding scheme (project number DE140101364) and support by the National Aeronautics and Space Administration under Grants NNX14AB92G and NNX17AF76G issued through the Kepler Participating Scientist Program and the K2 Guest Observer Program.
 

### 2018-07-11 paper/aldebaran.tex
Closes #8
Mention SONG in abstract.

@@
 
-\begin{abstract}
-The nearby red giant Aldebaran is known to host a gas giant planetary companion from decades of ground-based spectroscopic radial velocity measurements.
-Using Gaussian Process-based Continuous Auto-Regressive Moving Average (CARMA) models, we show that these historic data also contain evidence of acoustic oscillations in the star itself, and verify this result with further dedicated ground-based spectroscopy and space-based photometry with the \kepler Space Telescope.
-From the frequency of these oscillations we determine the mass of Aldebaran to be $1.16 \pm 0.07$~\msun, and note that this implies its planet will have been subject to insolation comparable to the Earth for some of the star's main sequence lifetime.
-Our approach to sparse, irregularly sampled time series astronomical observations has the potential to unlock asteroseismic measurements for thousands of stars in archival data, and push to lower-mass planets around red giant stars.
+\begin{abstract} The nearby red giant Aldebaran is known to host a gas giant
+planetary companion from decades of ground-based spectroscopic radial velocity
+measurements. Using Gaussian Process-based Continuous Auto-Regressive Moving
+Average (CARMA) models, we show that these historic data also contain evidence
+of acoustic oscillations in the star itself, and verify this result with further
+dedicated ground-based spectroscopy using the SONG telescope and space-based
+photometry with the \kepler Space Telescope. From the frequency of these
+oscillations we determine the mass of Aldebaran to be $1.16 \pm 0.07$~\msun, and
+note that this implies its planet will have been subject to insolation
+comparable to the Earth for some of the star's main sequence lifetime. Our
+approach to sparse, irregularly sampled time series astronomical observations
+has the potential to unlock asteroseismic measurements for thousands of stars in
+archival data, and push to lower-mass planets around red giant stars.
 \end{abstract}

### 2018-07-14 paper/aldebaran.tex
Closes #10
Clarify that "CARMA" in Fig 4 caption refers to our analysis, not the 
radio telescope.
addressed scaling relation comment by referee

@@
 
-While the accuracy of these scaling relations is still a matter of ongoing work \citep[e.g.,][]{2017ApJ...844..102H}, stellar parameters can be estimated by comparing observables to parameters from models of stellar evolution.
+The accuracy of these scaling relations is still a matter of ongoing work, but recent comparisons using Gaia parallaxes has shown encouraging agreement for red giants \citep[e.g.,][]{2017ApJ...844..102H,zinn18}.
+\citet{viani17} recently used models to demonstrate that the \numax\ scaling relation may show a dependence on the mean molecular weight.
+However, this effect is typically at the 0.01--0.02\,dex level for red giants, which is a factor $\sim$\,2--4 smaller than our measurement uncertainties (see Section 4) and thus does not significantly affect the conclusions in this paper.
 
@@
 \includegraphics[width=\textwidth]{psd-zoom.pdf}
-\caption{CARMA-derived power spectral density from archival and SONG RV observations of Aldebaran, showing the clear peak at around 2\,\muHz.
-The dark and light shaded regions represent $1\,\sigma$~and $2\,\sigma$ posterior probability contours.
-}
+%
+\caption{Power spectral density derived from the CARMA analysis of archival and
+SONG RV observations of Aldebaran, showing the clear peak at around 2\,\muHz.
+The dark and light shaded regions represent $1\,\sigma$~and $2\,\sigma$
+posterior probability contours. }
+%
 \label{psd}

### 2018-08-03 paper/aldebaran.tex
Close #12 addressing the medalist scale in PARAM fits.
updated acknowledgments

@@
 Our new approach to asteroseismic data analysis, based on Continuous Auto-Regressive Moving Average (CARMA) models, can extract exoplanet signals together with measures of \numax from sparse and irregularly-sampled time series.
+Studies of small numbers planet-hosting giants to determine their asteroseismic parameters ordinarily require dedicated observing campaigns with hundreds or thousands of epochs \citep{2017MNRAS.472.4110S}.
 An all-sky survey to find planetary companions and to precisely measure the masses of all nearby red giant stars is feasible with this new approach, and the required data either already exist in large radial velocity exoplanet surveys, or are easy to obtain with ground-based telescopes.
@@
 D.H.
-acknowledges support by the Australian Research Council's Discovery Projects funding scheme (project number DE140101364) and support by the National Aeronautics and Space Administration under Grants NNX14AB92G and NNX17AF76G issued through the Kepler Participating Scientist Program and the K2 Guest Observer Program.
+acknowledges support by the National Aeronautics and Space Administration under Grants NNX14AB92G, NNX17AF76G and 80NSSC18K0362 issued through the Kepler Participating Scientist Program and the K2 Guest Observer Program.
 
@@
 \begin{itemize}%GUY THIS BIT COULD PROBABLY BE DROPPED AND REFERENCED OUT TO \citep{2017Rod}
+\item Medalist scale from \citet{Grevesse93} as described in \citet{2017Rod}.
 \item Heavy element partitioning from \cite{1993Grevesse}.

### 2018-08-08 ref1/response.txt
Adding ref response and report.

Response to Reviewer's Comments:

First, we thank the referee for their report; we also apologize for the delay in
response.  We will address the comments of the referee in-line below.  In
addition to the changes discussed below we have modified the acknowledgements
section to add several new acknowledgements.

> Reviewer's Comments:

> This is an interesting paper and a clever use of statistical techniques to use
> historical data. I would like a few comments addressed and some revisions made
> before I recommend publication.

> I do find the title somewhat over the top given the uncertainties in Fig. 6! To
> me the paper is more about applying a novel (to astronomers!) statistical method
> to measure parameters that we otherwise would not be able to. [Ed: A little over
> the top is OK for an ApJL title... ]

We have not changed the title.

> Comments and concerns:

> (1) Since Kepler is mentioned so prominently in the abstract, SONG should be
> too, and judging by Fig 1, TLS may be worth a mention too.

We have done this; we now mention SONG in the abstract.  We have left TLS out of
the abstract since the archival data we use come only partially from TLS.

> (2) Discussion of the scaling relations: Given that there is work that shows
> that there might be a mean-molecular weight term in the numax scaling relation,
> a discussion of that and the reference to the paper to be given. This is
> relevant both after Eq. 3 and Eq. C7.

We have added two sentences with one reference addressing this point, in the
paragraph following Eq. (3).

> (3) I understand that the statistical method used has the acronym CARMA,
> however, for most astronomers CARMA is a radio array (Combined Array for
> Research in Millimeter-wave Astronomy). The caption of Fig 4 was thus very
> confusing when I first read it (I, like many others, look at figures before
> reading), and for a few moments I thought we could use radio observations to
> study pulsations. Please change that to Something like "Power spectral density
> derived from the CARMA analysis ..."

Done.

> (4) Were the same set of stellar models used for calculating Aldebaran's
> properties using PARAM and the results in Fig 6? The text is not clear about
> this and for consistent results, the same models should be used.

Originally a separate set of models were used; now the same MESA models are used
for both the PARAM code and the insolation calculation.  This is explicitly
mentioned in the text (paragraph beginning "With the determination of the
stellar mass...").

> (5) The medalist scale of the models is not mentioned in Section C.1. I.e., what
> solar metallicity and abundance mixture was assumed? Grevesse & Sauval (1998),
> Asplund et al. (2009)? Please specify. The difference is enough to cause a large
> effective change in Z/X and hence deduced stellar properties.

We now state the medalist scale explicitly, with reference to Grevesse & Noels
(1993), from which it is derived and Rodrigues, et al. (2017) which describes
its use.  (First bullet point of C.1.)

> (6) There are other (non-seismic) works that deal with modelling Aldebaran
> (e.g., Abia et al. 2012). How do the results obtained in this paper compare with
> literature values?

We have added a sentence in Section 4.4 pointing to Abia+2012 and noting that
our non-asteroseismic mass value and uncertainty are consistent with that work.
The asteroseismic mass estimate is, of course, improved in uncertainty but
remains consistent.

> (7) Massarotti et al. are not the only ones who report a low metallicity for
> Aldebaran, Liu et al. 2017 report [Fe/H]=-0.27. Sharma et al. (2016) claim
> -0.16. The choice of values in Table~2 should therefore be justified.

The referee is correct that there is a large scatter in the reported metallicity
results for Aldebaran. The focus of this work is the new analysis of historical
RV data and how this allows new inference. The treatment this scatter of, or at
least justification of the selected metallicity, is beyond the scope of this
work. In the text we already indicate that there is a large scatter and hence we
select a number of metallicity values. While we agree there is little science
behind this selection it is still broadly representative and should allow
readers to interpret the impact of metallicity.

--------
Data Editor's comments:

The authors should provide the RV data shown in Figure 3 as Data behind the
Figure (DbF). The readers would surely appreciate having this important data set
available with the published article.

### 2018-08-16 paper/aldebaran.tex

@@
 the stellar properties as $M = 1.27^{+0.24}_{-0.20} \, \mathrm{M_{\odot}}$ and
-age $4.9^{+3.6}_{-2.0} \, \rm Gyr$.  With the inclusion of \numax we find $M =
-1.16\, \pm \, {0.07} \, \mathrm{M_{\odot}}$ and age $6.4^{+1.4}_{-1.1} \, \rm
-Gyr$.  It is clear that the addition of \numax provides substantially more
-precise estimates of mass and age. With this stellar mass, the radial velocity
-signal translates to an inclination-dependent planetary mass of $M\,\sin\,{i} =
-5.8 \pm 0.7~M_J$ \citep[Equation~1]{Torres2008}, a massive giant planet; for
-sufficiently high inclinations ($\gtrsim 45^\circ$), the mass of Aldebaran~b may
-exceed $13~M_J$, making it a possible low-mass brown dwarf candidate. A recent
-study by \citet{2018arXiv180105239H} has cast doubt on the validity of the
-detection of the massive planet $\gamma$~Draconis~b, and by extension similar
-planets around similar K~giants such as Aldebaran. We remain convinced that the
+age $4.9^{+3.6}_{-2.0} \, \rm Gyr$.  This value and uncertainty are consistent
+with those obtained from other non-asteroseismic analyses of Aldebaran
+\cite[e.g.][]{Abia2012,Hatzes2015}. With the inclusion of \numax we find $M = 1.16\, \pm
+\, {0.07} \, \mathrm{M_{\odot}}$ and age $6.4^{+1.4}_{-1.1} \, \rm Gyr$.  It is
+clear that the addition of \numax provides substantially more precise estimates
+of mass and age. With this stellar mass, the radial velocity signal translates
+to an inclination-dependent planetary mass of $M\,\sin\,{i} = 5.8 \pm 0.7~M_J$
+\citep[Equation~1]{Torres2008}, a massive giant planet; for sufficiently high
+inclinations ($\gtrsim 45^\circ$), the mass of Aldebaran~b may exceed $13~M_J$,
+making it a possible low-mass brown dwarf candidate. A recent study by
+\citet{2018arXiv180105239H} has cast doubt on the validity of the detection of
+the massive planet $\gamma$~Draconis~b, and by extension similar planets around
+similar K~giants such as Aldebaran. We remain convinced that the
 $\gamma$~Draconis problem is not an issue in this case, and Aldebaran~b is a
@@
 
-We would like to thank Tim Bedding, David Hogg and Dan Foreman-Mackey for their helpful discussions.
+We would like to thank Tim Bedding, David Hogg and Dan Foreman-Mackey for their
+helpful discussions.  We would like to thank the anonymous referee for their
+efforts and their comments.
 
